### 2022-11-28 src/tex/ms.tex
[showyourwork] first commit

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@rmnldwg}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
Donec vehicula augue eu neque, morbi tristique senectus et netus et.
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
Phasellus eu tellus sit amet tortor gravida placerat.
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
Praesent eget sem vel leo ultrices bibendum.
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
Curabitur auctor semper nulla donec varius orci eget risus.
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2022-11-28 src/scripts/corner_plot_base_graph.py
add test plot and get everything connected

@@ -1,0 +1,33 @@
+"""
+Render a corner plot of all 1D and 2D marginals of the sampled posterior distribution
+over the spread parameters of the base graph model.
+"""
+import matplotlib.pyplot as plt
+from corner import corner
+from emcee.backends import HDFBackend
+
+import paths
+
+
+class Namespace:
+    """Make script runnable without snakemake by creating a simple namespace."""
+    def __init__(self, **kwargs) -> None:
+        self.__dict__.update(**kwargs)
+
+
+if "snakemake" not in locals():
+    snakemake = Namespace(
+        input=paths.data / "extended-base-v1-samples.hdf5",
+        thin=10,
+        output=paths.figures / "extended-base-v1-corner.png"
+    )
+
+
+if __name__ == "__main__":
+    # open samples
+    backend = HDFBackend(paths.data / snakemake.input, read_only=True)
+    samples = backend.get_chain(flat=True, thin=snakemake.thin)
+
+    # create plot
+    corner(samples)
+    plt.savefig(snakemake.output, dpi=300)

### 2022-11-28 src/tex/ms.tex
add test plot and get everything connected
import bibliography & write abstract draft

@@
 \usepackage{showyourwork}
+
+% Glossaries
+\usepackage[acronym,toc]{glossaries}
+\newcommand*\myglsentry[1]{%
+  \protect\ifglsused{#1}{%
+    \glsentryshort{#1}%
+  }{%
+    \glsentrylong{#1}%
+  }%
+}
+\newacronym{scc}{SCC}{squamous cell carcinoma}
+\newacronym{hnscc}{HNSCC}{head and neck \myglsentry{scc}}
+\newacronym{opscc}{OPSCC}{oropharyngeal \myglsentry{scc}}
+\newacronym{ctv}{CTV}{clinical target volume}
+\newacronym{ctv-n}{CTV-N}{elective \myglsentry{ctv}}
+\newacronym{lnl}{LNL}{lymph node level}
+\newacronym{hmm}{HMM}{hidden Markov model}
+\newacronym{rv}{RV}{random variable}
+\newacronym{dag}{DAG}{directed acyclic graph}
+\newacronym{mcmc}{MCMC}{Markov chain Monte Carlo}
+\newacronym{ti}{TI}{thermodynamic integration}
+\newacronym{bic}{BIC}{Bayesian information criterion}
 
@@
 % Title
-\title{An open source scientific article}
+\title{Modelling the Lymphatic Metastatic Progression Pathways of OPSCC from Multi-Institutional Datasets}
 
 % Author list
-\author{@rmnldwg}
+\author{Roman Ludwig}
+\author{Jean-Marc Hoffmann}
+\author{Bertrand Pouymayou}
+\author{Panagiotis Balermpas}
+\author{Lauence Bauwens}
+\author{Vincent Grégoire}
+\author{Roland Giger}
+\author{Jan Unkelbach}
 
@@
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-    Phasellus eu tellus sit amet tortor gravida placerat.
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-    Praesent eget sem vel leo ultrices bibendum.
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-    Curabitur auctor semper nulla donec varius orci eget risus.
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-    Duis eget orci sit amet orci dignissim rutrum.
+    The \gls{ctv-n} definition in \gls{hnscc} is currently based mostly on the prevalence of lymphatic metastases in different \glspl{lnl} based on the primary tumor location.
+In this work, we present an extension to a probabilistic model for lymphatic metastatic spread developed earlier that can quantify the risk for microscopic nodal involvement based on individual diagnoses.
+The extension is based on the same formalism of \glspl{hmm} as the original model, but in addition to the \glspl{lnl} I, II, III, and IV, it also covers the \glspl{lnl} V and VII of the ipsilateral neck.
+Moreover, we infer which pathways of lymphatic spread to model from clinical lymphatic progression patterns of 681 patients from three institutions.
+The extended model may allow for a more personalized \gls{ctv-n} definition based on a patient's individual state of disease.
+The \gls{hmm} uses a collection of hidden binary \glspl{rv} -- one for each \gls{lnl} -- to model a patient's state of lymphatic involvement.
+Clinical diagnoses and/or pathological examinations of resected \glspl{lnl} represent the observed binary \glspl{rv} corresponding to the unobservable true state of nodal disease.
+A \gls{dag} with parametrized edges is used to compute the transition matrix of the \gls{hmm} and consequently the (log-)likelihood of the diagnoses of 681 patients, given a set of spread parameters.
+Based on the defined likelihood function, \gls{mcmc} sampling is then used to infer a distribution over the model's parameters.
+T-category is model by assuming that, on average, patients with late T-category tumors are diagnosed at a later time compared to patients with early T-category tumors.
+Using \gls{ti} and the \gls{bic} we compare models based on different \glspl{dag} and demonstrate the accuracy and precision of the best-performing model.
+When extended to the contralateral side, this model may form the basis of future guidelines of elective \gls{ctv} definition.
 \end{abstract}
@@
 Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
-Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
-Pellentesque cursus luctus mauris \citep{Luger2021}.
+Aliquam tincidunt urna, nulla ullamcorper vestibulum.
+Pellentesque cursus luctus mauris \cite{luger_mapping_2021}.
+
+\section{Previous Work}
+\label{sec:previous_work}
+
+The \gls{hmm} presented in this work is based on the
+
+\begin{figure}
+    \script{corner_plot_base_graph.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/extended-base-v1-corner.png}
+        \caption{Testing this showyourwork thingy}
+        \label{fig:corner_plot_base_graph}
+    \end{centering}
+\end{figure}
 

### 2022-11-29 src/scripts/base_graph_corner_plot.py
rename some files

@@ -2,6 +2,9 @@
 Render a corner plot of all 1D and 2D marginals of the sampled posterior distribution
 over the spread parameters of the base graph model.
 """
+import warnings
+from pathlib import Path
+
 import matplotlib.pyplot as plt
 from corner import corner
 from emcee.backends import HDFBackend
@@ -15,12 +18,13 @@ class Namespace:
         self.__dict__.update(**kwargs)
 
 
-if "snakemake" not in locals():
-    snakemake = Namespace(
-        input=paths.data / "extended-base-v1-samples.hdf5",
-        thin=10,
-        output=paths.figures / "extended-base-v1-corner.png"
-    )
+# if "snakemake" not in locals():
+#     warnings.warn("Snakemake did not provide params.")
+#     snakemake = Namespace(
+#         input=paths.data / "base_graph_samples.hdf5",
+#         thin=10,
+#         output=paths.figures / Path(__file__).with_suffix(".png")
+#     )
 
 
 if __name__ == "__main__":

### 2022-11-29 src/scripts/base_graph_corner_plot.py
try to add another figure

@@ -2,7 +2,6 @@
 Render a corner plot of all 1D and 2D marginals of the sampled posterior distribution
 over the spread parameters of the base graph model.
 """
-import warnings
 from pathlib import Path
 
 import matplotlib.pyplot as plt
@@ -12,26 +11,16 @@ from emcee.backends import HDFBackend
 import paths
 
 
-class Namespace:
-    """Make script runnable without snakemake by creating a simple namespace."""
-    def __init__(self, **kwargs) -> None:
-        self.__dict__.update(**kwargs)
-
-
-# if "snakemake" not in locals():
-#     warnings.warn("Snakemake did not provide params.")
-#     snakemake = Namespace(
-#         input=paths.data / "base_graph_samples.hdf5",
-#         thin=10,
-#         output=paths.figures / Path(__file__).with_suffix(".png")
-#     )
+INPUT = paths.data / "base_graph_samples.hdf5"
+THIN = 10
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
 
 
 if __name__ == "__main__":
     # open samples
-    backend = HDFBackend(paths.data / snakemake.input, read_only=True)
-    samples = backend.get_chain(flat=True, thin=snakemake.thin)
+    backend = HDFBackend(paths.data / INPUT, read_only=True)
+    samples = backend.get_chain(flat=True, thin=THIN)
 
     # create plot
     corner(samples)
-    plt.savefig(snakemake.output, dpi=300)
+    plt.savefig(OUTPUT, dpi=300)

### 2022-11-29 src/scripts/conn12_prevalences_low.py

@@ -1,0 +1,49 @@
+"""
+Plot histograms of predicted and posteriors over observed prevalences for high-risk
+scenarios as computed by models with different arcs between LNL I and II.
+"""
+from pathlib import Path
+
+import matplotlib.pyplot as plt
+from lyscripts.plot.utils import Histogram, Posterior, draw, COLORS, get_size
+
+import paths
+from constants import linewidth_in_cm
+
+
+BASE_GRAPH = paths.data / "base_graph_prevalences.hdf5"
+ADD12 = paths.data / "add12_prevalences.hdf5"
+ADD21 = paths.data / "add21_prevalences.hdf5"
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
+
+
+if __name__ == "__main__":
+    fig, axes = plt.subplots(figsize=get_size(width=linewidth_in_cm))
+
+    early_12_panel = [
+        Histogram.from_hdf5(
+            filename=BASE_GRAPH,
+            dataname="I/early",
+            color="gray",
+            hatch=r"////",
+        ),
+        Histogram.from_hdf5(
+            filename=BASE_GRAPH,
+            dataname="InotII/early",
+            color="gray",
+            hatch=r"\\\\",
+        ),
+        Histogram.from_hdf5(
+            filename=ADD12,
+            dataname="I/early",
+            color=COLORS["blue"],
+        ),
+        Histogram.from_hdf5(
+            filename=ADD12,
+            dataname="InotII/early",
+            color=COLORS["green"],
+        ),
+    ]
+
+    draw(axes, contents=early_12_panel)
+    plt.savefig(OUTPUT, dpi=400)

### 2022-11-29 src/scripts/constants.py

@@ -1,0 +1,6 @@
+"""
+Definition of some constants that might be useful.
+"""
+linewidth_in_pt = 242.26653
+cm_per_pt = 0.03528
+linewidth_in_cm = linewidth_in_pt * cm_per_pt

### 2022-11-29 src/scripts/conn12_prevalences_low.py
use tueplots for pyplot's rcParams settings

@@ -6,6 +6,7 @@ from pathlib import Path
 
 import matplotlib.pyplot as plt
 from lyscripts.plot.utils import Histogram, Posterior, draw, COLORS, get_size
+from tueplots import bundles
 
 import paths
 from constants import linewidth_in_cm
@@ -18,6 +19,7 @@ OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
 
 
 if __name__ == "__main__":
+    plt.rcParams.update(bundles.icml2022())
     fig, axes = plt.subplots(figsize=get_size(width=linewidth_in_cm))
 
     early_12_panel = [
@@ -26,12 +28,18 @@ if __name__ == "__main__":
             dataname="I/early",
             color="gray",
             hatch=r"////",
+            histtype="step",
+            linewidth=1.5,
+            alpha=1.,
         ),
         Histogram.from_hdf5(
             filename=BASE_GRAPH,
             dataname="InotII/early",
             color="gray",
             hatch=r"\\\\",
+            histtype="step",
+            linewidth=1.5,
+            alpha=1.,
         ),
         Histogram.from_hdf5(
             filename=ADD12,
@@ -46,4 +54,4 @@ if __name__ == "__main__":
     ]
 
     draw(axes, contents=early_12_panel)
-    plt.savefig(OUTPUT, dpi=400)
+    plt.savefig(OUTPUT, dpi=300)

### 2022-11-30 src/tex/ms.tex
[showyourwork] trick SYW to sync with Overleaf
[showyourwork] overleaf sync

@@
 \usepackage{showyourwork}
+\usepackage{multirow}
 
@@
 \newcommand*\myglsentry[1]{%
-  \protect\ifglsused{#1}{%
+  \protected\ifglsused{#1}{%
     \glsentryshort{#1}%
@@
 \newacronym{bic}{BIC}{Bayesian information criterion}
+\newacronym{bn}{BN}{Bayesian network}
+\newacronym{ct}{CT}{computed tomography}
+\newacronym{mri}{MRI}{magnetic resonance imaging}
+\newacronym{pet}{PET}{positron emission tomography}
+\newacronym{fna}{FNA}{fine needle aspiration}
 
@@
 
-% Abstract with filler text
+% Abstract
 \begin{abstract}
-    The \gls{ctv-n} definition in \gls{hnscc} is currently based mostly on the prevalence of lymphatic metastases in different \glspl{lnl} based on the primary tumor location.
+    The \gls{ctv-n}\glsunset{ctv} definition in \gls{hnscc}\glsunset{scc} is currently based mostly on the prevalence of lymphatic metastases in different \glspl{lnl} based on the primary tumor location.
 In this work, we present an extension to a probabilistic model for lymphatic metastatic spread developed earlier that can quantify the risk for microscopic nodal involvement based on individual diagnoses.
@@
 
-% Main body with filler text
+% Main body
 \section{Introduction}
@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-Phasellus eu tellus sit amet tortor gravida placerat.
-Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-Praesent eget sem vel leo ultrices bibendum.
-Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-Curabitur auctor semper nulla donec varius orci eget risus.
-Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
+When treating cancer, either with radiotherapy or via surgery, the aim is to irradiate or resect as much of the present malignancies as possible.
+This includes the primary tumor mass and metastases that are sufficiently large to be detected using in-vivo imaging modalities such as \gls{ct}, \gls{mri}, or \gls{pet}.
+But it also includes regions of possible microscopic spread, which these modalities cannot detect, to increase the patient's probability of cure \cite{poortmans_internal_2020,murthy_prostate-only_2021}.
+Microscopic disease can currently only be observed by a pathological examination of the tissue.
+Consequently, clinicians that are presented with cancer patients need to routinely assess the risk of microscopic involvement in parts of the body that appear unsuspicious.
 
-Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
-Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
-Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
-Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
-Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
-Aliquam tincidunt urna, nulla ullamcorper vestibulum.
-Pellentesque cursus luctus mauris \cite{luger_mapping_2021}.
+Since in this work we consider \gls{hnscc}, which frequently spreads through the lymphatic system to form regional nodal metastases,
 
@@
 
-The \gls{hmm} presented in this work is based on the
+We have introduced a probabilistic model for lymphatic metastatic tumor progression based on \glspl{bn} in \cite{pouymayou_bayesian_2019} and based off of this an improved model using \glspl{hmm} in \cite{ludwig_hidden_2021}.
+We will briefly recap the \acrlong{hmm} below.
+
+A patient's state of (hidden) lymphatic involvement at time $t$ is described as a collection of binary \glspl{rv}, one for each of the $V$ \glspl{lnl}:
+%
+\begin{equation}
+    \mathbf{X}[t] = \left( X_v[t] \right) \qquad v \in \left\{ 1,2, \ldots, V \right\}
+\end{equation}
+%
+Where each of the \glspl{lnl} can be in the state $X_v=0$ (\texttt{FALSE}), meaning \gls{lnl} $v$ is healthy, or in the state $X_v=1$ (\texttt{TRUE}), indicating the \gls{lnl} harbors metastases.
+
+The transition from one time-step to another is governed by the transition probability $P\left( \mathbf{X}[t+1]=\boldsymbol{\xi}_i \mid \mathbf{X}[t]=\boldsymbol{\xi}_j \right)$, which can conveniently be collected into a transition matrix when we enumerate all $2^V$ distinct possible states $\boldsymbol{\xi}_i$ with $i \in \left\{ 1,2, \ldots, 2^V \right\}$ of lymphatic involvement:
+%
+\begin{equation}
+    \mathbf{A} = \left( A_{ij} \right) = \left( P\left( \mathbf{X}[t+1]=\boldsymbol{\xi}_i \mid \mathbf{X}[t]=\boldsymbol{\xi}_j \right) \right)
+\end{equation}
+%
+Diagnosis and true state of a patient are formally connected via the sensitivity $s_N$ and specificity $s_P$ of the used diagnostic modality.
+In clinical practice, these modalities are \gls{ct}, \gls{mri}, or \gls{pet} scan, but it may also include information from biopsies after a \gls{fna} or other techniques to detect lymphatic metastases.
+For each \gls{lnl} $v$ the conditional probability table of $P\left( Z_v \mid X_v \right)$ looks like this:
+
+\begin{center}
+\begin{tabular}{|cc|cc|}
+    \hline
+    & & \multicolumn{2}{c|}{$X$} \\
+    & & 0 & 1 \\
+    \hline
+    \multirow{2}{*}{$Z$} & 0 & $s_p$ & $1 - s_N$ \\
+    & 1 & $1 - s_P$ & $s_N$ \\
+    \hline
+\end{tabular}
+\end{center}
+
+Consequently, the conditional probability to observe a diagnosis $\mathbf{Z}=\boldsymbol{\zeta}_\ell$, given a hidden involvement state $\mathbf{X}=\boldsymbol{\xi}_k$ is a matrix $\mathbf{B}$ made up of products of terms from the table above:
+%
+\begin{equation}
+    \mathbf{B} = \left( B_{k\ell} \right) = \prod_{v=1}^V P\left( Z_v = \zeta_{\ell v} \mid X_v[t_\text{D}] = \xi_{kv} \right)
+\end{equation}
+%
+We define the time $t=0$ to be the moment just before a patient's tumor formed, and hence $X_v[t=0]=0 \,\,\, \forall v$.
+However, using this definition, we cannot know how many time-steps have passed until $t_\text{D}$, when the patient was diagnosed with cancer.
+We can only make the assumption that a patients with an earlier T-category tumor was \emph{probably} diagnosed after fewer time-steps than a patient with a late T-category tumor.
+We can use this information by marginalizing over the diagnose times $t_\text{D}$ of patients in different T-categories using different prior distributions over the diagnose time.
+E.g., $P_\text{early}\left( t_\text{D} \right)$ for early T-category patients (T1 \& T2) and $P_\text{late}\left( t_\text{D} \right)$ for late T-category patients (T3 \& T4).
+
+The main task of of personalizing the \gls{ctv-n} definition is now to infer the probability of each of the possible states $\boldsymbol{\xi}_k$ given a diagnosis $\mathbf{Z}=\boldsymbol{\zeta}_\ell$.
+Using Bayes' theorem, we get
+%
+\begin{equation}
+    P\left( \mathbf{X}=\boldsymbol{\xi}_k \mid \mathbf{Z}=\boldsymbol{\zeta}_\ell \right) = \frac{P\left( \boldsymbol{\zeta}_\ell \mid \boldsymbol{\xi}_k \right) P\left( \boldsymbol{\xi}_k \right)}{\sum_{r=1}^{2^V} P\left( \boldsymbol{\zeta}_\ell \mid \boldsymbol{\xi}_r \right) P\left( \boldsymbol{\xi}_r \right) }
+\end{equation}
+%
+The model we are describing here will yield an estimate of $P\left( \boldsymbol{\xi}_k \right)$
 
 \begin{figure}
-    \script{corner_plot_base_graph.py}
+    \script{conn12_prevalences_low.py}
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/extended-base-v1-corner.png}
+        \includegraphics[width=\textwidth]{figures/conn12_prevalences_low.png}
         \caption{Testing this showyourwork thingy}
-        \label{fig:corner_plot_base_graph}
+        \label{fig:conn12_prevalences_low}
     \end{centering}

### 2022-11-30 src/scripts/conn12_prevalences_low.py
finish low-prev plot about 12 vs 21 connection

@@ -4,54 +4,94 @@ scenarios as computed by models with different arcs between LNL I and II.
 """
 from pathlib import Path
 
+import numpy as np
 import matplotlib.pyplot as plt
-from lyscripts.plot.utils import Histogram, Posterior, draw, COLORS, get_size
-from tueplots import bundles
+from lyscripts.plot.utils import Histogram, Posterior, draw, COLORS
+from tueplots import figsizes, fontsizes
 
 import paths
-from constants import linewidth_in_cm
 
 
 BASE_GRAPH = paths.data / "base_graph_prevalences.hdf5"
-ADD12 = paths.data / "add12_prevalences.hdf5"
-ADD21 = paths.data / "add21_prevalences.hdf5"
+ADD = {
+    "12": paths.data / "add12_prevalences.hdf5",
+    "21": paths.data / "add21_prevalences.hdf5",
+}
 OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
 
 
 if __name__ == "__main__":
-    plt.rcParams.update(bundles.icml2022())
-    fig, axes = plt.subplots(figsize=get_size(width=linewidth_in_cm))
+    plt.rcParams.update(figsizes.icml2022_full(nrows=2, ncols=2))
+    plt.rcParams.update(fontsizes.icml2022())
+    fig, axes = plt.subplots(nrows=2, ncols=2, sharex="col", sharey="row")
 
-    early_12_panel = [
-        Histogram.from_hdf5(
-            filename=BASE_GRAPH,
-            dataname="I/early",
-            color="gray",
-            hatch=r"////",
-            histtype="step",
-            linewidth=1.5,
-            alpha=1.,
-        ),
-        Histogram.from_hdf5(
-            filename=BASE_GRAPH,
-            dataname="InotII/early",
-            color="gray",
-            hatch=r"\\\\",
-            histtype="step",
-            linewidth=1.5,
-            alpha=1.,
-        ),
-        Histogram.from_hdf5(
-            filename=ADD12,
-            dataname="I/early",
-            color=COLORS["blue"],
-        ),
-        Histogram.from_hdf5(
-            filename=ADD12,
-            dataname="InotII/early",
-            color=COLORS["green"],
-        ),
-    ]
+    panels = {
+        "early": {"12": [], "21": []},
+        "late": {"12": [], "21": []},
+    }
 
-    draw(axes, contents=early_12_panel)
+    for stage in ["early", "late"]:
[…]

### 2022-12-05 src/scripts/base_graph_prevalences.py
compute prevalences now here, not in `lynference`

@@ -1,0 +1,120 @@
+"""
+Plot histograms of predicted and posteriors over observed prevalences for basic
+scenarios involving the LNLs II, III and IV. This is to show that the base graph is
+already capable of capturing the frequency of these involvement, as well as their
+correlations.
+"""
+from fileinput import filename
+from pathlib import Path
+
+import matplotlib.pyplot as plt
+from lyscripts.plot.utils import Histogram, Posterior, draw, COLORS
+from tueplots import figsizes, fontsizes
+
+import paths
+
+
+INPUT = paths.data / "base_graph_prevalences.hdf5"
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
+NROWS, NCOLS = 2, 1
+
+
+if __name__ == "__main__":
+    plt.rcParams.update(figsizes.icml2022_full(
+        nrows=NROWS,
+        ncols=NCOLS,
+        height_to_width_ratio=0.3,
+    ))
+    plt.rcParams.update(fontsizes.icml2022())
+    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
+
+    panels = {
+        "early": [],
+        "late": [],
+    }
+
+    for stage in ["early", "late"]:
+        panels[stage] = [
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"II/{stage}",
+                label="LNL II overall",
+                color=COLORS["red"],
+            ),
+            Posterior.from_hdf5(
+                filename=INPUT,
+                dataname=f"II/{stage}",
+                color=COLORS["red"],
+            ),
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"III/{stage}",
+                label="LNL III overall",
+                color=COLORS["blue"],
+            ),
+            Posterior.from_hdf5(
+                filename=INPUT,
+                dataname=f"III/{stage}",
+                color=COLORS["blue"],
+            ),
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"IV/{stage}",
+                label="LNL IV overall",
+                color=COLORS["green"],
+            ),
+            Posterior.from_hdf5(
+                filename=INPUT,
+                dataname=f"IV/{stage}",
+                color=COLORS["green"],
+            ),
+        ]
+
[…]

### 2022-12-05 src/scripts/conn12_prevalences_low.py

@@ -1,97 +1,0 @@
-"""
-Plot histograms of predicted and posteriors over observed prevalences for high-risk
-scenarios as computed by models with different arcs between LNL I and II.
-"""
-from pathlib import Path
-
-import numpy as np
-import matplotlib.pyplot as plt
-from lyscripts.plot.utils import Histogram, Posterior, draw, COLORS
-from tueplots import figsizes, fontsizes
-
-import paths
-
-
-BASE_GRAPH = paths.data / "base_graph_prevalences.hdf5"
-ADD = {
-    "12": paths.data / "add12_prevalences.hdf5",
-    "21": paths.data / "add21_prevalences.hdf5",
-}
-OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
-
-
-if __name__ == "__main__":
-    plt.rcParams.update(figsizes.icml2022_full(nrows=2, ncols=2))
-    plt.rcParams.update(fontsizes.icml2022())
-    fig, axes = plt.subplots(nrows=2, ncols=2, sharex="col", sharey="row")
-
-    panels = {
-        "early": {"12": [], "21": []},
-        "late": {"12": [], "21": []},
-    }
-
-    for stage in ["early", "late"]:
-        for direction in ["12", "21"]:
-            panels[stage][direction] = [
-                Histogram.from_hdf5(
-                    filename=ADD[direction],
-                    dataname=f"I/{stage}",
-                    label="LNL I overall",
-                    color=COLORS["blue"],
-                ),
-                Histogram.from_hdf5(
-                    filename=BASE_GRAPH,
-                    dataname=f"I/{stage}",
-                    label="base graph",
-                    color="black",
-                    hatch=r"////",
-                    histtype="step",
-                    linewidth=1.5,
-                    alpha=1.,
-                ),
-                Posterior.from_hdf5(
-                    filename=ADD[direction],
-                    dataname=f"I/{stage}",
-                    color=COLORS["blue"],
-                ),
-                Histogram.from_hdf5(
-                    filename=ADD[direction],
-                    dataname=f"InotII/{stage}",
-                    label="LNL I without II",
-                    color=COLORS["green"],
-                ),
-                Histogram.from_hdf5(
-                    filename=BASE_GRAPH,
-                    dataname=f"InotII/{stage}",
-                    label="base graph",
[…]

### 2022-12-05 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \newcommand*\myglsentry[1]{%
-  \protected\ifglsused{#1}{%
+  \ifglsused{#1}{%
     \glsentryshort{#1}%
@@
 
+\subsection*{Lymphatic Spread as Hidden Markov Model}
+\label{subsec:previous_work:spread_as_hmm}
+
 We have introduced a probabilistic model for lymphatic metastatic tumor progression based on \glspl{bn} in \cite{pouymayou_bayesian_2019} and based off of this an improved model using \glspl{hmm} in \cite{ludwig_hidden_2021}.
@@
 %
+The term inside the brackets describes the probability to transition from the hidden state of lymphatic involvement $\boldsymbol{\xi}_j$ to the state $\boldsymbol{\xi}_i$ within between the time $t$ and $t+1$.
+We can use a \gls{dag} as depicted in
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{figures/graph_with_obs.png}
+    \caption{\Gls{dag} representing a possible abstraction of the lymphatic network comprising the tumor (red shaded circle) and \glspl{lnl} I through IV as hidden binary \glspl{rv} (blue outlined circles).
+Attached to each of these is the corresponding observed \gls{rv} (orange shaded squares).
+Lymphatic flow is depicted in the form of parameterized arrows (red and blue) that represent the probability of spread along the respective arc per time-step.
+Sensitivity and specificity connect the hidden \glspl{rv} to the diagnosis (orange arrows).}
+    \label{fig:graph_with_obs}
+\end{figure}
+
 Diagnosis and true state of a patient are formally connected via the sensitivity $s_N$ and specificity $s_P$ of the used diagnostic modality.
@@
 %
+If multiple diagnostic modalities $\mathcal{O} \in \{ \text{MRI}, \text{CT}, \ldots \}$ are used, one can compute an observation matrix $\mathbf{B}^\mathcal{O}$ for each of these modalities and combine them via an outer product into a joint observations matrix
+%
+\begin{equation}
+    \mathbf{B} = \mathbf{B}^\text{MRI} \otimes \mathbf{B}^\text{CT} \otimes \ldots
+\end{equation}
+%
 We define the time $t=0$ to be the moment just before a patient's tumor formed, and hence $X_v[t=0]=0 \,\,\, \forall v$.
@@
 %
-The model we are describing here will yield an estimate of $P\left( \boldsymbol{\xi}_k \right)$
+The model we are describing here will yield an estimate of $P\left( \boldsymbol{\xi}_k \right)$ and in the next subsection we will briefly explain how this probability distribution can be efficiently parameterized.
 
-\begin{figure}
-    \script{conn12_prevalences_low.py}
+\subsection*{Parameterization of the Hidden Markov Model}
+\label{subsec:previous_work:parametrization}
+
+blabla
+
+\begin{figure*}
+    \script{base_graph_prevalences.py}
     \begin{centering}
-        \includegraphics[width=\textwidth]{figures/conn12_prevalences_low.png}
+        \includegraphics[width=\textwidth]{figures/base_graph_prevalences.png}
         \caption{Testing this showyourwork thingy}
-        \label{fig:conn12_prevalences_low}
+        \label{fig:base_graph_prevalences}
     \end{centering}
-\end{figure}
+\end{figure*}
 

### 2022-12-05 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{multirow}
+\usepackage{hyperref}
+\usepackage{cleveref}
 
@@
 %
-The term inside the brackets describes the probability to transition from the hidden state of lymphatic involvement $\boldsymbol{\xi}_j$ to the state $\boldsymbol{\xi}_i$ within between the time $t$ and $t+1$.
-We can use a \gls{dag} as depicted in
-
+The term $P\left( \boldsymbol{\xi}_i \mid \boldsymbol{\xi}_j \right)$ describes the probability to transition from the hidden state of lymphatic involvement $\boldsymbol{\xi}_j$ to the state $\boldsymbol{\xi}_i$ within between the time $t$ and $t+1$.
+Using a \gls{dag} as depicted in \cref{fig:graph_with_obs}, we can parameterize this transition probability in the following way:
+%
+\begin{equation}
+    P\left( \boldsymbol{\xi}_i \mid \boldsymbol{\xi}_j \right) = \prod_{v \leq V} Q\left( \xi_{iv} ; \xi_{jv} \right) P \left( \xi_{iv} \mid \left\{ \xi_{jr} \right\}_{r \in \operatorname{pa}(v)} \right)^{1 - \xi_{jv}}
+\end{equation}
+%
+In this equation, we have denoted \glspl{lnl} that are parents of \gls{lnl} $v$ with the symbol $r\in\operatorname{pa}(v)$.
+Also, $\xi_{iv}$ denotes the value that \gls{lnl} $v$ takes on when the patient is in state $\boldsymbol{\xi}_i$.
 \begin{figure}
     \centering
-    \includegraphics[width=\linewidth]{figures/graph_with_obs.png}
+    \includegraphics[width=0.6\linewidth]{figures/graph_with_obs.png}
     \caption{\Gls{dag} representing a possible abstraction of the lymphatic network comprising the tumor (red shaded circle) and \glspl{lnl} I through IV as hidden binary \glspl{rv} (blue outlined circles).
@@
 Lymphatic flow is depicted in the form of parameterized arrows (red and blue) that represent the probability of spread along the respective arc per time-step.
-Sensitivity and specificity connect the hidden \glspl{rv} to the diagnosis (orange arrows).}
+Sensitivity and specificity (orange arrows) connect the hidden \glspl{rv} to the diagnosis.}
     \label{fig:graph_with_obs}
@@
 
-\begin{center}
+\begin{table}[h!]
+    \centering
     \begin{tabular}{|cc|cc|}
@@
     \end{tabular}
-\end{center}
+\end{table}
 

### 2022-12-09 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{multirow}
-\usepackage{hyperref}
-\usepackage{cleveref}
+
+%% References
+\usepackage[%
+    style=numeric-comp,%
+    sorting=none,%
+    maxnames=20,%
+    url=false,%
+]{biblatex}
+\addbibresource{bib.bib}
 
@@
 \newacronym{fna}{FNA}{fine needle aspiration}
+
+% Cross-referencing
+\usepackage{hyperref}
+\AtEndPreamble{\usepackage{cleveref}}
 
@@
 
-When treating cancer, either with radiotherapy or via surgery, the aim is to irradiate or resect as much of the present malignancies as possible.
+When treating \gls{hnscc}, either with radiotherapy or via surgery, the aim is to irradiate or resect as much of the present malignancies as possible.
 This includes the primary tumor mass and metastases that are sufficiently large to be detected using in-vivo imaging modalities such as \gls{ct}, \gls{mri}, or \gls{pet}.
@@
 Microscopic disease can currently only be observed by a pathological examination of the tissue.
-Consequently, clinicians that are presented with cancer patients need to routinely assess the risk of microscopic involvement in parts of the body that appear unsuspicious.
+Consequently, clinicians that are presented with cancer patients need to routinely assess the risk of microscopic involvement in parts of the body that appear clinically node negative.
 
-Since in this work we consider \gls{hnscc}, which frequently spreads through the lymphatic system to form regional nodal metastases,
+This work concerns itself with \gls{opscc}, which commonly spreads lymphatically.
+The only two publicly available datasets on nodal involvement patterns in \gls{opscc}, at the time of writing, show that 80\% of patients where either cN+ or -- if available -- pN+ \cite{ludwig_dataset_2022}.
+Since the exact number and location of lymph nodes varies from patient to patient, the nodes are grouped into anatomically defined \glspl{lnl}.
+These levels are then often irradiated or resected prophylactically due to the risk of harboring occult metastases despite negative findings from imaging modalities.
+In current clinical practice, \gls{ctv-n} definition is instead mostly based on guidelines \cite{gregoire_ct-based_2003,gregoire_delineation_2014,gregoire_delineation_2018,eisbruch_intensity-modulated_2002,biau_selection_2019,chao_determination_2002,vorwerk_guidelines_2011,ferlito_elective_2009} that are derived from the observed prevalence of involvement in an \gls{lnl} for a given tumor location.
+
+These guidelines, however, do not account for the personal risk of the patients that may depend greatly on their diagnosis.
+E.g., macroscopic metastases detected via \gls{pet} in the \glspl{lnl} II and III may increase the risk for occult disease in \gls{lnl} IV over the case of a patient who presents with a clinically N0 neck.
+
+A model for estimating the risk of microscopic disease, given a personal diagnosis based on \glspl{bn} was previously developed \cite{pouymayou_bayesian_2019} and subsequently adapted to make use of \glspl{hmm} to include T-category in an intuitive manner \cite{ludwig_hidden_2021}.
+These models were trained on a dataset of early T-category (T1 and T2) patients that could be reconstructed from \cite{sanguineti_defining_2009}.
+Since this publication only includes details about the involvement of the ipsilateral \glspl{lnl} I, II, III, and IV, the models were limited to these nodal levels as well.
+
+With the publication of larger and more detailed datasets \cite{ludwig_dataset_2022}, the risk estimation models may be extended as well to include the additionally reported \glspl{lnl} V and VII.
+This work aims to do that and at the same time demonstrate the capability
+of the previously introduced \gls{hmm} to adapt to real patient data from a multi-centric cohort of XXX \gls{opscc} patients.
 
@@
 
-\subsection*{Lymphatic Spread as Hidden Markov Model}
+\subsection{Lymphatic Spread as Hidden Markov Model}
 \label{subsec:previous_work:spread_as_hmm}
@@
 \begin{equation}
+    \label{eq:transition_prob}
     P\left( \boldsymbol{\xi}_i \mid \boldsymbol{\xi}_j \right) = \prod_{v \leq V} Q\left( \xi_{iv} ; \xi_{jv} \right) P \left( \xi_{iv} \mid \left\{ \xi_{jr} \right\}_{r \in \operatorname{pa}(v)} \right)^{1 - \xi_{jv}}
@@
 Also, $\xi_{iv}$ denotes the value that \gls{lnl} $v$ takes on when the patient is in state $\boldsymbol{\xi}_i$.
+Lastly, the term $Q(a;b) \in \{ 0,1 \}$ is there only to
+
 \begin{figure}
@@
 
-\subsection*{Parameterization of the Hidden Markov Model}
+\subsection{Parameterization of the Hidden Markov Model}
 \label{subsec:previous_work:parametrization}
 
-blabla
+The term $P \big( X_v = \xi_{iv} \mid \left\{ X_r = \xi_{jr} \right\}_{r \in \operatorname{pa}(v)} \big)$ in \cref{eq:transition_prob} implicitly depends on how we parameterize the arcs of the underlying \gls{dag}.
+If we use the graph displayed in \cref{fig:graph_with_obs}), we can write down a conditional probability table for the probability of $X_3$ being healthy or metastatic, based on the state of the parent \gls{lnl} $\operatorname{pa}(3) = 2$:
+
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{|cc|cc|}
+        \hline
+        & & \multicolumn{2}{c|}{$X_2$} \\
+        & & 0 & 1 \\
+        \hline
+        \multirow{2}{*}{$X_3$} & 0 & $1 - b_3$ & $(1 - b_3)(1 - t_{23})$ \\
+        & 1 & $b_3$ & $1 - b_3 - t_{23} + b_3 t_{23}$ \\
+        \hline
+    \end{tabular}
+\end{table}
 
@@
 
-\bibliography{bib}
+\printbibliography[%
+    title={References},%
+]
 

### 2022-12-09 src/scripts/base_graph_prevalences.py
forgot one label in base_graph_prevalences plot

@@ -115,6 +115,6 @@ if __name__ == "__main__":
     draw(axes[1], contents=panels["late"], xlims=(0., 85.), hist_kwargs={"nbins": 120})
     axes[1].set_ylabel("late T-cat.", fontweight="bold")
     axes[1].set_xlabel("prevalence [%]")
-    axes[1].legend(handles=axes[1].get_children()[6:10],)
+    axes[1].legend(handles=axes[1].get_children()[5:10],)
 
     plt.savefig(OUTPUT, dpi=300)

### 2022-12-20 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{multirow}
-
-%% References
-\usepackage[%
-    style=numeric-comp,%
-    sorting=none,%
-    maxnames=20,%
-    url=false,%
-]{biblatex}
-\addbibresource{bib.bib}
 
@@
 %
-\begin{equation}
+\begin{equation} \label{eq:transition_matrix}
     \mathbf{B} = \left( B_{k\ell} \right) = \prod_{v=1}^V P\left( Z_v = \zeta_{\ell v} \mid X_v[t_\text{D}] = \xi_{kv} \right)
@@
 We can use this information by marginalizing over the diagnose times $t_\text{D}$ of patients in different T-categories using different prior distributions over the diagnose time.
-E.g., $P_\text{early}\left( t_\text{D} \right)$ for early T-category patients (T1 \& T2) and $P_\text{late}\left( t_\text{D} \right)$ for late T-category patients (T3 \& T4).
-
-The main task of of personalizing the \gls{ctv-n} definition is now to infer the probability of each of the possible states $\boldsymbol{\xi}_k$ given a diagnosis $\mathbf{Z}=\boldsymbol{\zeta}_\ell$.
-Using Bayes' theorem, we get
+E.g., $P \left( t=t_D \mid \text{early} \right)$ for early T-category patients (T1 \& T2) and $P \left( t=t_D \mid \text{late} \right)$ for late T-category patients (T3 \& T4).
+Throughout this work we will use binomial distributions for these probability mass functions, mainly because they have a plausible shape for this purpose and only a single parameter.
 %
 \begin{equation}
-    P\left( \mathbf{X}=\boldsymbol{\xi}_k \mid \mathbf{Z}=\boldsymbol{\zeta}_\ell \right) = \frac{P\left( \boldsymbol{\zeta}_\ell \mid \boldsymbol{\xi}_k \right) P\left( \boldsymbol{\xi}_k \right)}{\sum_{r=1}^{2^V} P\left( \boldsymbol{\zeta}_\ell \mid \boldsymbol{\xi}_r \right) P\left( \boldsymbol{\xi}_r \right) }
+    P \left( t = t_D \mid \text{T-cat.} \right) = \operatorname{B}(t_\text{max},p_\text{T-cat.})
 \end{equation}
 %
-The model we are describing here will yield an estimate of $P\left( \boldsymbol{\xi}_k \right)$ and in the next subsection we will briefly explain how this probability distribution can be efficiently parameterized.
+We will use as a latest time-step $t_\text{max} = 10$.
 
-\subsection{Parameterization of the Hidden Markov Model}
-\label{subsec:previous_work:parametrization}
+\subsection{The Likelihood Function}
+\label{subsec:previous_work:likelihood}
 
-The term $P \big( X_v = \xi_{iv} \mid \left\{ X_r = \xi_{jr} \right\}_{r \in \operatorname{pa}(v)} \big)$ in \cref{eq:transition_prob} implicitly depends on how we parameterize the arcs of the underlying \gls{dag}.
-If we use the graph displayed in \cref{fig:graph_with_obs}), we can write down a conditional probability table for the probability of $X_3$ being healthy or metastatic, based on the state of the parent \gls{lnl} $\operatorname{pa}(3) = 2$:
+Using the definitions up to this point, we can compute a vector of likelihoods for every possible diagnosis:
+%
+\begin{equation}
+    \boldsymbol{\mathcal{L}} = \sum_{t=0}^{t_{max}} \left[ \boldsymbol{\pi} \cdot \mathbf{A}^t \cdot \mathbf{B} \right] \cdot P \left( t \mid \text{T-cat.} \right)
+\end{equation}
+%
+This likelihood implicitly depends on how we parameterize the arcs of the \gls{dag} underlying the model and the parameterization of the distribution over diagnosis times.
+If we use the graph displayed in \cref{fig:graph_with_obs}), we can write down a conditional probability table for the probability of \gls{lnl} $X_3$ being healthy or metastatic, based on the state of the parent \gls{lnl} $\operatorname{pa}(3) = 2$:
 
@@
 
+Similar tables can be written for all other \glspl{lnl}.
+The terms in these conditional probability tables can then be used to compute the transition probability between to particular states $\boldsymbol{\xi}_i$ and $\boldsymbol{\xi}_j$ as in \cref{eq:transition_prob}.
+Computing this quantity for all possible transitions subsequently yields the transition matrix $\mathbf{A}$ as described in \cref{eq:transition_matrix}.
+
+Together with the parametrizations of the distributions over the diagnosis time, the parameters that make up the transition matrix $\mathbf{A}$ comprises the set of model parameters
+%
+\begin{equation}
+    \boldsymbol{\theta} = \left( \left\{ b_v \right\}, \left\{ t_{vr} \right\}, p_\text{early}, p_\text{late} \right) \quad \text{with} \quad \genfrac{}{}{0pt}{2}{v\leq V}{r\in\operatorname{pa}(v)}
+\end{equation}
+%
+The main task of of personalizing the \gls{ctv-n} definition is now to infer the probability of each of the possible states $\boldsymbol{\xi}_k$ given a diagnosis $\mathbf{Z}=\boldsymbol{\zeta}_\ell$.
+Using Bayes' theorem, we get
+%
+\begin{equation}
+    P\left( \mathbf{X}=\boldsymbol{\xi}_k \mid \mathbf{Z}=\boldsymbol{\zeta}_\ell \right) = \frac{P\left( \boldsymbol{\zeta}_\ell \mid \boldsymbol{\xi}_k \right) P\left( \boldsymbol{\xi}_k \right)}{\sum_{r=1}^{2^V} P\left( \boldsymbol{\zeta}_\ell \mid \boldsymbol{\xi}_r \right) P\left( \boldsymbol{\xi}_r \right) }
+\end{equation}
+%
+The model we are describing here will yield an estimate of $P\left( \boldsymbol{\xi}_k \right)$ that depends on the parameters used to define the transition matrix $\mathbf{A}$ and the prior over the diagnosis time $P \left( t \mid \text{T-cat.} \right)$.
+
+\subsection{Parameterization of the Hidden Markov Model}
+\label{subsec:previous_work:parametrization}
+
 \begin{figure*}
@@
 
-\printbibliography[%
-    title={References},%
-]
+\bibliography{bib}
 

### 2022-12-20 src/scripts/bg_core_prevs.py
use more concise naming scheme
`bg` in file names now stands for "base graph", while `wg` means
"winning graph".

@@ -4,7 +4,6 @@ scenarios involving the LNLs II, III and IV. This is to show that the base graph
 already capable of capturing the frequency of these involvement, as well as their
 correlations.
 """
-from fileinput import filename
 from pathlib import Path
 
 import matplotlib.pyplot as plt
@@ -14,7 +13,7 @@ from tueplots import figsizes, fontsizes
 import paths
 
 
-INPUT = paths.data / "base_graph_prevalences.hdf5"
+INPUT = paths.data / "bg_core_prevs.hdf5"
 OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
 NROWS, NCOLS = 2, 1
 

### 2022-12-20 src/scripts/bg_corner_plot.py

@@ -11,7 +11,7 @@ from emcee.backends import HDFBackend
 import paths
 
 
-INPUT = paths.data / "base_graph_samples.hdf5"
+INPUT = paths.data / "bg_samples.hdf5"
 THIN = 10
 OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
 

### 2022-12-20 src/tex/ms.tex
use more concise naming scheme
`bg` in file names now stands for "base graph", while `wg` means
"winning graph".
[showyourwork] trick SYW to get sync working again
[showyourwork] overleaf sync

@@
 
-\begin{table}[h!]
-    \centering
+\noindent
+\begin{center}
     \begin{tabular}{|cc|cc|}
@@
     \end{tabular}
-\end{table}
+\end{center}
 
@@
 \end{equation}
-%
-If multiple diagnostic modalities $\mathcal{O} \in \{ \text{MRI}, \text{CT}, \ldots \}$ are used, one can compute an observation matrix $\mathbf{B}^\mathcal{O}$ for each of these modalities and combine them via an outer product into a joint observations matrix
+
+If multiple diagnostic modalities $\mathcal{O} \in \{ \text{MRI}, \text{CT}, \ldots \}$ are used, one can compute an observation matrix $\mathbf{B}^\mathcal{O}$ for each of these modalities and combine them via an outer product into a joint observations matrix:
 %
@@
 \end{equation}
-%
+
 We define the time $t=0$ to be the moment just before a patient's tumor formed, and hence $X_v[t=0]=0 \,\,\, \forall v$.
@@
 %
+Here, the parameter $p_\text{early}$ can be interpreted as the probability that the patient will be diagnosed at time-step $t+1$ given they are in time-step $t$.
 We will use as a latest time-step $t_\text{max} = 10$.
@@
 %
-\begin{equation}
-    \boldsymbol{\mathcal{L}} = \sum_{t=0}^{t_{max}} \left[ \boldsymbol{\pi} \cdot \mathbf{A}^t \cdot \mathbf{B} \right] \cdot P \left( t \mid \text{T-cat.} \right)
+\begin{equation} \label{eq:likelihood_vec}
+\begin{split}
+    \boldsymbol{\ell} &= \big( P\left( \mathbf{Z} = \boldsymbol{\zeta}_i \right) \big) \\
+    &= \sum_{t=0}^{t_{max}} \left[ \boldsymbol{\pi} \cdot \mathbf{A}^t \cdot \mathbf{B} \right] \cdot P \left( t \mid \text{T-cat.} \right)
+\end{split}
 \end{equation}
@@
 This likelihood implicitly depends on how we parameterize the arcs of the \gls{dag} underlying the model and the parameterization of the distribution over diagnosis times.
-If we use the graph displayed in \cref{fig:graph_with_obs}), we can write down a conditional probability table for the probability of \gls{lnl} $X_3$ being healthy or metastatic, based on the state of the parent \gls{lnl} $\operatorname{pa}(3) = 2$:
+If we use the graph displayed in \cref{fig:graph_with_obs}, we can write down a conditional probability table for \gls{lnl} $X_3$ being healthy or metastatic, based on the state of the parent \gls{lnl} $\operatorname{pa}(3) = 2$:
 
-\begin{table}[h!]
-    \centering
+\noindent
+\begin{center}
     \begin{tabular}{|cc|cc|}
@@
 
-Together with the parametrizations of the distributions over the diagnosis time, the parameters that make up the transition matrix $\mathbf{A}$ comprises the set of model parameters
+Together with the parametrizations of the distributions over the diagnosis time, the parameters that make up the transition matrix $\mathbf{A}$ comprises the set of model parameters:
 %
@@
 \end{equation}
+
+To infer these parameters from a dataset of $N$ \gls{hnscc} patients $\boldsymbol{\mathcal{D}} = \left( d_1, d_2, \ldots, d_N \right)$, we compute the data log-likelihood:
 %
-The main task of of personalizing the \gls{ctv-n} definition is now to infer the probability of each of the possible states $\boldsymbol{\xi}_k$ given a diagnosis $\mathbf{Z}=\boldsymbol{\zeta}_\ell$.
+\begin{equation}
+    \log\mathcal{L} \left( \boldsymbol{\mathcal{D}} \mid \boldsymbol{\theta} \right) = \sum_{i=1}^N \log P \left( \mathbf{Z} = d_i \right)
+\end{equation}
+%
+Which effectively amounts to computing the element-wise logarithm of the likelihood vector $\boldsymbol{\ell}$ from \cref{eq:likelihood_vec} and summing up the entries that correspond to each of the patients $d_i$ for $i\leq N$.
+
+Using this log-likelihood function one may now employ a variety of inference methods to learn the parameters of the model that best describe the observed data.
+After that, the main task of of personalizing the \gls{ctv-n} definition is to predict the probability of the hidden possible states $\boldsymbol{\xi}_k$ given the diagnosis $d^\star=\boldsymbol{\zeta}_\ell$ of a new patient.
 Using Bayes' theorem, we get
@@
 \begin{equation}
-    P\left( \mathbf{X}=\boldsymbol{\xi}_k \mid \mathbf{Z}=\boldsymbol{\zeta}_\ell \right) = \frac{P\left( \boldsymbol{\zeta}_\ell \mid \boldsymbol{\xi}_k \right) P\left( \boldsymbol{\xi}_k \right)}{\sum_{r=1}^{2^V} P\left( \boldsymbol{\zeta}_\ell \mid \boldsymbol{\xi}_r \right) P\left( \boldsymbol{\xi}_r \right) }
+    P\left( \mathbf{X}=\boldsymbol{\xi}_k \mid \mathbf{Z}=\boldsymbol{\zeta}_\ell \right) = \frac{P\left( \boldsymbol{\zeta}_\ell \mid \boldsymbol{\xi}_k \right) P\left( \boldsymbol{\xi}_k \mid \boldsymbol{\theta} \right)}{\sum_{r=1}^{2^V} P\left( \boldsymbol{\zeta}_\ell \mid \boldsymbol{\xi}_r \right) P\left( \boldsymbol{\xi}_r \mid \boldsymbol{\theta} \right) }
 \end{equation}
 %
-The model we are describing here will yield an estimate of $P\left( \boldsymbol{\xi}_k \right)$ that depends on the parameters used to define the transition matrix $\mathbf{A}$ and the prior over the diagnosis time $P \left( t \mid \text{T-cat.} \right)$.
-
-\subsection{Parameterization of the Hidden Markov Model}
-\label{subsec:previous_work:parametrization}
+The described model along with the inferred parameters $\boldsymbol{\hat{\theta}}$ will yield an estimate (or multiple estimates) for the ``prior'' in the above equation $P\big( \boldsymbol{\xi}_k \mid \boldsymbol{\hat{\theta}} \big)$.
 
 \begin{figure*}
-    \script{base_graph_prevalences.py}
+    \script{bg_core_prevs.py}
     \begin{centering}
-        \includegraphics[width=\textwidth]{figures/base_graph_prevalences.png}
+        \includegraphics[width=\textwidth]{figures/bg_core_prevs.png}
         \caption{Testing this showyourwork thingy}
-        \label{fig:base_graph_prevalences}
+        \label{fig:bg_prevalences}
     \end{centering}

### 2022-12-21 src/scripts/comp_IandII_prevs.py
add comparisons of LNLs I and II (base vs winning)

@@ -1,0 +1,92 @@
+"""
+Plot histograms of predicted and posteriors over observed prevalences for scenarios
+that involve the LNLs I and II. Compare these prevalence predictions for the base-graph
+model to the winning-graph model.
+"""
+from pathlib import Path
+
+import matplotlib.pyplot as plt
+from lyscripts.plot.utils import Histogram, Posterior, draw
+from lyscripts.plot.utils import COLORS as USZ
+from tueplots import figsizes, fontsizes
+
+import paths
+
+
+BASE_GRAPH_INPUT = paths.data / "bg_IandII_prevs.hdf5"
+WIN_GRAPH_INPUT = paths.data / "wg_IandII_prevs.hdf5"
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
+NROWS, NCOLS = 2, 1
+
+
+if __name__ == "__main__":
+    plt.rcParams.update(figsizes.icml2022_half(
+        nrows=NROWS,
+        ncols=NCOLS,
+        # height_to_width_ratio=0.3,
+    ))
+    plt.rcParams.update(fontsizes.icml2022())
+    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
+
+    panels = {
+        "early": [],
+        "late": [],
+    }
+
+    for stage in ["early", "late"]:
+        panels[stage] = [
+            Histogram.from_hdf5(
+                filename=WIN_GRAPH_INPUT,
+                dataname=f"I/{stage}",
+                label="LNL I overall",
+                color=USZ["blue"],
+            ),
+            Histogram.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"I/{stage}",
+                label="base graph",
+                histtype="step",
+                color="darkgray",
+                linewidth=2.,
+                hatch=r"////",
+            ),
+            Posterior.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"I/{stage}",
+                color=USZ["blue"],
+            ),
+
+            Histogram.from_hdf5(
+                filename=WIN_GRAPH_INPUT,
+                dataname=f"InotII/{stage}",
+                label="LNL I without II",
+                color=USZ["green"],
+            ),
+            Histogram.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"InotII/{stage}",
[…]

### 2022-12-21 src/scripts/comp_IandII_high_prevs.py
complete prevs comparison plots of LNLs I & II

@@ -37,13 +37,13 @@ if __name__ == "__main__":
         panels[stage] = [
             Histogram.from_hdf5(
                 filename=WIN_GRAPH_INPUT,
-                dataname=f"I/{stage}",
-                label="LNL I overall",
-                color=USZ["blue"],
+                dataname=f"II/{stage}",
+                label="LNL II overall",
+                color=USZ["red"],
             ),
             Histogram.from_hdf5(
                 filename=BASE_GRAPH_INPUT,
-                dataname=f"I/{stage}",
+                dataname=f"II/{stage}",
                 label="base graph",
                 histtype="step",
                 color="darkgray",
@@ -52,39 +52,39 @@ if __name__ == "__main__":
             ),
             Posterior.from_hdf5(
                 filename=BASE_GRAPH_INPUT,
-                dataname=f"I/{stage}",
-                color=USZ["blue"],
+                dataname=f"II/{stage}",
+                color=USZ["red"],
             ),
 
             Histogram.from_hdf5(
                 filename=WIN_GRAPH_INPUT,
-                dataname=f"InotII/{stage}",
-                label="LNL I without II",
-                color=USZ["green"],
+                dataname=f"IInotI/{stage}",
+                label="LNL II without I",
+                color=USZ["orange"],
             ),
             Histogram.from_hdf5(
                 filename=BASE_GRAPH_INPUT,
-                dataname=f"InotII/{stage}",
+                dataname=f"IInotI/{stage}",
                 label="base graph",
                 histtype="step",
                 color="darkgray",
                 linewidth=2.,
-                hatch=r"////",
+                hatch=r"\\\\",
             ),
             Posterior.from_hdf5(
                 filename=BASE_GRAPH_INPUT,
-                dataname=f"InotII/{stage}",
-                color=USZ["green"],
+                dataname=f"IInotI/{stage}",
+                color=USZ["orange"],
             ),
         ]
 
-    draw(axes[0], contents=panels["early"], xlims=(0., 15.), hist_kwargs={"nbins": 40})
-    axes[0].set_ylim(bottom=0., top=1.5)
+    draw(axes[0], contents=panels["early"], xlims=(50., 90.), hist_kwargs={"nbins": 40})
+    # axes[0].set_ylim(bottom=0., top=1.5)
[…]

### 2022-12-21 src/scripts/comp_IandII_low_prevs.py

@@ -1,0 +1,92 @@
+"""
+Plot histograms of predicted and posteriors over observed prevalences for scenarios
+that involve the LNLs I and II. Compare these prevalence predictions for the base-graph
+model to the winning-graph model.
+"""
+from pathlib import Path
+
+import matplotlib.pyplot as plt
+from lyscripts.plot.utils import Histogram, Posterior, draw
+from lyscripts.plot.utils import COLORS as USZ
+from tueplots import figsizes, fontsizes
+
+import paths
+
+
+BASE_GRAPH_INPUT = paths.data / "bg_IandII_prevs.hdf5"
+WIN_GRAPH_INPUT = paths.data / "wg_IandII_prevs.hdf5"
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
+NROWS, NCOLS = 2, 1
+
+
+if __name__ == "__main__":
+    plt.rcParams.update(figsizes.icml2022_half(
+        nrows=NROWS,
+        ncols=NCOLS,
+        # height_to_width_ratio=0.3,
+    ))
+    plt.rcParams.update(fontsizes.icml2022())
+    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
+
+    panels = {
+        "early": [],
+        "late": [],
+    }
+
+    for stage in ["early", "late"]:
+        panels[stage] = [
+            Histogram.from_hdf5(
+                filename=WIN_GRAPH_INPUT,
+                dataname=f"I/{stage}",
+                label="LNL I overall",
+                color=USZ["blue"],
+            ),
+            Histogram.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"I/{stage}",
+                label="base graph",
+                histtype="step",
+                color="darkgray",
+                linewidth=2.,
+                hatch=r"////",
+            ),
+            Posterior.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"I/{stage}",
+                color=USZ["blue"],
+            ),
+
+            Histogram.from_hdf5(
+                filename=WIN_GRAPH_INPUT,
+                dataname=f"InotII/{stage}",
+                label="LNL I without II",
+                color=USZ["green"],
+            ),
+            Histogram.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"InotII/{stage}",
[…]

### 2022-12-21 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
+\begin{figure}
+    \script{comp_IandII_low_prevs.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/comp_IandII_low_prevs.png}
+        \caption{This should be a one-column wide figure.}
+        \label{fig:IandII_low_prevs}
+    \end{centering}
+\end{figure}
+
+\begin{figure}
+    \script{comp_IandII_high_prevs.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/comp_IandII_high_prevs.png}
+        \caption{This should be a one-column wide figure.}
+        \label{fig:IandII_high_prevs}
+    \end{centering}
+\end{figure}
+
+\section{Results}
+\label{sec:results}
+
+\begin{center}
+    \begin{tabular}{|l|l|l|l|}
+        \hline
+        \textbf{Graph} & \textbf{-- BIC/2} & \textbf{max. llh} & \textbf{mean llh} \\
+        \hline
+        base graph    & \variable{output/bg_BIC.txt} & \variable{output/bg_max_llh.txt} & \variable{output/bg_mean_llh.txt} \\
+        winning graph & \variable{output/wg_BIC.txt} & \variable{output/wg_max_llh.txt} & \variable{output/wg_mean_llh.txt} \\
+        \hline
+    \end{tabular}
+\end{center}
+
 \bibliography{bib}

### 2022-12-21 src/scripts/compile_metrics_table.py
add results table as a variable

@@ -1,0 +1,38 @@
+"""
+Populate a Jinja2 LaTeX table template with the metrics from the experiment runs.
+"""
+import json
+from jinja2 import Environment, FileSystemLoader
+
+import paths
+
+
+BASE_GRAPH_INPUT = paths.data / "bg_metrics.json"
+WIN_GRAPH_INPUT = paths.data / "wg_metrics.json"
+TEMPLATE_NAME = "results_table.temp"
+OUTPUT = paths.output / "results_table.tex"
+
+
+environment = Environment(
+    loader=FileSystemLoader(paths.scripts)
+)
+table_template = environment.get_template(TEMPLATE_NAME)
+context = {}
+
+
+with open(BASE_GRAPH_INPUT, mode="r") as bg_file:
+    bg_metrics = json.load(bg_file)
+
+for key,val in bg_metrics.items():
+    context[f"bg_{key}"] = val
+
+
+with open(WIN_GRAPH_INPUT, mode="r") as wg_file:
+    wg_metrics = json.load(wg_file)
+
+for key,val in wg_metrics.items():
+    context[f"wg_{key}"] = val
+
+
+with open(OUTPUT, mode="w") as table_file:
+    table_file.write(table_template.render(**context))

### 2022-12-21 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\begin{center}
-    \begin{tabular}{|l|l|l|l|}
-        \hline
-        \textbf{Graph} & \textbf{-- BIC/2} & \textbf{max. llh} & \textbf{mean llh} \\
-        \hline
-        base graph    & \variable{output/bg_BIC.txt} & \variable{output/bg_max_llh.txt} & \variable{output/bg_mean_llh.txt} \\
-        winning graph & \variable{output/wg_BIC.txt} & \variable{output/wg_max_llh.txt} & \variable{output/wg_mean_llh.txt} \\
-        \hline
-    \end{tabular}
-\end{center}
+\begin{table}
+    \centering
+    \variable{output/results_table.tex}
+    \caption{Model comparison results from the base and the winning graph.}
+    \label{table:evidence}
+\end{table}
 

### 2022-12-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \newacronym{lnl}{LNL}{lymph node level}
+\newacronym{hpv}{HPV}{human papilloma virus}
 \newacronym{hmm}{HMM}{hidden Markov model}
@@
 \newacronym{fna}{FNA}{fine needle aspiration}
+\newacronym{clb}{CLB}{Centre Leon Bérard}
+\newacronym{usz}{USZ}{University Hospital Zurich}
+\newacronym{isb}{ISB}{Inselspital Bern}
 
@@
 %
-\begin{equation}
+\begin{equation} \label{eq:log_likelihood}
     \log\mathcal{L} \left( \boldsymbol{\mathcal{D}} \mid \boldsymbol{\theta} \right) = \sum_{i=1}^N \log P \left( \mathbf{Z} = d_i \right)
@@
 
+\section{Methods}
+\label{sec:methods}
+
+We use \gls{mcmc} sampling to draw parameter samples $\boldsymbol{\hat{\theta}}_i$ for $i \leq S$ from the likelihood described in \cref{eq:log_likelihood}.
+More specifically, we use the Python implementation \texttt{emcee} \cite{foreman-mackey_emcee_2013} and two sample proposal mechanisms based on differential evolution moves \cite{ter_braak_differential_2008,nelson_run_2013} for sampling.
+The $S$ parameter estimates are then used to compute distributions of estimates that depend on the parameters $\boldsymbol{\theta}$.
+
+The entire methodology used in this work is publicly available in the GitHub repository \href{https://github.com/rmnldwg/lynference}{\texttt{rmnldwg/lynference}}.
+Tagged references to specific versions of an inference pipeline allow reproducing the inferred parameters of the models described here.
+Prevalence and risk predictions, as well as the figures depicting them, can be reproduced following the instructions of the GitHub repository underlying this publication \href{https://github.com/rmnldwg/graph-extension}{\texttt{rmnldwg/graph-extension}}.
+
+\section{Multicentric Dataset}
+\label{sec:data}
+
+The dataset $\boldsymbol{\mathcal{D}}$ that we used for inference is comprised of the detailed reports on lymphatic involvement patterns in \gls{opscc} patients treated at three different institutions in France and Switzerland: The \gls{clb} in Lyon (France), the \gls{isb}, and the \gls{usz} (both in Swiztzerland).
+We have previously published the patterns of nodal involvement for the \gls{usz} cohort \cite{ludwig_dataset_2022} and described its characteristics in detail \cite{ludwig_detailed_2022}.
+A large part of the \gls{clb} data underlies a publication on \gls{hpv} status in \gls{opscc} \cite{bauwens_prevalence_2021} that was kindly provided to us by Vincent Grégoire.
+The patient information on lymphatic metastatic involvement from the \gls{isb} cohort was sent to us by Roland Giger through a collaboration.
+
+In total, the dataset contains 681 patients newly diagnosed with \gls{opscc} that were treated with either definitive radiotherapy or neck dissection.
+Pathologically assessed post-operative \gls{lnl} involvement was available for 393 of patients, while for the remainder the nodal involvement was assessed based on available diagnostic modalities.
+If multiple modalities were used to diagnose a patient's lymphatic involvement and conflicting conclusions were drawn from this, the conflicts were resolved by inferring the most likely state (healthy or metastatic) for each \gls{lnl} separately.
+To do so, we used literature values for the sensitivity and specificity of the diagnostic modalities \cite{de_bondt_detection_2007,kyzas_18f-fluorodeoxyglucose_2008}, which we also tabulated below.
+
+\noindent
+\begin{center}
+    \begin{tabular}{|l|rr|}
+        \hline
+        \textbf{Modality} & \textbf{Specificity} & \textbf{Sensitivity} \\
+        \hline
+        \acrshort{ct} & 76\% & 81\% \\
+        \acrshort{pet} & 86\% & 79\% \\
+        \gls{mri} & 63\% & 81\% \\
+        \acrshort{fna} & 98\% & 80\% \\
+        Pathology & $\approx$ 100\% & $\approx$ 100\% \\
+        \hline
+    \end{tabular}
+\end{center}
+
+\section{Results}
+\label{sec:results}
+
 \begin{figure*}
@@
         \includegraphics[width=\textwidth]{figures/bg_core_prevs.png}
-        \caption{Testing this showyourwork thingy}
+        \caption{Prevalence of involvement as predicted by the base graph model for different scenarios involving the most commonly metastatic \glspl{lnl} II, III and IV (shaded histograms).
+The model's predictions are compared to Beta posteriors over the prevalence based on the frequency of the same scenarios and a uniform prior (slid lines).
+The top panel shows some selected scenarios with early T-category tumors and the bottom panel with late T-category.}
         \label{fig:bg_prevalences}
@@
 \end{figure*}
+
+For the base graph, we have plotted the predicted prevalence of involvement patterns in the investigated patient cohort for scenarios involving the most commonly metastatic \glspl{lnl} II, III and IV in \cref{fig:bg_prevalences}.
+This figure shows that this minimal graph is already capable of describing the most important parts of the observed data very well.
+The subsequent changes to the graph connections we investigated therefore focus on the improvements w.r.t.
+involvement patterns that include the \glspl{lnl} I, V, and VII that more rarely harbor occult disease.
 
@@
         \includegraphics[width=\linewidth]{figures/comp_IandII_low_prevs.png}
-        \caption{This should be a one-column wide figure.}
+        \caption{Prevalence predictions of the winning extended graph (colored histograms) compared to the base graph (hatched histograms) for scenarios of nodal involvement in the levels I and II.
+The solid lines depict the Beta posterior over the prevalence based on a uniform prior and the number of patients matching the respective scenario in the data.
+The top and bottom panel show the same scenarios, but for early and late T-category respectively.}
         \label{fig:IandII_low_prevs}
@@
         \includegraphics[width=\linewidth]{figures/comp_IandII_high_prevs.png}
-        \caption{This should be a one-column wide figure.}
+        \caption{Histograms and Beta posteriors over the prevalence of involvement for \gls{lnl} II (with and without \gls{lnl} I being metastatic as well).
+The colored histograms depict the predictions of the extended graph, while the hatched ones show the same prediction for the base graph.
+The top panel displays the predictions and observations for early T-category patients, the bottom panel for late T-category ones.}
         \label{fig:IandII_high_prevs}
@@
 
-\section{Results}
-\label{sec:results}
-
-\begin{table}
+\begin{table*}
     \centering
     \variable{output/results_table.tex}
-    \caption{Model comparison results from the base and the winning graph.}
+    \caption{Model comparison results from the base graph and the extended graph we chose as the ``winnning'' model.
+For both \glspl{dag} we show the log-evidence, computed via thermodynamic integration, the negative one half of the \gls{bic}, as well as the maximum log-likelihood that was encountered during the final \gls{mcmc} sampling round.}
     \label{table:evidence}
-\end{table}
+\end{table*}
 

### 2022-12-23 src/scripts/comp_IIIandV_prevs.py
add prev plots investigating additional edges

@@ -1,0 +1,93 @@
+"""
+Plot histograms of predicted and posteriors over observed prevalences for scenarios
+that involve the LNLs III and V. Compare these prevalence predictions for the
+base-graph model to the winning-graph model.
+"""
+from pathlib import Path
+
+import matplotlib.pyplot as plt
+from matplotlib.lines import Line2D
+from lyscripts.plot.utils import Histogram, Posterior, draw
+from lyscripts.plot.utils import COLORS as USZ
+from tueplots import figsizes, fontsizes
+
+import paths
+
+
+BASE_GRAPH_INPUT = paths.data / "bg_IIIandV_prevs.hdf5"
+WIN_GRAPH_INPUT = paths.data / "wg_IIIandV_prevs.hdf5"
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
+NROWS, NCOLS = 2, 1
+
+
+if __name__ == "__main__":
+    plt.rcParams.update(figsizes.icml2022_half(
+        nrows=NROWS,
+        ncols=NCOLS,
+        # height_to_width_ratio=0.3,
+    ))
+    plt.rcParams.update(fontsizes.icml2022())
+    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
+
+    panels = {
+        "early": [],
+        "late": [],
+    }
+
+    for stage in ["early", "late"]:
+        panels[stage] = [
+            Histogram.from_hdf5(
+                filename=WIN_GRAPH_INPUT,
+                dataname=f"VnotIII/{stage}",
+                label="LNL V without III",
+                color=USZ["green"],
+            ),
+            Histogram.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"VnotIII/{stage}",
+                label="base graph",
+                histtype="step",
+                color="darkgray",
+                linewidth=2.,
+                hatch=r"////",
+            ),
+            Posterior.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"VnotIII/{stage}",
+                color=USZ["green"],
+            ),
+
+            Histogram.from_hdf5(
+                filename=WIN_GRAPH_INPUT,
+                dataname=f"IIIandV/{stage}",
+                label="LNL III and V",
+                color=USZ["red"],
+            ),
+            Histogram.from_hdf5(
[…]

### 2022-12-23 src/scripts/comp_IIandVII_prevs.py

@@ -1,0 +1,93 @@
+"""
+Plot histograms of predicted and posteriors over observed prevalences for scenarios
+that involve the LNLs II and VII. Compare these prevalence predictions for the
+base-graph model to the winning-graph model.
+"""
+from pathlib import Path
+
+import matplotlib.pyplot as plt
+from matplotlib.lines import Line2D
+from lyscripts.plot.utils import Histogram, Posterior, draw
+from lyscripts.plot.utils import COLORS as USZ
+from tueplots import figsizes, fontsizes
+
+import paths
+
+
+BASE_GRAPH_INPUT = paths.data / "bg_IIandVII_prevs.hdf5"
+WIN_GRAPH_INPUT = paths.data / "wg_IIandVII_prevs.hdf5"
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
+NROWS, NCOLS = 2, 1
+
+
+if __name__ == "__main__":
+    plt.rcParams.update(figsizes.icml2022_half(
+        nrows=NROWS,
+        ncols=NCOLS,
+        # height_to_width_ratio=0.3,
+    ))
+    plt.rcParams.update(fontsizes.icml2022())
+    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
+
+    panels = {
+        "early": [],
+        "late": [],
+    }
+
+    for stage in ["early", "late"]:
+        panels[stage] = [
+            Histogram.from_hdf5(
+                filename=WIN_GRAPH_INPUT,
+                dataname=f"IIandVII/{stage}",
+                label="LNL II with VII",
+                color=USZ["red"],
+            ),
+            Histogram.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"IIandVII/{stage}",
+                label="base graph",
+                histtype="step",
+                color="darkgray",
+                linewidth=2.,
+                hatch=r"////",
+            ),
+            Posterior.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"IIandVII/{stage}",
+                color=USZ["red"],
+            ),
+
+            Histogram.from_hdf5(
+                filename=WIN_GRAPH_INPUT,
+                dataname=f"VIInotII/{stage}",
+                label="LNL VII without II",
+                color=USZ["blue"],
+            ),
+            Histogram.from_hdf5(
[…]

### 2022-12-23 src/scripts/comp_IVandV_prevs.py

@@ -1,0 +1,93 @@
+"""
+Plot histograms of predicted and posteriors over observed prevalences for scenarios
+that involve the LNLs VI and V. Compare these prevalence predictions for the
+base-graph model to the winning-graph model.
+"""
+from pathlib import Path
+
+import matplotlib.pyplot as plt
+from matplotlib.lines import Line2D
+from lyscripts.plot.utils import Histogram, Posterior, draw
+from lyscripts.plot.utils import COLORS as USZ
+from tueplots import figsizes, fontsizes
+
+import paths
+
+
+BASE_GRAPH_INPUT = paths.data / "bg_IVandV_prevs.hdf5"
+WIN_GRAPH_INPUT = paths.data / "wg_IVandV_prevs.hdf5"
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
+NROWS, NCOLS = 2, 1
+
+
+if __name__ == "__main__":
+    plt.rcParams.update(figsizes.icml2022_half(
+        nrows=NROWS,
+        ncols=NCOLS,
+        # height_to_width_ratio=0.3,
+    ))
+    plt.rcParams.update(fontsizes.icml2022())
+    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
+
+    panels = {
+        "early": [],
+        "late": [],
+    }
+
+    for stage in ["early", "late"]:
+        panels[stage] = [
+            Histogram.from_hdf5(
+                filename=WIN_GRAPH_INPUT,
+                dataname=f"VnotIV/{stage}",
+                label="LNL V without IV",
+                color=USZ["orange"],
+            ),
+            Histogram.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"VnotIV/{stage}",
+                label="base graph",
+                histtype="step",
+                color="darkgray",
+                linewidth=2.,
+                hatch=r"////",
+            ),
+            Posterior.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"VnotIV/{stage}",
+                color=USZ["orange"],
+            ),
+
+            Histogram.from_hdf5(
+                filename=WIN_GRAPH_INPUT,
+                dataname=f"IVandV/{stage}",
+                label="LNL IV and V",
+                color=USZ["blue"],
+            ),
+            Histogram.from_hdf5(
[…]

### 2022-12-30 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{multirow}
+\usepackage{nicefrac}
 
@@
 
-We use \gls{mcmc} sampling to draw parameter samples $\boldsymbol{\hat{\theta}}_i$ for $i \leq S$ from the likelihood described in \cref{eq:log_likelihood}.
+\subsection{Parameter Inference}
+\label{subsec:methods:inference}
+
+We use \gls{mcmc} sampling to draw parameter samples $\boldsymbol{\hat{\theta}}_i$ for $i \leq S$ from the likelihood described in \cref{eq:log_likelihood} (i.e.
+the unnormalized posterior distribution over the parameters $\boldsymbol{\theta}$, since we used a uniform prior in this work).
+
 More specifically, we use the Python implementation \texttt{emcee} \cite{foreman-mackey_emcee_2013} and two sample proposal mechanisms based on differential evolution moves \cite{ter_braak_differential_2008,nelson_run_2013} for sampling.
-The $S$ parameter estimates are then used to compute distributions of estimates that depend on the parameters $\boldsymbol{\theta}$.
+Instead of proposing and then accepting or rejecting individual parameter samples one after the other (as in the classical Metropolis-Hastings algorithm), the \texttt{emcee} implementation makes use of an ensemble of $W$ so-called ``walkers''.
+This gives rise to $W$ parallel chains of samples that mutually influence each others proposal such that the sampling proceedure overall is \emph{affine invariant}.
+This means that scaling the parameter space along any dimension has no effect on the performance of the \gls{mcmc}.
+
+For the experiments in this work, we used $W = 20 \cdot k$ walkers, where $k$ is the dimensionaliry of the parameter space $\boldsymbol{\Theta}$.
+After an initial ``burn-in'' phase, during which all drawn samples are discarded because they are not yet independent from the initial state, we continued sampling for another 200 steps of which we discarded every 10th to be left with $S = 20 \cdot W$ samples.
+
+These $S$ parameter estimates are then used to compute expectation values of estimates that depend on the parameters $\boldsymbol{\theta}$ through an integral over the parameter space $\boldsymbol{\Theta}$:
+%
+\begin{equation}
+    \begin{aligned}
+        \mathbb{E}_p \left[ f \right] &= \int_{\boldsymbol{\Theta}} p(\boldsymbol{\theta}) f(\boldsymbol{\theta}) d\boldsymbol{\theta} \\
+        &\approx \frac{1}{S} \sum_{i=1}^S f \big( \boldsymbol{\hat{\theta}}_i \big)
+    \end{aligned}
+\end{equation}
+%
+Alternatively, the individual $\hat{f}_i = f\big( \boldsymbol{\hat{\theta}}_i \big)$ can be used to plot histograms over the distribution of $f$.
+We will do so in \cref{sec:results} to show distributions over prevalence predictions and risk computations.
+
+Another relevant model parameter that needs to be set for the inference process, is the maximum number of time-steps we used for the evolution of the system.
+We set this value to $t_\text{max} = 10$, such that $t \in \{ 0, 1, 2, \ldots, 10 \}$.
+The binomial ``success probability'' used to fix the shape of the early T-category's time-prior was set to $p_\text{early} = 0.3$.
+
+\subsection{Model Comparison}
+\label{subsec:methods:comparison}
+
+\begin{table*}
+    \centering
+    \begin{tabular}{ | c | c | c | }
+        \hline
+        $K_\text{1v2}$ & $\ln{K_\text{1v2}}$ & support for $\mathcal{M}_1$ \\
+        \hline
+        $< 10^0$ & $< 0$ & negative evidence (supports $\mathcal{M}_2$) \\
+        $10^0$ to $10^{\nicefrac{1}{2}}$ & 0 to 1.15 & barely worth a mention \\
+        $10^{\nicefrac{1}{2}}$ to $10^1$ & 1.15 to 2.3 & substantial \\
+        $10^1$ to $10^{\nicefrac{3}{2}}$ & 2.3 to 3.45 & strong \\
+        $10^{\nicefrac{3}{2}}$ to $10^2$ & 3.45 to 4.6 & very strong \\
+        $> 10^2$ & $> 4.6$ & decisive \\
+        \hline
+    \end{tabular}
+    \caption{Interpretation of Bayes factors and their natural logarithms in terms of their support for or against one of the two compared models as introduced by Harold Jeffreys \cite{jeffreys_theory_1998}.}
+    \label{table:bayes_factor}
+\end{table*}
+
+The aim of this work is to refine the graph structure underlying our risk model introduced in the previous section.
+This \gls{dag} determines the number of parameters of the model as well as how exactly the transition matrix $\mathbf{A}$ is parameterized.
+To compare different models that are based on different \glspl{dag}, e.g.
+models $\mathcal{M}_1$ and $\mathcal{M}_2$, in a Bayesian setting, we need to compute the probabilities of these models, given the data $\boldsymbol{\mathcal{D}}$:
+%
+\begin{equation}
+    P \left( \mathcal{M}_i \mid \boldsymbol{\mathcal{D}} \right) = \frac{P \left( \boldsymbol{\mathcal{D}} \mid \mathcal{M}_i \right) P \left( \mathcal{M}_i \right)}{ P \left( \boldsymbol{\mathcal{D}} \right) }
+\end{equation}
+%
+If we assume all models $\mathcal{M}_i$ for $i \in \{ 1,2 \}$ to have the same \emph{a priori} probability -- meaning in this case $P (\mathcal{M}_1) = P (\mathcal{M}_2)$ -- then we can compute the so-called \emph{Bayes factor} of the two models as the ratio of their likelihoods.
+The interpretation of the values for different Bayes factors is given in \cref{table:bayes_factor}.
+It is defined as follows:
+%
+\begin{equation}
+    K_\text{1v2} = \frac{P \left( \mathcal{M}_1 \mid \boldsymbol{\mathcal{D}} \right)}{P \left( \mathcal{M}_2 \mid \boldsymbol{\mathcal{D}} \right)} = \frac{P \left( \boldsymbol{\mathcal{D}} \mid \mathcal{M}_1 \right)}{P \left( \boldsymbol{\mathcal{D}} \mid \mathcal{M}_2 \right)}
+\end{equation}
+%
+These likelihoods are commonly called the \emph{model evidence} or \emph{marginal likelihood}. The latter because computing it involves marginalizing the data likelihood over all model parameters:
+%
+\begin{equation} \label{eq:evidence}
+    E_\mathcal{M} = P \left( \boldsymbol{\mathcal{D}} \mid \mathcal{M} \right) = \int_{\boldsymbol{\Theta}} P\left( \boldsymbol{\mathcal{D}} \mid \boldsymbol{\theta}, \mathcal{M} \right) p(\boldsymbol{\theta} \mid \mathcal{M}) d\boldsymbol{\theta}
+\end{equation}
+%
+However, this quantity is often very hard to compute or even intractable, due to the high dimensionality of the parameter space $\boldsymbol{\Theta}$.
+In our case, the number of dimensions ranges from $k=9$ for the \emph{base graph} to $k=11$ for the \emph{winning graph}.
+A brute-force integration over a unit cube with this many dimensions is inefficient and error prone, which is why we resorted to \gls{ti} for computing the (log-)evidence.
+
+Below, we will briefly outline the main concept behind this algorithm. An intuitive and extensive derivation of \gls{ti} is given by \cite{aponte_introduction_2022}.
+
+We start by taking the logarithm of the model evidence $E$ and subtract a zero from it in the form of the term $0 = \ln \int p(\boldsymbol{\theta} \mid \mathcal{M}) d\boldsymbol{\theta}$.
+Further, we can multiply the distribution over the parameters $\boldsymbol{\theta}$ inside this integral by $1 = P \left( \boldsymbol{\mathcal{D}} \mid \boldsymbol{\theta}, \mathcal{M} \right)^{\beta=0}$.
+Subsequently, we can write the logarithm of the evidence as an integral over a derivative:
+%
+\begin{equation} \label{eq:ti}
+    \begin{aligned}
+        \ln E &= \ln \int P \left( \boldsymbol{\mathcal{D}} \mid \boldsymbol{\theta}, \mathcal{M} \right)^{\beta=1} p \left( \boldsymbol{\theta} \mid \mathcal{M} \right) d\boldsymbol{\theta} - \ln E_0 \\
+        &= \int_0^1 \frac{d}{d\beta} \ln E_\beta d\beta
+    \end{aligned}
+\end{equation}
+%
+Where we have used the (unnormalized) \emph{power posterior} $p_\beta \left( \boldsymbol{\theta} \mid \boldsymbol{\mathcal{D}}, \mathcal{M} \right) = P \left( \boldsymbol{\mathcal{D}} \mid \boldsymbol{\theta}, \mathcal{M} \right)^\beta p \left( \boldsymbol{\theta} \mid \mathcal{M} \right)$ to compute the respective evidence $E_\beta = \int p_\beta \left( \boldsymbol{\theta} \mid \boldsymbol{\mathcal{D}}, \mathcal{M} \right) d\boldsymbol{\theta}$.
+
+The derivatives of the log-evidences $\ln E_\beta$ are essentially expectation values of the data log-likelihood under the power posteriors of the corresponding value for $\beta$.
+They can be computed using \gls{mcmc}:
+%
+\begin{equation}
+    \begin{aligned}
+        \frac{d}{d\beta} \ln E_\beta &= \int p_\beta \left( \boldsymbol{\theta} \mid \boldsymbol{\mathcal{D}}, \mathcal{M} \right) \ln P \left( \boldsymbol{\mathcal{D}} \mid \boldsymbol{\theta}, \mathcal{M} \right) d\boldsymbol{\theta} \\
+        &= \mathbb{E} \left[ \ln P \left( \boldsymbol{\mathcal{D}} \mid \boldsymbol{\theta}, \mathcal{M} \right) \right]_{ p_\beta \left( \boldsymbol{\theta} \mid \boldsymbol{\mathcal{D}}, \mathcal{M} \right) } \\
+        &\approx \frac{1}{S} \sum_{i=1}^S \ln P \left( \boldsymbol{D} \mid \hat{\boldsymbol{\theta}}_{\beta i}, \mathcal{M} \right) =: \mathcal{A}_\text{MC} \left( \beta \right)
+    \end{aligned}
+\end{equation}
+%
+The integral in \cref{eq:ti} can then be computed via a trapezoidal rule using the $\mathcal{A}_\text{MC}$ to yield a  numerical approximation of the model evidence:
+%
+\begin{equation}
+    \ln E \approx \frac{1}{2} \sum_{j=0}^{R-1} \left( \beta_{j+1} - \beta_j \right) \cdot \big( \mathcal{A}_\text{MC} (\beta_{j+1}) + \mathcal{A}_\text{MC} (\beta_j) \big)
+\end{equation}
+%
+This estimate gets better for more samples $S$ per sampling from the power posterior $p_\beta$ but more importantly it gets better for a tighter spacing of the values for $\beta$ within the interval $[0,1]$.
+The variable $\beta$ is also often referred to as an \emph{inverse temperature}, due to its origins in statistical physics.
+Often when performing \gls{ti}, the most drastic changes in the values of the $\mathcal{A}_\text{MC}$ occur at high temperatures (meaning $\beta$ very close to zero), while the changes become smaller and smaller for lower temparatures ($\beta$ towards one).
+It is therefore efficient to space the \emph{temperature ladder} unevenly, e.g.
+according to a fifth order power rule:
+%
+\begin{equation} \label{eq:power_rule}
+    \beta_j = \left( j / R \right)^5 \qquad j \in \{ 0, 1, 2, \ldots, R \}
+\end{equation}
+%
+For the \glspl{ti} that were performed in this work we used such a fifth order power rule with 64 steps, meaning that $R=63$.
+
+The process of computing the log-evidence using \gls{ti} was as follows: We randomly initialized the starting positions of the $W$ samplers in the ensemble within the $k$ dimensional unit cube $\boldsymbol{\Theta}$.
+Subsequently, for each $j \in \{ 0, 1, 2, \ldots, R=63 \}$ we drew samples from the corresponding power posterior with the value of $\beta_j$ set according to the power rule in \cref{eq:power_rule}.
+This sampling at point $j$ consisted of 1000 burn-in steps, followed by 200 steps, of which only every tenth was kept.
+The last position of the $W$ chains for the $j$-th $\beta$ value in the ladder was used to initialize the subsequent sampling round with $\beta_{j+1}$.
+Hence, after the computations are finished, we are left with $S = 20 \cdot W$ samples $\boldsymbol{\hat{\theta}}_{i,j}$ and respective log-likelihood $\hat{\ell}_{i,j}$ from each of the 64 power posteriors corresponding to the respective $\beta_j$.
+Subsequently, we numerically integrated the following quantity $S$ times:
+%
+\begin{equation}
+    \ln \hat{E}_i = \frac{1}{2} \sum_{j=0}^{R-1} \left( \beta_{j+1} - \beta_j \right) \cdot \left( \hat{\ell}_{i,j} + \hat{\ell}_{i,j+1} \right)
+\end{equation}
+%
+And then computed the mean and standard deviation of all the integrated $\ln \hat{E}_i$. We then used this for the log-evidence and its error.
+
+Without derivation or insight, we would like to mention that the model evidence naturally balances a model's accuracy against its complexity.
+The value of $\ln E$ will generally be larger (i.e., less negative) if a model fits the data better than another while being similarily complex.
+On the other hand, if e.g.
+additional parameters are introduced without sufficiently improving how well the model explains the data, the evidence will penalize the increase in complexity.
+
+An approximation to the evidence that also attempts to balance accuracy and complexity against each other is the heuristic called \gls{bic}.
+The negative one half of the \gls{bic} approximates the $\ln E$ via Lagrange's method \cite{bhat_derivation_2010} and yields an easy to compute estimate that may also be used to compare models, as long as its underlying assumptions are valid:
+%
+\begin{equation} \label{eq:bic}
+    - \text{BIC} / 2 = \ln{\hat{\mathcal{L}}} - \frac{k}{2} \ln{N} \approx \ln{E}
+\end{equation}
+%
+Here, $\hat{\mathcal{L}} = \max_{\boldsymbol{\theta}}{\left( \ln P \left( \boldsymbol{\mathcal{D}} \mid \boldsymbol{\theta} \right)\right)}$ is the maximum log-likelihood.
+The approximation is good, when the posterior distribution over the parameters $p\left( \boldsymbol{\theta} \mid \boldsymbol{\mathcal{D}} \right)$ is single-modal and falls quickly to zero from the maximum.
+Also, the number of datapoints $N$ needs to be much larger than the number of parameters $k$.
+We will see that for the models we consider here, the \gls{bic} is generally a good approximation and the conclusions drawn from comparing models using this metric can be reproduced reliably using the true model evidence computed with \gls{ti}.
+
+\subsection{Reproducibility}
+\label{subsec:methods:reproducibility}
 
@@
 Tagged references to specific versions of an inference pipeline allow reproducing the inferred parameters of the models described here.
-Prevalence and risk predictions, as well as the figures depicting them, can be reproduced following the instructions of the GitHub repository underlying this publication \href{https://github.com/rmnldwg/graph-extension}{\texttt{rmnldwg/graph-extension}}.
+Every parameter necessary to reproduce such a pipeline is specified there in designated configuration files.
+It also defines the sequences of computations that constitute the pipeline that are mostly calls to commands of the \href{https://pypi.org/project/lyscripts/}{\texttt{lyscripts}} we published.
+
+All figures in this work, like the risk predictions and prevalences we show histograms of, can be reproduced following the instructions of the GitHub repository underlying this publication \href{\GitHubURL}{\texttt{rmnldwg/graph-extension}}.
+It is based off of the \showyourwork project.
 
@@
 
-The dataset $\boldsymbol{\mathcal{D}}$ that we used for inference is comprised of the detailed reports on lymphatic involvement patterns in \gls{opscc} patients treated at three different institutions in France and Switzerland: The \gls{clb} in Lyon (France), the \gls{isb}, and the \gls{usz} (both in Swiztzerland).
+The dataset $\boldsymbol{\mathcal{D}}$ that we used for inference is comprised of the detailed reports on lymphatic involvement patterns in \gls{opscc} patients treated at three different institutions in France and Switzerland: The \gls{clb} in Lyon (France), the \gls{isb}, and the \gls{usz} (both in Switzerland).
 We have previously published the patterns of nodal involvement for the \gls{usz} cohort \cite{ludwig_dataset_2022} and described its characteristics in detail \cite{ludwig_detailed_2022}.
@@
 
+\begin{table*}
+    \centering
+    \variable{output/results_table.tex}
+    \caption{Model comparison results from the base graph and the extended graph we chose as the ``winnning'' model.
+For both \glspl{dag} we show the log-evidence, computed via thermodynamic integration, the negative one half of the \gls{bic}, as well as the maximum log-likelihood that was encountered during the final \gls{mcmc} sampling round.}
+    \label{table:evidence}
+\end{table*}
+
 \begin{figure*}
@@
 \end{figure*}
+
+\subsection{Base Graph}
+\label{subsec:results:base_graph}
 
@@
 
-\begin{table*}
-    \centering
-    \variable{output/results_table.tex}
-    \caption{Model comparison results from the base graph and the extended graph we chose as the ``winnning'' model.
-For both \glspl{dag} we show the log-evidence, computed via thermodynamic integration, the negative one half of the \gls{bic}, as well as the maximum log-likelihood that was encountered during the final \gls{mcmc} sampling round.}
-    \label{table:evidence}
-\end{table*}
+\begin{figure}
+    \script{comp_IIandVII_prevs.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/comp_IIandVII_prevs.png}
+        \caption{Placeholder.}
+        \label{fig:IIandVII_prevs}
+    \end{centering}
+\end{figure}
+
+\begin{figure}
+    \script{comp_IIIandV_prevs.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/comp_IIIandV_prevs.png}
+        \caption{Placeholder.}
+        \label{fig:IIIandV_prevs}
+    \end{centering}
+\end{figure}
+
+\begin{figure}
+    \script{comp_IVandV_prevs.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/comp_IVandV_prevs.png}
+        \caption{Placeholder.}
+        \label{fig:IVandV_prevs}
+    \end{centering}
+\end{figure}
 

### 2022-12-31 src/scripts/compile_data_table.py
finish script for prevalence table

@@ -5,66 +5,74 @@ the paper.
 """
 import re
 
-from jinja2 import Environment, FileSystemLoader
-import requests
-from bs4 import BeautifulSoup
+from jinja2 import Environment, FileSystemLoader, exceptions
 import yaml
+from lyscripts.plot.utils import Posterior
 
 import paths
 
 
-BASE_URL = "https://lyprox.org/dashboard"
-URL_PARAMS_PATH = paths.scripts / "default_dashboard_params.yaml"
-TSTAGE_PARAMS = {
-    "early": {"t_stage__in": 1, "t_stage__in": 2},
-    "late" : {"t_stage__in": 3, "t_stage__in": 4},
-}
-SCENARIO_PARAMS = {
-    "I": {"ipsi_I": 1},
-}
 TEMPLATE_NAME = "data_table.temp"
 OUTPUT = paths.output / "data_table.tex"
+REGEX_PATTERN = r"(^[wb]g_[IV]{1,3}and[IV]{1,3})_.+\.yaml$"
 
 
-with open(URL_PARAMS_PATH, mode="r", encoding="utf-8") as url_params_file:
-    default_url_params = yaml.safe_load(url_params_file)
+scenarios = {"early": [], "late": []}
+yaml_file_paths = paths.scripts.glob("*.yaml")
+for yaml_file_path in yaml_file_paths:
+    regex_match = re.match(REGEX_PATTERN, yaml_file_path.name)
 
-for stage, stage_params in TSTAGE_PARAMS.items():
-    for scenario, scenario_params in SCENARIO_PARAMS.items():
-        url_params = default_url_params.copy()
-        url_params.update(stage_params)
-        url_params.update(scenario_params)
+    if regex_match is None:
+        continue
 
-        response = requests.get(url=BASE_URL, params=url_params, timeout=1000)
-        if not response.ok:
-            raise ConnectionError("Request failed.")
-        soup = BeautifulSoup(response.text, "html.parser")
-        button_text = soup.html.find(
-            "button",
-            class_="button is-medium is-warning is-light is-outlined"
-        ).span.text
-        number = re.match(r"[0-9]{1,3}", button_text)
+    with open(yaml_file_path, mode="r", encoding="utf-8") as yaml_file:
+        yaml_params = yaml.safe_load(yaml_file)
+
+    hdf_file_path = paths.data / f"{regex_match[1]}_prevs.hdf5"
+    for scenario in yaml_params["prevalences"]:
+        posterior = Posterior.from_hdf5(
[…]

### 2023-01-01 src/scripts/thermo_int.py
finish plot of thermodynamic integration

@@ -18,7 +18,7 @@ ACCURACY_PATHS = {
     "win" : paths.data / "win_accuracies.csv",
 }
 OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
-POWER = 4
+POWER = 5
 
 
 if __name__ == "__main__":
@@ -54,13 +54,17 @@ if __name__ == "__main__":
         x="β", y=["base graph", "win graph"], ax=axes,
         color=[USZ["red"], USZ["green"]],
     )
+
     xticks = np.linspace(0., 1., 7)
     xticklabels = [f"{x**POWER:.2g}" for x in xticks]
     axes.set_xticks(xticks)
     axes.set_xticklabels(xticklabels)
     axes.set_xlim(left=0., right=1.)
+    axes.set_xlabel(r"inverse temperature $\beta$")
     axes.set_yscale("symlog",)
     axes.get_yaxis().set_major_locator(matplotlib.ticker.MultipleLocator(800))
     axes.get_yaxis().set_major_formatter(matplotlib.ticker.ScalarFormatter())
     axes.ticklabel_format(axis="y", style="sci", scilimits=(2,2))
+    axes.set_ylabel(r"$\ln \mathcal{A}_{MC}(\beta)$")
+
     fig.savefig(OUTPUT, dpi=400)

### 2023-01-03 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{nicefrac}
+\usepackage{wasysym}
 
@@
 \label{sec:intro}
+
+\begin{table*}
+    \centering
+    \input{output/data_table.tex}
+    \caption{Prevalence of involvement patterns in the multi-centric dataset.
+An involvement pattern is characterized by the state of the six \glspl{lnl}: A red dot means the \gls{lnl} was reported to be metastatic, a green dot means it was determined to be healthy and a question mark means that the prevalence was marginalized over the state of this \gls{lnl}.}
+    \label{table:data_prevalence}
+\end{table*}
 
@@
 
+\begin{figure}
+    \script{thermo_int.py}
+    \centering
+    \includegraphics[width=\linewidth]{figures/thermo_int.png}
+    \caption{Placeholder}
+    \label{fig:thermo_int}
+\end{figure}
+
 \begin{table*}
@@
     \centering
-    \variable{output/results_table.tex}
+    \input{output/results_table.tex}
     \caption{Model comparison results from the base graph and the extended graph we chose as the ``winnning'' model.

### 2023-01-03 src/scripts/compile_data_table.py
make data table prettier and remove duplicates

@@ -5,7 +5,7 @@ the paper.
 """
 import re
 
-from jinja2 import Environment, FileSystemLoader, exceptions
+from jinja2 import Environment, FileSystemLoader
 import yaml
 from lyscripts.plot.utils import Posterior
 
@@ -39,6 +39,10 @@ for yaml_file_path in yaml_file_paths:
             "num_total": posterior.num_total,
         })
         scenario["pattern"] = scenario["pattern"]["ipsi"]
+        
+        if scenario in scenarios[scenario["t_stage"]]:
+            continue
+
         scenarios[scenario["t_stage"]].append(scenario)
 
 
@@ -61,7 +65,7 @@ def get_lnl(value_dict, lnl):
 
 def prev(value):
     percent = int(100. * value["num_success"] / value["num_total"])
-    return f"{value['num_success']}/{value['num_total']} ({percent:d}\%)"
+    return f"{percent:d}\%"
 
 
 environment = Environment(

### 2023-01-06 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\begin{table*}
-    \centering
-    \input{output/data_table.tex}
-    \caption{Prevalence of involvement patterns in the multi-centric dataset.
-An involvement pattern is characterized by the state of the six \glspl{lnl}: A red dot means the \gls{lnl} was reported to be metastatic, a green dot means it was determined to be healthy and a question mark means that the prevalence was marginalized over the state of this \gls{lnl}.}
-    \label{table:data_prevalence}
-\end{table*}
-
 When treating \gls{hnscc}, either with radiotherapy or via surgery, the aim is to irradiate or resect as much of the present malignancies as possible.
@@
 These levels are then often irradiated or resected prophylactically due to the risk of harboring occult metastases despite negative findings from imaging modalities.
-In current clinical practice, \gls{ctv-n} definition is instead mostly based on guidelines \cite{gregoire_ct-based_2003,gregoire_delineation_2014,gregoire_delineation_2018,eisbruch_intensity-modulated_2002,biau_selection_2019,chao_determination_2002,vorwerk_guidelines_2011,ferlito_elective_2009} that are derived from the observed prevalence of involvement in an \gls{lnl} for a given tumor location.
+In current clinical practice, \gls{ctv-n} definition is mostly based on guidelines \cite{gregoire_ct-based_2003,gregoire_delineation_2014,gregoire_delineation_2018,eisbruch_intensity-modulated_2002,biau_selection_2019,chao_determination_2002,vorwerk_guidelines_2011,ferlito_elective_2009} that are derived from the observed prevalence of involvement in an \gls{lnl} for a given tumor location.
 
@@
 
-A model for estimating the risk of microscopic disease, given a personal diagnosis based on \glspl{bn} was previously developed \cite{pouymayou_bayesian_2019} and subsequently adapted to make use of \glspl{hmm} to include T-category in an intuitive manner \cite{ludwig_hidden_2021}.
+A model for estimating the risk of microscopic disease, given a personal diagnosis based on \glspl{bn} was previously developed \cite{pouymayou_bayesian_2019}.
+It was subsequently formulated as a \glspl{hmm} to include T-category in an intuitive manner \cite{ludwig_hidden_2021}.
 These models were trained on a dataset of early T-category (T1 and T2) patients that could be reconstructed from \cite{sanguineti_defining_2009}.
@@
     \includegraphics[width=\linewidth]{figures/thermo_int.png}
-    \caption{Placeholder}
+    \caption{The expected log-accuracy under the power posterior plotted against the value of the inverse temperature $\beta$.
+These values were computed during a \gls{ti} run for the base graph model (red line), as well as the winning graph model (green line).
+Note that the scaling of the x-axis is chosen such that the 64 points on the temperature ladder appear evenly spaced.
+This is to stress how the log-accuracy develops in the range from $\beta=0$ to around $\beta=0.1$.
+The area under these two curves yields an approximation of the log-evidence of the respective model.
+In this plot one can see that due to the simplicity of the base graph model its accuracy increases earlier and faster that the winning graph model's.
+However, already around $\beta=0.1$ the model based on the winning graph outperforms the base graph model.
+Ultimately, the winning graph's log-evidence is better than the base graph model's by a value of around 3 (see \cref{table:evidence}).}
     \label{fig:thermo_int}
@@
 
-\begin{table*}
+\begin{table}
     \centering
@@
         \hline
-        $< 10^0$ & $< 0$ & negative evidence (supports $\mathcal{M}_2$) \\
+        $< 10^0$ & $< 0$ & negative (supports $\mathcal{M}_2$) \\
         $10^0$ to $10^{\nicefrac{1}{2}}$ & 0 to 1.15 & barely worth a mention \\
@@
     \label{table:bayes_factor}
-\end{table*}
+\end{table}
 
@@
 
+\subsection{Investigating Spread Graphs}
+\label{subsec:methods:graphs}
+
+The \gls{dag} shown in \cref{fig:graph_with_obs} includes the \glspl{lnl} I, II, III, and IV since previous models \cite{pouymayou_bayesian_2019,ludwig_hidden_2021} had to rely on a reconstructed dataset of \gls{opscc} patients that contained only information on these four levels \cite{sanguineti_defining_2009}.
+Now, with datasets available that almost always report the \glspl{lnl} I, II, III, IV, V, and VII, we may add more \glspl{rv} to extend the graph and investigate how these \glspl{rv} should be connected among each other.
+
+First, we notice that the direct arcs from tumor to each of the \glspl{lnl} must be present, since every \gls{lnl} appears metastatic in isolation at least once.
+For example, some patients presented with metastases in \gls{lnl} I, while the other levels appeared healthy.
+If no spread was allowed from the tumor to $X_1$ (i.e., $b_i = 0$), the likelihood of observing this patient would be zero.
+
+We have more freedom in choosing how to connect the \glspl{lnl} to each other.
+
 \subsection{Reproducibility}
@@
 \label{sec:data}
+
+\begin{figure}
+    \centering
+    \includegraphics{figures/extended_graph.png}
+    \caption{Extended \gls{dag} representing different possible spread graphs underlying the \gls{hmm}.
+As in \cref{fig:graph_with_obs}, red arcs are parametrized with probabilities of spread from the tumor (red circle) to the \glspl{lnl} (blue circles).
+These red arcs, together with the blue arcs fron \gls{lnl} to \gls{lnl}, make up the \emph{base graph}.
+One after the other, each of the green arcs was added to the base graph.
+Subsequently, the performance of the resulting models in terms of its \gls{bic} was compared to the base graph to asssess whether the additional edge should be kept in the \emph{winning graph} or not.}
+    \label{fig:extended_graph}
+\end{figure}
+
+\begin{table}
+    \centering
+    \variable{output/data_table.tex}
+    \caption{Prevalence of involvement patterns in the multi-centric dataset.
+An involvement pattern is characterized by the state of the six \glspl{lnl}: A red dot means the \gls{lnl} was reported to be metastatic, a green dot means it was determined to be healthy and a question mark means that the prevalence was marginalized over the state of this \gls{lnl}.}
+    \label{table:data_prevalence}
+\end{table}
 
@@
 
-\begin{table*}
+\begin{table}
     \centering
-    \input{output/results_table.tex}
+    \variable{output/results_table.tex}
     \caption{Model comparison results from the base graph and the extended graph we chose as the ``winnning'' model.
@@
     \label{table:evidence}
-\end{table*}
+\end{table}
 
@@
 involvement patterns that include the \glspl{lnl} I, V, and VII that more rarely harbor occult disease.
+
+In \cref{fig:IandII_low_prevs,fig:IandII_high_prevs,fig:IIandVII_prevs,fig:IIIandV_prevs,fig:IVandV_prevs} the base graph's predictions for involvement patterns that include the \glspl{lnl} I, V, and VII have been plotted as gray, hatched histograms alongside the corresponding prevalence in the data and the prediction by the winning graph.
+This allows us to compare how the winning graph improves its fit to the data in these particular scenarios around low-risk \glspl{lnl}.
+How frequently each of the investigated lymphatic metastatic patterns were observed in the data is additionally tabulated in \cref{table:data_prevalence}, both for early and for late T-category cases.
 
@@
         \includegraphics[width=\linewidth]{figures/comp_IIandVII_prevs.png}
-        \caption{Placeholder.}
+        \caption{Prevalences for involvement patterns where \gls{lnl} VII is involved with (red) and without (blue) \gls{lnl} II.
+The solid lines depict a Beta posterior over the prevalence as observed in the data.
+The colored histograms show these prevalences as predicted by the winning graph.
+The hatched histograms show the same predictions from the base graph for comparison.
+Since the winning graph does not include an additional arc between \glspl{lnl} II and VII, the fit w.r.t.
+these two \glspl{LNL} does not improve.}
         \label{fig:IIandVII_prevs}
@@
         \includegraphics[width=\linewidth]{figures/comp_IIIandV_prevs.png}
-        \caption{Placeholder.}
+        \caption{Observations of prevalences in the data (solid lines) compared to predictions from the base (hatched histograms) and the winning graph model (colored histograms).
+The investigated involvement patterns here are how often is \gls{lnl} V involved together with \gls{lnl} III (red) and how often is level V involved while \gls{lnl} III is healthy (green).
+For early T-category cases, the fit of both models to the data is good.
+For late T-category, however, the winning graph shows a better ability to split the two investigated patterns of involvement.
+This is presumably due to the added connection from \gls{lnl} III to V.}
         \label{fig:IIIandV_prevs}
@@
         \includegraphics[width=\linewidth]{figures/comp_IVandV_prevs.png}
-        \caption{Placeholder.}
+        \caption{Predictions vs observations of the prevalence of two patterns of involvement: One for the case when both \gls{lnl} VI and V are metastatic (blue) and one for the case where \gls{lnl} V is involved with level IV being healthy (orange).
+Again, the solid lies depict the Beta posterior over the prevalence of these patterns in the data.
+Colored histograms show the predicted distribution over the two patterns from the winning graph model.
+Hatched histograms show the same predictions from the base graph model for comparison.
+Although no arc was added from \gls{lnl} IV to V, the winning graph slightly improves its predictions both for early (top panel) and late T-category cases (bottom panel).}
         \label{fig:IVandV_prevs}

### 2023-01-06 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.6\linewidth]{figures/extended_graph.png}
+    \caption{Extended \gls{dag} representing different possible spread graphs underlying the \gls{hmm}.
+As in \cref{fig:graph_with_obs}, red arcs are parametrized with probabilities of spread from the tumor (red circle) to the \glspl{lnl} (blue circles).
+These red arcs, together with the blue arcs fron \gls{lnl} to \gls{lnl}, make up the \emph{base graph}.
+One after the other, each of the green arcs was added to the base graph.
+Subsequently, the performance of the resulting models in terms of its \gls{bic} was compared to the base graph to asssess whether the additional edge should be kept in the \emph{winning graph} or not.}
+    \label{fig:extended_graph}
+\end{figure}
+
 The \gls{dag} shown in \cref{fig:graph_with_obs} includes the \glspl{lnl} I, II, III, and IV since previous models \cite{pouymayou_bayesian_2019,ludwig_hidden_2021} had to rely on a reconstructed dataset of \gls{opscc} patients that contained only information on these four levels \cite{sanguineti_defining_2009}.
@@
 We have more freedom in choosing how to connect the \glspl{lnl} to each other.
+To investigate which connections to add we start by establishing a baseline from a model using a minimal \emph{base graph}.
+It contains only the connections from \gls{lnl} II to III and an arc from \gls{lnl} III to IV, as motivated by the \emph{main lymphatic pathway} \cite{lengele_anatomical_2007}.
+In contrast to the \gls{dag} used in \cite{pouymayou_bayesian_2019,ludwig_hidden_2021}, we decided not to include the edge from level I to II into the base graph, since \gls{lnl} I is rarely involved and inferring $t_{12}$ is therefore difficult.
+Rather, we wanted to use the model comparison methodology outlined in \cref{subsec:methods:comparison} to see if the arc from \gls{lnl} I to II can be motivated from our data of if the arc should be flipped and go from \gls{lnl} II to I.
 
@@
 \label{sec:data}
-
-\begin{figure}
-    \centering
-    \includegraphics{figures/extended_graph.png}
-    \caption{Extended \gls{dag} representing different possible spread graphs underlying the \gls{hmm}.
-As in \cref{fig:graph_with_obs}, red arcs are parametrized with probabilities of spread from the tumor (red circle) to the \glspl{lnl} (blue circles).
-These red arcs, together with the blue arcs fron \gls{lnl} to \gls{lnl}, make up the \emph{base graph}.
-One after the other, each of the green arcs was added to the base graph.
-Subsequently, the performance of the resulting models in terms of its \gls{bic} was compared to the base graph to asssess whether the additional edge should be kept in the \emph{winning graph} or not.}
-    \label{fig:extended_graph}
-\end{figure}
 
@@
 Since the winning graph does not include an additional arc between \glspl{lnl} II and VII, the fit w.r.t.
-these two \glspl{LNL} does not improve.}
+these two \glspl{lnl} does not improve.}
         \label{fig:IIandVII_prevs}

### 2023-01-06 src/scripts/compile_data_table.py
include bg_core prevs in table

@@ -14,7 +14,7 @@ import paths
 
 TEMPLATE_NAME = "data_table.temp"
 OUTPUT = paths.output / "data_table.tex"
-REGEX_PATTERN = r"(^[wb]g_[IV]{1,3}and[IV]{1,3})_.+\.yaml$"
+REGEX_PATTERN = r"(^[wb]g_([IV]{1,3}and[IV]{1,3}|core))_.+\.yaml$"
 
 
 scenarios = {"early": [], "late": []}

### 2023-01-11 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %
-The term $P\left( \boldsymbol{\xi}_i \mid \boldsymbol{\xi}_j \right)$ describes the probability to transition from the hidden state of lymphatic involvement $\boldsymbol{\xi}_j$ to the state $\boldsymbol{\xi}_i$ within between the time $t$ and $t+1$.
+The term $P\left( \boldsymbol{\xi}_i \mid \boldsymbol{\xi}_j \right)$ describes the probability to transition from the hidden state of lymphatic involvement $\boldsymbol{\xi}_j$ to the state $\boldsymbol{\xi}_i$ between the time $t$ and $t+1$.
 Using a \gls{dag} as depicted in \cref{fig:graph_with_obs}, we can parameterize this transition probability in the following way:
@@
 Also, $\xi_{iv}$ denotes the value that \gls{lnl} $v$ takes on when the patient is in state $\boldsymbol{\xi}_i$.
-Lastly, the term $Q(a;b) \in \{ 0,1 \}$ is there only to
+Lastly, the term $Q(a;b) \in \{ 0,1 \}$ is there to prohibit self-healing.
+It is always one, except if \gls{lnl} $v$ is healthy in state $\xi_i$, but was metastatic in the previous state $\xi_j$.
+In that case the function becomes $Q(0;1) = 0$, making the transition back to healthier states impossible.
 
@@
 
+Anatomically, the \emph{posterior accessory pathway} that drains \gls{lnl} II through \gls{lnl} V motivates investigating and arc from $X_2$ into $X_5$ \cite{lengele_anatomical_2007}.
+And, although no lymphatic pathway is described that directly drains \gls{lnl} III or IV into level V, due to their proximity to each other, we will also investigate additional arcs from \gls{lnl} III and IV into level V.
+
+Lastly, we also compared how the model's performance improves with an added connection from the most commonly involved level II to one of the most rarely metastatic \glspl{lnl} VII.
+
 \subsection{Reproducibility}
@@
 In \cref{fig:IandII_low_prevs,fig:IandII_high_prevs,fig:IIandVII_prevs,fig:IIIandV_prevs,fig:IVandV_prevs} the base graph's predictions for involvement patterns that include the \glspl{lnl} I, V, and VII have been plotted as gray, hatched histograms alongside the corresponding prevalence in the data and the prediction by the winning graph.
+Every scenario is show for early and late T-category.
 This allows us to compare how the winning graph improves its fit to the data in these particular scenarios around low-risk \glspl{lnl}.
-How frequently each of the investigated lymphatic metastatic patterns were observed in the data is additionally tabulated in \cref{table:data_prevalence}, both for early and for late T-category cases.
+How frequently each of the investigated lymphatic metastatic patterns were observed in the data is additionally tabulated in \cref{table:data_prevalence}, again both for early and for late T-category cases.
 

### 2023-01-11 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 This work aims to do that and at the same time demonstrate the capability
-of the previously introduced \gls{hmm} to adapt to real patient data from a multi-centric cohort of XXX \gls{opscc} patients.
+of the previously introduced \gls{hmm} to adapt to real patient data from a multi-centric cohort of \variable{output/num_patients.txt} \gls{opscc} patients.
 

### 2023-01-11 src/scripts/get_num_patients.py
insert total number of patients dynamically

@@ -1,0 +1,18 @@
+"""
+Extract the number of patients used for training.
+"""
+import pandas as pd
+
+import paths
+
+
+INPUT = paths.data / "patients.csv"
+OUTPUT = paths.output / "num_patients.txt"
+
+
+if __name__ == "__main__":
+    patients = pd.read_csv(INPUT, header=[0,1])
+    num_patients = len(patients)
+
+    with open(OUTPUT, mode="w", encoding="utf-8") as txt_file:
+        txt_file.write(str(num_patients))

### 2023-03-09 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2023-01-11 src/tex/ms.tex
@@
 \usepackage{wasysym}
+\usepackage{amssymb}
 
@@
 \begin{abstract}
-    The \gls{ctv-n}\glsunset{ctv} definition in \gls{hnscc}\glsunset{scc} is currently based mostly on the prevalence of lymphatic metastases in different \glspl{lnl} based on the primary tumor location.
-In this work, we present an extension to a probabilistic model for lymphatic metastatic spread developed earlier that can quantify the risk for microscopic nodal involvement based on individual diagnoses.
+    The \gls{ctv-n}\glsunset{ctv} in \gls{hnscc}\glsunset{scc} is currently based mostly on the prevalence of lymphatic metastases in different \glspl{lnl} for a given primary tumor location.
+In this work, we present an extension to a probabilistic model for lymphatic metastatic spread developed earlier that can quantify the risk for microscopic nodal involvement based on an individual patient diagnosis.
 The extension is based on the same formalism of \glspl{hmm} as the original model, but in addition to the \glspl{lnl} I, II, III, and IV, it also covers the \glspl{lnl} V and VII of the ipsilateral neck.
@@
 
-With the publication of larger and more detailed datasets \cite{ludwig_dataset_2022}, the risk estimation models may be extended as well to include the additionally reported \glspl{lnl} V and VII.
-This work aims to do that and at the same time demonstrate the capability
-of the previously introduced \gls{hmm} to adapt to real patient data from a multi-centric cohort of XXX \gls{opscc} patients.
+With the publication of larger and more detailed datasets \cite{ludwig_dataset_2022}, we are thus able to
+\begin{enumerate}
+    \item include all relevant \glspl{lnl} into the model, namely the levels I, II, III, IV, V, and VII.
+    \item train the probabilistic model based on a multi-centric dataset consisting of \variable{output/num_patients.txt} patients from %\variable{output/num_institutions.txt} institutions.
+    \item present exaples of results on personalized risk estimations performed with the trained model.
+\end{enumerate}
 
-\section{Previous Work}
-\label{sec:previous_work}
+\section{HMM Formalism and Notation}
+\label{sec:formalism}
 
-\subsection{Lymphatic Spread as Hidden Markov Model}
-\label{subsec:previous_work:spread_as_hmm}
+\glsreset{bn}
 
-We have introduced a probabilistic model for lymphatic metastatic tumor progression based on \glspl{bn} in \cite{pouymayou_bayesian_2019} and based off of this an improved model using \glspl{hmm} in \cite{ludwig_hidden_2021}.
-We will briefly recap the \acrlong{hmm} below.
+\subsection{State of the Hidden Markov Model}
+\label{subsec:formalism:state}
+
+We have introduced a probabilistic model for lymph node involvement based on \glspl{bn} in \cite{pouymayou_bayesian_2019}.
+The model was extended using \glspl{hmm} in \cite{ludwig_hidden_2021}.
+We will briefly recap the \acrlong{hmm} to introduce the notation used throughout the work.
 
@@
 %
-Where each of the \glspl{lnl} can be in the state $X_v=0$ (\texttt{FALSE}), meaning \gls{lnl} $v$ is healthy, or in the state $X_v=1$ (\texttt{TRUE}), indicating the \gls{lnl} harbors metastases.
+Each the \gls{lnl} can be in the state $X_v=0$ (\texttt{FALSE}), meaning \gls{lnl} $v$ is healthy, or in the state $X_v=1$ (\texttt{TRUE}), indicating the \gls{lnl} harbors metastases.
+The involved state includes occult disease.
 
@@
 The term $P\left( \boldsymbol{\xi}_i \mid \boldsymbol{\xi}_j \right)$ describes the probability to transition from the hidden state of lymphatic involvement $\boldsymbol{\xi}_j$ to the state $\boldsymbol{\xi}_i$ between the time $t$ and $t+1$.
-Using a \gls{dag} as depicted in \cref{fig:graph_with_obs}, we can parameterize this transition probability in the following way:
+Using a \gls{dag} as depicted in \cref{fig:graph_with_obs}, we can formulate this transition probability in the following way:
 %
@@
 Also, $\xi_{iv}$ denotes the value that \gls{lnl} $v$ takes on when the patient is in state $\boldsymbol{\xi}_i$.
-Lastly, the term $Q(a;b) \in \{ 0,1 \}$ is there to prohibit self-healing.
+The term $Q(a;b) \in \{ 0,1 \}$ is there to prohibit self-healing.
 It is always one, except if \gls{lnl} $v$ is healthy in state $\xi_i$, but was metastatic in the previous state $\xi_j$.
 In that case the function becomes $Q(0;1) = 0$, making the transition back to healthier states impossible.
+
+The terms of the form $P \left( \xi_{iv} \mid \left\{ \xi_{jr} \right\}_{r \in \operatorname{pa}(v)} \right)$ implicitly depend on how we parametrize e.g.
+the arcs of \cref{fig:graph_with_obs}.
+For example, if we look at the probability of spread to the \gls{lnl} $X_3$ depending on the state of that level's parent -- which, in this case, is $\operatorname{pa}(3) = 2$ -- we can write the different combinations into a conditional probability table as below.
+
+\noindent
+\begin{center}
+    \begin{tabular}{|cc|cc|}
+        \hline
+        & & \multicolumn{2}{c|}{$X_2$} \\
+        & & 0 & 1 \\
+        \hline
+        \multirow{2}{*}{$X_3$} & 0 & $1 - b_3$ & $(1 - b_3)(1 - t_{23})$ \\
+        & 1 & $b_3$ & $1 - b_3 - t_{23} + b_3 t_{23}$ \\
+        \hline
+    \end{tabular}
+\end{center}
+
+The variable $b_3$ denotes the probability of lymphatic spread from the tumor to \gls{lnl} III during one time-step, and $t_{23}$ is the probability of spread from an involved level II further down the lymphatic chain into \gls{lnl} III.
 
@@
     \includegraphics[width=0.6\linewidth]{figures/graph_with_obs.png}
-    \caption{\Gls{dag} representing a possible abstraction of the lymphatic network comprising the tumor (red shaded circle) and \glspl{lnl} I through IV as hidden binary \glspl{rv} (blue outlined circles).
+    \caption{\Gls{dag} representing a possible abstraction of the lymphatic network comprising the tumor (red shaded circle) and \glspl{lnl} II through IV as hidden binary \glspl{rv} (blue outlined circles).
 Attached to each of these is the corresponding observed \gls{rv} (orange shaded squares).
-Lymphatic flow is depicted in the form of parameterized arrows (red and blue) that represent the probability of spread along the respective arc per time-step.
+Lymphatic flow is depicted in the form of parametrized arrows (red and blue) that represent the probability of spread along the respective arc per time-step.
 Sensitivity and specificity (orange arrows) connect the hidden \glspl{rv} to the diagnosis.}
@@
 
+\subsection{Diagnostic Observation}
+\label{subsec:formalism:diagnosis}
+
+We also need to introduce a separate collection of \glspl{rv} that describe the diagnostic observation of a patient's involvement.
+In analogy to the hidden true state $\mathbf{X}[t]$ at time $t$, we write this diagnosis as
+%
+\begin{equation}
+    \mathbf{Z} = \left( Z_v \right) \qquad v \in \left\{ 1,2, \ldots, V \right\}
+\end{equation}
+%
+We do not need to differentiate between different times $t$ here, since a patient is ever only diagnosed once, after which treatment usually starts timely.
 Diagnosis and true state of a patient are formally connected via the sensitivity $s_N$ and specificity $s_P$ of the used diagnostic modality.
@@
 
-If multiple diagnostic modalities $\mathcal{O} \in \{ \text{MRI}, \text{CT}, \ldots \}$ are used, one can compute an observation matrix $\mathbf{B}^\mathcal{O}$ for each of these modalities and combine them via an outer product into a joint observations matrix:
-%
-\begin{equation}
-    \mathbf{B} = \mathbf{B}^\text{MRI} \otimes \mathbf{B}^\text{CT} \otimes \ldots
-\end{equation}
-
 We define the time $t=0$ to be the moment just before a patient's tumor formed, and hence $X_v[t=0]=0 \,\,\, \forall v$.
@@
 %
-\begin{equation}
-    P \left( t = t_D \mid \text{T-cat.} \right) = \operatorname{B}(t_\text{max},p_\text{T-cat.})
+\begin{equation} \label{eq:time_dist}
+    P \left( t = t_D \mid \text{T}x \right) = \operatorname{\mathfrak{B}}(t_\text{max},p_{\text{T}x})
 \end{equation}
 %
-Here, the parameter $p_\text{early}$ can be interpreted as the probability that the patient will be diagnosed at time-step $t+1$ given they are in time-step $t$.
-We will use as a latest time-step $t_\text{max} = 10$.
+Here, the parameter $p_{\text{T}x}$ can be interpreted as the probability that the patient with a tumor of T-category $x$ will be diagnosed at time-step $t+1$ given they are in time-step $t$.
+We will use as the latest time-step $t_\text{max} = 10$, which will therefore give us a distribution over the diagnosis time that has its mean at $\mathbb{E}[t_D] = 10 \cdot p_{\text{T}x}$.
 
 \subsection{The Likelihood Function}
-\label{subsec:previous_work:likelihood}
+\label{subsec:formalism:likelihood}
 
@@
     \boldsymbol{\ell} &= \big( P\left( \mathbf{Z} = \boldsymbol{\zeta}_i \right) \big) \\
-    &= \sum_{t=0}^{t_{max}} \left[ \boldsymbol{\pi} \cdot \mathbf{A}^t \cdot \mathbf{B} \right] \cdot P \left( t \mid \text{T-cat.} \right)
+    &= \sum_{t=0}^{t_{max}} \left[ \boldsymbol{\pi} \cdot \mathbf{A}^t \cdot \mathbf{B} \right] \cdot P \left( t \mid \text{T} \right)
 \end{split}
@@
 %
-This likelihood implicitly depends on how we parameterize the arcs of the \gls{dag} underlying the model and the parameterization of the distribution over diagnosis times.
-If we use the graph displayed in \cref{fig:graph_with_obs}, we can write down a conditional probability table for \gls{lnl} $X_3$ being healthy or metastatic, based on the state of the parent \gls{lnl} $\operatorname{pa}(3) = 2$:
+This likelihood implicitly depends on how we parameterize the arcs of the \gls{dag} underlying the model -- see \cref{eq:transition_prob} -- and the parameterization of the distribution over diagnosis times -- e.g., as in \cref{eq:time_dist}.
 
-\noindent
-\begin{center}
-    \begin{tabular}{|cc|cc|}
-        \hline
-        & & \multicolumn{2}{c|}{$X_2$} \\
-        & & 0 & 1 \\
-        \hline
-        \multirow{2}{*}{$X_3$} & 0 & $1 - b_3$ & $(1 - b_3)(1 - t_{23})$ \\
-        & 1 & $b_3$ & $1 - b_3 - t_{23} + b_3 t_{23}$ \\
-        \hline
-    \end{tabular}
-\end{center}
-
-Similar tables can be written for all other \glspl{lnl}.
-The terms in these conditional probability tables can then be used to compute the transition probability between to particular states $\boldsymbol{\xi}_i$ and $\boldsymbol{\xi}_j$ as in \cref{eq:transition_prob}.
-Computing this quantity for all possible transitions subsequently yields the transition matrix $\mathbf{A}$ as described in \cref{eq:transition_matrix}.
-
-Together with the parametrizations of the distributions over the diagnosis time, the parameters that make up the transition matrix $\mathbf{A}$ comprises the set of model parameters:
+Together with the parametrizations of the distributions over the diagnosis time, the parameters $b_v$ and $t_{vr}$ that make up the transition matrix $\mathbf{A}$ comprises the set of model parameters:
 %
@@
 Using this log-likelihood function one may now employ a variety of inference methods to learn the parameters of the model that best describe the observed data.
-After that, the main task of of personalizing the \gls{ctv-n} definition is to predict the probability of the hidden possible states $\boldsymbol{\xi}_k$ given the diagnosis $d^\star=\boldsymbol{\zeta}_\ell$ of a new patient.
-Using Bayes' theorem, we get
-%
-\begin{equation}
-    P\left( \mathbf{X}=\boldsymbol{\xi}_k \mid \mathbf{Z}=\boldsymbol{\zeta}_\ell \right) = \frac{P\left( \boldsymbol{\zeta}_\ell \mid \boldsymbol{\xi}_k \right) P\left( \boldsymbol{\xi}_k \mid \boldsymbol{\theta} \right)}{\sum_{r=1}^{2^V} P\left( \boldsymbol{\zeta}_\ell \mid \boldsymbol{\xi}_r \right) P\left( \boldsymbol{\xi}_r \mid \boldsymbol{\theta} \right) }
-\end{equation}
-%
-The described model along with the inferred parameters $\boldsymbol{\hat{\theta}}$ will yield an estimate (or multiple estimates) for the ``prior'' in the above equation $P\big( \boldsymbol{\xi}_k \mid \boldsymbol{\hat{\theta}} \big)$.
-
-\section{Methods}
-\label{sec:methods}
 
 \subsection{Parameter Inference}
-\label{subsec:methods:inference}
+\label{subsec:formalism:inference}
 
@@
 
-For the experiments in this work, we used $W = 20 \cdot k$ walkers, where $k$ is the dimensionaliry of the parameter space $\boldsymbol{\Theta}$.
-After an initial ``burn-in'' phase, during which all drawn samples are discarded because they are not yet independent from the initial state, we continued sampling for another 200 steps of which we discarded every 10th to be left with $S = 20 \cdot W$ samples.
+For the experiments in this work, we used $W = 20 \cdot k$ walkers, where $k$ is the dimensionality of the parameter space $\boldsymbol{\Theta}$.
+After an initial ``burn-in'' phase, during which all drawn samples are discarded because they are not yet independent of the initial state, we continued sampling for another 200 steps of which we discarded every 10th to be left with $S = 20 \cdot W$ samples.
 
@@
 
+\subsection{Risk estimation}
+\label{subsec:formalism:risk}
+
+Now, the main task of personalizing the \gls{ctv-n} definition is to predict the probability of the hidden possible states $\boldsymbol{\xi}_k$ given the diagnosis $d^\star=\boldsymbol{\zeta}_\ell$ of a new patient.
+Using Bayes' theorem, we get
+%
+\begin{equation}
+    P\left( \mathbf{X}=\boldsymbol{\xi}_k \mid \mathbf{Z}=\boldsymbol{\zeta}_\ell \right) = \frac{P\left( \boldsymbol{\zeta}_\ell \mid \boldsymbol{\xi}_k \right) P\left( \boldsymbol{\xi}_k \mid \boldsymbol{\theta} \right)}{\sum_{r=1}^{2^V} P\left( \boldsymbol{\zeta}_\ell \mid \boldsymbol{\xi}_r \right) P\left( \boldsymbol{\xi}_r \mid \boldsymbol{\theta} \right) }
+\end{equation}
+%
+The described model along with the inferred parameters $\boldsymbol{\hat{\theta}}$ will yield an estimate (or multiple estimates) for the ``prior'' in the above equation $P\big( \boldsymbol{\xi}_k \mid \boldsymbol{\hat{\theta}} \big)$.
+
+\section{Complete Model of Ipsilateral Spread in OPSCC}
+\label{sec:complete_model}
+
+\subsection{Investigating Spread Graphs}
+\label{subsec:complete_model:graphs}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.6\linewidth]{figures/extended_graph.png}
+    \caption{Extended \gls{dag} representing different possible spread graphs underlying the \gls{hmm}.
+As in \cref{fig:graph_with_obs}, red arcs are parametrized with probabilities of spread from the tumor (red circle) to the \glspl{lnl} (blue circles).
+These red arcs, together with the blue arcs fron \gls{lnl} to \gls{lnl}, make up the \emph{base graph}.
+One after the other, each of the green arcs was added to the base graph.
+Subsequently, the performance of the resulting models in terms of its \gls{bic} was compared to the base graph to asssess whether the additional edge should be kept in the \emph{winning graph} or not.}
+    \label{fig:extended_graph}
+\end{figure}
+
+The \gls{dag} shown in \cref{fig:graph_with_obs} includes the \glspl{lnl} II, III, and IV since previous models \cite{pouymayou_bayesian_2019,ludwig_hidden_2021} had to rely on a reconstructed dataset of \gls{opscc} patients that contained only information on \gls{lnl} I and these three levels \cite{sanguineti_defining_2009}.
+Now, with datasets available that almost always report the \glspl{lnl} I, II, III, IV, V, and VII, we may add more \glspl{rv} to extend the graph and investigate how these \glspl{rv} should be connected among each other.
+
+First, we notice that the direct arcs from tumor to each of the \glspl{lnl} must be present, since every \gls{lnl} appears metastatic in isolation at least once.
+For example, some patients presented with metastases in \gls{lnl} I, while the other levels appeared healthy.
+If no spread was allowed from the tumor to $X_1$ (i.e., $b_i = 0$), the likelihood of observing this patient would be zero.
+
+We have more freedom in choosing how to connect the \glspl{lnl} to each other.
+To investigate which connections to add we start by establishing a baseline from a model using a minimal \emph{base graph}.
+It contains only the connections from \gls{lnl} II to III and an arc from \gls{lnl} III to IV, as motivated by the \emph{main lymphatic pathway} \cite{lengele_anatomical_2007}.
+In contrast to the \gls{dag} used in \cite{pouymayou_bayesian_2019,ludwig_hidden_2021}, we decided not to include the edge from level I to II into the base graph, since \gls{lnl} I is rarely involved and inferring $t_{12}$ is therefore difficult.
+Rather, we wanted to use the model comparison methodology outlined in \cref{subsec:complete_model:comparison} to see if the arc from \gls{lnl} I to II can be motivated from our data or if the arc should be flipped and go from \gls{lnl} II to I.
+
+Anatomically, the \emph{posterior accessory pathway} that drains \gls{lnl} II through \gls{lnl} V motivates investigating and arc from $X_2$ into $X_5$ \cite{lengele_anatomical_2007}.
+And, although no lymphatic pathway is described that directly drains \gls{lnl} III or IV into level V, due to their proximity to each other, we will also investigate additional arcs from \gls{lnl} III and IV into level V.
+
+Lastly, we also compared how the model's performance improves with an added connection from the most commonly involved level II to one of the most rarely metastatic \glspl{lnl} VII.
+
 \subsection{Model Comparison}
-\label{subsec:methods:comparison}
+\label{subsec:complete_model:comparison}
 
@@
 In our case, the number of dimensions ranges from $k=9$ for the \emph{base graph} to $k=11$ for the \emph{winning graph}.
-A brute-force integration over a unit cube with this many dimensions is inefficient and error prone, which is why we resorted to \gls{ti} for computing the (log-)evidence.
+A brute-force integration over a unit cube with this many dimensions is inefficient and error-prone, which is why we resorted to \gls{ti} for computing the (log-)evidence.
 
@@
 The variable $\beta$ is also often referred to as an \emph{inverse temperature}, due to its origins in statistical physics.
-Often when performing \gls{ti}, the most drastic changes in the values of the $\mathcal{A}_\text{MC}$ occur at high temperatures (meaning $\beta$ very close to zero), while the changes become smaller and smaller for lower temparatures ($\beta$ towards one).
+Often when performing \gls{ti}, the most drastic changes in the values of the $\mathcal{A}_\text{MC}$ occur at high temperatures (meaning $\beta$ very close to zero), while the changes become smaller and smaller for lower temperatures ($\beta$ towards one).
 It is therefore efficient to space the \emph{temperature ladder} unevenly, e.g.
@@
 
-\subsection{Investigating Spread Graphs}
-\label{subsec:methods:graphs}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.6\linewidth]{figures/extended_graph.png}
-    \caption{Extended \gls{dag} representing different possible spread graphs underlying the \gls{hmm}.
-As in \cref{fig:graph_with_obs}, red arcs are parametrized with probabilities of spread from the tumor (red circle) to the \glspl{lnl} (blue circles).
-These red arcs, together with the blue arcs fron \gls{lnl} to \gls{lnl}, make up the \emph{base graph}.
-One after the other, each of the green arcs was added to the base graph.
-Subsequently, the performance of the resulting models in terms of its \gls{bic} was compared to the base graph to asssess whether the additional edge should be kept in the \emph{winning graph} or not.}
-    \label{fig:extended_graph}
-\end{figure}
-
-The \gls{dag} shown in \cref{fig:graph_with_obs} includes the \glspl{lnl} I, II, III, and IV since previous models \cite{pouymayou_bayesian_2019,ludwig_hidden_2021} had to rely on a reconstructed dataset of \gls{opscc} patients that contained only information on these four levels \cite{sanguineti_defining_2009}.
-Now, with datasets available that almost always report the \glspl{lnl} I, II, III, IV, V, and VII, we may add more \glspl{rv} to extend the graph and investigate how these \glspl{rv} should be connected among each other.
-
-First, we notice that the direct arcs from tumor to each of the \glspl{lnl} must be present, since every \gls{lnl} appears metastatic in isolation at least once.
-For example, some patients presented with metastases in \gls{lnl} I, while the other levels appeared healthy.
-If no spread was allowed from the tumor to $X_1$ (i.e., $b_i = 0$), the likelihood of observing this patient would be zero.
-
-We have more freedom in choosing how to connect the \glspl{lnl} to each other.
-To investigate which connections to add we start by establishing a baseline from a model using a minimal \emph{base graph}.
-It contains only the connections from \gls{lnl} II to III and an arc from \gls{lnl} III to IV, as motivated by the \emph{main lymphatic pathway} \cite{lengele_anatomical_2007}.
-In contrast to the \gls{dag} used in \cite{pouymayou_bayesian_2019,ludwig_hidden_2021}, we decided not to include the edge from level I to II into the base graph, since \gls{lnl} I is rarely involved and inferring $t_{12}$ is therefore difficult.
-Rather, we wanted to use the model comparison methodology outlined in \cref{subsec:methods:comparison} to see if the arc from \gls{lnl} I to II can be motivated from our data of if the arc should be flipped and go from \gls{lnl} II to I.
-
-Anatomically, the \emph{posterior accessory pathway} that drains \gls{lnl} II through \gls{lnl} V motivates investigating and arc from $X_2$ into $X_5$ \cite{lengele_anatomical_2007}.
-And, although no lymphatic pathway is described that directly drains \gls{lnl} III or IV into level V, due to their proximity to each other, we will also investigate additional arcs from \gls{lnl} III and IV into level V.
-
-Lastly, we also compared how the model's performance improves with an added connection from the most commonly involved level II to one of the most rarely metastatic \glspl{lnl} VII.
-
 \subsection{Reproducibility}
-\label{subsec:methods:reproducibility}
+\label{subsec:complete_model:reproducibility}
 
@@
 
-\subsection{Base Graph}
-\label{subsec:results:base_graph}
+\subsection{Involvement of Levels II, III, and IV}
+\label{subsec:results:lnls_II_III_IV}
 
@@
 
+\subsection{The Optimal Extended Graph}
+\label{subsec:results:optimal_graph}
+
+\subsection{Risk Prediction for Occult Disease}
+\label{subsec:results:risk_prediction}
+
 \bibliography{bib}

### 2023-03-09 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     \item include all relevant \glspl{lnl} into the model, namely the levels I, II, III, IV, V, and VII.
-    \item train the probabilistic model based on a multi-centric dataset consisting of \variable{output/num_patients.txt} patients from %\variable{output/num_institutions.txt} institutions.
+    \item train the probabilistic model based on a multi-centric dataset consisting of \variable{output/num_patients.txt} patients from four institutions.
     \item present exaples of results on personalized risk estimations performed with the trained model.
@@
 For the base graph, we have plotted the predicted prevalence of involvement patterns in the investigated patient cohort for scenarios involving the most commonly metastatic \glspl{lnl} II, III and IV in \cref{fig:bg_prevalences}.
-This figure shows that this minimal graph is already capable of describing the most important parts of the observed data very well.
-The subsequent changes to the graph connections we investigated therefore focus on the improvements w.r.t.
-involvement patterns that include the \glspl{lnl} I, V, and VII that more rarely harbor occult disease.
+It shows -- for each pattern of lymphatic involvement -- two plots overlaid:
 
-In \cref{fig:IandII_low_prevs,fig:IandII_high_prevs,fig:IIandVII_prevs,fig:IIIandV_prevs,fig:IVandV_prevs} the base graph's predictions for involvement patterns that include the \glspl{lnl} I, V, and VII have been plotted as gray, hatched histograms alongside the corresponding prevalence in the data and the prediction by the winning graph.
-Every scenario is show for early and late T-category.
+\begin{enumerate}
+    \item The colored histograms over the base graph model's prediction for the prevalence of the respective pattern of involvement.
+These histograms are obtained by computing the same prevalence with different samples from the inference process, thus providing us with a measure of uncertainty for the prediction.
+    \item Colored lines, depicting the Beta posterior over the same involvement pattern's prevalence, given a uniform prior and the observed data as likelihood.
+The maximum of the Beta distributions always coincides with the data prevalence but we additionally gain an intuition into how statistically significant the data is.
+E.g., Observing 3 out of 10 patients with a particular pattern of nodal metastases is less convincing than 300 out of a cohort of 1000 patients.
+A Beta posterior over these prevalences reflects that in its variance.
+\end{enumerate}
+
+\Cref{fig:bg_prevalences} shows that this minimal graph is already capable of describing the most important parts of the observed data very well.
+Notably, the model is not only accurate in its predictions, it also correctly estimates the variance stemming from the limited amount of data.
+The spread between involvement prevalences of early and late T-category tumors is also represented well.
+This shows that expecting later diagnosis times, on average, for patients with advanced T-category tumors is able to explain more severe lymphatic involvement.
+
+\subsection{The Optimal Extended Graph}
+\label{subsec:results:optimal_graph}
+
+The changes to the graph connections we investigate in this work focus on the improvements w.r.t.
+involvement patterns that include the \glspl{lnl} I, V, and VII, that more rarely harbor occult disease.
+
+In \cref{fig:IandII_low_prevs,fig:IandII_high_prevs,fig:IIIandV_prevs} the base graph's estimations for prevalences of involvement patterns that include the \glspl{lnl} I, V, and VII have been plotted as gray, hatched histograms.
+Alongside, we show again the Beta posterior over the prevalence of the respective involvement pattern, given the data, as colored line.
+The prediction by the winning graph are shown as colored histograms.
+Every scenario is shown for early and late T-category.
 This allows us to compare how the winning graph improves its fit to the data in these particular scenarios around low-risk \glspl{lnl}.
@@
 
-\begin{figure}
-    \script{comp_IIandVII_prevs.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/comp_IIandVII_prevs.png}
-        \caption{Prevalences for involvement patterns where \gls{lnl} VII is involved with (red) and without (blue) \gls{lnl} II.
+% \begin{figure}
+%     \script{comp_IIandVII_prevs.py}
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{figures/comp_IIandVII_prevs.png}
+%         \caption{Prevalences for involvement patterns where \gls{lnl} VII is involved with (red) and without (blue) \gls{lnl} II.
 The solid lines depict a Beta posterior over the prevalence as observed in the data.
@@
 these two \glspl{lnl} does not improve.}
-        \label{fig:IIandVII_prevs}
-    \end{centering}
-\end{figure}
+%         \label{fig:IIandVII_prevs}
+%     \end{centering}
+% \end{figure}
 
@@
 
-\begin{figure}
-    \script{comp_IVandV_prevs.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/comp_IVandV_prevs.png}
-        \caption{Predictions vs observations of the prevalence of two patterns of involvement: One for the case when both \gls{lnl} VI and V are metastatic (blue) and one for the case where \gls{lnl} V is involved with level IV being healthy (orange).
+% \begin{figure}
+%     \script{comp_IVandV_prevs.py}
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{figures/comp_IVandV_prevs.png}
+%         \caption{Predictions vs observations of the prevalence of two patterns of involvement: One for the case when both \gls{lnl} VI and V are metastatic (blue) and one for the case where \gls{lnl} V is involved with level IV being healthy (orange).
 Again, the solid lies depict the Beta posterior over the prevalence of these patterns in the data.
@@
 Although no arc was added from \gls{lnl} IV to V, the winning graph slightly improves its predictions both for early (top panel) and late T-category cases (bottom panel).}
-        \label{fig:IVandV_prevs}
-    \end{centering}
-\end{figure}
-
-\subsection{The Optimal Extended Graph}
-\label{subsec:results:optimal_graph}
+%         \label{fig:IVandV_prevs}
+%     \end{centering}
+% \end{figure}
 

### 2023-03-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-In total, the dataset contains 681 patients newly diagnosed with \gls{opscc} that were treated with either definitive radiotherapy or neck dissection.
+In total, the dataset contains \variable{output/num_patients.txt} patients newly diagnosed with \gls{opscc} that were treated with either definitive radiotherapy or neck dissection.
 Pathologically assessed post-operative \gls{lnl} involvement was available for 393 of patients, while for the remainder the nodal involvement was assessed based on available diagnostic modalities.
 If multiple modalities were used to diagnose a patient's lymphatic involvement and conflicting conclusions were drawn from this, the conflicts were resolved by inferring the most likely state (healthy or metastatic) for each \gls{lnl} separately.
-To do so, we used literature values for the sensitivity and specificity of the diagnostic modalities \cite{de_bondt_detection_2007,kyzas_18f-fluorodeoxyglucose_2008}, which we also tabulated below.
+To do so, we used literature values for the sensitivity and specificity of the diagnostic modalities \cite{de_bondt_detection_2007,kyzas_18f-fluorodeoxyglucose_2008}, which we also tabulated in \cref{table:sens_spec}.
 
-\noindent
-\begin{center}
+\begin{table}
+\centering
     \begin{tabular}{|l|rr|}
@@
     \end{tabular}
-\end{center}
+    \caption{Literature sensitivity and specificity values that we used to infer the most likely involvement for a patient when multiple diagnostic modalities reported conflicting nodal involvement \cite{de_bondt_detection_2007,kyzas_18f-fluorodeoxyglucose_2008}.}
+    \label{table:sens_spec}
+\end{table}
 

### 2023-03-14 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
+There are only a few involvement patterns for which the model's prediction does not match the prevalence of the data with perfect accuracy.
+But in these instances, e.g.
+the case where \gls{lnl} III is metastatic while \gls{lnl} IV is healthy in an early T-category patient (\cref{fig:bg_prevalences}, third row, green line and histogram), the model slightly overestimates the prevalence.
+An overestimation by a few percent however, does not pose a risk to the model's clinical applicability, since in the context of volume-deescalation, this can be considered cautious.
+
+It is interesting to note that not all involvement patterns become more prevalent with advanced T-category.
+For example, a healthy \gls{lnl} III is observed slightly less often together with a metastatic level II in advanced, compared to early T-category.
+This is because in the latter instance, it is more likely the disease has already spread to \gls{lnl} III.
+Our model captures this accurately and precisely.
+
 \subsection{The Optimal Extended Graph}
@@
 \begin{figure}
+    \script{comp_IIandV_prevs.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/comp_IIandV_prevs.png}
+        \caption{TODO: Write Caption.}
+        \label{fig:IIandV_prevs}
+    \end{centering}
+\end{figure}
+
+\begin{figure}
     \script{comp_IIIandV_prevs.py}

### 2023-03-14 src/scripts/comp_IIandV_prevs.py
add bg-wg comp for LNLs II and V

@@ -1,0 +1,93 @@
+"""
+Plot histograms of predicted and posteriors over observed prevalences for scenarios
+that involve the LNLs II and V. Compare these prevalence predictions for the
+base-graph model to the winning-graph model.
+"""
+from pathlib import Path
+
+import matplotlib.pyplot as plt
+from matplotlib.lines import Line2D
+from lyscripts.plot.utils import Histogram, Posterior, draw
+from lyscripts.plot.utils import COLORS as USZ
+from tueplots import figsizes, fontsizes
+
+import paths
+
+
+BASE_GRAPH_INPUT = paths.data / "bg_IIandV_prevs.hdf5"
+WIN_GRAPH_INPUT = paths.data / "wg_IIandV_prevs.hdf5"
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
+NROWS, NCOLS = 2, 1
+
+
+if __name__ == "__main__":
+    plt.rcParams.update(figsizes.icml2022_half(
+        nrows=NROWS,
+        ncols=NCOLS,
+        # height_to_width_ratio=0.3,
+    ))
+    plt.rcParams.update(fontsizes.icml2022())
+    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
+
+    panels = {
+        "early": [],
+        "late": [],
+    }
+
+    for stage in ["early", "late"]:
+        panels[stage] = [
+            Histogram.from_hdf5(
+                filename=WIN_GRAPH_INPUT,
+                dataname=f"VnotII/{stage}",
+                label="LNL V without II",
+                color=USZ["blue"],
+            ),
+            Histogram.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"VnotII/{stage}",
+                label="base graph",
+                histtype="step",
+                color="darkgray",
+                linewidth=2.,
+                hatch=r"////",
+            ),
+            Posterior.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"VnotII/{stage}",
+                color=USZ["blue"],
+            ),
+
+            Histogram.from_hdf5(
+                filename=WIN_GRAPH_INPUT,
+                dataname=f"IIandV/{stage}",
+                label="LNL II and V",
+                color=USZ["orange"],
+            ),
+            Histogram.from_hdf5(
[…]

### 2023-03-14 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-In \cref{fig:IandII_low_prevs,fig:IandII_high_prevs,fig:IIIandV_prevs} the base graph's estimations for prevalences of involvement patterns that include the \glspl{lnl} I, V, and VII have been plotted as gray, hatched histograms.
+In \cref{fig:IandII_prevs,fig:IIandV_prevs,fig:IIIandV_prevs} the base graph's estimations for prevalences of involvement patterns that include the \glspl{lnl} I, V, and VII have been plotted as gray, hatched histograms.
 Alongside, we show again the Beta posterior over the prevalence of the respective involvement pattern, given the data, as colored line.
@@
 \begin{figure}
-    \script{comp_IandII_low_prevs.py}
+    \script{comp_IandII_prevs.py}
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/comp_IandII_low_prevs.png}
-        \caption{Prevalence predictions of the winning extended graph (colored histograms) compared to the base graph (hatched histograms) for scenarios of nodal involvement in the levels I and II.
-The solid lines depict the Beta posterior over the prevalence based on a uniform prior and the number of patients matching the respective scenario in the data.
-The top and bottom panel show the same scenarios, but for early and late T-category respectively.}
-        \label{fig:IandII_low_prevs}
-    \end{centering}
-\end{figure}
-
-\begin{figure}
-    \script{comp_IandII_high_prevs.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/comp_IandII_high_prevs.png}
-        \caption{Histograms and Beta posteriors over the prevalence of involvement for \gls{lnl} II (with and without \gls{lnl} I being metastatic as well).
-The colored histograms depict the predictions of the extended graph, while the hatched ones show the same prediction for the base graph.
-The top panel displays the predictions and observations for early T-category patients, the bottom panel for late T-category ones.}
-        \label{fig:IandII_high_prevs}
+        \includegraphics[width=\linewidth]{figures/comp_IandII_prevs.png}
+        \caption{TODO: Write Caption.}
+        \label{fig:IandII_prevs}
     \end{centering}

### 2023-03-14 src/scripts/bg_core_prevs.py
update & improve prevalence figures

@@ -57,7 +57,7 @@ def draw_early_late_panels(panels, lnl, xlims, ylims=None, idx=0):
     post_late_handles = late_handles[:num]
     post_late_labels = late_labels[:num]
     late_ax.set_yticks([])
-    late_ax.set_xlabel("prevalence [%]", fontweight="bold")
+    late_ax.set_xlabel("prevalence [%]")
     late_ax.set_ylabel("late T-cat.", fontweight="bold")
     late_ax.legend(post_late_handles, post_late_labels, labelspacing=0.3)
 

### 2023-03-14 src/scripts/comp_IIIandV_prevs.py

@@ -24,7 +24,7 @@ if __name__ == "__main__":
     plt.rcParams.update(figsizes.icml2022_half(
         nrows=NROWS,
         ncols=NCOLS,
-        # height_to_width_ratio=0.3,
+        height_to_width_ratio=0.45,
     ))
     plt.rcParams.update(fontsizes.icml2022())
     fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
@@ -82,11 +82,13 @@ if __name__ == "__main__":
     draw(axes[0], contents=panels["early"], xlims=(0., 8.), hist_kwargs={"nbins": 40})
     axes[0].set_ylim(bottom=0., top=0.9)
     axes[0].set_ylabel("early T-cat.", fontweight="bold")
+    axes[0].set_yticks([])
     axes[0].legend()
 
     draw(axes[1], contents=panels["late"], xlims=(0., 8.), hist_kwargs={"nbins": 40})
     # axes[1].set_ylim(bottom=0., top=1.2)
     axes[1].set_ylabel("late T-cat.", fontweight="bold")
+    axes[1].set_yticks([])
     axes[1].set_xlabel("prevalence [%]")
     axes[1].legend(handles=[h for h in axes[1].get_children() if isinstance(h, Line2D)])
 

### 2023-03-14 src/scripts/comp_IIandVII_prevs.py

@@ -1,93 +1,0 @@
-"""
-Plot histograms of predicted and posteriors over observed prevalences for scenarios
-that involve the LNLs II and VII. Compare these prevalence predictions for the
-base-graph model to the winning-graph model.
-"""
-from pathlib import Path
-
-import matplotlib.pyplot as plt
-from matplotlib.lines import Line2D
-from lyscripts.plot.utils import Histogram, Posterior, draw
-from lyscripts.plot.utils import COLORS as USZ
-from tueplots import figsizes, fontsizes
-
-import paths
-
-
-BASE_GRAPH_INPUT = paths.data / "bg_IIandVII_prevs.hdf5"
-WIN_GRAPH_INPUT = paths.data / "wg_IIandVII_prevs.hdf5"
-OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
-NROWS, NCOLS = 2, 1
-
-
-if __name__ == "__main__":
-    plt.rcParams.update(figsizes.icml2022_half(
-        nrows=NROWS,
-        ncols=NCOLS,
-        # height_to_width_ratio=0.3,
-    ))
-    plt.rcParams.update(fontsizes.icml2022())
-    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
-
-    panels = {
-        "early": [],
-        "late": [],
-    }
-
-    for stage in ["early", "late"]:
-        panels[stage] = [
-            Histogram.from_hdf5(
-                filename=WIN_GRAPH_INPUT,
-                dataname=f"IIandVII/{stage}",
-                label="LNL II with VII",
-                color=USZ["red"],
-            ),
-            Histogram.from_hdf5(
-                filename=BASE_GRAPH_INPUT,
-                dataname=f"IIandVII/{stage}",
-                label="base graph",
-                histtype="step",
-                color="darkgray",
-                linewidth=2.,
-                hatch=r"////",
-            ),
-            Posterior.from_hdf5(
-                filename=BASE_GRAPH_INPUT,
-                dataname=f"IIandVII/{stage}",
-                color=USZ["red"],
-            ),
-
-            Histogram.from_hdf5(
-                filename=WIN_GRAPH_INPUT,
-                dataname=f"VIInotII/{stage}",
-                label="LNL VII without II",
-                color=USZ["blue"],
-            ),
-            Histogram.from_hdf5(
[…]

### 2023-03-14 src/scripts/comp_IIandV_prevs.py

@@ -24,7 +24,7 @@ if __name__ == "__main__":
     plt.rcParams.update(figsizes.icml2022_half(
         nrows=NROWS,
         ncols=NCOLS,
-        # height_to_width_ratio=0.3,
+        height_to_width_ratio=0.45,
     ))
     plt.rcParams.update(fontsizes.icml2022())
     fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
@@ -82,11 +82,13 @@ if __name__ == "__main__":
     draw(axes[0], contents=panels["early"], xlims=(0., 12.), hist_kwargs={"nbins": 40})
     axes[0].set_ylim(bottom=0., top=0.9)
     axes[0].set_ylabel("early T-cat.", fontweight="bold")
+    axes[0].set_yticks([])
     axes[0].legend()
 
     draw(axes[1], contents=panels["late"], xlims=(0., 12.), hist_kwargs={"nbins": 40})
     # axes[1].set_ylim(bottom=0., top=1.2)
     axes[1].set_ylabel("late T-cat.", fontweight="bold")
+    axes[1].set_yticks([])
     axes[1].set_xlabel("prevalence [%]")
     axes[1].legend(handles=[h for h in axes[1].get_children() if isinstance(h, Line2D)])
 

### 2023-03-14 src/scripts/comp_IVandV_prevs.py

@@ -1,93 +1,0 @@
-"""
-Plot histograms of predicted and posteriors over observed prevalences for scenarios
-that involve the LNLs VI and V. Compare these prevalence predictions for the
-base-graph model to the winning-graph model.
-"""
-from pathlib import Path
-
-import matplotlib.pyplot as plt
-from matplotlib.lines import Line2D
-from lyscripts.plot.utils import Histogram, Posterior, draw
-from lyscripts.plot.utils import COLORS as USZ
-from tueplots import figsizes, fontsizes
-
-import paths
-
-
-BASE_GRAPH_INPUT = paths.data / "bg_IVandV_prevs.hdf5"
-WIN_GRAPH_INPUT = paths.data / "wg_IVandV_prevs.hdf5"
-OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
-NROWS, NCOLS = 2, 1
-
-
-if __name__ == "__main__":
-    plt.rcParams.update(figsizes.icml2022_half(
-        nrows=NROWS,
-        ncols=NCOLS,
-        # height_to_width_ratio=0.3,
-    ))
-    plt.rcParams.update(fontsizes.icml2022())
-    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
-
-    panels = {
-        "early": [],
-        "late": [],
-    }
-
-    for stage in ["early", "late"]:
-        panels[stage] = [
-            Histogram.from_hdf5(
-                filename=WIN_GRAPH_INPUT,
-                dataname=f"VnotIV/{stage}",
-                label="LNL V without IV",
-                color=USZ["orange"],
-            ),
-            Histogram.from_hdf5(
-                filename=BASE_GRAPH_INPUT,
-                dataname=f"VnotIV/{stage}",
-                label="base graph",
-                histtype="step",
-                color="darkgray",
-                linewidth=2.,
-                hatch=r"////",
-            ),
-            Posterior.from_hdf5(
-                filename=BASE_GRAPH_INPUT,
-                dataname=f"VnotIV/{stage}",
-                color=USZ["orange"],
-            ),
-
-            Histogram.from_hdf5(
-                filename=WIN_GRAPH_INPUT,
-                dataname=f"IVandV/{stage}",
-                label="LNL IV and V",
-                color=USZ["blue"],
-            ),
-            Histogram.from_hdf5(
[…]

### 2023-03-14 src/scripts/comp_IandII_high_prevs.py

@@ -1,92 +1,0 @@
-"""
-Plot histograms of predicted and posteriors over observed prevalences for scenarios
-that involve the LNLs I and II. Compare these prevalence predictions for the base-graph
-model to the winning-graph model.
-"""
-from pathlib import Path
-
-import matplotlib.pyplot as plt
-from lyscripts.plot.utils import Histogram, Posterior, draw
-from lyscripts.plot.utils import COLORS as USZ
-from tueplots import figsizes, fontsizes
-
-import paths
-
-
-BASE_GRAPH_INPUT = paths.data / "bg_IandII_prevs.hdf5"
-WIN_GRAPH_INPUT = paths.data / "wg_IandII_prevs.hdf5"
-OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
-NROWS, NCOLS = 2, 1
-
-
-if __name__ == "__main__":
-    plt.rcParams.update(figsizes.icml2022_half(
-        nrows=NROWS,
-        ncols=NCOLS,
-        # height_to_width_ratio=0.3,
-    ))
-    plt.rcParams.update(fontsizes.icml2022())
-    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
-
-    panels = {
-        "early": [],
-        "late": [],
-    }
-
-    for stage in ["early", "late"]:
-        panels[stage] = [
-            Histogram.from_hdf5(
-                filename=WIN_GRAPH_INPUT,
-                dataname=f"II/{stage}",
-                label="LNL II overall",
-                color=USZ["red"],
-            ),
-            Histogram.from_hdf5(
-                filename=BASE_GRAPH_INPUT,
-                dataname=f"II/{stage}",
-                label="base graph",
-                histtype="step",
-                color="darkgray",
-                linewidth=2.,
-                hatch=r"////",
-            ),
-            Posterior.from_hdf5(
-                filename=BASE_GRAPH_INPUT,
-                dataname=f"II/{stage}",
-                color=USZ["red"],
-            ),
-
-            Histogram.from_hdf5(
-                filename=WIN_GRAPH_INPUT,
-                dataname=f"IInotI/{stage}",
-                label="LNL II without I",
-                color=USZ["orange"],
-            ),
-            Histogram.from_hdf5(
-                filename=BASE_GRAPH_INPUT,
-                dataname=f"IInotI/{stage}",
[…]

### 2023-03-14 src/scripts/comp_IandII_low_prevs.py

@@ -1,92 +1,0 @@
-"""
-Plot histograms of predicted and posteriors over observed prevalences for scenarios
-that involve the LNLs I and II. Compare these prevalence predictions for the base-graph
-model to the winning-graph model.
-"""
-from pathlib import Path
-
-import matplotlib.pyplot as plt
-from lyscripts.plot.utils import Histogram, Posterior, draw
-from lyscripts.plot.utils import COLORS as USZ
-from tueplots import figsizes, fontsizes
-
-import paths
-
-
-BASE_GRAPH_INPUT = paths.data / "bg_IandII_prevs.hdf5"
-WIN_GRAPH_INPUT = paths.data / "wg_IandII_prevs.hdf5"
-OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
-NROWS, NCOLS = 2, 1
-
-
-if __name__ == "__main__":
-    plt.rcParams.update(figsizes.icml2022_half(
-        nrows=NROWS,
-        ncols=NCOLS,
-        # height_to_width_ratio=0.3,
-    ))
-    plt.rcParams.update(fontsizes.icml2022())
-    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
-
-    panels = {
-        "early": [],
-        "late": [],
-    }
-
-    for stage in ["early", "late"]:
-        panels[stage] = [
-            Histogram.from_hdf5(
-                filename=WIN_GRAPH_INPUT,
-                dataname=f"I/{stage}",
-                label="LNL I overall",
-                color=USZ["blue"],
-            ),
-            Histogram.from_hdf5(
-                filename=BASE_GRAPH_INPUT,
-                dataname=f"I/{stage}",
-                label="base graph",
-                histtype="step",
-                color="darkgray",
-                linewidth=2.,
-                hatch=r"////",
-            ),
-            Posterior.from_hdf5(
-                filename=BASE_GRAPH_INPUT,
-                dataname=f"I/{stage}",
-                color=USZ["blue"],
-            ),
-
-            Histogram.from_hdf5(
-                filename=WIN_GRAPH_INPUT,
-                dataname=f"InotII/{stage}",
-                label="LNL I without II",
-                color=USZ["green"],
-            ),
-            Histogram.from_hdf5(
-                filename=BASE_GRAPH_INPUT,
-                dataname=f"InotII/{stage}",
[…]

### 2023-03-14 src/scripts/comp_IandII_prevs.py

@@ -1,0 +1,167 @@
+"""
+Plot histograms of predicted and posteriors over observed prevalences for scenarios
+that involve the LNLs I and II. Compare these prevalence predictions for the base-graph
+model to the winning-graph model.
+"""
+from pathlib import Path
+
+import matplotlib.pyplot as plt
+from lyscripts.plot.utils import Histogram, Posterior, draw
+from lyscripts.plot.utils import COLORS as USZ
+from tueplots import figsizes, fontsizes
+
+import paths
+
+
+BASE_GRAPH_INPUT = paths.data / "bg_IandII_prevs.hdf5"
+WIN_GRAPH_INPUT = paths.data / "wg_IandII_prevs.hdf5"
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
+NROWS, NCOLS = 2, 2
+
+
+if __name__ == "__main__":
+    plt.rcParams.update(figsizes.icml2022_full(
+        nrows=NROWS,
+        ncols=NCOLS,
+        height_to_width_ratio=0.45,
+    ))
+    plt.rcParams.update(fontsizes.icml2022())
+    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col")
+
+
+    left_panels = {
+        "early": [],
+        "late": [],
+    }
+
+    for stage in ["early", "late"]:
+        left_panels[stage] = [
+            Histogram.from_hdf5(
+                filename=WIN_GRAPH_INPUT,
+                dataname=f"I/{stage}",
+                label="LNL I overall",
+                color=USZ["blue"],
+            ),
+            Histogram.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"I/{stage}",
+                label="base graph",
+                histtype="step",
+                color="darkgray",
+                linewidth=2.,
+                hatch=r"////",
+            ),
+            Posterior.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"I/{stage}",
+                color=USZ["blue"],
+            ),
+
+            Histogram.from_hdf5(
+                filename=WIN_GRAPH_INPUT,
+                dataname=f"InotII/{stage}",
+                label="LNL I without II",
+                color=USZ["green"],
+            ),
+            Histogram.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"InotII/{stage}",
[…]

### 2023-03-14 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 These histograms are obtained by computing the same prevalence with different samples from the inference process, thus providing us with a measure of uncertainty for the prediction.
-    \item Colored lines, depicting the Beta posterior over the same involvement pattern's prevalence, given a uniform prior and the observed data as likelihood.
-The maximum of the Beta distributions always coincides with the data prevalence but we additionally gain an intuition into how statistically significant the data is.
+    \item Colored lines, depicting the beta posterior over the same involvement pattern's prevalence, given a uniform beta prior and the binomial likelihood of the observed data.
+The maximum of the beta distributions always coincides with the data prevalence but we additionally gain an intuition into how statistically significant the data is.
 E.g., Observing 3 out of 10 patients with a particular pattern of nodal metastases is less convincing than 300 out of a cohort of 1000 patients.
-A Beta posterior over these prevalences reflects that in its variance.
+A beta posterior over these prevalences reflects that in its variance.
 \end{enumerate}
@@
 
-\begin{figure}
+\begin{figure*}
     \script{comp_IandII_prevs.py}
@@
         \includegraphics[width=\linewidth]{figures/comp_IandII_prevs.png}
-        \caption{TODO: Write Caption.}
+        \caption{Comparison of observed and predicted prevalences of \gls{lnl} I and II involvement patterns.
+The top and bottom panels show the prevalences for early and late T-category, respectively.
+The solid lines are Beta posteriors from the data, while the histograms are predicted prevalences (colored: winning graph, gray-hatched: base graph).
+Blue and red plots indicate overall \gls{lnl} I and II involvement, respectively.
+Green plots indicate \gls{lnl} I involvement without level II, while orange plots indicate the opposite (\gls{lnl} II without level I).
+The winning graph has an added edge from \gls{lnl} II to I, which improves the prediction of the rare green pattern and reduces the overestimation of the orange pattern in late T-category.
+However, it slightly underestimates the orange pattern in early T-category.}
         \label{fig:IandII_prevs}
     \end{centering}
-\end{figure}
-
-% \begin{figure}
-%     \script{comp_IIandVII_prevs.py}
-%     \begin{centering}
-%         \includegraphics[width=\linewidth]{figures/comp_IIandVII_prevs.png}
-%         \caption{Prevalences for involvement patterns where \gls{lnl} VII is involved with (red) and without (blue) \gls{lnl} II.
-The solid lines depict a Beta posterior over the prevalence as observed in the data.
-The colored histograms show these prevalences as predicted by the winning graph.
-The hatched histograms show the same predictions from the base graph for comparison.
-Since the winning graph does not include an additional arc between \glspl{lnl} II and VII, the fit w.r.t.
-these two \glspl{lnl} does not improve.}
-%         \label{fig:IIandVII_prevs}
-%     \end{centering}
-% \end{figure}
+\end{figure*}
 
@@
         \includegraphics[width=\linewidth]{figures/comp_IIandV_prevs.png}
-        \caption{TODO: Write Caption.}
+        \caption{Beta posterior over the prevalence of the involvement pattern where \gls{lnl} II \emph{and} V show metastases (orange line), as well as the pattern with \gls{lnl} V involved \emph{without} level II (blue line).
+Both prevalences are shown for early (top panel) and late T-category (bottom panel).
+The correspondingly colored histograms show the same prevalences as predicted by the winning graph model.
+The hatched gray modalities display the same predictions as made by the base graph model.
+The winning graph slightly improves its fit to the data in the case of early T-category prevalence of \gls{lnl} V without \gls{lnl} II, and in the case of late T-category co-involvement of the levels II and V.
+This likely stems from the additional arc the winning graph has from \gls{lnl} II to V.}
         \label{fig:IIandV_prevs}
@@
 
-% \begin{figure}
-%     \script{comp_IVandV_prevs.py}
-%     \begin{centering}
-%         \includegraphics[width=\linewidth]{figures/comp_IVandV_prevs.png}
-%         \caption{Predictions vs observations of the prevalence of two patterns of involvement: One for the case when both \gls{lnl} VI and V are metastatic (blue) and one for the case where \gls{lnl} V is involved with level IV being healthy (orange).
-Again, the solid lies depict the Beta posterior over the prevalence of these patterns in the data.
-Colored histograms show the predicted distribution over the two patterns from the winning graph model.
-Hatched histograms show the same predictions from the base graph model for comparison.
-Although no arc was added from \gls{lnl} IV to V, the winning graph slightly improves its predictions both for early (top panel) and late T-category cases (bottom panel).}
-%         \label{fig:IVandV_prevs}
-%     \end{centering}
-% \end{figure}
-
 \subsection{Risk Prediction for Occult Disease}

### 2023-03-14 src/scripts/wg_IV_risks.py
add some first risk plots

@@ -1,0 +1,75 @@
+"""
+Plot histograms of predicted risks, given different diagnoses.
+"""
+from pathlib import Path
+
+import matplotlib.pyplot as plt
+from lyscripts.plot.utils import Histogram, draw
+from lyscripts.plot.utils import COLORS as USZ
+from tueplots import figsizes, fontsizes
+
+import paths
+
+
+INPUT = paths.data / "wg_risks.hdf5"
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
+NROWS, NCOLS = 2, 1
+
+
+if __name__ == "__main__":
+    plt.rcParams.update(figsizes.icml2022_half(
+        nrows=NROWS,
+        ncols=NCOLS,
+        height_to_width_ratio=0.45,
+    ))
+    plt.rcParams.update(fontsizes.icml2022())
+    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
+
+    panels = {
+        "early": [],
+        "late": [],
+    }
+
+    for stage in ["early", "late"]:
+        panels[stage] = [
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"{stage}/IV/N0",
+                label=None,
+                color=USZ["gray"],
+            ),
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"{stage}/IV/N0",
+                label="LNL IV, given N0",
+                color="#97a3a7",
+                histtype="step",
+                linewidth=2.,
+                alpha=1.0,
+                zorder=1.,
+            ),
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"{stage}/IV/II",
+                label="LNL IV, given LNL II involved",
+                color=USZ["blue"],
+            ),
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"{stage}/IV/IIandIII",
+                label="LNL IV, given LNL II & III involved",
+                color=USZ["orange"],
+            ),
+        ]
+
+    draw(axes[0], contents=panels["early"], xlims=(0., 7.), hist_kwargs={"nbins": 40})
+    axes[0].set_ylabel("early T-cat.", fontweight="bold")
+    axes[0].set_yticks([])
+    axes[0].legend()
+
[…]

### 2023-03-14 src/scripts/wg_V_risks.py

@@ -1,0 +1,65 @@
+"""
+Plot histograms of predicted risks, given different diagnoses.
+"""
+from pathlib import Path
+
+import matplotlib.pyplot as plt
+from lyscripts.plot.utils import Histogram, draw
+from lyscripts.plot.utils import COLORS as USZ
+from tueplots import figsizes, fontsizes
+
+import paths
+
+
+INPUT = paths.data / "wg_risks.hdf5"
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
+NROWS, NCOLS = 2, 1
+
+
+if __name__ == "__main__":
+    plt.rcParams.update(figsizes.icml2022_half(
+        nrows=NROWS,
+        ncols=NCOLS,
+        height_to_width_ratio=0.45,
+    ))
+    plt.rcParams.update(fontsizes.icml2022())
+    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
+
+    panels = {
+        "early": [],
+        "late": [],
+    }
+
+    for stage in ["early", "late"]:
+        panels[stage] = [
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"{stage}/V/N0",
+                label="LNL V, given N0",
+                color=USZ["green"],
+            ),
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"{stage}/V/II",
+                label="LNL V, given LNL II involved",
+                color=USZ["orange"],
+            ),
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"{stage}/V/IIandIII",
+                label="LNL V, given LNL II & III involved",
+                color=USZ["red"],
+            ),
+        ]
+
+    draw(axes[0], contents=panels["early"], xlims=(0., 5.), hist_kwargs={"nbins": 40})
+    axes[0].set_ylabel("early T-cat.", fontweight="bold")
+    axes[0].set_yticks([])
+    axes[0].legend()
+
+    draw(axes[1], contents=panels["late"], xlims=(0., 5.), hist_kwargs={"nbins": 40})
+    axes[1].set_ylabel("late T-cat.", fontweight="bold")
+    axes[1].set_yticks([])
+    axes[1].set_xlabel("risk [%]")
+
+    plt.savefig(OUTPUT, dpi=300)

### 2023-03-15 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{amssymb}
+\usepackage[inline]{enumitem}
 
@@
 
+\begin{figure}
+    \script{wg_IV_risks.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/wg_IV_risks.png}
+        \caption{Distributions over the risk for microscopic involvement in \gls{lnl} IV as predicted by the winning graph model, given early (top panel) or late T-category (bottom panel), and three different CT-based diagnoses:
+        \begin{enumerate*}[label=\arabic*)]
+            \item A clinical N0 patient (gray histogram),
+            \item visible metastases in \gls{lnl} II, but otherwise healthy-looking lymph nodes (blue histogram), and
+            \item macroscopic metastases in the \glspl{lnl} II \& III, whith the rest of the neck still being clinically ode negative (orange histogram).
+        \end{enumerate*}
+        }
+        \label{fig:wg_IV_risks}
+    \end{centering}
+\end{figure}
+
+\begin{figure}
+    \script{wg_V_risks.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/wg_V_risks.png}
+        \caption{The winning graph's predictions over the risk for occult disease in \gls{lnl} V for early (top panel) and late T-category (bottom panel).
+Each of the risks is computed, given one of three diagnostic scenarios that were obtained by a CT scan:
+        \begin{enumerate*}[label=\arabic*)]
+            \item A clinically healthy neck (green),
+            \item on CT imaging visible involvement in the \glspl{lnl} II \& III (orange) with the remaining levels still clinically negative, and
+            \item observed involvement in the levels II, III, and IV (red).
+        \end{enumerate*}}
+        \label{fig:wg_V_risks}
+    \end{centering}
+\end{figure}
+
 \bibliography{bib}

### 2023-03-16 src/scripts/wg_III_risks.py
add more risk plots

@@ -1,0 +1,59 @@
+"""
+Plot histograms of predicted risks, given different diagnoses.
+"""
+from pathlib import Path
+
+import matplotlib.pyplot as plt
+from lyscripts.plot.utils import Histogram, draw
+from lyscripts.plot.utils import COLORS as USZ
+from tueplots import figsizes, fontsizes
+
+import paths
+
+
+INPUT = paths.data / "wg_risks.hdf5"
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
+NROWS, NCOLS = 2, 1
+
+
+if __name__ == "__main__":
+    plt.rcParams.update(figsizes.icml2022_half(
+        nrows=NROWS,
+        ncols=NCOLS,
+        height_to_width_ratio=0.45,
+    ))
+    plt.rcParams.update(fontsizes.icml2022())
+    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
+
+    panels = {
+        "early": [],
+        "late": [],
+    }
+
+    for stage in ["early", "late"]:
+        panels[stage] = [
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"{stage}/III/N0",
+                label="LNL III, given N0",
+                color=USZ["orange"],
+            ),
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"{stage}/III/II",
+                label="LNL III, given LNL II involved",
+                color=USZ["red"],
+            ),
+        ]
+
+    draw(axes[0], contents=panels["early"], xlims=(0., 15.), hist_kwargs={"nbins": 40})
+    axes[0].set_ylabel("early T-cat.", fontweight="bold")
+    axes[0].set_yticks([])
+    axes[0].legend()
+
+    draw(axes[1], contents=panels["late"], xlims=(0., 15.), hist_kwargs={"nbins": 40})
+    axes[1].set_ylabel("late T-cat.", fontweight="bold")
+    axes[1].set_yticks([])
+    axes[1].set_xlabel("risk [%]")
+
+    plt.savefig(OUTPUT, dpi=300)

### 2023-03-16 src/scripts/wg_I_risks.py

@@ -1,0 +1,65 @@
+"""
+Plot histograms of predicted risks, given different diagnoses.
+"""
+from pathlib import Path
+
+import matplotlib.pyplot as plt
+from lyscripts.plot.utils import Histogram, draw
+from lyscripts.plot.utils import COLORS as USZ
+from tueplots import figsizes, fontsizes
+
+import paths
+
+
+INPUT = paths.data / "wg_risks.hdf5"
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
+NROWS, NCOLS = 2, 1
+
+
+if __name__ == "__main__":
+    plt.rcParams.update(figsizes.icml2022_half(
+        nrows=NROWS,
+        ncols=NCOLS,
+        height_to_width_ratio=0.45,
+    ))
+    plt.rcParams.update(fontsizes.icml2022())
+    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
+
+    panels = {
+        "early": [],
+        "late": [],
+    }
+
+    for stage in ["early", "late"]:
+        panels[stage] = [
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"{stage}/I/N0",
+                label="LNL I, given N0",
+                color=USZ["green"],
+            ),
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"{stage}/I/II",
+                label="LNL I, given LNL II involved",
+                color=USZ["blue"],
+            ),
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"{stage}/I/IIandIIIandIV",
+                label="LNL I, given LNL II, III, and IV involved",
+                color=USZ["red"],
+            ),
+        ]
+
+    draw(axes[0], contents=panels["early"], xlims=(0., 7.), hist_kwargs={"nbins": 40})
+    axes[0].set_ylabel("early T-cat.", fontweight="bold")
+    axes[0].set_yticks([])
+    axes[0].legend()
+
+    draw(axes[1], contents=panels["late"], xlims=(0., 7.), hist_kwargs={"nbins": 40})
+    axes[1].set_ylabel("late T-cat.", fontweight="bold")
+    axes[1].set_yticks([])
+    axes[1].set_xlabel("risk [%]")
+
+    plt.savefig(OUTPUT, dpi=300)

### 2023-03-16 src/scripts/wg_III_risks.py
add sens spec map

@@ -49,11 +49,11 @@ if __name__ == "__main__":
     draw(axes[0], contents=panels["early"], xlims=(0., 15.), hist_kwargs={"nbins": 40})
     axes[0].set_ylabel("early T-cat.", fontweight="bold")
     axes[0].set_yticks([])
-    axes[0].legend()
 
     draw(axes[1], contents=panels["late"], xlims=(0., 15.), hist_kwargs={"nbins": 40})
     axes[1].set_ylabel("late T-cat.", fontweight="bold")
     axes[1].set_yticks([])
     axes[1].set_xlabel("risk [%]")
+    axes[1].legend()
 
     plt.savefig(OUTPUT, dpi=300)

### 2023-03-16 src/scripts/wg_sens_spec_risks.py

@@ -1,0 +1,99 @@
+"""
+Create a plot that investigates how sensitive the risk predictions are
+to the sensitivity & specificity of the diagnostic modality.
+"""
+from pathlib import Path
+from matplotlib.colors import LinearSegmentedColormap
+
+import scipy as sp
+import numpy as np
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+from lymph import Unilateral
+from lyscripts.plot.utils import COLORS as USZ
+from lyscripts.utils import load_hdf5_samples
+from tueplots import figsizes, fontsizes
+
+import paths
+
+
+SAMPLES_PATH = paths.data / "wg_samples.hdf5"
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
+WINNING_GRAPH = {
+    ("tumor", "primary"): ["I", "II", "III", "IV", "V", "VII"],
+    ("lnl", "I"): [],
+    ("lnl", "II"): ["I", "III", "V"],
+    ("lnl", "III"): ["IV", "V"],
+    ("lnl", "IV"): [],
+    ("lnl", "V"): [],
+    ("lnl", "VII"): [],
+}
+PATTERN_OF_INTEREST = {"IV": True}
+DIAGNOSIS = {
+    "risk": {"I": False, "II": True, "III": True, "IV": False, "V": False, "VII": False}
+}
+T_STAGE = "late"
+NSTEP = 11
+MAX_T = 10
+TIME_STEPS = np.arange(MAX_T + 1)
+CMAP = LinearSegmentedColormap.from_list(
+    "usz_cmap", [USZ["green"], USZ["gray"], USZ["red"]]
+)
+
+
+if __name__ == "__main__":
+    model = Unilateral(WINNING_GRAPH)
+    samples = load_hdf5_samples(SAMPLES_PATH)
+
+    samples_mean = np.mean(samples, axis=0)
+    spread_probs = samples_mean[:-1]
+    time_dist_prob = samples_mean[-1]
+
+    model.spread_probs = spread_probs
+    model.diag_time_dists[T_STAGE] = sp.stats.binom.pmf(
+        TIME_STEPS, MAX_T, time_dist_prob,
+    )
+
+    risk_map = np.empty(shape = (NSTEP, NSTEP), dtype=float)
+
+    for i,sens in enumerate(np.linspace(0.5, 1., NSTEP)):
+        for j,spec in enumerate(np.linspace(0.5, 1., NSTEP)):
[…]

### 2023-03-29 src/scripts/wg_any_risks.py
add risk for any of the LNLs I, IV, V, or VII

@@ -1,0 +1,65 @@
+"""
+Plot histograms of predicted risks, given different diagnoses.
+"""
+from pathlib import Path
+
+import matplotlib.pyplot as plt
+from lyscripts.plot.utils import Histogram, draw
+from lyscripts.plot.utils import COLORS as USZ
+from tueplots import figsizes, fontsizes
+
+import paths
+
+
+INPUT = paths.data / "wg_risks.hdf5"
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
+NROWS, NCOLS = 2, 1
+
+
+if __name__ == "__main__":
+    plt.rcParams.update(figsizes.icml2022_half(
+        nrows=NROWS,
+        ncols=NCOLS,
+        height_to_width_ratio=0.45,
+    ))
+    plt.rcParams.update(fontsizes.icml2022())
+    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
+
+    panels = {
+        "early": [],
+        "late": [],
+    }
+
+    for stage in ["early", "late"]:
+        panels[stage] = [
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"{stage}/anyIorIVorVorVII/N0",
+                label="I, IV, V, or VII, given N0",
+                color=USZ["green"],
+            ),
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"{stage}/anyIorIVorVorVII/II",
+                label="I, IV, V, or VII, given LNL II involved",
+                color=USZ["orange"],
+            ),
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"{stage}/anyIorIVorVorVII/IIandIII",
+                label="I, IV, V, or VII, given LNL II, and III involved",
+                color=USZ["red"],
+            ),
+        ]
+
+    draw(axes[0], contents=panels["early"], xlims=(2., 16.), hist_kwargs={"nbins": 40})
+    axes[0].set_ylabel("early T-cat.", fontweight="bold")
+    axes[0].set_yticks([])
+    axes[0].legend()
+
+    draw(axes[1], contents=panels["late"], xlims=(2., 16.), hist_kwargs={"nbins": 40})
+    axes[1].set_ylabel("late T-cat.", fontweight="bold")
+    axes[1].set_yticks([])
+    axes[1].set_xlabel("risk [%]")
+
+    plt.savefig(OUTPUT, dpi=300)

### 2023-08-11 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
+\newcommand{\ju}[1]{{\textcolor{red}{#1}}}
+
 % Cross-referencing
@@
 \author{Bertrand Pouymayou}
-\author{Panagiotis Balermpas}
 \author{Lauence Bauwens}
@@
 \author{Roland Giger}
+\author{Panagiotis Balermpas}
 \author{Jan Unkelbach}
@@
 \begin{abstract}
-    The \gls{ctv-n}\glsunset{ctv} in \gls{hnscc}\glsunset{scc} is currently based mostly on the prevalence of lymphatic metastases in different \glspl{lnl} for a given primary tumor location.
+    Objective: The \gls{ctv-n}\glsunset{ctv} in \gls{hnscc}\glsunset{scc} is currently based mostly on the prevalence of lymph node metastases in different \glspl{lnl} for a given primary tumor location.
+In this work, we present a probabilistic model for ipsilateral lymphatic metastatic spread that can quantify the risk for microscopic nodal involvement based on an individual patient's T-stage and clinical involvement of \glspl{lnl} at diagnosis.
+\\
+    Methods: The model is based on the formalism of \glspl{hmm} as conceptually introduced in a previous publication.
+The \gls{hmm} uses a collection of hidden binary \glspl{rv} -- one for each of the \glspl{lnl} I, II, III, IV, V, and VII -- to model a patient's true state of lymph node involvement corresponding to pathological examinations.
+Clinical involvement at diagnosis represent the observed binary \glspl{rv} linked to the true state via sensitivity and specificity.
+A \gls{dag} with is used to compute the transition matrix of the \gls{hmm}, where each edge is associated with a parameter for the probability of tumor spread from the primary tumor and between \gls{lnl}.
+To learn the parameters, we collected a dataset of 681 \gls{opscc} patients from three institutions.
+To determine the best graph that balances model complexity and accurate description of \gls{lnl} involvement patterns in the data, we compare graphs based on the model evidence calculated via \gls{ti}.
+\\
+    Results: The graph maximizing the model evidence contains edges for lymphatic spread from \gls{lnl} II to I, III and V, and from \gls{lnl} III to IV and V, and is thus consistent with commonly assumed lymphatic drainage in the neck.
+Application of the model suggests that the risk of occult disease in level IV is below 5\% if level III is clinically negative, and that the risk of occult disease in level V is below 5\% except for late T-category patients with clinical involvement of levels II, III, and IV.
+    Conclusion: We provide a statistical model of lymph node involvement in \gls{opscc} patients trained on a multi-institutional dataset, which may guide the design of clinical trials on volume-deescalated treatment of \gls{opscc} and eventually contribute to further personalized guidelines on elective nodal treatment.
+%    The \gls{ctv-n}\glsunset{ctv} in \gls{hnscc}\glsunset{scc} is currently based mostly on the prevalence of lymphatic metastases in different \glspl{lnl} for a given primary tumor location.
 In this work, we present an extension to a probabilistic model for lymphatic metastatic spread developed earlier that can quantify the risk for microscopic nodal involvement based on an individual patient diagnosis.
@@
 
-When treating \gls{hnscc}, either with radiotherapy or via surgery, the aim is to irradiate or resect as much of the present malignancies as possible.
-This includes the primary tumor mass and metastases that are sufficiently large to be detected using in-vivo imaging modalities such as \gls{ct}, \gls{mri}, or \gls{pet}.
-But it also includes regions of possible microscopic spread, which these modalities cannot detect, to increase the patient's probability of cure \cite{poortmans_internal_2020,murthy_prostate-only_2021}.
-Microscopic disease can currently only be observed by a pathological examination of the tissue.
+When treating \gls{hnscc} with radiotherapy or surgery, the aim is to irradiate or resect as much of the malignant tissue as possible.
+This includes the primary tumor mass and clinically detected lymph node metastases.
+However, to reduce the risk of locoregional failure, treatment also includes regions of the lymph drainage system of the neck with possible microscopic tumor spread, which in-vivo imaging modalities such as \gls{ct}, \gls{mri}, or \gls{pet} cannot detect.
+This is referred to as elective nodal irradiation or prophylactic neck dissection.
+Treatment decisions regarding the \gls{ctv-n} or the extent of neck dissection must balance the conflicting goals of treating regions at risk of occult lymph node metastases to avoid recurrences while avoiding toxicity related to unnecessary treatment of healthy tissues.
+
+%\cite{poortmans_internal_2020,murthy_prostate-only_2021}. 
+%Microscopic disease can currently only be observed by a pathological examination of the tissue.
 Consequently, clinicians that are presented with cancer patients need to routinely assess the risk of microscopic involvement in parts of the body that appear clinically node negative.
 
-This work concerns itself with \gls{opscc}, which commonly spreads lymphatically.
+This work concerns itself with \gls{opscc}, where approximately 70-80\% of patients present with lymph node metastases at the time of diagnosis.  
+%which commonly spreads lymphatically.
 The only two publicly available datasets on nodal involvement patterns in \gls{opscc}, at the time of writing, show that 80\% of patients where either cN+ or -- if available -- pN+ \cite{ludwig_dataset_2022}.
@@
 These levels are then often irradiated or resected prophylactically due to the risk of harboring occult metastases despite negative findings from imaging modalities.
-In current clinical practice, \gls{ctv-n} definition is mostly based on guidelines \cite{gregoire_ct-based_2003,gregoire_delineation_2014,gregoire_delineation_2018,eisbruch_intensity-modulated_2002,biau_selection_2019,chao_determination_2002,vorwerk_guidelines_2011,ferlito_elective_2009} that are derived from the observed prevalence of involvement in an \gls{lnl} for a given tumor location.
+In clinical practice, \gls{ctv-n} definition in radiotherapy is based on guidelines \cite{gregoire_ct-based_2003,gregoire_delineation_2014,gregoire_delineation_2018,eisbruch_intensity-modulated_2002,biau_selection_2019,chao_determination_2002,vorwerk_guidelines_2011,ferlito_elective_2009} that are mostly derived from the observed prevalence of involvement in an \gls{lnl} for a given tumor location.
+These guidelines currently suggest extensive irradiation of both sides of the neck for most patients.
+In the ipsilateral neck, the \gls{ctv-n} includes \glspl{lnl} II, III and IV for all patients, and levels I and V for the majority of patients.
+These guidelines, however, do not account for the personal risk of the patients that may depend greatly on their state of tumor progression at diagnosis.
+E.g., a patient with macroscopic metastases detected via \gls{pet} in both \glspl{lnl} II and III may have a substantial risk for occult disease in \gls{lnl} IV.
+Instead, patients who present with a clinically N0 neck or a single metastasis in \gls{lnl} II may have a much smaller risk for occult disease in \gls{lnl} IV.
 
-These guidelines, however, do not account for the personal risk of the patients that may depend greatly on their diagnosis.
-E.g., macroscopic metastases detected via \gls{pet} in the \glspl{lnl} II and III may increase the risk for occult disease in \gls{lnl} IV over the case of a patient who presents with a clinically N0 neck.
-
-A model for estimating the risk of microscopic disease, given a personal diagnosis based on \glspl{bn} was previously developed \cite{pouymayou_bayesian_2019}.
-It was subsequently formulated as a \glspl{hmm} to include T-category in an intuitive manner \cite{ludwig_hidden_2021}.
-These models were trained on a dataset of early T-category (T1 and T2) patients that could be reconstructed from \cite{sanguineti_defining_2009}.
-Since this publication only includes details about the involvement of the ipsilateral \glspl{lnl} I, II, III, and IV, the models were limited to these nodal levels as well.
-
-With the publication of larger and more detailed datasets \cite{ludwig_dataset_2022}, we are thus able to
+We previously developed a model of lymphatic metastatic progression for estimating the risk of microscopic disease, given a patient's personal diagnosis.
+The initial model was based on the methodology of Bayesian networks \cite{pouymayou_bayesian_2019}.
+It was subsequently extended and formulated as a \glspl{hmm} to include T-category in an intuitive manner \cite{ludwig_hidden_2021}.
+However, these models were introduced based on only a small dataset of approximately 100 early T-category patients \cite{sanguineti_defining_2009} available at that time.
+The limited data did not allow us to quantify the probability of metastases in the rarely involved \gls{lnl} I, V, and VII, nor did the data allow us to verify that the \gls{hmm} is adequate to describe the dependence on lymph node involvement on T-category.
+%Since this publication only includes details about the involvement of the ipsilateral \glspl{lnl} I, II, III, and IV, the models were limited to these nodal levels as well.
+%With the publication of larger and more detailed datasets \cite{ludwig_dataset_2022}, we are thus able to
+In this paper, we extend the previous work \cite{ludwig_hidden_2021} by making the following contributions:
 \begin{enumerate}
-    \item include all relevant \glspl{lnl} into the model, namely the levels I, II, III, IV, V, and VII.
-    \item train the probabilistic model based on a multi-centric dataset consisting of \variable{output/num_patients.txt} patients from four institutions.
-    \item present exaples of results on personalized risk estimations performed with the trained model.
+    \item We provide a \glspl{hmm} of ipsilateral lymph node involvement including all relevant \glspl{lnl}, namely the levels I, II, III, IV, V, and VII.
+To determine the underlying \gls{dag} we compare different graphs by calculating the model evidence through \gls{ti}.
+    \item We collect a multi-centric dataset consisting of \variable{output/num_patients.txt} patients from three institutions, allowing us to train the model based on a sizable dataset.
+    \item We use the trained model to provide personalized risk estimations for occult metastases for typical clinical states of tumor progression at diagnosis, illustrating its potential for guiding volume-deescalated treatment strategies in the future.
 \end{enumerate}
@@
 
-Now, the main task of personalizing the \gls{ctv-n} definition is to predict the probability of the hidden possible states $\boldsymbol{\xi}_k$ given the diagnosis $d^\star=\boldsymbol{\zeta}_\ell$ of a new patient.
+The main task for personalizing the \gls{ctv-n} definition is to predict the probability of the hidden possible states $\boldsymbol{\xi}_k$ given the diagnosis $d^\star=\boldsymbol{\zeta}_\ell$ of a new patient.
 Using Bayes' theorem, we get
@@
 The described model along with the inferred parameters $\boldsymbol{\hat{\theta}}$ will yield an estimate (or multiple estimates) for the ``prior'' in the above equation $P\big( \boldsymbol{\xi}_k \mid \boldsymbol{\hat{\theta}} \big)$.
+
+\ju{Also say that the above equation gives the probability of a hidden state. To get the probability of involvement in a given level one has to marginalize over all corresponding states}
 
@@
 One after the other, each of the green arcs was added to the base graph.
-Subsequently, the performance of the resulting models in terms of its \gls{bic} was compared to the base graph to asssess whether the additional edge should be kept in the \emph{winning graph} or not.}
+Subsequently, the performance of the resulting models in terms of its \gls{bic} was compared to the base graph to assess whether the additional edge should be kept in the \emph{winning graph} or not.}
     \label{fig:extended_graph}
@@
 
-The \gls{dag} shown in \cref{fig:graph_with_obs} includes the \glspl{lnl} II, III, and IV since previous models \cite{pouymayou_bayesian_2019,ludwig_hidden_2021} had to rely on a reconstructed dataset of \gls{opscc} patients that contained only information on \gls{lnl} I and these three levels \cite{sanguineti_defining_2009}.
-Now, with datasets available that almost always report the \glspl{lnl} I, II, III, IV, V, and VII, we may add more \glspl{rv} to extend the graph and investigate how these \glspl{rv} should be connected among each other.
+The \gls{dag} shown in \cref{fig:graph_with_obs} includes the \glspl{lnl} II, III, and IV, which represent the most relevant lymph node levels for oropharyngeal SCC.
+It includes the arcs from II to III and from III to IV, representing the main direction of lymphatic drainage, which is well motivated anatomically and by the data on lymph node involvement.
+Previous publications  \cite{pouymayou_bayesian_2019,ludwig_hidden_2021} focused on these levels because they relied on a limited reconstructed dataset of \gls{opscc} patients \cite{sanguineti_defining_2009}.
+Now, with the datasets available for this work, we can extend the graph to include random variables for all \glspl{lnl} that are relevant for oropharyngeal SCC: I, II, III, IV, V, and VII.
+The main question to answer is: which arcs between \glspl{lnl} are needed to accurately model the data on lymph node involvement without increasing the model's complexity unnecessarily.
 
-First, we notice that the direct arcs from tumor to each of the \glspl{lnl} must be present, since every \gls{lnl} appears metastatic in isolation at least once.
+First, we notice that the direct arcs from tumor to each of the \glspl{lnl} must be present, since every \gls{lnl} appears metastatic in isolation at least once in the dataset.
 For example, some patients presented with metastases in \gls{lnl} I, while the other levels appeared healthy.
-If no spread was allowed from the tumor to $X_1$ (i.e., $b_i = 0$), the likelihood of observing this patient would be zero.
+If no spread was allowed from the tumor to $X_1$ (i.e., $b_1 = 0$), the likelihood of observing this patient would be zero.
 
@@
 It contains only the connections from \gls{lnl} II to III and an arc from \gls{lnl} III to IV, as motivated by the \emph{main lymphatic pathway} \cite{lengele_anatomical_2007}.
-In contrast to the \gls{dag} used in \cite{pouymayou_bayesian_2019,ludwig_hidden_2021}, we decided not to include the edge from level I to II into the base graph, since \gls{lnl} I is rarely involved and inferring $t_{12}$ is therefore difficult.
-Rather, we wanted to use the model comparison methodology outlined in \cref{subsec:complete_model:comparison} to see if the arc from \gls{lnl} I to II can be motivated from our data or if the arc should be flipped and go from \gls{lnl} II to I.
+The base graph is illustrated in \cref{fig:extended_graph} via the red and blue arcs.
+The lymphatic drainage to or from levels I, V, and VII is not as clearly defined.
+Therefore, we define a set of candidate arcs (green arcs in \cref{fig:extended_graph}) and use the model comparison methodology outlined in \cref{subsec:complete_model:comparison} to see which arc is motivated from our data.
+
+To connect level I, we investigate two candidate arcs: from I to II and from II to I.
+An arc from I to II was used in \cite{pouymayou_bayesian_2019,ludwig_hidden_2021} and is anatomically motivated.
+However, since \gls{lnl} I is rarely involved compared to level II, the associated parameter $t_{12}$ is mostly undetermined.
+Therefore, we also consider the flipped arc from \gls{lnl} II to I and investigate if it helps to describe the correlations between the involvement of levels I and II.
 
@@
 
-The dataset $\boldsymbol{\mathcal{D}}$ that we used for inference is comprised of the detailed reports on lymphatic involvement patterns in \gls{opscc} patients treated at three different institutions in France and Switzerland: The \gls{clb} in Lyon (France), the \gls{isb}, and the \gls{usz} (both in Switzerland).
-We have previously published the patterns of nodal involvement for the \gls{usz} cohort \cite{ludwig_dataset_2022} and described its characteristics in detail \cite{ludwig_detailed_2022}.
-A large part of the \gls{clb} data underlies a publication on \gls{hpv} status in \gls{opscc} \cite{bauwens_prevalence_2021} that was kindly provided to us by Vincent Grégoire.
-The patient information on lymphatic metastatic involvement from the \gls{isb} cohort was sent to us by Roland Giger through a collaboration.
+The dataset $\boldsymbol{\mathcal{D}}$ that we used for inference is comprised of the detailed reports on lymph node involvement patterns in \gls{opscc} patients treated at three different institutions in France and Switzerland: The \gls{clb} in Lyon (France), the \gls{isb}, and the \gls{usz} (both in Switzerland).
+We have previously published the patterns of nodal involvement for the \gls{usz} cohort (287 patients) \cite{ludwig_dataset_2022} and described its characteristics in detail \cite{ludwig_detailed_2022}.
+A large part of the \gls{clb} data (263 patients) underlies a publication on \gls{hpv} status in \gls{opscc} \cite{bauwens_prevalence_2021}.
+The data from \gls{clb} and \gls{isb} are described in details in a data set publication [lets cite it as under review and attach the submitted draft of the DiB].
 
-In total, the dataset contains \variable{output/num_patients.txt} patients newly diagnosed with \gls{opscc} that were treated with either definitive radiotherapy or neck dissection.
-Pathologically assessed post-operative \gls{lnl} involvement was available for 393 of patients, while for the remainder the nodal involvement was assessed based on available diagnostic modalities.
-If multiple modalities were used to diagnose a patient's lymphatic involvement and conflicting conclusions were drawn from this, the conflicts were resolved by inferring the most likely state (healthy or metastatic) for each \gls{lnl} separately.
+In total, the dataset contains \variable{output/num_patients.txt} patients with newly diagnosed \gls{opscc}.
+It includes patients treated with definitive (chemo)radiotherapy, adjuvant (chemo)radiotherapy following neck dissection, or neck dissection alone.
+Pathologically assessed \gls{lnl} involvement was available for 393 surgically treated patients, while for the remainder the nodal involvement was assessed based on available diagnostic modalities (FDG-PET-CT, CT, MRI, FNA).
+If multiple modalities were used to diagnose a patient's lymph node involvement, the available modalities were combined into a consensus decision.
+When different modalities were conflicting, the conflicts were resolved by inferring the most likely state (healthy or metastatic) for each \gls{lnl} separately.
 To do so, we used literature values for the sensitivity and specificity of the diagnostic modalities \cite{de_bondt_detection_2007,kyzas_18f-fluorodeoxyglucose_2008}, which we also tabulated in \cref{table:sens_spec}.
+Practically, this means that, whenever pathology after neck dissection was available, the pathology result was taken as the consensus, overruling any other clinical diagnostic modality.
+If, for example, PET-CT and MRI ware available and conflicting, PET-CT was taken as the consensus, overruling MRI.
+
+The dataset containing the consensus decision for the involvement of each level in every patient was then used for model parameter learning, assuming that it represents an observation of the true hidden state $\boldsymbol{\xi}_k$.
+The frequencies of some of the most important combinations of involved lymph node levels are listed in \cref{table:data_prevalence}.
 
@@
 \label{sec:results}
-
-\begin{table}
-    \centering
-    \variable{output/results_table.tex}
-    \caption{Model comparison results from the base graph and the extended graph we chose as the ``winnning'' model.
-For both \glspl{dag} we show the log-evidence, computed via thermodynamic integration, the negative one half of the \gls{bic}, as well as the maximum log-likelihood that was encountered during the final \gls{mcmc} sampling round.}
-    \label{table:evidence}
-\end{table}
 
@@
 Notably, the model is not only accurate in its predictions, it also correctly estimates the variance stemming from the limited amount of data.
-The spread between involvement prevalences of early and late T-category tumors is also represented well.
+The separation between involvement prevalences of early and late T-category tumors is also reproduced well by the model.
+This is remarkable because the model introduces only a single parameter to describe the differences between early and late T-category for all involvement patterns.
 This shows that expecting later diagnosis times, on average, for patients with advanced T-category tumors is able to explain more severe lymphatic involvement.
 
-There are only a few involvement patterns for which the model's prediction does not match the prevalence of the data with perfect accuracy.
+%I wouldn't even describe these minor deviations
+
+%There are only a few involvement patterns for which the model's prediction does not match the prevalence of the data with perfect accuracy.
 But in these instances, e.g.
@@
 It is interesting to note that not all involvement patterns become more prevalent with advanced T-category.
-For example, a healthy \gls{lnl} III is observed slightly less often together with a metastatic level II in advanced, compared to early T-category.
-This is because in the latter instance, it is more likely the disease has already spread to \gls{lnl} III.
+For example, a healthy \gls{lnl} III together with a metastatic level II is observed slightly less often for advanced T-category tumors compared to early T-category (yellow histogram, row 1 versus 2).
+This is because, for advanced T-category, it is more likely the disease has already spread to \gls{lnl} III (blue histogram, row 1 versus 2).
 Our model captures this accurately and precisely.
@@
 
-The changes to the graph connections we investigate in this work focus on the improvements w.r.t.
-involvement patterns that include the \glspl{lnl} I, V, and VII, that more rarely harbor occult disease.
+\begin{table}
+    \centering
+    \variable{output/results_table.tex}
+    \caption{Model comparison results from the base graph and the extended graph we chose as the ``winnning'' model.
+For both \glspl{dag} we show the log-evidence, computed via thermodynamic integration, the negative one half of the \gls{bic}, as well as the maximum log-likelihood that was encountered during the final \gls{mcmc} sampling round.}
+    \label{table:evidence}
+\end{table}
 
-In \cref{fig:IandII_prevs,fig:IIandV_prevs,fig:IIIandV_prevs} the base graph's estimations for prevalences of involvement patterns that include the \glspl{lnl} I, V, and VII have been plotted as gray, hatched histograms.
+\Cref{table:evidence} reports the model evidence for all extended graphs that were investigated.
+The results is additionally visualized in Figure xxx.
+The winning graph adds arcs from level II to I, from level II to V, and from level III to level V to the base graph.
+Below, we further analyze why these connections improve the model's ability to describe the involvement patterns seen in the data.
+
+The most likely model parameters for the winning graph, corresponding to the mean of the marginals of the sampled posterior distributions, are given by \ju{as placeholder taken from figure 7.5 of your thesis}: $b_1=?$, $b_2=0.37$, $b_3=0.07$, $b_4=0.01$, $b_5=?$, $b_7=0.02$ for the base probabilities, and $t_{23}=0.13$, $t_{34}=0.14$, $t_{21}=?$, $t_{25}=?$, $t_{35}=?$, and $p_{late}=0.4$ for the time prior parameter for advanced T-category patients (where we have fixed $p_{early}=0.3$ for early T-category tumors, and $t_{max}=10$ time steps).
+The result that $t_{23}$ and $t_{34}$ are relatively large compared to $b_3$ and $b_4$ reflects the observation that skip metastases in levels III and IV without involvement of the upstream level is rare.
+Since level II has no parent nodes except for the primary tumor, $b_2$ can be easily related to the prevalence of involvement in level II.
+The probability for no involvement of level II when the patient is diagnosed after $t$ time steps is $\left(1-b_2\right)^t$.
+The prevalence of level II involvement for late T-category patients is thus $1-\sum_{t=0}^{10} \left(1-0.37\right)^t 0.4^t \, (1-0.4)^{(10-t)}\binom{10}{t}\approx 0.8$, which agrees with \cref{fig:bg_prevalences}.
+
+\subsection{Involvement of levels I, V, and VII}
+
+First, we observe that the winning graph describes the involvement of levels II, III, and IV equally well as the base graph, a result that is expected and not further shown.
+We thus focus on the improvements w.r.t.
+involvement patterns that include the \glspl{lnl} I, V, and VII, that more rarely harbor metastases.
+
+{\it Level V:} In \cref{fig:IIandV_prevs,fig:IIIandV_prevs} we compare the base graph's and the winning graph's estimations for prevalences of involvement patterns that include \gls{lnl} V.
+
+, and VII have been plotted as gray, hatched histograms.
 Alongside, we show again the Beta posterior over the prevalence of the respective involvement pattern, given the data, as colored line.
@@
 
-\begin{figure*}
-    \script{comp_IandII_prevs.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/comp_IandII_prevs.png}
-        \caption{Comparison of observed and predicted prevalences of \gls{lnl} I and II involvement patterns.
-The top and bottom panels show the prevalences for early and late T-category, respectively.
-The solid lines are Beta posteriors from the data, while the histograms are predicted prevalences (colored: winning graph, gray-hatched: base graph).
-Blue and red plots indicate overall \gls{lnl} I and II involvement, respectively.
-Green plots indicate \gls{lnl} I involvement without level II, while orange plots indicate the opposite (\gls{lnl} II without level I).
-The winning graph has an added edge from \gls{lnl} II to I, which improves the prediction of the rare green pattern and reduces the overestimation of the orange pattern in late T-category.
-However, it slightly underestimates the orange pattern in early T-category.}
-        \label{fig:IandII_prevs}
-    \end{centering}
-\end{figure*}
-
 \begin{figure}
@@
         \caption{Beta posterior over the prevalence of the involvement pattern where \gls{lnl} II \emph{and} V show metastases (orange line), as well as the pattern with \gls{lnl} V involved \emph{without} level II (blue line).
-Both prevalences are shown for early (top panel) and late T-category (bottom panel).
+Both   are shown for early (top panel) and late T-category (bottom panel).
 The correspondingly colored histograms show the same prevalences as predicted by the winning graph model.
@@
 
+In \cref{fig:IandII_prevs,fig:IIandV_prevs,fig:IIIandV_prevs}
+
+\begin{figure*}
+    \script{comp_IandII_prevs.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/comp_IandII_prevs.png}
+        \caption{Comparison of observed and predicted prevalences of \gls{lnl} I and II involvement patterns.
+The top and bottom panels show the prevalences for early and late T-category, respectively.
+The solid lines are Beta posteriors from the data, while the histograms are predicted prevalences (colored: winning graph, gray-hatched: base graph).
+Blue and red plots indicate overall \gls{lnl} I and II involvement, respectively.
+Green plots indicate \gls{lnl} I involvement without level II, while orange plots indicate the opposite (\gls{lnl} II without level I).
+The winning graph has an added edge from \gls{lnl} II to I, which improves the prediction of the rare green pattern and reduces the overestimation of the orange pattern in late T-category.
+However, it slightly underestimates the orange pattern in early T-category.}
+        \label{fig:IandII_prevs}
+    \end{centering}
+\end{figure*}
+
 \subsection{Risk Prediction for Occult Disease}
 \label{subsec:results:risk_prediction}
+In this section, the model corresponding to the winning graph is applied to estimating the risk of occult metastases in clinically negative \glspl{lnl}.
+We assume a sensitivity of 0.76 and a specificity of 0.81 for the clinical diagnosis of lymph node metastases, corresponding to CT imaging in \cref{table:sens_spec}.
+\\
+
+{\it Level III:} \Cref{fig:wg_III_risks} compares the risk of occult disease in level III between patients that are clinically N0 (orange) and patients with clinically diagnosed involvement of only level II (red), for early T-category (upper panel) and advanced T-category (bottom panel).
+The histograms represent the uncertainty in model's risk prediction arising from the uncertainty in the model parameters and are generated by randomly drawing 1000 samples from the model parameter's joint posterior distribution.
+The model predicts a risk of 5\% for early T-stage tumors and 8\% for advanced T-stage.
+For patients with involvement of level II, the risk in level III increases to approximately 9\% and 12\%, respectively.
+
+\ju{perhaps one could use the same color in each risk figure for a given clinical involvement, e.g. always green for N0}
+
+\begin{figure}
+    \script{wg_III_risks.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/wg_III_risks.png}
+        \caption{TODO: Write caption.}
+        \label{fig:wg_III_risks}
+    \end{centering}
+\end{figure}
+
+{\it Level IV:} \Cref{fig:wg_IV_risks} compares the risk of occult disease in level IV for the typical clinical presentations: clinically N0 (grey), metastases in level II (blue), and metastases in levels II and III (orange).
+The model predicts a low risk of 1-2\% in level IV for patients with clinically healthy level III.
+For patients with clinical involvement of level III, the risk of occult disease in level IV increases to approximately 3\% for early T-stage and 5\% for advanced T-stage tumors.
 
@@
             \item visible metastases in \gls{lnl} II, but otherwise healthy-looking lymph nodes (blue histogram), and
-            \item macroscopic metastases in the \glspl{lnl} II \& III, whith the rest of the neck still being clinically ode negative (orange histogram).
+            \item macroscopic metastases in the \glspl{lnl} II \& III, with the rest of the neck still being clinically node negative (orange histogram).
         \end{enumerate*}
@@
 \end{figure}
+
+{\it Level V:} \Cref{fig:wg_V_risks} shows the risk of occult disease in level V depending on T-category and the clinical involvement of levels II-IV.
+For clinically N0 patients, the risk in level V is estimated to be around 1\%.
+Extensive nodal involvement of levels II-IV increases the risk in level V to approximately 4\% for advanced T-stage tumors.
 
@@
 
+{\it Level I:} \Cref{fig:wg_I_risks} shows the risk of occult disease in level I depending on T-category and the clinical involvement of levels II-IV.
+For clinically N0 patients, the risk in level I is estimated to be in the order of 1-2\%.
+Extensive nodal involvement of levels II-IV increases the risk in level I to approximately 4-5\% for advanced T-stage tumors.
+It is pointed out that the winning graph does not contain arcs from levels III or IV to level I (and anatomically we do not assume that there is lymphatic drainage from levels III or IV to level I).
+Thus, the increased risk in level I is related to the time evolution: Involvement of levels III and IV corresponds to a more advances state of disease that is likely diagnosed at a later time step, such that the tumor had more time to spread to level I.
+
+\begin{figure}
+    \script{wg_I_risks.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/wg_I_risks.png}
+        \caption{TODO: Write caption.}
+        \label{fig:wg_I_risks}
+    \end{centering}
+\end{figure}
+
+\begin{figure}
+    \script{wg_any_risks.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/wg_any_risks.png}
+        \caption{TODO: Write caption.}
+        \label{fig:wg_any_risks}
+    \end{centering}
+\end{figure}
+
+\section{Discussion}
+\label{sec:discussion}
+\subsection{Summary} 
+In this publication we present a statistical model of ipsilateral lymph node involvement in oropharyngeal SCC patients.
+Although the basic \gls{hmm} of lymphatic progression has been conceptually introduced in a previous work \cite{ludwig_hidden_2021}, this is the first publication that evaluates the model based on a large multi-institutional dataset containing \variable{output/num_patients.txt} patients.
+It is demonstrated for the first time that the model can accurately describe the patterns of lymph node involvement observed in the data, including the correlations between levels and its dependence on T-stage.
+Furthermore, state-of-the-art techniques are applied to calculate the model evidence for Bayesian model comparison.
+This yields a complete model including all \glspl{lnl} relevant for \gls{opscc}: I, II, III, IV, V, and VII with an parameterization that balances accuracy and model complexity.
+
+\subsection{Implications for elective nodal treatment} 
+Risk predictions obtained by the model may be used to design clinical trials on volume-deescalated treatment of \gls{opscc}.
+In the context of radiotherapy, this corresponds to excluding \glspl{lnl} from the CTV-N, which are irradiated according to the current guidelines.
+The list below should be seen as a summary of \cref{subsec:results:risk_prediction} and the limitations discussed in \cref{subsec:disc:limitation} should be taken into account in its interpretation.
+Assuming that one accepts approximately a 5\% risk of occult metastases per \gls{lnl}, the statistical model presented in this paper would suggest to:
+\begin{itemize}
+    \item[$\bullet$] Irradiate level II for all patients.
+\ju{maybe we should actually show the risk in level II for N0 patients, at least state the risk in the text, are even have the risk histogram for completeness.}
+    \item[$\bullet$] Irradiate level III for most patients. Only for clinically N0, early T-category patients, not irradiating level III could be considered. 
+    \item[$\bullet$] Exclude level IV from the CTV-N for patients with clinically negative level III.
+For advanced T-category patients with involvement of level III, level IV should be irradiated.
+%For early T-category patients with involvement of level III, not irradiating level IV could be considered.
+    \item[$\bullet$] Exclude level V from the CTV-N for most patients. Only for patients with extensive involvement of levels II, III, and IV, irradiation of level V can be considered.
+    \item[$\bullet$] Exclude level I from the CTV-N for early T-category patients with limited metastatic disease.
+For advanced T-category patients with extensive nodal involvement of levels II, III, and IV, level I should be irradiated (see also the limitations discussed below).
+    \item[$\bullet$] \ju{Level VII}
+\end{itemize}
+
+\subsection{Limitations and future work}
+\label{subsec:disc:limitation}
+{\it T-stage dependence of level I involvement:} As shown in \cref{fig:bg_prevalences}, the model describes the involvement of \glspl{lnl} II, III, and IV depending on T-stage very well despite having only a single parameter related to T-stage.
+\Cref{fig:IandII_prevs} shows that the model does not perfectly describe the T-stage dependence of level I.
+It adjusts the parameters such that level I involvement is correctly described for the set of all patients combined, but it overestimates level I involvement for early T-stage and underestimates it for advanced T-stage.
+A possible explanation is that advanced T-stage tumors are more likely to have grown into a regions with direct lymph drainage to level I.
+The more severe involvement in levels II, III, IV for advanced T-stage can be explained by tumors having more time to spread while keeping the spread probability rates $b_2$, $b_3$, $b_4$ constant.
+Regarding level I, early versus advanced T-stage tumors may have different spread probability rates $b_1$.
+
+\begin{figure}
+    \script{wg_sens_spec_risks.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/wg_sens_spec_risks.png}
+        \caption{Dependence of risk of occult disease on sensitivity and specificity.
+The figure shows the mean risk (i.e.
+averaged over the risk calculated for different model parameter samples \ju{is that what you do?}) in level IV in patients with clinical involvement of levels II and III.}
+        \label{fig:wg_sens_spec_risks}
+    \end{centering}
+\end{figure}
+
+{\it Sensitivity and Specificity:} Estimating the risk of occult metastases depends on the assumed parameter values for sensitivity and specificity of clinical detection of lymph node metastases.
+For this work, we adopted literature values for sensitivity and specificity.
+However, different authors have estimated these values using different criteria and different methods.
+Consequently, these values need to be considered with caution.
+\cref{fig:wg_sens_spec_risks} illustrates for one example how the risk of occult disease depends on sensitivity and specificity.
+Here, we consider the risk in level IV in patients with clinically detected metastases in levels II and III.
+For our default parameters of 81\% specificity and 76 \% sensitivity, the risk is 5\%, but it increases to x\% for a a sensitivity of 66\%.
+Future work may aim at developing new methods to model the difference between pathological and clinical lymph node involvement based on surgically treated patients in whom both is reported.
+
 \bibliography{bib}

### 2023-08-11 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-The most likely model parameters for the winning graph, corresponding to the mean of the marginals of the sampled posterior distributions, are given by \ju{as placeholder taken from figure 7.5 of your thesis}: $b_1=?$, $b_2=0.37$, $b_3=0.07$, $b_4=0.01$, $b_5=?$, $b_7=0.02$ for the base probabilities, and $t_{23}=0.13$, $t_{34}=0.14$, $t_{21}=?$, $t_{25}=?$, $t_{35}=?$, and $p_{late}=0.4$ for the time prior parameter for advanced T-category patients (where we have fixed $p_{early}=0.3$ for early T-category tumors, and $t_{max}=10$ time steps).
+The most likely model parameters for the winning graph, corresponding to the mean of the marginals of the sampled posterior distributions, are given by \ju{as placeholder taken from figure 7.5 of your thesis}: $b_1=\variable{output/wg_mean_b1.txt}$, $b_2=0.37$, $b_3=0.07$, $b_4=0.01$, $b_5=?$, $b_7=0.02$ for the base probabilities, and $t_{23}=0.13$, $t_{34}=0.14$, $t_{21}=?$, $t_{25}=?$, $t_{35}=?$, and $p_{late}=0.4$ for the time prior parameter for advanced T-category patients (where we have fixed $p_{early}=0.3$ for early T-category tumors, and $t_{max}=10$ time steps).
 The result that $t_{23}$ and $t_{34}$ are relatively large compared to $b_3$ and $b_4$ reflects the observation that skip metastases in levels III and IV without involvement of the upstream level is rare.

### 2023-08-11 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2023-08-11 src/tex/ms.tex
@@
 
-The most likely model parameters for the winning graph, corresponding to the mean of the marginals of the sampled posterior distributions, are given by \ju{as placeholder taken from figure 7.5 of your thesis}: $b_1=?$, $b_2=0.37$, $b_3=0.07$, $b_4=0.01$, $b_5=?$, $b_7=0.02$ for the base probabilities, and $t_{23}=0.13$, $t_{34}=0.14$, $t_{21}=?$, $t_{25}=?$, $t_{35}=?$, and $p_{late}=0.4$ for the time prior parameter for advanced T-category patients (where we have fixed $p_{early}=0.3$ for early T-category tumors, and $t_{max}=10$ time steps).
+The most likely model parameters for the winning graph, corresponding to the mean of the marginals of the sampled posterior distributions, are given by \ju{as placeholder taken from figure 7.5 of your thesis}: $b_1=\variable{output/wg_mean_b1.txt}\%$, $b_2=\variable{output/wg_mean_b2.txt}\%$, $b_3=\variable{output/wg_mean_b3.txt}\%$, $b_4=\variable{output/wg_mean_b4.txt}\%$, $b_5=\variable{output/wg_mean_b5.txt}\%$, $b_7=\variable{output/wg_mean_b7.txt}\%$ for the base probabilities, and $t_{23}=\variable{output/wg_mean_t23.txt}\%$, $t_{34}=\variable{output/wg_mean_t34.txt}\%$, $t_{21}=\variable{output/wg_mean_t21.txt}\%$, $t_{25}=\variable{output/wg_mean_t25.txt}\%$, $t_{35}=\variable{output/wg_mean_t35.txt}\%$, and $p_\text{late}=\variable{output/wg_mean_late_p.txt}\%$ for the time prior parameter for advanced T-category patients (where we have fixed $p_{early}=0.3$ for early T-category tumors, and $t_{max}=10$ time steps).
 The result that $t_{23}$ and $t_{34}$ are relatively large compared to $b_3$ and $b_4$ reflects the observation that skip metastases in levels III and IV without involvement of the upstream level is rare.

### 2023-08-11 src/scripts/wg_mean_b2.py
add mean param variable scripts

@@ -1,0 +1,26 @@
+"""
+Write mean spread probability from the tumor to LNL II to a file.
+"""
+from pathlib import Path
+
+from emcee.backends import HDFBackend
+
+import paths
+
+
+INPUT = paths.data / "wg_samples.hdf5"
+OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
+THIN = 10
+
+
+if __name__ == "__main__":
+    # open samples
+    backend = HDFBackend(paths.data / INPUT, read_only=True)
+    samples = backend.get_chain(flat=True, thin=THIN)
+
+    # compute mean
+    mean = samples[:, 1].mean()
+
+    # write to file
+    with open(OUTPUT, "w", encoding="utf-8") as f:
+        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_b3.py

@@ -1,0 +1,26 @@
+"""
+Write mean spread probability from the tumor to LNL IV to a file.
+"""
+from pathlib import Path
+
+from emcee.backends import HDFBackend
+
+import paths
+
+
+INPUT = paths.data / "wg_samples.hdf5"
+OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
+THIN = 10
+
+
+if __name__ == "__main__":
+    # open samples
+    backend = HDFBackend(paths.data / INPUT, read_only=True)
+    samples = backend.get_chain(flat=True, thin=THIN)
+
+    # compute mean
+    mean = samples[:, 3].mean()
+
+    # write to file
+    with open(OUTPUT, "w", encoding="utf-8") as f:
+        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_b4.py

@@ -1,0 +1,26 @@
+"""
+Write mean spread probability from the tumor to LNL I to a file.
+"""
+from pathlib import Path
+
+from emcee.backends import HDFBackend
+
+import paths
+
+
+INPUT = paths.data / "wg_samples.hdf5"
+OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
+THIN = 10
+
+
+if __name__ == "__main__":
+    # open samples
+    backend = HDFBackend(paths.data / INPUT, read_only=True)
+    samples = backend.get_chain(flat=True, thin=THIN)
+
+    # compute mean
+    mean = samples[:, 0].mean()
+
+    # write to file
+    with open(OUTPUT, "w", encoding="utf-8") as f:
+        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_b5.py

@@ -1,0 +1,26 @@
+"""
+Write mean spread probability from the tumor to LNL V to a file.
+"""
+from pathlib import Path
+
+from emcee.backends import HDFBackend
+
+import paths
+
+
+INPUT = paths.data / "wg_samples.hdf5"
+OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
+THIN = 10
+
+
+if __name__ == "__main__":
+    # open samples
+    backend = HDFBackend(paths.data / INPUT, read_only=True)
+    samples = backend.get_chain(flat=True, thin=THIN)
+
+    # compute mean
+    mean = samples[:, 4].mean()
+
+    # write to file
+    with open(OUTPUT, "w", encoding="utf-8") as f:
+        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_b7.py

@@ -1,0 +1,26 @@
+"""
+Write mean spread probability from the tumor to LNL VII to a file.
+"""
+from pathlib import Path
+
+from emcee.backends import HDFBackend
+
+import paths
+
+
+INPUT = paths.data / "wg_samples.hdf5"
+OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
+THIN = 10
+
+
+if __name__ == "__main__":
+    # open samples
+    backend = HDFBackend(paths.data / INPUT, read_only=True)
+    samples = backend.get_chain(flat=True, thin=THIN)
+
+    # compute mean
+    mean = samples[:, 5].mean()
+
+    # write to file
+    with open(OUTPUT, "w", encoding="utf-8") as f:
+        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_late_p.py

@@ -1,0 +1,26 @@
+"""
+Write mean late T-stage's binomial parameter to a file.
+"""
+from pathlib import Path
+
+from emcee.backends import HDFBackend
+
+import paths
+
+
+INPUT = paths.data / "wg_samples.hdf5"
+OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
+THIN = 10
+
+
+if __name__ == "__main__":
+    # open samples
+    backend = HDFBackend(paths.data / INPUT, read_only=True)
+    samples = backend.get_chain(flat=True, thin=THIN)
+
+    # compute mean
+    mean = samples[:, 11].mean()
+
+    # write to file
+    with open(OUTPUT, "w", encoding="utf-8") as f:
+        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_t21.py

@@ -1,0 +1,26 @@
+"""
+Write mean spread probability from LNL II to LNL I to a file.
+"""
+from pathlib import Path
+
+from emcee.backends import HDFBackend
+
+import paths
+
+
+INPUT = paths.data / "wg_samples.hdf5"
+OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
+THIN = 10
+
+
+if __name__ == "__main__":
+    # open samples
+    backend = HDFBackend(paths.data / INPUT, read_only=True)
+    samples = backend.get_chain(flat=True, thin=THIN)
+
+    # compute mean
+    mean = samples[:, 6].mean()
+
+    # write to file
+    with open(OUTPUT, "w", encoding="utf-8") as f:
+        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_t23.py

@@ -1,0 +1,26 @@
+"""
+Write mean spread probability from LNL II to LNL III to a file.
+"""
+from pathlib import Path
+
+from emcee.backends import HDFBackend
+
+import paths
+
+
+INPUT = paths.data / "wg_samples.hdf5"
+OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
+THIN = 10
+
+
+if __name__ == "__main__":
+    # open samples
+    backend = HDFBackend(paths.data / INPUT, read_only=True)
+    samples = backend.get_chain(flat=True, thin=THIN)
+
+    # compute mean
+    mean = samples[:, 7].mean()
+
+    # write to file
+    with open(OUTPUT, "w", encoding="utf-8") as f:
+        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_t25.py

@@ -1,0 +1,26 @@
+"""
+Write mean spread probability from LNL II to LNL V to a file.
+"""
+from pathlib import Path
+
+from emcee.backends import HDFBackend
+
+import paths
+
+
+INPUT = paths.data / "wg_samples.hdf5"
+OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
+THIN = 10
+
+
+if __name__ == "__main__":
+    # open samples
+    backend = HDFBackend(paths.data / INPUT, read_only=True)
+    samples = backend.get_chain(flat=True, thin=THIN)
+
+    # compute mean
+    mean = samples[:, 8].mean()
+
+    # write to file
+    with open(OUTPUT, "w", encoding="utf-8") as f:
+        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_t34.py

@@ -1,0 +1,26 @@
+"""
+Write mean spread probability from LNL III to LNL IV to a file.
+"""
+from pathlib import Path
+
+from emcee.backends import HDFBackend
+
+import paths
+
+
+INPUT = paths.data / "wg_samples.hdf5"
+OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
+THIN = 10
+
+
+if __name__ == "__main__":
+    # open samples
+    backend = HDFBackend(paths.data / INPUT, read_only=True)
+    samples = backend.get_chain(flat=True, thin=THIN)
+
+    # compute mean
+    mean = samples[:, 9].mean()
+
+    # write to file
+    with open(OUTPUT, "w", encoding="utf-8") as f:
+        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_t35.py

@@ -1,0 +1,26 @@
+"""
+Write mean spread probability from LNL III to LNL V to a file.
+"""
+from pathlib import Path
+
+from emcee.backends import HDFBackend
+
+import paths
+
+
+INPUT = paths.data / "wg_samples.hdf5"
+OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
+THIN = 10
+
+
+if __name__ == "__main__":
+    # open samples
+    backend = HDFBackend(paths.data / INPUT, read_only=True)
+    samples = backend.get_chain(flat=True, thin=THIN)
+
+    # compute mean
+    mean = samples[:, 10].mean()
+
+    # write to file
+    with open(OUTPUT, "w", encoding="utf-8") as f:
+        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean.py
dynamically compute parameter means

@@ -1,0 +1,29 @@
+"""
+Write all mean parameter samples to a file.
+"""
+from pathlib import Path
+
+from emcee.backends import HDFBackend
+
+import paths
+
+
+INPUT = paths.data / "wg_samples.hdf5"
+OUTPUT = paths.output
+THIN = 10
+NAMES = ["b1", "b2", "b3", "b4", "b5", "b7", "t21", "t23", "t25", "t34", "t35", "late_p"]
+
+
+if __name__ == "__main__":
+    # open samples
+    backend = HDFBackend(paths.data / INPUT, read_only=True)
+    samples = backend.get_chain(flat=True, thin=THIN)
+
+    for i in range(samples.shape[1]):
+        # compute mean
+        mean = samples[:, i].mean()
+
+        # write to file
+        filename = f"wg_mean_{NAMES[i]}.txt"
+        with open(OUTPUT / filename, "w", encoding="utf-8") as f:
+            f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_b1.py

@@ -1,26 +1,0 @@
-"""
-Write mean spread probability from the tumor to LNL I to a file.
-"""
-from pathlib import Path
-
-from emcee.backends import HDFBackend
-
-import paths
-
-
-INPUT = paths.data / "wg_samples.hdf5"
-OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
-THIN = 10
-
-
-if __name__ == "__main__":
-    # open samples
-    backend = HDFBackend(paths.data / INPUT, read_only=True)
-    samples = backend.get_chain(flat=True, thin=THIN)
-
-    # compute mean
-    mean = samples[:, 0].mean()
-
-    # write to file
-    with open(OUTPUT, "w", encoding="utf-8") as f:
-        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_b2.py

@@ -1,26 +1,0 @@
-"""
-Write mean spread probability from the tumor to LNL II to a file.
-"""
-from pathlib import Path
-
-from emcee.backends import HDFBackend
-
-import paths
-
-
-INPUT = paths.data / "wg_samples.hdf5"
-OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
-THIN = 10
-
-
-if __name__ == "__main__":
-    # open samples
-    backend = HDFBackend(paths.data / INPUT, read_only=True)
-    samples = backend.get_chain(flat=True, thin=THIN)
-
-    # compute mean
-    mean = samples[:, 1].mean()
-
-    # write to file
-    with open(OUTPUT, "w", encoding="utf-8") as f:
-        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_b3.py

@@ -1,26 +1,0 @@
-"""
-Write mean spread probability from the tumor to LNL IV to a file.
-"""
-from pathlib import Path
-
-from emcee.backends import HDFBackend
-
-import paths
-
-
-INPUT = paths.data / "wg_samples.hdf5"
-OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
-THIN = 10
-
-
-if __name__ == "__main__":
-    # open samples
-    backend = HDFBackend(paths.data / INPUT, read_only=True)
-    samples = backend.get_chain(flat=True, thin=THIN)
-
-    # compute mean
-    mean = samples[:, 3].mean()
-
-    # write to file
-    with open(OUTPUT, "w", encoding="utf-8") as f:
-        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_b5.py

@@ -1,26 +1,0 @@
-"""
-Write mean spread probability from the tumor to LNL V to a file.
-"""
-from pathlib import Path
-
-from emcee.backends import HDFBackend
-
-import paths
-
-
-INPUT = paths.data / "wg_samples.hdf5"
-OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
-THIN = 10
-
-
-if __name__ == "__main__":
-    # open samples
-    backend = HDFBackend(paths.data / INPUT, read_only=True)
-    samples = backend.get_chain(flat=True, thin=THIN)
-
-    # compute mean
-    mean = samples[:, 4].mean()
-
-    # write to file
-    with open(OUTPUT, "w", encoding="utf-8") as f:
-        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_b7.py

@@ -1,26 +1,0 @@
-"""
-Write mean spread probability from the tumor to LNL VII to a file.
-"""
-from pathlib import Path
-
-from emcee.backends import HDFBackend
-
-import paths
-
-
-INPUT = paths.data / "wg_samples.hdf5"
-OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
-THIN = 10
-
-
-if __name__ == "__main__":
-    # open samples
-    backend = HDFBackend(paths.data / INPUT, read_only=True)
-    samples = backend.get_chain(flat=True, thin=THIN)
-
-    # compute mean
-    mean = samples[:, 5].mean()
-
-    # write to file
-    with open(OUTPUT, "w", encoding="utf-8") as f:
-        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_late_p.py

@@ -1,26 +1,0 @@
-"""
-Write mean late T-stage's binomial parameter to a file.
-"""
-from pathlib import Path
-
-from emcee.backends import HDFBackend
-
-import paths
-
-
-INPUT = paths.data / "wg_samples.hdf5"
-OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
-THIN = 10
-
-
-if __name__ == "__main__":
-    # open samples
-    backend = HDFBackend(paths.data / INPUT, read_only=True)
-    samples = backend.get_chain(flat=True, thin=THIN)
-
-    # compute mean
-    mean = samples[:, 11].mean()
-
-    # write to file
-    with open(OUTPUT, "w", encoding="utf-8") as f:
-        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_t21.py

@@ -1,26 +1,0 @@
-"""
-Write mean spread probability from LNL II to LNL I to a file.
-"""
-from pathlib import Path
-
-from emcee.backends import HDFBackend
-
-import paths
-
-
-INPUT = paths.data / "wg_samples.hdf5"
-OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
-THIN = 10
-
-
-if __name__ == "__main__":
-    # open samples
-    backend = HDFBackend(paths.data / INPUT, read_only=True)
-    samples = backend.get_chain(flat=True, thin=THIN)
-
-    # compute mean
-    mean = samples[:, 6].mean()
-
-    # write to file
-    with open(OUTPUT, "w", encoding="utf-8") as f:
-        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_t23.py

@@ -1,26 +1,0 @@
-"""
-Write mean spread probability from LNL II to LNL III to a file.
-"""
-from pathlib import Path
-
-from emcee.backends import HDFBackend
-
-import paths
-
-
-INPUT = paths.data / "wg_samples.hdf5"
-OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
-THIN = 10
-
-
-if __name__ == "__main__":
-    # open samples
-    backend = HDFBackend(paths.data / INPUT, read_only=True)
-    samples = backend.get_chain(flat=True, thin=THIN)
-
-    # compute mean
-    mean = samples[:, 7].mean()
-
-    # write to file
-    with open(OUTPUT, "w", encoding="utf-8") as f:
-        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_t25.py

@@ -1,26 +1,0 @@
-"""
-Write mean spread probability from LNL II to LNL V to a file.
-"""
-from pathlib import Path
-
-from emcee.backends import HDFBackend
-
-import paths
-
-
-INPUT = paths.data / "wg_samples.hdf5"
-OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
-THIN = 10
-
-
-if __name__ == "__main__":
-    # open samples
-    backend = HDFBackend(paths.data / INPUT, read_only=True)
-    samples = backend.get_chain(flat=True, thin=THIN)
-
-    # compute mean
-    mean = samples[:, 8].mean()
-
-    # write to file
-    with open(OUTPUT, "w", encoding="utf-8") as f:
-        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_t34.py

@@ -1,26 +1,0 @@
-"""
-Write mean spread probability from LNL III to LNL IV to a file.
-"""
-from pathlib import Path
-
-from emcee.backends import HDFBackend
-
-import paths
-
-
-INPUT = paths.data / "wg_samples.hdf5"
-OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
-THIN = 10
-
-
-if __name__ == "__main__":
-    # open samples
-    backend = HDFBackend(paths.data / INPUT, read_only=True)
-    samples = backend.get_chain(flat=True, thin=THIN)
-
-    # compute mean
-    mean = samples[:, 9].mean()
-
-    # write to file
-    with open(OUTPUT, "w", encoding="utf-8") as f:
-        f.write(f"{mean:.2%}")

### 2023-08-11 src/scripts/wg_mean_t35.py

@@ -1,26 +1,0 @@
-"""
-Write mean spread probability from LNL III to LNL V to a file.
-"""
-from pathlib import Path
-
-from emcee.backends import HDFBackend
-
-import paths
-
-
-INPUT = paths.data / "wg_samples.hdf5"
-OUTPUT = (paths.output / Path(__file__).name).with_suffix(".txt")
-THIN = 10
-
-
-if __name__ == "__main__":
-    # open samples
-    backend = HDFBackend(paths.data / INPUT, read_only=True)
-    samples = backend.get_chain(flat=True, thin=THIN)
-
-    # compute mean
-    mean = samples[:, 10].mean()
-
-    # write to file
-    with open(OUTPUT, "w", encoding="utf-8") as f:
-        f.write(f"{mean:.2%}")

### 2023-08-11 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2023-08-11 src/tex/ms.tex
@@
 \newcommand{\ju}[1]{{\textcolor{red}{#1}}}
+\newcommand{\rl}[1]{{\textcolor{blue}{#1}}}
 
@@
 
-The most likely model parameters for the winning graph, corresponding to the mean of the marginals of the sampled posterior distributions, are given by \ju{as placeholder taken from figure 7.5 of your thesis}: $b_1=?$, $b_2=0.37$, $b_3=0.07$, $b_4=0.01$, $b_5=?$, $b_7=0.02$ for the base probabilities, and $t_{23}=0.13$, $t_{34}=0.14$, $t_{21}=?$, $t_{25}=?$, $t_{35}=?$, and $p_{late}=0.4$ for the time prior parameter for advanced T-category patients (where we have fixed $p_{early}=0.3$ for early T-category tumors, and $t_{max}=10$ time steps).
+The most likely model parameters for the winning graph, corresponding to the mean of the marginals of the sampled posterior distributions, are given by: $b_1=\variable{output/wg_mean_b1.txt}\%$, $b_2=\variable{output/wg_mean_b2.txt}\%$, $b_3=\variable{output/wg_mean_b3.txt}\%$, $b_4=\variable{output/wg_mean_b4.txt}\%$, $b_5=\variable{output/wg_mean_b5.txt}\%$, $b_7=\variable{output/wg_mean_b7.txt}\%$ for the base probabilities, and $t_{23}=\variable{output/wg_mean_t23.txt}\%$, $t_{34}=\variable{output/wg_mean_t34.txt}\%$, $t_{21}=\variable{output/wg_mean_t21.txt}\%$, $t_{25}=\variable{output/wg_mean_t25.txt}\%$, $t_{35}=\variable{output/wg_mean_t35.txt}\%$, and $p_\text{late}=\variable{output/wg_mean_late_p.txt}\%$ for the time prior parameter for advanced T-category patients (where we have fixed $p_{early}=0.3$ for early T-category tumors, and $t_{max}=10$ time steps).
 The result that $t_{23}$ and $t_{34}$ are relatively large compared to $b_3$ and $b_4$ reflects the observation that skip metastases in levels III and IV without involvement of the upstream level is rare.
@@
         \caption{Dependence of risk of occult disease on sensitivity and specificity.
-The figure shows the mean risk (i.e.
-averaged over the risk calculated for different model parameter samples \ju{is that what you do?}) in level IV in patients with clinical involvement of levels II and III.}
+The figure shows the risk of involvement in level IV, given the mean of the parameter samples, drawn during sampling, in patients with clinical involvement of levels II and III.}
         \label{fig:wg_sens_spec_risks}

### 2023-08-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-The most likely model parameters for the winning graph, corresponding to the mean of the marginals of the sampled posterior distributions, are given by: $b_1=\variable{output/wg_mean_b1.txt}\%$, $b_2=\variable{output/wg_mean_b2.txt}\%$, $b_3=\variable{output/wg_mean_b3.txt}\%$, $b_4=\variable{output/wg_mean_b4.txt}\%$, $b_5=\variable{output/wg_mean_b5.txt}\%$, $b_7=\variable{output/wg_mean_b7.txt}\%$ for the base probabilities, and $t_{23}=\variable{output/wg_mean_t23.txt}\%$, $t_{34}=\variable{output/wg_mean_t34.txt}\%$, $t_{21}=\variable{output/wg_mean_t21.txt}\%$, $t_{25}=\variable{output/wg_mean_t25.txt}\%$, $t_{35}=\variable{output/wg_mean_t35.txt}\%$, and $p_\text{late}=\variable{output/wg_mean_late_p.txt}\%$ for the time prior parameter for advanced T-category patients (where we have fixed $p_{early}=0.3$ for early T-category tumors, and $t_{max}=10$ time steps).
+The most likely model parameters for the winning graph, corresponding to the mean of the marginals of the sampled posterior distributions, are tabulated in \cref{table:params}.
+We have fixed $p_\text{early}=0.3$ for early T-category tumors, and $t_\text{max}=10$ time steps).
 The result that $t_{23}$ and $t_{34}$ are relatively large compared to $b_3$ and $b_4$ reflects the observation that skip metastases in levels III and IV without involvement of the upstream level is rare.
@@
 The prevalence of level II involvement for late T-category patients is thus $1-\sum_{t=0}^{10} \left(1-0.37\right)^t 0.4^t \, (1-0.4)^{(10-t)}\binom{10}{t}\approx 0.8$, which agrees with \cref{fig:bg_prevalences}.
+
+\begin{table}[]
+    \centering
+    % \variable{output/wg_means.tex}
+    \caption{Mean of sampled parameters.}
+    \label{tab:params}
+\end{table}
 

### 2023-08-18 src/scripts/wg_mean.py
put parameter means in separate table

@@ -1,29 +1,0 @@
-"""
-Write all mean parameter samples to a file.
-"""
-from pathlib import Path
-
-from emcee.backends import HDFBackend
-
-import paths
-
-
-INPUT = paths.data / "wg_samples.hdf5"
-OUTPUT = paths.output
-THIN = 10
-NAMES = ["b1", "b2", "b3", "b4", "b5", "b7", "t21", "t23", "t25", "t34", "t35", "late_p"]
-
-
-if __name__ == "__main__":
-    # open samples
-    backend = HDFBackend(paths.data / INPUT, read_only=True)
-    samples = backend.get_chain(flat=True, thin=THIN)
-
-    for i in range(samples.shape[1]):
-        # compute mean
-        mean = samples[:, i].mean()
-
-        # write to file
-        filename = f"wg_mean_{NAMES[i]}.txt"
-        with open(OUTPUT / filename, "w", encoding="utf-8") as f:
-            f.write(f"{mean:.2%}")

### 2023-08-18 src/scripts/wg_means.py

@@ -1,0 +1,44 @@
+"""
+Write all mean parameter samples to a markdown table.
+"""
+import pandas as pd
+from emcee.backends import HDFBackend
+
+import paths
+
+
+INPUT = paths.data / "wg_samples.hdf5"
+OUTPUT = paths.output
+THIN = 10
+NAMES = [
+    r"b_\text{I}",
+    r"b_\text{II}",
+    r"b_\text{III}",
+    r"b_\text{IV}",
+    r"b_\text{V}",
+    r"b_\text{VII}",
+    r"t_{\text{II} \rightarrow {I}}",
+    r"t_{\text{II} \rightarrow {III}}",
+    r"t_{\text{II} \rightarrow {V}}",
+    r"t_{\text{III} \rightarrow {IV}}",
+    r"t_{\text{III} \rightarrow {V}}",
+    r"p_\text{late}",
+]
+
+
+if __name__ == "__main__":
+    # open samples
+    backend = HDFBackend(paths.data / INPUT, read_only=True)
+    samples = backend.get_chain(flat=True, thin=THIN)
+    means = samples.mean(axis=0)
+    stddevs = samples.std(axis=0)
+
+    means_dict = {
+        f"${name}$": [f"{100 * mean:.2f} \%", f"$\pm$ {100 * std:.2f} \%"]
+        for name, mean, std in zip(NAMES, means, stddevs)
+    }
+    df = pd.DataFrame.from_dict(means_dict, orient="index", columns=["mean", "std. dev."])
+    latex_table = df.to_latex(column_format="lrr")
+
+    with open(OUTPUT / "wg_means.tex", "w") as f:
+        f.write(latex_table)

### 2023-08-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage[inline]{enumitem}
+\usepackage{booktabs}
 
@@
 We have fixed $p_\text{early}=0.3$ for early T-category tumors, and $t_\text{max}=10$ time steps).
-The result that $t_{23}$ and $t_{34}$ are relatively large compared to $b_3$ and $b_4$ reflects the observation that skip metastases in levels III and IV without involvement of the upstream level is rare.
-Since level II has no parent nodes except for the primary tumor, $b_2$ can be easily related to the prevalence of involvement in level II.
-The probability for no involvement of level II when the patient is diagnosed after $t$ time steps is $\left(1-b_2\right)^t$.
-The prevalence of level II involvement for late T-category patients is thus $1-\sum_{t=0}^{10} \left(1-0.37\right)^t 0.4^t \, (1-0.4)^{(10-t)}\binom{10}{t}\approx 0.8$, which agrees with \cref{fig:bg_prevalences}.
+The result that $t_{\text{II} \rightarrow \text{III}}$ and $t_{\text{III} \rightarrow \text{IV}}$ are relatively large compared to $b_\text{III}$ and $b_\text{IV}$ reflects the observation that skip metastases in levels III and IV without involvement of the upstream level is rare.
+Since level II has no parent nodes except for the primary tumor, $b_\text{II}$ can be easily related to the prevalence of involvement in level II.
+The probability for no involvement of level II when the patient is diagnosed after $t$ time steps is $\left(1-b_\text{II}\right)^t$.
+The prevalence of level II involvement for late T-category patients is thus
+%
+\begin{equation}
+    \begin{aligned}
+        \text{prev}_\text{late}^\text{II} &= 1-\sum_{t=0}^{10} \left(1-b_\text{II}\right)^t \cdot p_\text{late}^t \, (1-p_\text{late})^{(10-t)}\binom{10}{t} \\
+        &\approx 0.8
+    \end{aligned}
+\end{equation}
+%
+which agrees with \cref{fig:bg_prevalences}.
 
@@
     \centering
-    % \variable{output/wg_means.tex}
-    \caption{Mean of sampled parameters.}
+    \variable{output/wg_means.tex}
+    \caption{Mean of sampled parameters in percent.}
     \label{tab:params}
@@
 \ju{perhaps one could use the same color in each risk figure for a given clinical involvement, e.g. always green for N0}
+\rl{Done}
 

### 2023-08-18 src/scripts/wg_III_risks.py
use consistent colors accross risk plots

@@ -36,13 +36,13 @@ if __name__ == "__main__":
                 filename=INPUT,
                 dataname=f"{stage}/III/N0",
                 label="LNL III, given N0",
-                color=USZ["orange"],
+                color=USZ["green"],
             ),
             Histogram.from_hdf5(
                 filename=INPUT,
                 dataname=f"{stage}/III/II",
                 label="LNL III, given LNL II involved",
-                color=USZ["red"],
+                color=USZ["blue"],
             ),
         ]
 

### 2023-08-18 src/scripts/wg_IV_risks.py

@@ -35,18 +35,8 @@ if __name__ == "__main__":
             Histogram.from_hdf5(
                 filename=INPUT,
                 dataname=f"{stage}/IV/N0",
-                label=None,
-                color=USZ["gray"],
-            ),
-            Histogram.from_hdf5(
-                filename=INPUT,
-                dataname=f"{stage}/IV/N0",
                 label="LNL IV, given N0",
-                color="#97a3a7",
-                histtype="step",
-                linewidth=2.,
-                alpha=1.0,
-                zorder=1.,
+                color=USZ["green"],
             ),
             Histogram.from_hdf5(
                 filename=INPUT,

### 2023-08-18 src/scripts/wg_any_risks.py

@@ -42,22 +42,22 @@ if __name__ == "__main__":
                 filename=INPUT,
                 dataname=f"{stage}/anyIorIVorVorVII/II",
                 label="I, IV, V, or VII, given LNL II involved",
-                color=USZ["orange"],
+                color=USZ["blue"],
             ),
             Histogram.from_hdf5(
                 filename=INPUT,
                 dataname=f"{stage}/anyIorIVorVorVII/IIandIII",
                 label="I, IV, V, or VII, given LNL II, and III involved",
-                color=USZ["red"],
+                color=USZ["orange"],
             ),
         ]
 
-    draw(axes[0], contents=panels["early"], xlims=(2., 16.), hist_kwargs={"nbins": 40})
+    draw(axes[0], contents=panels["early"], xlims=(0., 16.), hist_kwargs={"nbins": 40})
     axes[0].set_ylabel("early T-cat.", fontweight="bold")
     axes[0].set_yticks([])
     axes[0].legend()
 
-    draw(axes[1], contents=panels["late"], xlims=(2., 16.), hist_kwargs={"nbins": 40})
+    draw(axes[1], contents=panels["late"], xlims=(0., 16.), hist_kwargs={"nbins": 40})
     axes[1].set_ylabel("late T-cat.", fontweight="bold")
     axes[1].set_yticks([])
     axes[1].set_xlabel("risk [%]")

### 2023-08-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-{\it Level IV:} \Cref{fig:wg_IV_risks} compares the risk of occult disease in level IV for the typical clinical presentations: clinically N0 (grey), metastases in level II (blue), and metastases in levels II and III (orange).
+{\it Level IV:} \Cref{fig:wg_IV_risks} compares the risk of occult disease in level IV for the typical clinical presentations: clinically N0 (green), metastases in level II (blue), and metastases in levels II and III (orange).
 The model predicts a low risk of 1-2\% in level IV for patients with clinically healthy level III.
@@
 
+{\it Level II:} As can be seen in \cref{tab:params} the model infers the spread from the tumor to \gls{lnl} II to be the most probable transition at any given time step.
+As a consequence, even for an early T-category patient that presents with a clinical N0 neck, our model predicts a \variable{output/wg_II_risks.tex} risk for microscopic metastases in \gls{lnl} II.
+
 \begin{figure}
@@
         \includegraphics[width=\linewidth]{figures/wg_I_risks.png}
-        \caption{TODO: Write caption.}
+        \caption{Distributions over predicted risk for involvement in \gls{lnl} I, given different clinical diagnosis scenarios: For N0 patients (green), patients with macroscopic involvement in \gls{lnl} II (blue), and for the case where the \glspl{lnl} II, III, and IV show involvement.
+The top panel shows these risks for early T-category and the bottom row for late T-category.}
         \label{fig:wg_I_risks}
@@
         \includegraphics[width=\linewidth]{figures/wg_any_risks.png}
-        \caption{TODO: Write caption.}
+        \caption{TODO: Shown are the historgams over the predicted risk for involvement in any of the \glspl{lnl} I, IV, V, or VII.
+The risk is plotted given a clinical N0 diagnosis (green), macroscopically detected metastases in \gls{lnl} II (blue), and lastly given visible involvement in both \gls{lnl} II and III (orange).
+The top row shows these risks for early T-category diagnoses and the bottom row for late T-category.}
         \label{fig:wg_any_risks}

### 2023-08-18 src/scripts/wg_II_risks.py
add variable on LNL II risk

@@ -1,0 +1,24 @@
+"""
+Write the expected risk of LNL II and its standard deviation to a file.
+"""
+import h5py
+
+import paths
+
+
+INPUT = paths.data / "wg_risks.hdf5"
+OUTPUT = paths.output / "wg_II_risks.tex"
+
+
+if __name__ == "__main__":
+    # open risks
+    with h5py.File(INPUT, "r") as f:
+        risks = f["early/II/N0"][:]
+
+    # compute mean and standard deviation
+    mean = 100 * risks.mean()
+    stddev = 100 * risks.std()
+
+    # write to file
+    with open(OUTPUT, "w") as f:
+        f.write(f"{mean:.2f} \% $\pm$ {stddev:.2f} \%")

### 2023-08-18 src/scripts/wg_means.py
remove unnecessary thin param

@@ -9,7 +9,6 @@ import paths
 
 INPUT = paths.data / "wg_samples.hdf5"
 OUTPUT = paths.output
-THIN = 10
 NAMES = [
     r"b_\text{I}",
     r"b_\text{II}",
@@ -29,7 +28,7 @@ NAMES = [
 if __name__ == "__main__":
     # open samples
     backend = HDFBackend(paths.data / INPUT, read_only=True)
-    samples = backend.get_chain(flat=True, thin=THIN)
+    samples = backend.get_chain(flat=True)
     means = samples.mean(axis=0)
     stddevs = samples.std(axis=0)
 

### 2023-08-29 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     Conclusion: We provide a statistical model of lymph node involvement in \gls{opscc} patients trained on a multi-institutional dataset, which may guide the design of clinical trials on volume-deescalated treatment of \gls{opscc} and eventually contribute to further personalized guidelines on elective nodal treatment.
-%    The \gls{ctv-n}\glsunset{ctv} in \gls{hnscc}\glsunset{scc} is currently based mostly on the prevalence of lymphatic metastases in different \glspl{lnl} for a given primary tumor location.
-In this work, we present an extension to a probabilistic model for lymphatic metastatic spread developed earlier that can quantify the risk for microscopic nodal involvement based on an individual patient diagnosis.
-The extension is based on the same formalism of \glspl{hmm} as the original model, but in addition to the \glspl{lnl} I, II, III, and IV, it also covers the \glspl{lnl} V and VII of the ipsilateral neck.
-Moreover, we infer which pathways of lymphatic spread to model from clinical lymphatic progression patterns of 681 patients from three institutions.
-The extended model may allow for a more personalized \gls{ctv-n} definition based on a patient's individual state of disease.
-The \gls{hmm} uses a collection of hidden binary \glspl{rv} -- one for each \gls{lnl} -- to model a patient's state of lymphatic involvement.
-Clinical diagnoses and/or pathological examinations of resected \glspl{lnl} represent the observed binary \glspl{rv} corresponding to the unobservable true state of nodal disease.
-A \gls{dag} with parametrized edges is used to compute the transition matrix of the \gls{hmm} and consequently the (log-)likelihood of the diagnoses of 681 patients, given a set of spread parameters.
-Based on the defined likelihood function, \gls{mcmc} sampling is then used to infer a distribution over the model's parameters.
-T-category is model by assuming that, on average, patients with late T-category tumors are diagnosed at a later time compared to patients with early T-category tumors.
-Using \gls{ti} and the \gls{bic} we compare models based on different \glspl{dag} and demonstrate the accuracy and precision of the best-performing model.
-When extended to the contralateral side, this model may form the basis of future guidelines of elective \gls{ctv} definition.
 \end{abstract}
@@
 
-%\cite{poortmans_internal_2020,murthy_prostate-only_2021}. 
-%Microscopic disease can currently only be observed by a pathological examination of the tissue.
-Consequently, clinicians that are presented with cancer patients need to routinely assess the risk of microscopic involvement in parts of the body that appear clinically node negative.
-
 This work concerns itself with \gls{opscc}, where approximately 70-80\% of patients present with lymph node metastases at the time of diagnosis.
-%which commonly spreads lymphatically.
-The only two publicly available datasets on nodal involvement patterns in \gls{opscc}, at the time of writing, show that 80\% of patients where either cN+ or -- if available -- pN+ \cite{ludwig_dataset_2022}.
-Since the exact number and location of lymph nodes varies from patient to patient, the nodes are grouped into anatomically defined \glspl{lnl}.
-These levels are then often irradiated or resected prophylactically due to the risk of harboring occult metastases despite negative findings from imaging modalities.
 In clinical practice, \gls{ctv-n} definition in radiotherapy is based on guidelines \cite{gregoire_ct-based_2003,gregoire_delineation_2014,gregoire_delineation_2018,eisbruch_intensity-modulated_2002,biau_selection_2019,chao_determination_2002,vorwerk_guidelines_2011,ferlito_elective_2009} that are mostly derived from the observed prevalence of involvement in an \gls{lnl} for a given tumor location.
@@
 The limited data did not allow us to quantify the probability of metastases in the rarely involved \gls{lnl} I, V, and VII, nor did the data allow us to verify that the \gls{hmm} is adequate to describe the dependence on lymph node involvement on T-category.
-%Since this publication only includes details about the involvement of the ipsilateral \glspl{lnl} I, II, III, and IV, the models were limited to these nodal levels as well.
-%With the publication of larger and more detailed datasets \cite{ludwig_dataset_2022}, we are thus able to
 In this paper, we extend the previous work \cite{ludwig_hidden_2021} by making the following contributions:
@@
 
-The terms of the form $P \left( \xi_{iv} \mid \left\{ \xi_{jr} \right\}_{r \in \operatorname{pa}(v)} \right)$ implicitly depend on how we parametrize e.g.
+The terms of the form $P \left( \xi_{iv} \mid \left\{ \xi_{jr} \right\}_{r \in \operatorname{pa}(v)} \right)$ implicitly depend on how we parameterize e.g.
 the arcs of \cref{fig:graph_with_obs}.
@@
 Attached to each of these is the corresponding observed \gls{rv} (orange shaded squares).
-Lymphatic flow is depicted in the form of parametrized arrows (red and blue) that represent the probability of spread along the respective arc per time-step.
+Lymphatic flow is depicted in the form of parameterized arrows (red and blue) that represent the probability of spread along the respective arc per time-step.
 Sensitivity and specificity (orange arrows) connect the hidden \glspl{rv} to the diagnosis.}
@@
 
-\ju{Also say that the above equation gives the probability of a hidden state. To get the probability of involvement in a given level one has to marginalize over all corresponding states}
+From this probability for any possible hidden state, we can also compute the probability of, for example, involvement in \gls{lnl} IV.
+To that end, we marginalize over all states $\boldsymbol{\xi}_k$ where $\xi_{k4} = 1$, meaning those states in which \gls{lnl} IV harbors metastases.
+Formally, we can define a marginalization vector $\mathbf{m}$ that is one for every hidden state we want to include in the marginalization and zero elsewhere.
+In the example of the marginalized probability for \gls{lnl} IV involvement, the components would look like this:
+%
+\begin{equation}
+    m_{4k} = \operatorname{id}(\xi_{k4} = 1)
+\end{equation}
+%
+Subsequently, we can compute the marginalization as a dot product:
+%
+\begin{equation}
+    \begin{aligned}
+        P\left(\text{IV} = 1 \mid \mathbf{Z}=\boldsymbol{\zeta}_\ell \right) &= \sum_{k:\xi_{k4} = 1} P\left( \mathbf{X}=\boldsymbol{\xi}_k \mid \mathbf{Z}=\boldsymbol{\zeta}_\ell \right) \\
+        &= \mathbf{m}_4 \cdot P\left( \mathbf{X}=\boldsymbol{\xi}_k \mid \mathbf{Z}=\boldsymbol{\zeta}_\ell \right)
+    \end{aligned}
+\end{equation}
+%
+% \ju{Also say that the above equation gives the probability of a hidden state. To get the probability of involvement in a given level one has to marginalize over all corresponding states}
 
@@
 
-%I wouldn't even describe these minor deviations
-
-%There are only a few involvement patterns for which the model's prediction does not match the prevalence of the data with perfect accuracy.
-But in these instances, e.g.
-the case where \gls{lnl} III is metastatic while \gls{lnl} IV is healthy in an early T-category patient (\cref{fig:bg_prevalences}, third row, green line and histogram), the model slightly overestimates the prevalence.
-An overestimation by a few percent however, does not pose a risk to the model's clinical applicability, since in the context of volume-deescalation, this can be considered cautious.
-
 It is interesting to note that not all involvement patterns become more prevalent with advanced T-category.
@@
 
-\ju{perhaps one could use the same color in each risk figure for a given clinical involvement, e.g. always green for N0}
-\rl{Done}
+% \ju{perhaps one could use the same color in each risk figure for a given clinical involvement, e.g. always green for N0}
 
@@
         \includegraphics[width=\linewidth]{figures/wg_III_risks.png}
-        \caption{TODO: Write caption.}
+        \caption{Histograms over the risk for microscopic involvement in \gls{lnl} III, given that a patient presents as clinically N0 (green), or given that \gls{lnl} II shows clinical involvement (blue).
+The top panel shows these risks for early T-category, the bottom panel for late.}
         \label{fig:wg_III_risks}
@@
         \includegraphics[width=\linewidth]{figures/wg_any_risks.png}
-        \caption{TODO: Shown are the historgams over the predicted risk for involvement in any of the \glspl{lnl} I, IV, V, or VII.
+        \caption{Shown are the histograms over the predicted risk for involvement in any of the \glspl{lnl} I, IV, V, or VII.
 The risk is plotted given a clinical N0 diagnosis (green), macroscopically detected metastases in \gls{lnl} II (blue), and lastly given visible involvement in both \gls{lnl} II and III (orange).
@@
 \begin{itemize}
-    \item[$\bullet$] Irradiate level II for all patients.
-\ju{maybe we should actually show the risk in level II for N0 patients, at least state the risk in the text, are even have the risk histogram for completeness.}
+    \item[$\bullet$] Irradiate level II for all patients.%\ju{maybe we should actually show the risk in level II for N0 patients, at least state the risk in the text, are even have the risk histogram for completeness.}
     \item[$\bullet$] Irradiate level III for most patients. Only for clinically N0, early T-category patients, not irradiating level III could be considered. 
-    \item[$\bullet$] Exclude level IV from the CTV-N for patients with clinically negative level III.
-For advanced T-category patients with involvement of level III, level IV should be irradiated.
-%For early T-category patients with involvement of level III, not irradiating level IV could be considered.
+    \item[$\bullet$] Exclude level IV from the CTV-N for patients with clinically negative level III. For advanced T-category patients with involvement of level III, level IV should be irradiated.
     \item[$\bullet$] Exclude level V from the CTV-N for most patients. Only for patients with extensive involvement of levels II, III, and IV, irradiation of level V can be considered.
@@
 For advanced T-category patients with extensive nodal involvement of levels II, III, and IV, level I should be irradiated (see also the limitations discussed below).
-    \item[$\bullet$] \ju{Level VII}
+    \item[$\bullet$] Exclude level VII in all patients, unless it is clinically involved.
 \end{itemize}
@@
 
+\section{Acknowledgments}
+
+This work was supported by the Swiss Cancer Research Foundation under grant \href{https://www.krebsforschung.ch/unterstuetzen-sie-uns/stiftungen/-dl-/fileadmin/downloads/unterstuetzen-sie-uns/projekte-der-stiftung-krebsforschung-schweiz-2023.pdf}{KFS-5645-08-2022} and by the University Zürich under the Clinical Research Priority Program \href{https://www.crpp-ai-oncology.uzh.ch/en/Projects/Project-5.html}{Artificial Intelligence in Oncological Imaging}.
+
 \bibliography{bib}

### 2023-08-29 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 However, already around $\beta=0.1$ the model based on the winning graph outperforms the base graph model.
-Ultimately, the winning graph's log-evidence is better than the base graph model's by a value of around 3 (see \cref{table:evidence}).}
+Ultimately, the winning graph's log-evidence is better than the base graph model's by a value of around 0.8 (see \cref{table:evidence}).}
     \label{fig:thermo_int}
@@
     \caption{Mean of sampled parameters in percent.}
-    \label{tab:params}
+    \label{table:params}
 \end{table}
@@
 
-{\it Level V:} In \cref{fig:IIandV_prevs,fig:IIIandV_prevs} we compare the base graph's and the winning graph's estimations for prevalences of involvement patterns that include \gls{lnl} V.
-
-, and VII have been plotted as gray, hatched histograms.
+{\it Level V:} In \cref{fig:IIandIIIandV_prevs} we compare the base graph's and the winning graph's estimations for prevalences of involvement patterns that include \gls{lnl} V.
+The base graph's predictions have been plotted as gray, hatched histograms.
 Alongside, we show again the Beta posterior over the prevalence of the respective involvement pattern, given the data, as colored line.
@@
 \begin{figure}
-    \script{comp_IIandV_prevs.py}
+    \script{comp_IIandIIIandV_prevs.py}
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/comp_IIandV_prevs.png}
+        \includegraphics[width=\linewidth]{figures/comp_IIandIIIandV_prevs.png}
         \caption{Beta posterior over the prevalence of the involvement pattern where \gls{lnl} II \emph{and} V show metastases (orange line), as well as the pattern with \gls{lnl} V involved \emph{without} level II (blue line).
@@
 The winning graph slightly improves its fit to the data in the case of early T-category prevalence of \gls{lnl} V without \gls{lnl} II, and in the case of late T-category co-involvement of the levels II and V.
-This likely stems from the additional arc the winning graph has from \gls{lnl} II to V.}
-        \label{fig:IIandV_prevs}
-    \end{centering}
-\end{figure}
-
-\begin{figure}
-    \script{comp_IIIandV_prevs.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/comp_IIIandV_prevs.png}
-        \caption{Observations of prevalences in the data (solid lines) compared to predictions from the base (hatched histograms) and the winning graph model (colored histograms).
+This likely stems from the additional arc the winning graph has from \gls{lnl} II to V.
+Observations of prevalences in the data (solid lines) compared to predictions from the base (hatched histograms) and the winning graph model (colored histograms).
 The investigated involvement patterns here are how often is \gls{lnl} V involved together with \gls{lnl} III (red) and how often is level V involved while \gls{lnl} III is healthy (green).
@@
 This is presumably due to the added connection from \gls{lnl} III to V.}
-        \label{fig:IIIandV_prevs}
+        \label{fig:IIandIIIandV_prevs}
     \end{centering}
@@
 
-In \cref{fig:IandII_prevs,fig:IIandV_prevs,fig:IIIandV_prevs}
+{\it Level I:} In \cref{fig:IandII_prevs}, the same kind of comparisons are plotted for involvement patterns that inlcude metastatic disease in \gls{lnl} I.
+Here, we observe that the model does at times struggle to capture the spread between the prevalence of \gls{lnl} I involvement for early and late T-category, which is more pronounced than in the levels II and III.
 
@@
 Green plots indicate \gls{lnl} I involvement without level II, while orange plots indicate the opposite (\gls{lnl} II without level I).
-The winning graph has an added edge from \gls{lnl} II to I, which improves the prediction of the rare green pattern and reduces the overestimation of the orange pattern in late T-category.
-However, it slightly underestimates the orange pattern in early T-category.}
+The winning graph has an added edge from \gls{lnl} II to I, which improves the prediction of the rare green pattern and minimally reduces the overestimation of the orange pattern in late T-category.}
         \label{fig:IandII_prevs}
@@
 \end{figure*}
+
+{\it Level VII:} The two models' predictions w.r.t.
+the prevalence of involvement patterns that include \gls{lnl} VII are not compared in a dedicated figure, as this level is only connected to the primary tumor in both graphs.
 
@@
 We assume a sensitivity of 0.76 and a specificity of 0.81 for the clinical diagnosis of lymph node metastases, corresponding to CT imaging in \cref{table:sens_spec}.
-\\
 

### 2023-08-29 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\begin{figure}
+\begin{figure*}
     \script{comp_IIandIIIandV_prevs.py}
@@
         \includegraphics[width=\linewidth]{figures/comp_IIandIIIandV_prevs.png}
-        \caption{Beta posterior over the prevalence of the involvement pattern where \gls{lnl} II \emph{and} V show metastases (orange line), as well as the pattern with \gls{lnl} V involved \emph{without} level II (blue line).
+        \caption{{\it Left panel:} Beta posterior over the prevalence of the involvement pattern where \gls{lnl} II \emph{and} V show metastases (orange line), as well as the pattern with \gls{lnl} V involved \emph{without} level II (blue line).
 Both are shown for early (top panel) and late T-category (bottom panel).
@@
 The winning graph slightly improves its fit to the data in the case of early T-category prevalence of \gls{lnl} V without \gls{lnl} II, and in the case of late T-category co-involvement of the levels II and V.
-This likely stems from the additional arc the winning graph has from \gls{lnl} II to V.
-Observations of prevalences in the data (solid lines) compared to predictions from the base (hatched histograms) and the winning graph model (colored histograms).
-The investigated involvement patterns here are how often is \gls{lnl} V involved together with \gls{lnl} III (red) and how often is level V involved while \gls{lnl} III is healthy (green).
-For early T-category cases, the fit of both models to the data is good.
-For late T-category, however, the winning graph shows a better ability to split the two investigated patterns of involvement.
+This may stems from the added arc from \gls{lnl} II to V.
+\\
+        {\it Right panel:} The same plots, but for involvement patterns where \gls{lnl} V is involved together with \gls{lnl} III (red) and where level V is involved while \gls{lnl} III is healthy (green).
+For early T-category cases, the fit of the base graph model to the data is actually better than the winning graph's.
+For late T-category, however, the winning graph shows a better ability to reverse the prevalence of the two investigated patterns of involvement (note that the base graph model predicts the prevalences the wrong way around here).
 This is presumably due to the added connection from \gls{lnl} III to V.}
@@
     \end{centering}
-\end{figure}
+\end{figure*}
 
@@
 the prevalence of involvement patterns that include \gls{lnl} VII are not compared in a dedicated figure, as this level is only connected to the primary tumor in both graphs.
+
+\begin{figure}
+    \script{comp_VII_prevs.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/comp_VII_prevs.png}
+        \caption{TODO: Write caption}
+        \label{fig:VII_prevs}
+    \end{centering}
+\end{figure}
 

### 2023-08-29 src/scripts/comp_IIIandV_prevs.py
merge two prevs and add one about LNL VII

@@ -1,95 +1,0 @@
-"""
-Plot histograms of predicted and posteriors over observed prevalences for scenarios
-that involve the LNLs III and V. Compare these prevalence predictions for the
-base-graph model to the winning-graph model.
-"""
-from pathlib import Path
-
-import matplotlib.pyplot as plt
-from matplotlib.lines import Line2D
-from lyscripts.plot.utils import Histogram, Posterior, draw
-from lyscripts.plot.utils import COLORS as USZ
-from tueplots import figsizes, fontsizes
-
-import paths
-
-
-BASE_GRAPH_INPUT = paths.data / "bg_IIIandV_prevs.hdf5"
-WIN_GRAPH_INPUT = paths.data / "wg_IIIandV_prevs.hdf5"
-OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
-NROWS, NCOLS = 2, 1
-
-
-if __name__ == "__main__":
-    plt.rcParams.update(figsizes.icml2022_half(
-        nrows=NROWS,
-        ncols=NCOLS,
-        height_to_width_ratio=0.45,
-    ))
-    plt.rcParams.update(fontsizes.icml2022())
-    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
-
-    panels = {
-        "early": [],
-        "late": [],
-    }
-
-    for stage in ["early", "late"]:
-        panels[stage] = [
-            Histogram.from_hdf5(
-                filename=WIN_GRAPH_INPUT,
-                dataname=f"VnotIII/{stage}",
-                label="LNL V without III",
-                color=USZ["green"],
-            ),
-            Histogram.from_hdf5(
-                filename=BASE_GRAPH_INPUT,
-                dataname=f"VnotIII/{stage}",
-                label="base graph",
-                histtype="step",
-                color="darkgray",
-                linewidth=2.,
-                hatch=r"////",
-            ),
-            Posterior.from_hdf5(
-                filename=BASE_GRAPH_INPUT,
-                dataname=f"VnotIII/{stage}",
-                color=USZ["green"],
-            ),
-
-            Histogram.from_hdf5(
-                filename=WIN_GRAPH_INPUT,
-                dataname=f"IIIandV/{stage}",
-                label="LNL III and V",
-                color=USZ["red"],
-            ),
-            Histogram.from_hdf5(
[…]

### 2023-08-29 src/scripts/comp_IIandIIIandV_prevs.py

@@ -1,0 +1,164 @@
+"""
+Plot histograms of predicted and posteriors over observed prevalences for scenarios
+that involve the LNLs II, III, and V. Compare these prevalence predictions for the
+base-graph model to the winning-graph model.
+"""
+from pathlib import Path
+
+import matplotlib.pyplot as plt
+from matplotlib.lines import Line2D
+from lyscripts.plot.utils import Histogram, Posterior, draw
+from lyscripts.plot.utils import COLORS as USZ
+from tueplots import figsizes, fontsizes
+
+import paths
+
+
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
+NROWS, NCOLS = 2, 2
+
+
+if __name__ == "__main__":
+    plt.rcParams.update(figsizes.icml2022_full(
+        nrows=NROWS,
+        ncols=NCOLS,
+        height_to_width_ratio=0.45,
+    ))
+    plt.rcParams.update(fontsizes.icml2022())
+    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col")
+
+    # left panels: prevalences of LNLs II and V
+    BASE_GRAPH_INPUT = paths.data / "bg_IIandV_prevs.hdf5"
+    WIN_GRAPH_INPUT = paths.data / "wg_IIandV_prevs.hdf5"
+
+    left_panels = {
+        "early": [],
+        "late": [],
+    }
+
+    for stage in ["early", "late"]:
+        left_panels[stage] = [
+            Histogram.from_hdf5(
+                filename=WIN_GRAPH_INPUT,
+                dataname=f"IIandV/{stage}",
+                label="LNL II and V",
+                color=USZ["orange"],
+            ),
+            Histogram.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"IIandV/{stage}",
+                label="base graph",
+                histtype="step",
+                color="darkgray",
+                linewidth=2.,
+                hatch=r"\\\\",
+            ),
+            Posterior.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"IIandV/{stage}",
+                color=USZ["orange"],
+            ),
+
+            Histogram.from_hdf5(
+                filename=WIN_GRAPH_INPUT,
+                dataname=f"VnotII/{stage}",
+                label="LNL V without II",
+                color=USZ["blue"],
+            ),
[…]

### 2023-08-29 src/scripts/comp_IIandV_prevs.py

@@ -1,95 +1,0 @@
-"""
-Plot histograms of predicted and posteriors over observed prevalences for scenarios
-that involve the LNLs II and V. Compare these prevalence predictions for the
-base-graph model to the winning-graph model.
-"""
-from pathlib import Path
-
-import matplotlib.pyplot as plt
-from matplotlib.lines import Line2D
-from lyscripts.plot.utils import Histogram, Posterior, draw
-from lyscripts.plot.utils import COLORS as USZ
-from tueplots import figsizes, fontsizes
-
-import paths
-
-
-BASE_GRAPH_INPUT = paths.data / "bg_IIandV_prevs.hdf5"
-WIN_GRAPH_INPUT = paths.data / "wg_IIandV_prevs.hdf5"
-OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
-NROWS, NCOLS = 2, 1
-
-
-if __name__ == "__main__":
-    plt.rcParams.update(figsizes.icml2022_half(
-        nrows=NROWS,
-        ncols=NCOLS,
-        height_to_width_ratio=0.45,
-    ))
-    plt.rcParams.update(fontsizes.icml2022())
-    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
-
-    panels = {
-        "early": [],
-        "late": [],
-    }
-
-    for stage in ["early", "late"]:
-        panels[stage] = [
-            Histogram.from_hdf5(
-                filename=WIN_GRAPH_INPUT,
-                dataname=f"VnotII/{stage}",
-                label="LNL V without II",
-                color=USZ["blue"],
-            ),
-            Histogram.from_hdf5(
-                filename=BASE_GRAPH_INPUT,
-                dataname=f"VnotII/{stage}",
-                label="base graph",
-                histtype="step",
-                color="darkgray",
-                linewidth=2.,
-                hatch=r"////",
-            ),
-            Posterior.from_hdf5(
-                filename=BASE_GRAPH_INPUT,
-                dataname=f"VnotII/{stage}",
-                color=USZ["blue"],
-            ),
-
-            Histogram.from_hdf5(
-                filename=WIN_GRAPH_INPUT,
-                dataname=f"IIandV/{stage}",
-                label="LNL II and V",
-                color=USZ["orange"],
-            ),
-            Histogram.from_hdf5(
[…]

### 2023-08-29 src/scripts/comp_VII_prevs.py

@@ -1,0 +1,80 @@
+"""
+Plot histograms of predicted and posteriors over observed prevalences for scenarios
+that involve the LNL VII. Compare these prevalence predictions for the base-graph
+model to the winning-graph model.
+"""
+from pathlib import Path
+
+import matplotlib.pyplot as plt
+from lyscripts.plot.utils import Histogram, Posterior, draw
+from lyscripts.plot.utils import COLORS as USZ
+from tueplots import figsizes, fontsizes
+
+import paths
+
+
+BASE_GRAPH_INPUT = paths.data / "bg_VII_prevs.hdf5"
+WIN_GRAPH_INPUT = paths.data / "wg_VII_prevs.hdf5"
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
+NROWS, NCOLS = 2, 1
+
+
+if __name__ == "__main__":
+    plt.rcParams.update(figsizes.icml2022_half(
+        nrows=NROWS,
+        ncols=NCOLS,
+        height_to_width_ratio=0.45,
+    ))
+    plt.rcParams.update(fontsizes.icml2022())
+    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col")
+
+
+    panels = {
+        "early": [],
+        "late": [],
+    }
+
+    for stage in ["early", "late"]:
+        panels[stage] = [
+            Histogram.from_hdf5(
+                filename=WIN_GRAPH_INPUT,
+                dataname=f"VII/{stage}",
+                label="LNL VII overall",
+                color=USZ["blue"],
+            ),
+            Histogram.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"VII/{stage}",
+                label="base graph",
+                histtype="step",
+                color="darkgray",
+                linewidth=2.,
+                hatch=r"////",
+            ),
+            Posterior.from_hdf5(
+                filename=BASE_GRAPH_INPUT,
+                dataname=f"VII/{stage}",
+                color=USZ["blue"],
+            ),
+        ]
+
+    draw(axes[0], contents=panels["early"], xlims=(0., 10.), hist_kwargs={"nbins": 40})
+    h, l = axes[0].get_legend_handles_labels()
+    handles = [*h[2:4], h[0], *h[4:6], h[1]]
[…]

### 2023-08-29 src/scripts/compile_metrics_table.py

@@ -13,26 +13,27 @@ TEMPLATE_NAME = "results_table.temp"
 OUTPUT = paths.output / "results_table.tex"
 
 
-environment = Environment(
-    loader=FileSystemLoader(paths.scripts)
-)
-table_template = environment.get_template(TEMPLATE_NAME)
-context = {}
+if __name__ == "__main__":
+    environment = Environment(
+        loader=FileSystemLoader(paths.scripts)
+    )
+    table_template = environment.get_template(TEMPLATE_NAME)
+    context = {}
 
 
-with open(BASE_GRAPH_INPUT, mode="r") as bg_file:
-    bg_metrics = json.load(bg_file)
+    with open(BASE_GRAPH_INPUT, mode="r", encoding="utf-8") as bg_file:
+        bg_metrics = json.load(bg_file)
 
-for key,val in bg_metrics.items():
-    context[f"bg_{key}"] = val
-
-
-with open(WIN_GRAPH_INPUT, mode="r") as wg_file:
-    wg_metrics = json.load(wg_file)
-
-for key,val in wg_metrics.items():
-    context[f"wg_{key}"] = val
+    for key,val in bg_metrics.items():
+        context[f"bg_{key}"] = val
 
 
-with open(OUTPUT, mode="w") as table_file:
-    table_file.write(table_template.render(**context))
+    with open(WIN_GRAPH_INPUT, mode="r", encoding="utf-8") as wg_file:
+        wg_metrics = json.load(wg_file)
+
+    for key,val in wg_metrics.items():
+        context[f"wg_{key}"] = val
+
+
+    with open(OUTPUT, mode="w", encoding="utf-8") as table_file:
+        table_file.write(table_template.render(**context))

### 2023-08-29 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-{\it Level V:} In \cref{fig:IIandIIIandV_prevs} we compare the base graph's and the winning graph's estimations for prevalences of involvement patterns that include \gls{lnl} V.
-The base graph's predictions have been plotted as gray, hatched histograms.
-Alongside, we show again the Beta posterior over the prevalence of the respective involvement pattern, given the data, as colored line.
-The prediction by the winning graph are shown as colored histograms.
-Every scenario is shown for early and late T-category.
-This allows us to compare how the winning graph improves its fit to the data in these particular scenarios around low-risk \glspl{lnl}.
-How frequently each of the investigated lymphatic metastatic patterns were observed in the data is additionally tabulated in \cref{table:data_prevalence}, again both for early and for late T-category cases.
-
 \begin{figure*}
@@
 
+{\it Level V:} In \cref{fig:IIandIIIandV_prevs} we compare the base graph's and the winning graph's estimations for prevalences of involvement patterns that include \gls{lnl} V.
+The base graph's predictions have been plotted as gray, hatched histograms.
+Alongside, we show again the Beta posterior over the prevalence of the respective involvement pattern, given the data, as colored line.
+The prediction by the winning graph are shown as colored histograms.
+Every scenario is shown for early and late T-category.
+This allows us to compare how the winning graph improves its fit to the data in these particular scenarios around low-risk \glspl{lnl}.
+How frequently each of the investigated lymphatic metastatic patterns were observed in the data is additionally tabulated in \cref{table:data_prevalence}, again both for early and for late T-category cases.
+
 {\it Level I:} In \cref{fig:IandII_prevs}, the same kind of comparisons are plotted for involvement patterns that inlcude metastatic disease in \gls{lnl} I.
@@
 {\it Level VII:} The two models' predictions w.r.t.
-the prevalence of involvement patterns that include \gls{lnl} VII are not compared in a dedicated figure, as this level is only connected to the primary tumor in both graphs.
+the prevalence of involvement patterns that include \gls{lnl} VII are compared in \cref{fig:VII_prevs}.
+Here, the base and winning graph yield almost the same predictions, as no new edge connects from or to \gls{lnl} VII in the latter model.
+Notably, the prevalence of early T-category level VII involvement is slightly overestimated, while for late T-category, the fit is very good.
 
@@
         \includegraphics[width=\linewidth]{figures/comp_VII_prevs.png}
-        \caption{TODO: Write caption}
+        \caption{Predictions for the prevalence of \gls{lnl} VII involvement based on the base graph (hatched) and the winning graph (blue histogram) plotted agains the beta posterior over the prevalence in the data (solid line).}
         \label{fig:VII_prevs}
@@
 
-{\it Level II:} As can be seen in \cref{tab:params} the model infers the spread from the tumor to \gls{lnl} II to be the most probable transition at any given time step.
-As a consequence, even for an early T-category patient that presents with a clinical N0 neck, our model predicts a \variable{output/wg_II_risks.tex} risk for microscopic metastases in \gls{lnl} II.
-
 \begin{figure}
@@
 \end{figure}
+
+{\it Level II:} As can be seen in \cref{tab:params} the model infers the spread from the tumor to \gls{lnl} II to be the most probable transition at any given time step.
+As a consequence, even for an early T-category patient that presents with a clinical N0 neck, our model predicts a \variable{output/wg_II_risks.tex} risk for microscopic metastases in \gls{lnl} II.
 
@@
 
+{\it Sensitivity and Specificity:} Estimating the risk of occult metastases depends on the assumed parameter values for sensitivity and specificity of clinical detection of lymph node metastases.
+For this work, we adopted literature values for sensitivity and specificity.
+However, different authors have estimated these values using different criteria and different methods.
+Consequently, these values need to be considered with caution.
+\cref{fig:wg_sens_spec_risks} illustrates for one example how the risk of occult disease depends on sensitivity and specificity.
+Here, we consider the risk in level IV in patients with clinically detected metastases in levels II and III.
+For our default parameters of 81\% specificity and 76 \% sensitivity, the risk is 5\%, but it increases to x\% for a a sensitivity of 66\%.
+Future work may aim at developing new methods to model the difference between pathological and clinical lymph node involvement based on surgically treated patients in whom both is reported.
+
 \begin{figure}
@@
 
-{\it Sensitivity and Specificity:} Estimating the risk of occult metastases depends on the assumed parameter values for sensitivity and specificity of clinical detection of lymph node metastases.
-For this work, we adopted literature values for sensitivity and specificity.
-However, different authors have estimated these values using different criteria and different methods.
-Consequently, these values need to be considered with caution.
-\cref{fig:wg_sens_spec_risks} illustrates for one example how the risk of occult disease depends on sensitivity and specificity.
-Here, we consider the risk in level IV in patients with clinically detected metastases in levels II and III.
-For our default parameters of 81\% specificity and 76 \% sensitivity, the risk is 5\%, but it increases to x\% for a a sensitivity of 66\%.
-Future work may aim at developing new methods to model the difference between pathological and clinical lymph node involvement based on surgically treated patients in whom both is reported.
-
 \section{Acknowledgments}

### 2023-08-29 src/scripts/comp_VII_prevs.py
[showyourwork] overleaf sync

@@ -16,17 +16,17 @@ import paths
 BASE_GRAPH_INPUT = paths.data / "bg_VII_prevs.hdf5"
 WIN_GRAPH_INPUT = paths.data / "wg_VII_prevs.hdf5"
 OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
-NROWS, NCOLS = 2, 1
+NROWS, NCOLS = 1, 2
 
 
 if __name__ == "__main__":
     plt.rcParams.update(figsizes.icml2022_half(
         nrows=NROWS,
         ncols=NCOLS,
-        height_to_width_ratio=0.45,
+        height_to_width_ratio=0.3,
     ))
     plt.rcParams.update(fontsizes.icml2022())
-    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col")
+    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS)
 
 
     panels = {
@@ -58,21 +58,21 @@ if __name__ == "__main__":
             ),
         ]
 
-    draw(axes[0], contents=panels["early"], xlims=(0., 10.), hist_kwargs={"nbins": 40})
+    draw(axes[0], contents=panels["early"], xlims=(2., 10.), hist_kwargs={"nbins": 40})
     h, l = axes[0].get_legend_handles_labels()
     handles = [*h[2:4], h[0], *h[4:6], h[1]]
     labels = [*l[2:4], l[0], *l[4:6], l[1]]
-    axes[0].set_ylim(bottom=0., top=1.5)
-    axes[0].set_ylabel("early T-cat.", fontweight="bold")
+    axes[0].set_ylim(bottom=0.)
+    axes[0].set_title("early T-cat.", fontweight="bold")
     axes[0].set_yticks([])
     axes[0].legend(handles=handles, labels=labels)
 
-    draw(axes[1], contents=panels["late"], xlims=(0., 10.), hist_kwargs={"nbins": 40})
+    draw(axes[1], contents=panels["late"], xlims=(4., 12.), hist_kwargs={"nbins": 40})
     h, l = axes[1].get_legend_handles_labels()
     handles = h[:2]
     labels = l[:2]
-    axes[1].set_ylim(bottom=0., top=1.2)
-    axes[1].set_ylabel("late T-cat.", fontweight="bold")
+    axes[1].set_ylim(bottom=0.)
+    axes[1].set_title("late T-cat.", fontweight="bold")
[…]

### 2023-08-29 src/scripts/comp_VII_prevs.py
fine tune LNL VII plot

@@ -23,10 +23,10 @@ if __name__ == "__main__":
     plt.rcParams.update(figsizes.icml2022_half(
         nrows=NROWS,
         ncols=NCOLS,
-        height_to_width_ratio=0.3,
+        height_to_width_ratio=1.3,
     ))
     plt.rcParams.update(fontsizes.icml2022())
-    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS)
+    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharey="row")
 
 
     panels = {
@@ -58,20 +58,21 @@ if __name__ == "__main__":
             ),
         ]
 
-    draw(axes[0], contents=panels["early"], xlims=(2., 10.), hist_kwargs={"nbins": 40})
+    draw(axes[0], contents=panels["early"], xlims=(1., 9.), hist_kwargs={"nbins": 20})
     h, l = axes[0].get_legend_handles_labels()
     handles = [*h[2:4], h[0], *h[4:6], h[1]]
     labels = [*l[2:4], l[0], *l[4:6], l[1]]
     axes[0].set_ylim(bottom=0.)
     axes[0].set_title("early T-cat.", fontweight="bold")
     axes[0].set_yticks([])
+    axes[0].set_xlabel("prevalence [%]")
     axes[0].legend(handles=handles, labels=labels)
 
-    draw(axes[1], contents=panels["late"], xlims=(4., 12.), hist_kwargs={"nbins": 40})
+    draw(axes[1], contents=panels["late"], xlims=(4., 12.), hist_kwargs={"nbins": 20})
     h, l = axes[1].get_legend_handles_labels()
-    handles = h[:2]
-    labels = l[:2]
-    axes[1].set_ylim(bottom=0.)
+    handles = h[2:]
+    labels = l[2:]
+    axes[1].set_ylim(bottom=0., top=0.7)
     axes[1].set_title("late T-cat.", fontweight="bold")
     axes[1].set_yticks([])
     axes[1].set_xlabel("prevalence [%]")

### 2023-08-29 src/tex/ms.tex
[showyourwork] overleaf sync

@@
         \includegraphics[width=\linewidth]{figures/comp_VII_prevs.png}
-        \caption{Predictions for the prevalence of \gls{lnl} VII involvement based on the base graph (hatched) and the winning graph (blue histogram) plotted agains the beta posterior over the prevalence in the data (solid line).}
+        \caption{Predictions for the prevalence of \gls{lnl} VII involvement based on the base graph (hatched) and the winning graph (blue histogram) plotted against the beta posterior over the prevalence in the data (solid line).
+This is shown for early T-category (left panel) and late T-category (right panel).}
         \label{fig:VII_prevs}
@@
 \label{subsec:results:risk_prediction}
+
 In this section, the model corresponding to the winning graph is applied to estimating the risk of occult metastases in clinically negative \glspl{lnl}.
@@
 
-{\it Level II:} As can be seen in \cref{tab:params} the model infers the spread from the tumor to \gls{lnl} II to be the most probable transition at any given time step.
+{\it Level II:} As can be seen in \cref{table:params} the model infers the spread from the tumor to \gls{lnl} II to be the most probable transition at any given time step.
 As a consequence, even for an early T-category patient that presents with a clinical N0 neck, our model predicts a \variable{output/wg_II_risks.tex} risk for microscopic metastases in \gls{lnl} II.
@@
 \label{subsec:disc:limitation}
+
 {\it T-stage dependence of level I involvement:} As shown in \cref{fig:bg_prevalences}, the model describes the involvement of \glspl{lnl} II, III, and IV depending on T-stage very well despite having only a single parameter related to T-stage.

### 2023-09-14 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{hyperref}
-\AtEndPreamble{\usepackage{cleveref}}
+\usepackage{cleveref}
 
@@
 
-Lastly, we also compared how the model's performance improves with an added connection from the most commonly involved level II to one of the most rarely metastatic \glspl{lnl} VII.
+To determine the optimal graph describing the data at hand best, we considered six models, each with one of the green candidate arcs of \cref{fig:extended_graph} added to the base graph.
+Every one of these models was evaluated as described below in \cref{subsec:complete_model:comparison}.
 
@@
 However, already around $\beta=0.1$ the model based on the winning graph outperforms the base graph model.
-Ultimately, the winning graph's log-evidence is better than the base graph model's by a value of around 0.8 (see \cref{table:evidence}).}
+Ultimately, the winning graph's log-evidence is better than the base graph model's by a value of around \variable{output/evidence_diff.txt} (see \cref{table:evidence}).}
     \label{fig:thermo_int}
@@
 Without derivation or insight, we would like to mention that the model evidence naturally balances a model's accuracy against its complexity.
-The value of $\ln E$ will generally be larger (i.e., less negative) if a model fits the data better than another while being similarily complex.
+The value of $\ln E$ will generally be larger (i.e., less negative) if a model fits the data better than another while being similarly complex.
 On the other hand, if e.g.
@@
 The approximation is good, when the posterior distribution over the parameters $p\left( \boldsymbol{\theta} \mid \boldsymbol{\mathcal{D}} \right)$ is single-modal and falls quickly to zero from the maximum.
-Also, the number of datapoints $N$ needs to be much larger than the number of parameters $k$.
+Also, the number of data points $N$ needs to be much larger than the number of parameters $k$.
 We will see that for the models we consider here, the \gls{bic} is generally a good approximation and the conclusions drawn from comparing models using this metric can be reproduced reliably using the true model evidence computed with \gls{ti}.

### 2023-09-14 src/scripts/evidence_diff.py

@@ -1,0 +1,20 @@
+"""
+Compute the difference between the winning and the base graph model's evidence and
+write it to a file.
+"""
+import json
+
+import paths
+
+
+if __name__ == "__main__":
+    with open(paths.data / "wg_metrics.json") as f:
+        winning = json.load(f)
+
+    with open(paths.data / "bg_metrics.json") as f:
+        base = json.load(f)
+
+    diff = winning["evidence"] - base["evidence"]
+
+    with open(paths.output / "evidence_diff.txt", "w") as f:
+        f.write(f"{diff:.2f}")

### 2023-10-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \author{Vincent Grégoire}
-\author{Roland Giger}
+%\author{Roland Giger}
+\author{... TBD ...}
 \author{Panagiotis Balermpas}
@@
 To determine the optimal graph describing the data at hand best, we considered six models, each with one of the green candidate arcs of \cref{fig:extended_graph} added to the base graph.
-Every one of these models was evaluated as described below in \cref{subsec:complete_model:comparison}.
+Every one of these models was evaluated by computing its evidence via thermodynamic integration as described below in \cref{subsec:complete_model:comparison}.
+These evidences are shown in table XXX.
+The X most promising added edges were then considered in a combined model which was ultimately designated to be the "winning graph" which is in great detail compared to the base graph in the results (\cref{sec:results}).
+All considered graphs' performances are visualized in figure XXX.
 
@@
 
+\rl{\textbf{NOTE:} This subsection will be updated based on the data that will be finally included.}
+
 \begin{table}
@@
 
-The dataset $\boldsymbol{\mathcal{D}}$ that we used for inference is comprised of the detailed reports on lymph node involvement patterns in \gls{opscc} patients treated at three different institutions in France and Switzerland: The \gls{clb} in Lyon (France), the \gls{isb}, and the \gls{usz} (both in Switzerland).
+%The dataset $\boldsymbol{\mathcal{D}}$ that we used for inference is comprised of the detailed reports on lymph node involvement patterns in \gls{opscc} patients treated at three different institutions in France and Switzerland: The \gls{clb} in Lyon (France), the \gls{isb}, and the \gls{usz} (both in Switzerland).
 We have previously published the patterns of nodal involvement for the \gls{usz} cohort (287 patients) \cite{ludwig_dataset_2022} and described its characteristics in detail \cite{ludwig_detailed_2022}.
@@
 
+The dataset $\boldsymbol{\mathcal{D}}$ that we used for inference is comprised of the detailed reports on lymph node involvement patterns in \gls{opscc} patients treated at two different institutions in France and Switzerland: The \gls{clb} in Lyon (France) and the \gls{usz} in Zürich (Switzerland).
+We have previously published the patterns of nodal involvement for the \gls{usz} cohort (287 patients) \cite{ludwig_dataset_2022} and described its characteristics in detail \cite{ludwig_detailed_2022}.
+The \gls{clb} data (263 patients) underlies a publication on \gls{hpv} status in \gls{opscc} \cite{bauwens_prevalence_2021}.
+\rl{\textbf{NOTE:} For the ISB and second CLB dataset, we can cite the data in brief paper as under review}.
+
 In total, the dataset contains \variable{output/num_patients.txt} patients with newly diagnosed \gls{opscc}.
 It includes patients treated with definitive (chemo)radiotherapy, adjuvant (chemo)radiotherapy following neck dissection, or neck dissection alone.
-Pathologically assessed \gls{lnl} involvement was available for 393 surgically treated patients, while for the remainder the nodal involvement was assessed based on available diagnostic modalities (FDG-PET-CT, CT, MRI, FNA).
+Pathologically assessed \gls{lnl} involvement was available for 263 surgically treated patients, while for the remainder the nodal involvement was assessed based on available diagnostic modalities (FDG-PET-CT, CT, MRI, FNA).
 If multiple modalities were used to diagnose a patient's lymph node involvement, the available modalities were combined into a consensus decision.
@@
 
+\subsection{Graph Candidates}
+\label{subsec:results:graph_candidates}
+
+\rl{\textbf{NOTE:} This subsection is still unfinished an will be completed when the respective computations are ready. Qualitatively, we don't expect any major changes though.}
+
+The model evidence of the six models that were based on graphs with one additional arc compared to the base model's is plotted in \cref{table:evidence}.
+A visual ranking is provided in figure XXX.
+Evidently, adding a connection from \gls{lnl} II to I is strongly/substantially supported given this dataset.
+Similarly, there is strong/substantial evidence for introducing an arc from \gls{lnl} II to V.
+All other investigated additions lead to improvements that are at best barely worth a mention or do not justify the additional complexity by yielding a higher accuracy at all.
+
+Based on this table, we determined the candidate for the optimal graph to have both the new arc from level II to I, as well as arcs from \gls{lnl} II and \gls{lnl} III to level V.
+To ensure the evidence gain of these edges added individually carries over into one combined model, we performed another thermodynamic integration.
+
 \subsection{The Optimal Extended Graph}
@@
 \end{table}
-
-\Cref{table:evidence} reports the model evidence for all extended graphs that were investigated.
-The results is additionally visualized in Figure xxx.
-The winning graph adds arcs from level II to I, from level II to V, and from level III to level V to the base graph.
-Below, we further analyze why these connections improve the model's ability to describe the involvement patterns seen in the data.
 

### 2023-11-29 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Therefore, we also consider the flipped arc from \gls{lnl} II to I and investigate if it helps to describe the correlations between the involvement of levels I and II.
+
+\begin{figure}
+    \centering
+    \includegraphics{figures/ranking.pdf}
+    \caption{Visual ranking of the investigated graphs w.r.t.
+their model evidence, computed via thermodynamic integration.
+Missing are only the two graphs where the arc from \gls{lnl} III to IV was removed.
+Their respective model evidence fell below $-1528$ and did thus not fit on this figure.
+In the bottom left corner, we provide a visual reference in analogy to \cref{table:bayes_factor}: E.g., any difference in the model evidence shorter than the first of the three rulers indicates that the improvement is "barely worth a mention".}
+    \label{fig:ranking}
+\end{figure}
 

### 2023-11-29 src/scripts/wg_means.py
update references to new base & win graph

@@ -16,11 +16,10 @@ NAMES = [
     r"b_\text{IV}",
     r"b_\text{V}",
     r"b_\text{VII}",
-    r"t_{\text{II} \rightarrow {I}}",
+    r"t_{\text{I} \rightarrow {II}}",
     r"t_{\text{II} \rightarrow {III}}",
-    r"t_{\text{II} \rightarrow {V}}",
     r"t_{\text{III} \rightarrow {IV}}",
-    r"t_{\text{III} \rightarrow {V}}",
+    r"t_{\text{IV} \rightarrow {V}}",
     r"p_\text{late}",
 ]
 

### 2023-11-29 src/scripts/wg_sens_spec_risks.py

@@ -21,10 +21,10 @@ SAMPLES_PATH = paths.data / "wg_samples.hdf5"
 OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
 WINNING_GRAPH = {
     ("tumor", "primary"): ["I", "II", "III", "IV", "V", "VII"],
-    ("lnl", "I"): [],
-    ("lnl", "II"): ["I", "III", "V"],
-    ("lnl", "III"): ["IV", "V"],
-    ("lnl", "IV"): [],
+    ("lnl", "I"): ["II"],
+    ("lnl", "II"): ["III"],
+    ("lnl", "III"): ["IV"],
+    ("lnl", "IV"): ["V"],
     ("lnl", "V"): [],
     ("lnl", "VII"): [],
 }

### 2023-11-30 src/scripts/evidence_diff.py
try to fix missing evidence diff

@@ -8,13 +8,13 @@ import paths
 
 
 if __name__ == "__main__":
-    with open(paths.data / "wg_metrics.json") as f:
+    with open(paths.data / "wg_metrics.json", encoding="utf-8") as f:
         winning = json.load(f)
 
-    with open(paths.data / "bg_metrics.json") as f:
+    with open(paths.data / "bg_metrics.json", encoding="utf-8") as f:
         base = json.load(f)
 
     diff = winning["evidence"] - base["evidence"]
 
-    with open(paths.output / "evidence_diff.txt", "w") as f:
+    with open(paths.output / "evidence_diff.txt", "w", encoding="utf-8") as f:
         f.write(f"{diff:.2f}")

### 2023-11-30 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\begin{figure}
+\begin{figure*}
     \centering
-    \includegraphics{figures/ranking.pdf}
+    \includegraphics[width=\linewidth]{figures/ranking.pdf}
     \caption{Visual ranking of the investigated graphs w.r.t.
@@
     \label{fig:ranking}
-\end{figure}
+\end{figure*}
 
@@
 However, already around $\beta=0.1$ the model based on the winning graph outperforms the base graph model.
-Ultimately, the winning graph's log-evidence is better than the base graph model's by a value of around \variable{output/evidence_diff.txt} (see \cref{table:evidence}).}
+Ultimately, the winning graph's log-evidence is better than the base graph model's by a value of around \variable{output/test.txt} (see \cref{table:evidence}).}
     \label{fig:thermo_int}

### 2023-11-30 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 However, already around $\beta=0.1$ the model based on the winning graph outperforms the base graph model.
-Ultimately, the winning graph's log-evidence is better than the base graph model's by a value of around \variable{output/test.txt} (see \cref{table:evidence}).}
+Ultimately, the winning graph's log-evidence is better than the base graph model's by a value of around \variable{output/evidence_diff.tex} (see \cref{table:evidence}).}
     \label{fig:thermo_int}

### 2023-11-30 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2023-11-29 src/tex/ms.tex
@@
 \author{Roman Ludwig}
+\author{Adrian Schubert}
+\author{Dorothea Barbatei}
+\author{Lauence Bauwens}
 \author{Jean-Marc Hoffmann}
+\author{Sandrine Werlen}
+\author{Olgun Elicin}
+\author{Matthias Dettmer}
+\author{Philippe Zrounba}
 \author{Bertrand Pouymayou}
-\author{Lauence Bauwens}
+\author{Panagiotis Balermpas}
 \author{Vincent Grégoire}
-%\author{Roland Giger}
-\author{... TBD ...}
-\author{Panagiotis Balermpas}
+\author{Roland Giger}
 \author{Jan Unkelbach}
@@
     Results: The graph maximizing the model evidence contains edges for lymphatic spread from \gls{lnl} II to I, III and V, and from \gls{lnl} III to IV and V, and is thus consistent with commonly assumed lymphatic drainage in the neck.
-Application of the model suggests that the risk of occult disease in level IV is below 5\% if level III is clinically negative, and that the risk of occult disease in level V is below 5\% except for late T-category patients with clinical involvement of levels II, III, and IV.
+Application of the model suggests that the risk of occult disease in level IV is below 5\% if level III is clinically negative, and that the risk of occult disease in level V is below 5\% except for advanced T-category (T3 and T4) patients with clinical involvement of levels II, III, and IV.
     Conclusion: We provide a statistical model of lymph node involvement in \gls{opscc} patients trained on a multi-institutional dataset, which may guide the design of clinical trials on volume-deescalated treatment of \gls{opscc} and eventually contribute to further personalized guidelines on elective nodal treatment.
@@
 
-The \gls{dag} shown in \cref{fig:graph_with_obs} includes the \glspl{lnl} II, III, and IV, which represent the most relevant lymph node levels for oropharyngeal SCC.
+The \gls{dag} shown in \cref{fig:graph_with_obs} includes the \glspl{lnl} II, III, and IV, which represent the most relevant lymph node levels for \gls{opscc}.
 It includes the arcs from II to III and from III to IV, representing the main direction of lymphatic drainage, which is well motivated anatomically and by the data on lymph node involvement.
 Previous publications  \cite{pouymayou_bayesian_2019,ludwig_hidden_2021} focused on these levels because they relied on a limited reconstructed dataset of \gls{opscc} patients \cite{sanguineti_defining_2009}.
-Now, with the datasets available for this work, we can extend the graph to include random variables for all \glspl{lnl} that are relevant for oropharyngeal SCC: I, II, III, IV, V, and VII.
+Now, with the datasets available for this work, we can extend the graph to include random variables for all \glspl{lnl} that are relevant for \gls{opscc}: I, II, III, IV, V, and VII.
 The main question to answer is: which arcs between \glspl{lnl} are needed to accurately model the data on lymph node involvement without increasing the model's complexity unnecessarily.

### 2023-11-30 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     \caption{The expected log-accuracy under the power posterior plotted against the value of the inverse temperature $\beta$.
-These values were computed during a \gls{ti} run for the base graph model (red line), as well as the winning graph model (green line).
+These values were computed during a \gls{ti} run for the loosing graph model (red line), as well as the winning graph model (green line).
 Note that the scaling of the x-axis is chosen such that the 64 points on the temperature ladder appear evenly spaced.
@@
 The area under these two curves yields an approximation of the log-evidence of the respective model.
-In this plot one can see that due to the simplicity of the base graph model its accuracy increases earlier and faster that the winning graph model's.
-However, already around $\beta=0.1$ the model based on the winning graph outperforms the base graph model.
-Ultimately, the winning graph's log-evidence is better than the base graph model's by a value of around \variable{output/evidence_diff.txt} (see \cref{table:evidence}).}
+This plot one indicates that due to the simplicity of the loosing graph model its accuracy increases earlier and faster that the winning graph model's.
+However, already before $\beta=0.031$ the model based on the winning graph outperforms the loosing graph model.
+Ultimately, the winning graph's log-evidence is better than the loosing graph model's by a value of around \variable{output/evidence_diff.txt} (see \cref{table:evidence}).}
     \label{fig:thermo_int}

### 2023-11-30 src/scripts/evidence_diff.py
plot evidence diff between win & loose graph

@@ -1,5 +1,5 @@
 """
-Compute the difference between the winning and the base graph model's evidence and
+Compute the difference between the winning and the loosing graph model's evidence and
 write it to a file.
 """
 import json
@@ -11,10 +11,10 @@ if __name__ == "__main__":
     with open(paths.data / "wg_metrics.json", encoding="utf-8") as f:
         winning = json.load(f)
 
-    with open(paths.data / "bg_metrics.json", encoding="utf-8") as f:
-        base = json.load(f)
+    with open(paths.data / "lg_metrics.json", encoding="utf-8") as f:
+        loose = json.load(f)
 
-    diff = winning["evidence"] - base["evidence"]
+    diff = winning["evidence"] - loose["evidence"]
 
     with open(paths.output / "evidence_diff.txt", "w", encoding="utf-8") as f:
         f.write(f"{diff:.2f}")

### 2023-11-30 src/scripts/thermo_int.py

@@ -1,5 +1,5 @@
 """
-Plot how the accuracies evolved for the base vs the winning graph.
+Plot how the accuracies evolved for the lose vs the winning graph.
 """
 from pathlib import Path
 
@@ -14,7 +14,7 @@ import paths
 
 
 ACCURACY_PATHS = {
-    "base": paths.data / "base_accuracies.csv",
+    "lose": paths.data / "loose_accuracies.csv",
     "win" : paths.data / "win_accuracies.csv",
 }
 OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
@@ -28,9 +28,9 @@ if __name__ == "__main__":
     fig, axes = plt.subplots()
 
     accuracies = pd.read_csv(
-        ACCURACY_PATHS["base"]
+        ACCURACY_PATHS["lose"]
     ).rename(
-        columns={"accuracy": "base graph"}
+        columns={"accuracy": "lose graph"}
     ).join(
         pd.read_csv(
             ACCURACY_PATHS["win"]
@@ -43,16 +43,16 @@ if __name__ == "__main__":
     )
 
     errors = accuracies.copy().drop(
-        columns=["base graph", "win graph"]
+        columns=["lose graph", "win graph"]
     ).rename(
-        columns={"std_": "base graph", "std": "win graph"}
+        columns={"std_": "lose graph", "std": "win graph"}
     )
     errors["β"] = errors["β"] ** (1./POWER)
     accuracies = accuracies.drop(columns=["std_", "std"])
     accuracies["β"] = accuracies["β"] ** (1./POWER)
 
     accuracies.plot(
-        x="β", y=["base graph", "win graph"], ax=axes,
+        x="β", y=["lose graph", "win graph"], ax=axes,
         color=[USZ["red"], USZ["green"]],
     )
 

### 2023-11-30 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{figures/ranking.pdf}
-    \caption{Visual ranking of the investigated graphs w.r.t.
-their model evidence, computed via thermodynamic integration.
-Missing are only the two graphs where the arc from \gls{lnl} III to IV was removed.
-Their respective model evidence fell below $-1528$ and did thus not fit on this figure.
-In the bottom left corner, we provide a visual reference in analogy to \cref{table:bayes_factor}: E.g., any difference in the model evidence shorter than the first of the three rulers indicates that the improvement is "barely worth a mention".}
-    \label{fig:ranking}
-\end{figure*}
-
 Anatomically, the \emph{posterior accessory pathway} that drains \gls{lnl} II through \gls{lnl} V motivates investigating and arc from $X_2$ into $X_5$ \cite{lengele_anatomical_2007}.
@@
 Every one of these models was evaluated by computing its evidence via thermodynamic integration as described below in \cref{subsec:complete_model:comparison}.
-These evidences are shown in table XXX.
-The X most promising added edges were then considered in a combined model which was ultimately designated to be the "winning graph" which is in great detail compared to the base graph in the results (\cref{sec:results}).
-All considered graphs' performances are visualized in figure XXX.
+These evidences are shown, among others, in \cref{table:evidence}.
+Promising combination with two added edges were then considered in a combined model which was ultimately designated to be the "winning graph" which is in great detail compared to the base graph in the results (\cref{sec:results}).
+All considered graphs' performances are visualized in \cref{fig:ranking}.
 
@@
 
-\rl{\textbf{NOTE:} This subsection will be updated based on the data that will be finally included.}
-
 \begin{table}
@@
 
-%The dataset $\boldsymbol{\mathcal{D}}$ that we used for inference is comprised of the detailed reports on lymph node involvement patterns in \gls{opscc} patients treated at three different institutions in France and Switzerland: The \gls{clb} in Lyon (France), the \gls{isb}, and the \gls{usz} (both in Switzerland).
-We have previously published the patterns of nodal involvement for the \gls{usz} cohort (287 patients) \cite{ludwig_dataset_2022} and described its characteristics in detail \cite{ludwig_detailed_2022}.
-A large part of the \gls{clb} data (263 patients) underlies a publication on \gls{hpv} status in \gls{opscc} \cite{bauwens_prevalence_2021}.
-The data from \gls{clb} and \gls{isb} are described in details in a data set publication [lets cite it as under review and attach the submitted draft of the DiB].
+The dataset $\boldsymbol{\mathcal{D}}$ that we used for inference is comprised of the detailed reports on lymph node involvement patterns in \gls{opscc} patients treated at three different institutions in France and Switzerland: The \gls{clb} in Lyon (France), the \gls{isb} in Bern (Switzerland), and the \gls{usz} in Zürich (Switzerland).
+We have previously published the patterns of nodal involvement for the \gls{usz} cohort (287 patients) \cite{ludwig_dataset_2022,ludwig_lydata_2023} and described its characteristics in detail \cite{ludwig_detailed_2022}.
+The first \gls{clb} dataset (263 patients) \cite{ludwig_lydata_2022} underlies a publication on \gls{hpv} status in \gls{opscc} \cite{bauwens_prevalence_2021} and is made available in a separate ``Data in Brief'' article alongside the second dataset from France \cite{ludwig_lydata_2023-2} and the lymphatic progression patterns from the \gls{isb} \cite{ludwig_lydata_2023-1,roman_ludwig_multi-centric_nodate}.
 
-The dataset $\boldsymbol{\mathcal{D}}$ that we used for inference is comprised of the detailed reports on lymph node involvement patterns in \gls{opscc} patients treated at two different institutions in France and Switzerland: The \gls{clb} in Lyon (France) and the \gls{usz} in Zürich (Switzerland).
-We have previously published the patterns of nodal involvement for the \gls{usz} cohort (287 patients) \cite{ludwig_dataset_2022} and described its characteristics in detail \cite{ludwig_detailed_2022}.
-The \gls{clb} data (263 patients) underlies a publication on \gls{hpv} status in \gls{opscc} \cite{bauwens_prevalence_2021}.
-\rl{\textbf{NOTE:} For the ISB and second CLB dataset, we can cite the data in brief paper as under review}.
-
-In total, the dataset contains \variable{output/num_patients.txt} patients with newly diagnosed \gls{opscc}.
+In total, the dataset contains \variable{output/num_patients.txt}patients with newly diagnosed \gls{opscc}.
 It includes patients treated with definitive (chemo)radiotherapy, adjuvant (chemo)radiotherapy following neck dissection, or neck dissection alone.
@@
 
-\rl{\textbf{NOTE:} This subsection is still unfinished an will be completed when the respective computations are ready. Qualitatively, we don't expect any major changes though.}
+% \rl{\textbf{NOTE:} This subsection is still unfinished an will be completed when the respective computations are ready. Qualitatively, we don't expect any major changes though.}
 
-The model evidence of the six models that were based on graphs with one additional arc compared to the base model's is plotted in \cref{table:evidence}.
-A visual ranking is provided in figure XXX.
-Evidently, adding a connection from \gls{lnl} II to I is strongly/substantially supported given this dataset.
-Similarly, there is strong/substantial evidence for introducing an arc from \gls{lnl} II to V.
-All other investigated additions lead to improvements that are at best barely worth a mention or do not justify the additional complexity by yielding a higher accuracy at all.
+The model evidences of the six models that were based on graphs with one additional arc compared to the base model's are plotted in \cref{table:evidence}.
+A visual ranking is provided in \cref{fig:ranking}.
+Evidently, adding a connection from \gls{lnl} I to II is strongly supported given this dataset and slightly beats the reverse connection from \gls{lnl} II to I.
+Similarly, there is strong evidence for introducing an arc from \gls{lnl} IV to V.
+Furthermore, there is substantial evidence for an arc from \gls{lnl} III to V.
+All other investigated additions lead to improvements that are at best barely worth a mention or do not justify the additional complexity at all, indicated by a lower evidence.
 
-Based on this table, we determined the candidate for the optimal graph to have both the new arc from level II to I, as well as arcs from \gls{lnl} II and \gls{lnl} III to level V.
-To ensure the evidence gain of these edges added individually carries over into one combined model, we performed another thermodynamic integration.
+Based on this table, we determined the final candidates for the optimal graph should be combinations of the graphs with the added connections from level I to II, III to V, and IV to V.
+For the three resulting graphs we again computed the model evidence and added it to the table in \cref{table:evidence}.
+The best performing graph with decisive evidence over the base graph turned out to be the one which combines the arcs from \gls{lnl} I to II and IV to V.
+Interestingly, the evidence gain of this "winning graph" is roughly the sum of the gains seen in the two candidates where only one of these connections was added, respectively.
+This indicates that the two additional parameters are largely independent of each other and manage to describe different aspects of the data.
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{figures/ranking.pdf}
+    \caption{Visual ranking of the investigated graphs w.r.t.
+their model evidence, computed via thermodynamic integration.
+Missing are only the two graphs where the arc from \gls{lnl} III to IV was removed.
+Their respective model evidence fell below $-1528$ and did thus not fit on this figure.
+In the bottom left corner, we provide a visual reference in analogy to \cref{table:bayes_factor}: E.g., any difference in the model evidence shorter than the first of the three rulers indicates that the improvement is "barely worth a mention".}
+    \label{fig:ranking}
+\end{figure*}
 
@@
 The result that $t_{\text{II} \rightarrow \text{III}}$ and $t_{\text{III} \rightarrow \text{IV}}$ are relatively large compared to $b_\text{III}$ and $b_\text{IV}$ reflects the observation that skip metastases in levels III and IV without involvement of the upstream level is rare.
-Since level II has no parent nodes except for the primary tumor, $b_\text{II}$ can be easily related to the prevalence of involvement in level II.
+Since level II's parent node is rather rarely involved, $b_\text{II}$ can roughly be related to the prevalence of involvement in level II.
 The probability for no involvement of level II when the patient is diagnosed after $t$ time steps is $\left(1-b_\text{II}\right)^t$.
@@
         \text{prev}_\text{late}^\text{II} &= 1-\sum_{t=0}^{10} \left(1-b_\text{II}\right)^t \cdot p_\text{late}^t \, (1-p_\text{late})^{(10-t)}\binom{10}{t} \\
-        &\approx 0.8
+        &\approx 79 \%
     \end{aligned}
@@
 %
-which agrees with \cref{fig:bg_prevalences}.
+which agrees with the second panel from the top in \cref{fig:bg_prevalences}.
 
@@
 
-{\it T-stage dependence of level I involvement:} As shown in \cref{fig:bg_prevalences}, the model describes the involvement of \glspl{lnl} II, III, and IV depending on T-stage very well despite having only a single parameter related to T-stage.
+{\it T-category dependence of level I involvement:} As shown in \cref{fig:bg_prevalences}, the model describes the involvement of \glspl{lnl} II, III, and IV depending on T-stage very well despite having only a single parameter related to T-stage.
 \Cref{fig:IandII_prevs} shows that the model does not perfectly describe the T-stage dependence of level I.

### 2023-12-04 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 We have fixed $p_\text{early}=0.3$ for early T-category tumors, and $t_\text{max}=10$ time steps).
-The result that $t_{\text{II} \rightarrow \text{III}}$ and $t_{\text{III} \rightarrow \text{IV}}$ are relatively large compared to $b_\text{III}$ and $b_\text{IV}$ reflects the observation that skip metastases in levels III and IV without involvement of the upstream level is rare.
+The result that $t_{\text{II} \rightarrow \text{III}}$ and $t_{\text{III} \rightarrow \text{IV}}$ are relatively large compared to $b_\text{III}$ and $b_\text{IV}$ reflects the observation that skip metastases in levels III and IV without involvement of the upstream level are rare.
 Since level II's parent node is rather rarely involved, $b_\text{II}$ can roughly be related to the prevalence of involvement in level II.
@@
 The hatched gray modalities display the same predictions as made by the base graph model.
-The winning graph slightly improves its fit to the data in the case of early T-category prevalence of \gls{lnl} V without \gls{lnl} II, and in the case of late T-category co-involvement of the levels II and V.
-This may stems from the added arc from \gls{lnl} II to V.
+The winning graph barely improves its fit to the data in the case of early T-category prevalence of \gls{lnl} V without \gls{lnl} II, and in the case of late T-category co-involvement of the levels II and V.
 \\
@@
 For early T-category cases, the fit of the base graph model to the data is actually better than the winning graph's.
-For late T-category, however, the winning graph shows a better ability to reverse the prevalence of the two investigated patterns of involvement (note that the base graph model predicts the prevalences the wrong way around here).
-This is presumably due to the added connection from \gls{lnl} III to V.}
+For late T-category, however, the winning graph shows a better ability to reverse the prevalence of the two investigated patterns of involvement (note that the base graph model predicts the prevalences the wrong way around here).}
         \label{fig:IIandIIIandV_prevs}
@@
 Green plots indicate \gls{lnl} I involvement without level II, while orange plots indicate the opposite (\gls{lnl} II without level I).
-The winning graph has an added edge from \gls{lnl} II to I, which improves the prediction of the rare green pattern and minimally reduces the overestimation of the orange pattern in late T-category.}
+The winning graph has an added edge from \gls{lnl} I to II, which improves the prediction of the rare green pattern.}
         \label{fig:IandII_prevs}
@@
         \caption{Predictions for the prevalence of \gls{lnl} VII involvement based on the base graph (hatched) and the winning graph (blue histogram) plotted against the beta posterior over the prevalence in the data (solid line).
-This is shown for early T-category (left panel) and late T-category (right panel).}
+This is shown for early T-category (left panel) and late T-category (right panel).
+Since no connections are added to or from level VII, no change in the model's ability to describe this \gls{lnl}'s involvement is expected.}
         \label{fig:VII_prevs}

### 2023-12-04 src/scripts/compile_metrics_table.py
use tag for loose_accuracies

@@ -23,7 +23,7 @@ REVISIONS = {
     "add-III-to-V": "add III → V",
     "add-IV-to-V": "add IV → V",
     "del-II-to-III": "remove II → III",
-    "del-II-to-III-and-III-to-IV": "remove II → III & III → IV",
+    "loose-graph-v1": "remove II → III & III → IV",
     "del-III-to-IV": "remove III → IV",
 }
 

### 2023-12-05 src/scripts/comp_IIIandIVandV_prevs.py
change plots to more relevant with new win graph

@@ -1,6 +1,6 @@
 """
 Plot histograms of predicted and posteriors over observed prevalences for scenarios
-that involve the LNLs II, III, and V. Compare these prevalence predictions for the
+that involve the LNLs III, IV, and V. Compare these prevalence predictions for the
 base-graph model to the winning-graph model.
 """
 from pathlib import Path
@@ -27,26 +27,26 @@ if __name__ == "__main__":
     plt.rcParams.update(fontsizes.icml2022())
     fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col")
 
-    # left panels: prevalences of LNLs II and V
-    BASE_GRAPH_INPUT = paths.data / "bg_IIandV_prevs.hdf5"
-    WIN_GRAPH_INPUT = paths.data / "wg_IIandV_prevs.hdf5"
+    # prevalences of LNLs IV and V
+    BASE_GRAPH_INPUT = paths.data / "bg_IVandV_prevs.hdf5"
+    WIN_GRAPH_INPUT = paths.data / "wg_IVandV_prevs.hdf5"
 
-    left_panels = {
+    panels_IVandV = {
         "early": [],
         "late": [],
     }
 
     for stage in ["early", "late"]:
-        left_panels[stage] = [
+        panels_IVandV[stage] = [
             Histogram.from_hdf5(
                 filename=WIN_GRAPH_INPUT,
-                dataname=f"IIandV/{stage}",
-                label="LNL II and V",
+                dataname=f"IVandV/{stage}",
+                label="LNL IV and V",
                 color=USZ["orange"],
             ),
             Histogram.from_hdf5(
                 filename=BASE_GRAPH_INPUT,
-                dataname=f"IIandV/{stage}",
+                dataname=f"IVandV/{stage}",
                 label="base graph",
                 histtype="step",
                 color="darkgray",
@@ -55,19 +55,19 @@ if __name__ == "__main__":
             ),
             Posterior.from_hdf5(
                 filename=BASE_GRAPH_INPUT,
-                dataname=f"IIandV/{stage}",
+                dataname=f"IVandV/{stage}",
                 color=USZ["orange"],
             ),
 
             Histogram.from_hdf5(
                 filename=WIN_GRAPH_INPUT,
[…]

### 2023-12-05 src/scripts/comp_VII_prevs.py

@@ -1,81 +1,0 @@
-"""
-Plot histograms of predicted and posteriors over observed prevalences for scenarios
-that involve the LNL VII. Compare these prevalence predictions for the base-graph
-model to the winning-graph model.
-"""
-from pathlib import Path
-
-import matplotlib.pyplot as plt
-from lyscripts.plot.utils import Histogram, Posterior, draw
-from lyscripts.plot.utils import COLORS as USZ
-from tueplots import figsizes, fontsizes
-
-import paths
-
-
-BASE_GRAPH_INPUT = paths.data / "bg_VII_prevs.hdf5"
-WIN_GRAPH_INPUT = paths.data / "wg_VII_prevs.hdf5"
-OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
-NROWS, NCOLS = 1, 2
-
-
-if __name__ == "__main__":
-    plt.rcParams.update(figsizes.icml2022_half(
-        nrows=NROWS,
-        ncols=NCOLS,
-        height_to_width_ratio=1.3,
-    ))
-    plt.rcParams.update(fontsizes.icml2022())
-    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharey="row")
-
-
-    panels = {
-        "early": [],
-        "late": [],
-    }
-
-    for stage in ["early", "late"]:
-        panels[stage] = [
-            Histogram.from_hdf5(
-                filename=WIN_GRAPH_INPUT,
-                dataname=f"VII/{stage}",
-                label="LNL VII overall",
-                color=USZ["blue"],
-            ),
-            Histogram.from_hdf5(
-                filename=BASE_GRAPH_INPUT,
-                dataname=f"VII/{stage}",
-                label="base graph",
-                histtype="step",
-                color="darkgray",
-                linewidth=2.,
-                hatch=r"////",
-            ),
-            Posterior.from_hdf5(
-                filename=BASE_GRAPH_INPUT,
-                dataname=f"VII/{stage}",
-                color=USZ["blue"],
-            ),
-        ]
-
-    draw(axes[0], contents=panels["early"], xlims=(1., 9.), hist_kwargs={"nbins": 20})
-    h, l = axes[0].get_legend_handles_labels()
-    handles = [*h[2:4], h[0], *h[4:6], h[1]]
[…]

### 2023-12-05 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \begin{figure*}
-    \script{comp_IIandIIIandV_prevs.py}
+    \script{comp_IIIandIVandV_prevs.py}
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/comp_IIandIIIandV_prevs.png}
-        \caption{{\it Left panel:} Beta posterior over the prevalence of the involvement pattern where \gls{lnl} II \emph{and} V show metastases (orange line), as well as the pattern with \gls{lnl} V involved \emph{without} level II (blue line).
-Both are shown for early (top panel) and late T-category (bottom panel).
-The correspondingly colored histograms show the same prevalences as predicted by the winning graph model.
-The hatched gray modalities display the same predictions as made by the base graph model.
-The winning graph barely improves its fit to the data in the case of early T-category prevalence of \gls{lnl} V without \gls{lnl} II, and in the case of late T-category co-involvement of the levels II and V.
-\\
-        {\it Right panel:} The same plots, but for involvement patterns where \gls{lnl} V is involved together with \gls{lnl} III (red) and where level V is involved while \gls{lnl} III is healthy (green).
-For early T-category cases, the fit of the base graph model to the data is actually better than the winning graph's.
-For late T-category, however, the winning graph shows a better ability to reverse the prevalence of the two investigated patterns of involvement (note that the base graph model predicts the prevalences the wrong way around here).}
-        \label{fig:IIandIIIandV_prevs}
+        \includegraphics[width=\linewidth]{figures/comp_IIIandIVandV_prevs.png}
+        \caption{TODO}
+        \label{fig:IIIandIVandV_prevs}
     \end{centering}
@@
 
-{\it Level V:} In \cref{fig:IIandIIIandV_prevs} we compare the base graph's and the winning graph's estimations for prevalences of involvement patterns that include \gls{lnl} V.
+{\it Level V:} In \cref{fig:IIIandIVandV_prevs} we compare the base graph's and the winning graph's estimations for prevalences of involvement patterns that include \gls{lnl} V.
 The base graph's predictions have been plotted as gray, hatched histograms.
@@
 \end{figure*}
-
-{\it Level VII:} The two models' predictions w.r.t.
-the prevalence of involvement patterns that include \gls{lnl} VII are compared in \cref{fig:VII_prevs}.
-Here, the base and winning graph yield almost the same predictions, as no new edge connects from or to \gls{lnl} VII in the latter model.
-Notably, the prevalence of early T-category level VII involvement is slightly overestimated, while for late T-category, the fit is very good.
-
-\begin{figure}
-    \script{comp_VII_prevs.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/comp_VII_prevs.png}
-        \caption{Predictions for the prevalence of \gls{lnl} VII involvement based on the base graph (hatched) and the winning graph (blue histogram) plotted against the beta posterior over the prevalence in the data (solid line).
-This is shown for early T-category (left panel) and late T-category (right panel).
-Since no connections are added to or from level VII, no change in the model's ability to describe this \gls{lnl}'s involvement is expected.}
-        \label{fig:VII_prevs}
-    \end{centering}
-\end{figure}
 

### 2023-12-11 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
+Note that it is also possible to account for incomplete diagnoses, i.e., a diagnosis where the involvement information for one or more \glspl{lnl} is missing.
+In that case, we can sum over those elements of $\boldsymbol{\ell}$ that correspond to complete diagnoses which match the provided incomplete one.
+In this paper, for some patients involvement information of level VII was missing and hence marginalized over.
+A detailed explanation of this formalism can be found in \cite{zora231470}, section 6.2.7.
+
 Using this log-likelihood function one may now employ a variety of inference methods to learn the parameters of the model that best describe the observed data.
@@
 \end{equation}
-%
-% \ju{Also say that the above equation gives the probability of a hidden state. To get the probability of involvement in a given level one has to marginalize over all corresponding states}
 
@@
 The lymphatic drainage to or from levels I, V, and VII is not as clearly defined.
-Therefore, we define a set of candidate arcs (green arcs in \cref{fig:extended_graph}) and use the model comparison methodology outlined in \cref{subsec:complete_model:comparison} to see which arc is motivated from our data.
+Therefore, we define a set of candidate arcs (green arcs in \cref{fig:extended_graph}) and use the model comparison methodology described in \cref{subsec:complete_model:comparison} to determine which graph is most supported our data.
 
@@
 And, although no lymphatic pathway is described that directly drains \gls{lnl} III or IV into level V, due to their proximity to each other, we will also investigate additional arcs from \gls{lnl} III and IV into level V.
+Finally, an arc from \gls{lnl} II to \gls{lnl} VII is considered.
 
-To determine the optimal graph describing the data at hand best, we considered six models, each with one of the green candidate arcs of \cref{fig:extended_graph} added to the base graph.
+To determine the optimal graph, we first consider six models, each with one of the six green candidate arcs of \cref{fig:extended_graph} added to the base graph.
 Every one of these models was evaluated by computing its evidence via thermodynamic integration as described below in \cref{subsec:complete_model:comparison}.
-These evidences are shown, among others, in \cref{table:evidence}.
-Promising combination with two added edges were then considered in a combined model which was ultimately designated to be the "winning graph" which is in great detail compared to the base graph in the results (\cref{sec:results}).
-All considered graphs' performances are visualized in \cref{fig:ranking}.
+%These evidences are shown, among others, in \cref{table:evidence}. 
+Subsequently, graphs combining multiple arcs that individually improve the model evidence are considered.
+Thereby, the ``winning graph'' is determined, which yields the highest value of the model evidence.
+%which is in great detail compared to the base graph in the results (\cref{sec:results}). All considered graphs' performances are visualized in \cref{fig:ranking}.
 
@@
 This is remarkable because the model introduces only a single parameter to describe the differences between early and late T-category for all involvement patterns.
-This shows that expecting later diagnosis times, on average, for patients with advanced T-category tumors is able to explain more severe lymphatic involvement.
+This shows that expecting later diagnosis times, on average, for patients with advanced T-category tumors can explain more severe lymphatic involvement.
 
@@
 
-\subsection{Graph Candidates}
+\subsection{Comparison of candidates graphs}
 \label{subsec:results:graph_candidates}
@@
 
-The model evidences of the six models that were based on graphs with one additional arc compared to the base model's are plotted in \cref{table:evidence}.
+The model evidences of all candidate graphs are reported in \cref{table:evidence}.
 A visual ranking is provided in \cref{fig:ranking}.
-Evidently, adding a connection from \gls{lnl} I to II is strongly supported given this dataset and slightly beats the reverse connection from \gls{lnl} II to I.
-Similarly, there is strong evidence for introducing an arc from \gls{lnl} IV to V.
+Let us first consider the six models in which one of the candidate arcs is added to the base graph.
+Evidently, adding a connection from \gls{lnl} I to II is strongly supported given this dataset, and is slightly superior to the reverse connection from \gls{lnl} II to I.
+In addition, there is strong evidence for introducing an arc from \gls{lnl} IV to V.
 Furthermore, there is substantial evidence for an arc from \gls{lnl} III to V.
-All other investigated additions lead to improvements that are at best barely worth a mention or do not justify the additional complexity at all, indicated by a lower evidence.
+All other investigated additions lead to improvements that are barely worth a mention or do not justify the additional complexity at all, indicated by a lower evidence.
 
-Based on this table, we determined the final candidates for the optimal graph should be combinations of the graphs with the added connections from level I to II, III to V, and IV to V.
-For the three resulting graphs we again computed the model evidence and added it to the table in \cref{table:evidence}.
+Based on these results, we consider three additional candidates for the optimal graph that combine the added arc from \gls{lnl} I to II with the arc(s) $\text{III} \rightarrow \text{V}$ and/or $\text{IV} \rightarrow \text{V}$.
+The model evidence for these three graphs is reported in \cref{table:evidence}.
 The best performing graph with decisive evidence over the base graph turned out to be the one which combines the arcs from \gls{lnl} I to II and IV to V.
-Interestingly, the evidence gain of this "winning graph" is roughly the sum of the gains seen in the two candidates where only one of these connections was added, respectively.
+Interestingly, the evidence gain of this ``winning graph'' is roughly the sum of the gains seen in the two candidates where only one of these connections was added, respectively.
 This indicates that the two additional parameters are largely independent of each other and manage to describe different aspects of the data.
+
+\Cref{table:evidence} additionally shows the evidence of graphs in which the arc from level II to III and from level III to IV is removed.
+The low model evidence for these graphs confirms the importance of these connections and is consistent with the anatomical motivation.
+The connection from level III to IV is crucial for describing the observation that metastases in level IV are extremely rare without simultaneous involvement of the upstream level III.
 
@@
 their model evidence, computed via thermodynamic integration.
-Missing are only the two graphs where the arc from \gls{lnl} III to IV was removed.
-Their respective model evidence fell below $-1528$ and did thus not fit on this figure.
+Not shown are the two graphs where the arc from \gls{lnl} III to IV was removed.
+Their respective model evidence fell below $-1528$ and the two graphs would appear far left in the figure.
 In the bottom left corner, we provide a visual reference in analogy to \cref{table:bayes_factor}: E.g., any difference in the model evidence shorter than the first of the three rulers indicates that the improvement is "barely worth a mention".}
@@
 
-\subsection{The Optimal Extended Graph}
+\subsection{The winning graph}
 \label{subsec:results:optimal_graph}
@@
     \variable{output/results_table.tex}
-    \caption{Model comparison results from the base graph and the extended graph we chose as the ``winnning'' model.
+    \caption{Model comparison results from the base graph and the extended graph we chose as the ``winning'' model.
 For both \glspl{dag} we show the log-evidence, computed via thermodynamic integration, the negative one half of the \gls{bic}, as well as the maximum log-likelihood that was encountered during the final \gls{mcmc} sampling round.}
@@
 The most likely model parameters for the winning graph, corresponding to the mean of the marginals of the sampled posterior distributions, are tabulated in \cref{table:params}.
-We have fixed $p_\text{early}=0.3$ for early T-category tumors, and $t_\text{max}=10$ time steps).
+We have fixed $p_\text{early}=0.3$ for early T-category tumors (i.e., T0, T1, and T2), and $t_\text{max}=10$ time steps.
 The result that $t_{\text{II} \rightarrow \text{III}}$ and $t_{\text{III} \rightarrow \text{IV}}$ are relatively large compared to $b_\text{III}$ and $b_\text{IV}$ reflects the observation that skip metastases in levels III and IV without involvement of the upstream level are rare.
-Since level II's parent node is rather rarely involved, $b_\text{II}$ can roughly be related to the prevalence of involvement in level II.
+Since level II's parent node (level I) is rarely involved, $b_\text{II}$ can approximately be related to the prevalence of involvement in level II.
 The probability for no involvement of level II when the patient is diagnosed after $t$ time steps is $\left(1-b_\text{II}\right)^t$.
@@
 which agrees with the second panel from the top in \cref{fig:bg_prevalences}. 
+The large value for the parameter $t_{\text{I} \rightarrow \text{II}}$ reflects the observation that in almost all patients with level I involvement, level II is also involved.
+The large uncertainty in $t_{\text{I} \rightarrow \text{II}}$ is related to the fact that level I involvement is rare compared to level II.
 
@@
 
-\subsection{Involvement of levels I, V, and VII}
+\subsection{Involvement of levels I and V}
 
-First, we observe that the winning graph describes the involvement of levels II, III, and IV equally well as the base graph, a result that is expected and not further shown.
+We can observe that the winning graph describes the involvement of levels II, III, and IV equally well as the base graph, a result that is expected and not further shown.
 We thus focus on the improvements w.r.t.
-involvement patterns that include the \glspl{lnl} I, V, and VII, that more rarely harbor metastases.
+involvement patterns that include the \glspl{lnl} I and V, that more rarely harbor metastases.
 
@@
         \includegraphics[width=\linewidth]{figures/comp_IIIandIVandV_prevs.png}
-        \caption{TODO}
+        \caption{Observed (Beta posteriors as lines) vs.
+predicted (histograms) prevalences of involvement combinations that include \gls{lnl} V.
+We have plotted the predictions from the winning graph (colored histograms) and those of the base graph (gray, hatched histograms).
+The top two panels show scenarios for early T-category patients, the bottom two panels for advanced T-category.
+The left two panels consider combinations of \gls{lnl} III and V involvement, while the right two panels consider combinations of \gls{lnl} IV and V.
+The colored lines show the Beta posterior over the prevalence of the respective involvement pattern, given the data}
         \label{fig:IIIandIVandV_prevs}
@@
 {\it Level V:} In \cref{fig:IIIandIVandV_prevs} we compare the base graph's and the winning graph's estimations for prevalences of involvement patterns that include \gls{lnl} V.
-The base graph's predictions have been plotted as gray, hatched histograms.
-Alongside, we show again the Beta posterior over the prevalence of the respective involvement pattern, given the data, as colored line.
-The prediction by the winning graph are shown as colored histograms.
-Every scenario is shown for early and late T-category.
-This allows us to compare how the winning graph improves its fit to the data in these particular scenarios around low-risk \glspl{lnl}.
-How frequently each of the investigated lymphatic metastatic patterns were observed in the data is additionally tabulated in \cref{table:data_prevalence}, again both for early and for late T-category cases.
+The base graph underestimates the probability that level IV and V are simultaneously involved, and overestimates that probability the level V but not IV is involved.
+By introducing the arc from level IV to V, the winning graph can describe the observation that level V involvement is typically associated with severe involvement of level II-IV.
+In the dataset, 15 patients out of 68 patients with level IV involvement have metastases in level V.
+Instead only 40 patients out of 618 patients without level IV involvement have metastases in level V.
 
-{\it Level I:} In \cref{fig:IandII_prevs}, the same kind of comparisons are plotted for involvement patterns that inlcude metastatic disease in \gls{lnl} I.
-Here, we observe that the model does at times struggle to capture the spread between the prevalence of \gls{lnl} I involvement for early and late T-category, which is more pronounced than in the levels II and III.
+{\it Level I:} In \cref{fig:IandII_prevs}, analogous comparisons are shown for involvement patterns that include \gls{lnl} I.
+The base graph overestimates the probability of level I involvement without simultaneous involvement of level II.
+By introducing the arc from level I to II, the winning graph can capture the correlations between levels I and II.
+It can also be noted that both models overestimate level I involvement for early T-category patients and underestimate level I involvement for late T-category patients.
+This is further described in the discussion section below.
 
@@
 
+{\it Level II:} As can be seen in \cref{table:params}, spread from the tumor to \gls{lnl} II to be the most probable transition at any given time step.
+As a consequence, even for an early T-category patient that presents with a clinical N0 neck, our model predicts a \variable{output/wg_II_risks.tex} risk for microscopic metastases in \gls{lnl} II.
+
 {\it Level III:} \Cref{fig:wg_III_risks} compares the risk of occult disease in level III between patients that are clinically N0 (orange) and patients with clinically diagnosed involvement of only level II (red), for early T-category (upper panel) and advanced T-category (bottom panel).
@@
 For patients with involvement of level II, the risk in level III increases to approximately 9\% and 12\%, respectively.
-
-% \ju{perhaps one could use the same color in each risk figure for a given clinical involvement, e.g. always green for N0}
 
@@
 For clinically N0 patients, the risk in level V is estimated to be around 1\%.
-Extensive nodal involvement of levels II-IV increases the risk in level V to approximately 4\% for advanced T-stage tumors.
+Extensive nodal involvement of levels II-IV increases the risk in level V to more than 4\% for advanced T-stage tumors.
 
@@
 For clinically N0 patients, the risk in level I is estimated to be in the order of 1-2\%.
-Extensive nodal involvement of levels II-IV increases the risk in level I to approximately 4-5\% for advanced T-stage tumors.
+Extensive nodal involvement of levels II-IV increases the risk in level I to just below 4\% for advanced T-stage tumors.
 It is pointed out that the winning graph does not contain arcs from levels III or IV to level I (and anatomically we do not assume that there is lymphatic drainage from levels III or IV to level I).
 Thus, the increased risk in level I is related to the time evolution: Involvement of levels III and IV corresponds to a more advances state of disease that is likely diagnosed at a later time step, such that the tumor had more time to spread to level I.
+The correlation between the clinical involvement pattern, the likely time of diagnosis in the tumor's time frame based on it, and from that the risk of involvement is another benefit of the formulation of the model as a \gls{hmm}.
 
@@
 \end{figure}
-
-{\it Level II:} As can be seen in \cref{table:params} the model infers the spread from the tumor to \gls{lnl} II to be the most probable transition at any given time step.
-As a consequence, even for an early T-category patient that presents with a clinical N0 neck, our model predicts a \variable{output/wg_II_risks.tex} risk for microscopic metastases in \gls{lnl} II.
 
@@
 It is demonstrated for the first time that the model can accurately describe the patterns of lymph node involvement observed in the data, including the correlations between levels and its dependence on T-stage.
-Furthermore, state-of-the-art techniques are applied to calculate the model evidence for Bayesian model comparison.
-This yields a complete model including all \glspl{lnl} relevant for \gls{opscc}: I, II, III, IV, V, and VII with an parameterization that balances accuracy and model complexity.
+Furthermore, techniques from statistical physics are applied to calculate the model evidence for Bayesian model comparison.
+This yields a complete model including all \glspl{lnl} relevant for \gls{opscc}: I, II, III, IV, V, and VII with a parameterization that balances accuracy and model complexity.
 
@@
 Risk predictions obtained by the model may be used to design clinical trials on volume-deescalated treatment of \gls{opscc}.
-In the context of radiotherapy, this corresponds to excluding \glspl{lnl} from the CTV-N, which are irradiated according to the current guidelines.
+In the context of radiotherapy, this corresponds to excluding \glspl{lnl} from the \gls{ctv-n}, which are irradiated according to the current guidelines.
 The list below should be seen as a summary of \cref{subsec:results:risk_prediction} and the limitations discussed in \cref{subsec:disc:limitation} should be taken into account in its interpretation.
@@
 \begin{itemize}
-    \item[$\bullet$] Irradiate level II for all patients.%\ju{maybe we should actually show the risk in level II for N0 patients, at least state the risk in the text, are even have the risk histogram for completeness.}
+    \item[$\bullet$] Irradiate level II for all patients.
     \item[$\bullet$] Irradiate level III for most patients. Only for clinically N0, early T-category patients, not irradiating level III could be considered. 
-    \item[$\bullet$] Exclude level IV from the CTV-N for patients with clinically negative level III. For advanced T-category patients with involvement of level III, level IV should be irradiated.
-    \item[$\bullet$] Exclude level V from the CTV-N for most patients. Only for patients with extensive involvement of levels II, III, and IV, irradiation of level V can be considered.
-    \item[$\bullet$] Exclude level I from the CTV-N for early T-category patients with limited metastatic disease.
+    \item[$\bullet$] Exclude level IV from the \gls{ctv-n} for patients with clinically negative level III.
+For advanced T-category patients with involvement of level III, level IV should be irradiated.
+    \item[$\bullet$] Exclude level V from the \gls{ctv-n} for most patients. Only for patients with extensive involvement of levels II, III, and IV, irradiation of level V can be considered.
+    \item[$\bullet$] Exclude level I from the \gls{ctv-n} for early T-category patients with limited metastatic disease.
 For advanced T-category patients with extensive nodal involvement of levels II, III, and IV, level I should be irradiated (see also the limitations discussed below).
@@
 It adjusts the parameters such that level I involvement is correctly described for the set of all patients combined, but it overestimates level I involvement for early T-stage and underestimates it for advanced T-stage.
-A possible explanation is that advanced T-stage tumors are more likely to have grown into a regions with direct lymph drainage to level I.
+A possible explanation is that advanced T-stage tumors are more likely to have grown into regions with direct lymph drainage to level I.
 The more severe involvement in levels II, III, IV for advanced T-stage can be explained by tumors having more time to spread while keeping the spread probability rates $b_2$, $b_3$, $b_4$ constant.
@@
 Consequently, these values need to be considered with caution.
-\cref{fig:wg_sens_spec_risks} illustrates for one example how the risk of occult disease depends on sensitivity and specificity.
+\Cref{fig:wg_sens_spec_risks} illustrates for one example how the risk of occult disease depends on sensitivity and specificity.
 Here, we consider the risk in level IV in patients with clinically detected metastases in levels II and III.
-For our default parameters of 81\% specificity and 76 \% sensitivity, the risk is 5\%, but it increases to x\% for a a sensitivity of 66\%.
+For our default parameters of 81\% specificity and 76 \% sensitivity, the risk is 5\%, but it increases to around 8\% for a a sensitivity of 66\%.
 Future work may aim at developing new methods to model the difference between pathological and clinical lymph node involvement based on surgically treated patients in whom both is reported.

### 2023-12-14 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Cross-referencing
-\usepackage{hyperref}
-\usepackage{cleveref}
+% \usepackage{cleveref}
 
@@
 \begin{abstract}
-    Objective: The \gls{ctv-n}\glsunset{ctv} in \gls{hnscc}\glsunset{scc} is currently based mostly on the prevalence of lymph node metastases in different \glspl{lnl} for a given primary tumor location.
-In this work, we present a probabilistic model for ipsilateral lymphatic metastatic spread that can quantify the risk for microscopic nodal involvement based on an individual patient's T-stage and clinical involvement of \glspl{lnl} at diagnosis.
-\\
-    Methods: The model is based on the formalism of \glspl{hmm} as conceptually introduced in a previous publication.
-The \gls{hmm} uses a collection of hidden binary \glspl{rv} -- one for each of the \glspl{lnl} I, II, III, IV, V, and VII -- to model a patient's true state of lymph node involvement corresponding to pathological examinations.
-Clinical involvement at diagnosis represent the observed binary \glspl{rv} linked to the true state via sensitivity and specificity.
-A \gls{dag} with is used to compute the transition matrix of the \gls{hmm}, where each edge is associated with a parameter for the probability of tumor spread from the primary tumor and between \gls{lnl}.
-To learn the parameters, we collected a dataset of 681 \gls{opscc} patients from three institutions.
-To determine the best graph that balances model complexity and accurate description of \gls{lnl} involvement patterns in the data, we compare graphs based on the model evidence calculated via \gls{ti}.
-\\
-    Results: The graph maximizing the model evidence contains edges for lymphatic spread from \gls{lnl} II to I, III and V, and from \gls{lnl} III to IV and V, and is thus consistent with commonly assumed lymphatic drainage in the neck.
-Application of the model suggests that the risk of occult disease in level IV is below 5\% if level III is clinically negative, and that the risk of occult disease in level V is below 5\% except for advanced T-category (T3 and T4) patients with clinical involvement of levels II, III, and IV.
-    Conclusion: We provide a statistical model of lymph node involvement in \gls{opscc} patients trained on a multi-institutional dataset, which may guide the design of clinical trials on volume-deescalated treatment of \gls{opscc} and eventually contribute to further personalized guidelines on elective nodal treatment.
+    Objective: The \gls{ctv-n} in \gls{opscc} is currently based mostly on the prevalence of lymph node metastases in different \glspl{lnl} for a given primary tumor location.
+We present a probabilistic model for ipsilateral lymphatic spread that can quantify the risk for microscopic nodal involvement based on an individual patient's T-category and clinical involvement of \glspl{lnl} at diagnosis.
+
+    Methods: We extend a previously published \gls{hmm}, which models the \glspl{lnl} (I, II, III, IV, V, and VII) as hidden binary \glspl{rv}.
+Each represents a patient's true state of lymphatic involvement corresponding to pathological examinations in that level.
+Clinical involvement at diagnosis represents the observed binary \glspl{rv} linked to the true state via sensitivity and specificity.
+The primary tumor and the hidden \glspl{rv} are connected in a graph.
+Each edge represents the conditional probability of metastatic spread per abstract time-step, given involvement at the edge's starting node.
+To learn these probabilities, we draw \acrlong{mcmc} samples from the likelihood of a dataset (\variable{output/num_patients.txt} \gls{opscc} patients) from three institutions.
+We compute the model evidence using \acrlong{ti} for different graphs to determine which describes the data best.
+
+    Results: The graph maximizing the model evidence connects the tumor to each \gls{lnl} and the \glspl{lnl} I through V in order.
+It predicts the risk of occult disease in level IV is below 5\% if level III is clinically negative, and that the risk of occult disease in level V is below 5\% except for advanced T-category (T3 and T4) patients with clinical involvement of levels II, III, and IV.
+
+    Conclusion: The provided statistical model of nodal involvement in \gls{opscc} patients trained on multi-institutional data may guide the design of clinical trials on volume-deescalated treatment of \gls{opscc} and contribute to further personalized guidelines on elective nodal treatment.
 \end{abstract}
@@
 \label{sec:intro}
+
+\glsunset{ctv}
+\glsunset{scc}
 
@@
 We previously developed a model of lymphatic metastatic progression for estimating the risk of microscopic disease, given a patient's personal diagnosis.
-The initial model was based on the methodology of Bayesian networks \cite{pouymayou_bayesian_2019}.
-It was subsequently extended and formulated as a \glspl{hmm} to include T-category in an intuitive manner \cite{ludwig_hidden_2021}.
-However, these models were introduced based on only a small dataset of approximately 100 early T-category patients \cite{sanguineti_defining_2009} available at that time.
-The limited data did not allow us to quantify the probability of metastases in the rarely involved \gls{lnl} I, V, and VII, nor did the data allow us to verify that the \gls{hmm} is adequate to describe the dependence on lymph node involvement on T-category.
+The initial model was based on the methodology of \glspl{bn} \cite{pouymayou_bayesian_2019}.
+It was subsequently extended and formulated as an \gls{hmm} to include T-category in an intuitive manner \cite{ludwig_hidden_2021}.
+However, these models were introduced based on only a small dataset of approximately 100 early T-category patients available at that time \cite{sanguineti_defining_2009}.
+The limited data did not allow us to quantify the probability of metastases in the rarely involved \glspl{lnl} I, V, and VII, nor did the data allow us to verify that the \gls{hmm} is adequate to describe the dependence on lymph node involvement on T-category.
 In this paper, we extend the previous work \cite{ludwig_hidden_2021} by making the following contributions:
 \begin{enumerate}
-    \item We provide a \glspl{hmm} of ipsilateral lymph node involvement including all relevant \glspl{lnl}, namely the levels I, II, III, IV, V, and VII.
-To determine the underlying \gls{dag} we compare different graphs by calculating the model evidence through \gls{ti}.
-    \item We collect a multi-centric dataset consisting of \variable{output/num_patients.txt} patients from three institutions, allowing us to train the model based on a sizable dataset.
+    \item We provide a \gls{hmm} of ipsilateral lymph node involvement including all relevant \glspl{lnl}, namely the levels I, II, III, IV, V, and VII.
+To determine the optimal underlying \gls{dag} we compare different graphs by calculating the model evidence through \gls{ti}.
+    \item We collect a multi-centric dataset consisting of \variable{output/num_patients.txt} patients from three institutions, allowing us to train the model based on a sizable dataset \cite{ludwig_dataset_2022,ludwig_multi-centric_2023}.
     \item We use the trained model to provide personalized risk estimations for occult metastases for typical clinical states of tumor progression at diagnosis, illustrating its potential for guiding volume-deescalated treatment strategies in the future.
@@
 %
-Each the \gls{lnl} can be in the state $X_v=0$ (\texttt{FALSE}), meaning \gls{lnl} $v$ is healthy, or in the state $X_v=1$ (\texttt{TRUE}), indicating the \gls{lnl} harbors metastases.
+Each of the \glspl{lnl} can be in the state $X_v=0$ (\texttt{FALSE}), meaning \gls{lnl} $v$ is healthy, or in the state $X_v=1$ (\texttt{TRUE}), indicating the \gls{lnl} harbors metastases.
 The involved state includes occult disease.
@@
 The term $P\left( \boldsymbol{\xi}_i \mid \boldsymbol{\xi}_j \right)$ describes the probability to transition from the hidden state of lymphatic involvement $\boldsymbol{\xi}_j$ to the state $\boldsymbol{\xi}_i$ between the time $t$ and $t+1$.
-Using a \gls{dag} as depicted in \cref{fig:graph_with_obs}, we can formulate this transition probability in the following way:
+Using a \gls{dag} as depicted in \autoref{fig:graph_with_obs}, we can formulate this transition probability in the following way:
 %
@@
 
-The terms of the form $P \left( \xi_{iv} \mid \left\{ \xi_{jr} \right\}_{r \in \operatorname{pa}(v)} \right)$ implicitly depend on how we parameterize e.g.
-the arcs of \cref{fig:graph_with_obs}.
-For example, if we look at the probability of spread to the \gls{lnl} $X_3$ depending on the state of that level's parent -- which, in this case, is $\operatorname{pa}(3) = 2$ -- we can write the different combinations into a conditional probability table as below.
+The terms of the form $P \left( \xi_{iv} \mid \left\{ \xi_{jr} \right\}_{r \in \operatorname{pa}(v)} \right)$ implicitly depend on how we parameterize the arcs of \autoref{fig:graph_with_obs}.
+For example, if we look at the probability of spread to \gls{lnl} III ($X_3$) depending on the state of that level's parent -- which, in this case, is $\operatorname{pa}(3) = 2$ -- we can write the different combinations into a conditional probability table as below.
 
@@
     \includegraphics[width=0.6\linewidth]{figures/graph_with_obs.png}
-    \caption{\Gls{dag} representing a possible abstraction of the lymphatic network comprising the tumor (red shaded circle) and \glspl{lnl} II through IV as hidden binary \glspl{rv} (blue outlined circles).
+    \caption[x]{\Gls{dag} representing a possible abstraction of the lymphatic network comprising the tumor (red shaded circle) and \glspl{lnl} II through IV as hidden binary \glspl{rv} (blue outlined circles).
 Attached to each of these is the corresponding observed \gls{rv} (orange shaded squares).
@@
 We can only make the assumption that a patients with an earlier T-category tumor was \emph{probably} diagnosed after fewer time-steps than a patient with a late T-category tumor.
-We can use this information by marginalizing over the diagnose times $t_\text{D}$ of patients in different T-categories using different prior distributions over the diagnose time.
-E.g., $P \left( t=t_D \mid \text{early} \right)$ for early T-category patients (T1 \& T2) and $P \left( t=t_D \mid \text{late} \right)$ for late T-category patients (T3 \& T4).
+
+We can use this assumption by marginalizing over the diagnose times $t_\text{D}$ of patients in different T-categories using different prior distributions over the diagnose time.
+E.g., $P \left( t=t_\text{D} \mid \text{early} \right)$ for early T-category patients (T1 \& T2) and $P \left( t=t_\text{D} \mid \text{late} \right)$ for late T-category patients (T3 \& T4).
 Throughout this work we will use binomial distributions for these probability mass functions, mainly because they have a plausible shape for this purpose and only a single parameter.
@@
 \begin{equation} \label{eq:time_dist}
-    P \left( t = t_D \mid \text{T}x \right) = \operatorname{\mathfrak{B}}(t_\text{max},p_{\text{T}x})
+    P \left( t = t_\text{D} \mid \text{T}x \right) = \operatorname{\mathfrak{B}}(t_\text{max},p_{\text{T}x})
 \end{equation}
@@
 Here, the parameter $p_{\text{T}x}$ can be interpreted as the probability that the patient with a tumor of T-category $x$ will be diagnosed at time-step $t+1$ given they are in time-step $t$.
-We will use as the latest time-step $t_\text{max} = 10$, which will therefore give us a distribution over the diagnosis time that has its mean at $\mathbb{E}[t_D] = 10 \cdot p_{\text{T}x}$.
+We will use as the latest time-step $t_\text{max} = 10$, which will therefore give us a distribution over the diagnosis time that has its mean at $\mathbb{E}[t_\text{D}] = 10 \cdot p_{\text{T}x}$.
 
@@
 %
-This likelihood implicitly depends on how we parameterize the arcs of the \gls{dag} underlying the model -- see \cref{eq:transition_prob} -- and the parameterization of the distribution over diagnosis times -- e.g., as in \cref{eq:time_dist}.
+This likelihood implicitly depends on how we parameterize the arcs of the \gls{dag} underlying the model -- see \autoref{eq:transition_prob} -- and the parameterization of the distribution over diagnosis times -- e.g., as in \autoref{eq:time_dist}.
 
@@
 
-To infer these parameters from a dataset of $N$ \gls{hnscc} patients $\boldsymbol{\mathcal{D}} = \left( d_1, d_2, \ldots, d_N \right)$, we compute the data log-likelihood:
+To infer these parameters from a dataset of $N$ \gls{opscc} patients $\boldsymbol{\mathcal{D}} = \left( d_1, d_2, \ldots, d_N \right)$, we compute the data log-likelihood:
 %
@@
 %
-Which effectively amounts to computing the element-wise logarithm of the likelihood vector $\boldsymbol{\ell}$ from \cref{eq:likelihood_vec} and summing up the entries that correspond to each of the patients $d_i$ for $i\leq N$.
+Which effectively amounts to computing the element-wise logarithm of the likelihood vector $\boldsymbol{\ell}$ from \autoref{eq:likelihood_vec} and summing up the entries that correspond to each of the patients $d_i$ for $i\leq N$.
 
@@
 
-We use \gls{mcmc} sampling to draw parameter samples $\boldsymbol{\hat{\theta}}_i$ for $i \leq S$ from the likelihood described in \cref{eq:log_likelihood} (i.e.
+We use \gls{mcmc} sampling to draw parameter samples $\boldsymbol{\hat{\theta}}_i$ for $i \leq S$ from the likelihood described in \autoref{eq:log_likelihood} (i.e.
 the unnormalized posterior distribution over the parameters $\boldsymbol{\theta}$, since we used a uniform prior in this work).
@@
 This gives rise to $W$ parallel chains of samples that mutually influence each others proposal such that the sampling proceedure overall is \emph{affine invariant}.
-This means that scaling the parameter space along any dimension has no effect on the performance of the \gls{mcmc}.
+This means that scaling the parameter space along any dimension has no effect on the performance of the \gls{mcmc} sampling algorithm.
 
@@
 Alternatively, the individual $\hat{f}_i = f\big( \boldsymbol{\hat{\theta}}_i \big)$ can be used to plot histograms over the distribution of $f$.
-We will do so in \cref{sec:results} to show distributions over prevalence predictions and risk computations.
+We will do so in \autoref{sec:results} to show distributions over prevalence predictions and risk computations.
 
@@
     \includegraphics[width=0.6\linewidth]{figures/extended_graph.png}
-    \caption{Extended \gls{dag} representing different possible spread graphs underlying the \gls{hmm}.
-As in \cref{fig:graph_with_obs}, red arcs are parametrized with probabilities of spread from the tumor (red circle) to the \glspl{lnl} (blue circles).
+    \caption[x]{Extended \gls{dag} representing different possible spread graphs underlying the \gls{hmm}.
+As in \autoref{fig:graph_with_obs}, red arcs are parametrized with probabilities of spread from the tumor (red circle) to the \glspl{lnl} (blue circles).
 These red arcs, together with the blue arcs fron \gls{lnl} to \gls{lnl}, make up the \emph{base graph}.
@@
 
-The \gls{dag} shown in \cref{fig:graph_with_obs} includes the \glspl{lnl} II, III, and IV, which represent the most relevant lymph node levels for \gls{opscc}.
+The \gls{dag} shown in \autoref{fig:graph_with_obs} includes the \glspl{lnl} II, III, and IV, which represent the most relevant lymph node levels for \gls{opscc}.
 It includes the arcs from II to III and from III to IV, representing the main direction of lymphatic drainage, which is well motivated anatomically and by the data on lymph node involvement.
@@
 It contains only the connections from \gls{lnl} II to III and an arc from \gls{lnl} III to IV, as motivated by the \emph{main lymphatic pathway} \cite{lengele_anatomical_2007}.
-The base graph is illustrated in \cref{fig:extended_graph} via the red and blue arcs.
+The base graph is illustrated in \autoref{fig:extended_graph} via the red and blue arcs.
 The lymphatic drainage to or from levels I, V, and VII is not as clearly defined.
-Therefore, we define a set of candidate arcs (green arcs in \cref{fig:extended_graph}) and use the model comparison methodology described in \cref{subsec:complete_model:comparison} to determine which graph is most supported our data.
+Therefore, we define a set of candidate arcs (green arcs in \autoref{fig:extended_graph}) and use the model comparison methodology described in \autoref{subsec:complete_model:comparison} to determine which graph is most supported our data.
 
@@
 
-To determine the optimal graph, we first consider six models, each with one of the six green candidate arcs of \cref{fig:extended_graph} added to the base graph.
-Every one of these models was evaluated by computing its evidence via thermodynamic integration as described below in \cref{subsec:complete_model:comparison}.
-%These evidences are shown, among others, in \cref{table:evidence}. 
+To determine the optimal graph, we first consider six models, each with one of the six green candidate arcs of \autoref{fig:extended_graph} added to the base graph.
+Every one of these models was evaluated by computing its evidence via thermodynamic integration as described below in \autoref{subsec:complete_model:comparison}.
+%These evidences are shown, among others, in \autoref{table:evidence}. 
 Subsequently, graphs combining multiple arcs that individually improve the model evidence are considered.
 Thereby, the ``winning graph'' is determined, which yields the highest value of the model evidence.
-%which is in great detail compared to the base graph in the results (\cref{sec:results}). All considered graphs' performances are visualized in \cref{fig:ranking}.
+%which is in great detail compared to the base graph in the results (\autoref{sec:results}). All considered graphs' performances are visualized in \autoref{fig:ranking}.
 
@@
     \includegraphics[width=\linewidth]{figures/thermo_int.png}
-    \caption{The expected log-accuracy under the power posterior plotted against the value of the inverse temperature $\beta$.
+    \caption[x]{The expected log-accuracy under the power posterior plotted against the value of the inverse temperature $\beta$.
 These values were computed during a \gls{ti} run for the loosing graph model (red line), as well as the winning graph model (green line).
@@
 However, already before $\beta=0.031$ the model based on the winning graph outperforms the loosing graph model.
-Ultimately, the winning graph's log-evidence is better than the loosing graph model's by a value of around \variable{output/evidence_diff.txt} (see \cref{table:evidence}).}
+Ultimately, the winning graph's log-evidence is better than the loosing graph model's by a value of around \variable{output/evidence_diff.txt} (see \autoref{table:evidence}).}
     \label{fig:thermo_int}
@@
     \end{tabular}
-    \caption{Interpretation of Bayes factors and their natural logarithms in terms of their support for or against one of the two compared models as introduced by Harold Jeffreys \cite{jeffreys_theory_1998}.}
+    \caption[x]{Interpretation of Bayes factors and their natural logarithms in terms of their support for or against one of the two compared models as introduced by Harold Jeffreys \cite{jeffreys_theory_1998}.}
     \label{table:bayes_factor}
@@
 If we assume all models $\mathcal{M}_i$ for $i \in \{ 1,2 \}$ to have the same \emph{a priori} probability -- meaning in this case $P (\mathcal{M}_1) = P (\mathcal{M}_2)$ -- then we can compute the so-called \emph{Bayes factor} of the two models as the ratio of their likelihoods.
-The interpretation of the values for different Bayes factors is given in \cref{table:bayes_factor}.
+The interpretation of the values for different Bayes factors is given in \autoref{table:bayes_factor}.
 It is defined as follows:
@@
 %
-The integral in \cref{eq:ti} can then be computed via a trapezoidal rule using the $\mathcal{A}_\text{MC}$ to yield a  numerical approximation of the model evidence:
+The integral in \autoref{eq:ti} can then be computed via a trapezoidal rule using the $\mathcal{A}_\text{MC}$ to yield a  numerical approximation of the model evidence:
 %
@@
 The process of computing the log-evidence using \gls{ti} was as follows: We randomly initialized the starting positions of the $W$ samplers in the ensemble within the $k$ dimensional unit cube $\boldsymbol{\Theta}$.
-Subsequently, for each $j \in \{ 0, 1, 2, \ldots, R=63 \}$ we drew samples from the corresponding power posterior with the value of $\beta_j$ set according to the power rule in \cref{eq:power_rule}.
+Subsequently, for each $j \in \{ 0, 1, 2, \ldots, R=63 \}$ we drew samples from the corresponding power posterior with the value of $\beta_j$ set according to the power rule in \autoref{eq:power_rule}.
 This sampling at point $j$ consisted of 1000 burn-in steps, followed by 200 steps, of which only every tenth was kept.
@@
 An approximation to the evidence that also attempts to balance accuracy and complexity against each other is the heuristic called \gls{bic}.
-The negative one half of the \gls{bic} approximates the $\ln E$ via Lagrange's method \cite{bhat_derivation_2010} and yields an easy to compute estimate that may also be used to compare models, as long as its underlying assumptions are valid:
+The negative one half of the \gls{bic} approximates the $\ln E$ via Lagrange's method \cite{bhat_\text{D}erivation_2010} and yields an easy to compute estimate that may also be used to compare models, as long as its underlying assumptions are valid:
 %
@@
     \variable{output/data_table.tex}
-    \caption{Prevalence of involvement patterns in the multi-centric dataset.
+    \caption[x]{Prevalence of involvement patterns in the multi-centric dataset.
 An involvement pattern is characterized by the state of the six \glspl{lnl}: A red dot means the \gls{lnl} was reported to be metastatic, a green dot means it was determined to be healthy and a question mark means that the prevalence was marginalized over the state of this \gls{lnl}.}
@@
 When different modalities were conflicting, the conflicts were resolved by inferring the most likely state (healthy or metastatic) for each \gls{lnl} separately.
-To do so, we used literature values for the sensitivity and specificity of the diagnostic modalities \cite{de_bondt_detection_2007,kyzas_18f-fluorodeoxyglucose_2008}, which we also tabulated in \cref{table:sens_spec}.
+To do so, we used literature values for the sensitivity and specificity of the diagnostic modalities \cite{de_bondt_\text{D}etection_2007,kyzas_18f-fluorodeoxyglucose_2008}, which we also tabulated in \autoref{table:sens_spec}.
 Practically, this means that, whenever pathology after neck dissection was available, the pathology result was taken as the consensus, overruling any other clinical diagnostic modality.
@@
 The dataset containing the consensus decision for the involvement of each level in every patient was then used for model parameter learning, assuming that it represents an observation of the true hidden state $\boldsymbol{\xi}_k$.
-The frequencies of some of the most important combinations of involved lymph node levels are listed in \cref{table:data_prevalence}.
+The frequencies of some of the most important combinations of involved lymph node levels are listed in \autoref{table:data_prevalence}.
 
@@
     \end{tabular}
-    \caption{Literature sensitivity and specificity values that we used to infer the most likely involvement for a patient when multiple diagnostic modalities reported conflicting nodal involvement \cite{de_bondt_detection_2007,kyzas_18f-fluorodeoxyglucose_2008}.}
+    \caption[x]{Literature sensitivity and specificity values that we used to infer the most likely involvement for a patient when multiple diagnostic modalities reported conflicting nodal involvement \cite{de_bondt_\text{D}etection_2007,kyzas_18f-fluorodeoxyglucose_2008}.}
     \label{table:sens_spec}
@@
 
-\section{Results}
-\label{sec:results}
+\section{Results} \label{sec:results}
 
@@
         \includegraphics[width=\textwidth]{figures/bg_core_prevs.png}
-        \caption{Prevalence of involvement as predicted by the base graph model for different scenarios involving the most commonly metastatic \glspl{lnl} II, III and IV (shaded histograms).
+        \caption[x]{Prevalence of involvement as predicted by the base graph model for different scenarios involving the most commonly metastatic \glspl{lnl} II, III and IV (shaded histograms).
 The model's predictions are compared to Beta posteriors over the prevalence based on the frequency of the same scenarios and a uniform prior (slid lines).
@@
 
-For the base graph, we have plotted the predicted prevalence of involvement patterns in the investigated patient cohort for scenarios involving the most commonly metastatic \glspl{lnl} II, III and IV in \cref{fig:bg_prevalences}.
+For the base graph, we have plotted the predicted prevalence of involvement patterns in the investigated patient cohort for scenarios involving the most commonly metastatic \glspl{lnl} II, III and IV in \autoref{fig:bg_prevalences}.
 It shows -- for each pattern of lymphatic involvement -- two plots overlaid:
@@
 
-\Cref{fig:bg_prevalences} shows that this minimal graph is already capable of describing the most important parts of the observed data very well.
+\autoref{fig:bg_prevalences} shows that this minimal graph is already capable of describing the most important parts of the observed data very well.
 Notably, the model is not only accurate in its predictions, it also correctly estimates the variance stemming from the limited amount of data.
@@
 
-The model evidences of all candidate graphs are reported in \cref{table:evidence}.
-A visual ranking is provided in \cref{fig:ranking}.
+The model evidences of all candidate graphs are reported in \autoref{table:evidence}.
+A visual ranking is provided in \autoref{fig:ranking}.
 Let us first consider the six models in which one of the candidate arcs is added to the base graph.
@@
 Based on these results, we consider three additional candidates for the optimal graph that combine the added arc from \gls{lnl} I to II with the arc(s) $\text{III} \rightarrow \text{V}$ and/or $\text{IV} \rightarrow \text{V}$.
-The model evidence for these three graphs is reported in \cref{table:evidence}.
+The model evidence for these three graphs is reported in \autoref{table:evidence}.
 The best performing graph with decisive evidence over the base graph turned out to be the one which combines the arcs from \gls{lnl} I to II and IV to V.
@@
 
-\Cref{table:evidence} additionally shows the evidence of graphs in which the arc from level II to III and from level III to IV is removed.
+\autoref{table:evidence} additionally shows the evidence of graphs in which the arc from level II to III and from level III to IV is removed.
 The low model evidence for these graphs confirms the importance of these connections and is consistent with the anatomical motivation.
@@
     \includegraphics[width=\linewidth]{figures/ranking.pdf}
-    \caption{Visual ranking of the investigated graphs w.r.t.
+    \caption[x]{Visual ranking of the investigated graphs w.r.t.
 their model evidence, computed via thermodynamic integration.
@@
 Their respective model evidence fell below $-1528$ and the two graphs would appear far left in the figure.
-In the bottom left corner, we provide a visual reference in analogy to \cref{table:bayes_factor}: E.g., any difference in the model evidence shorter than the first of the three rulers indicates that the improvement is "barely worth a mention".}
+In the bottom left corner, we provide a visual reference in analogy to \autoref{table:bayes_factor}: E.g., any difference in the model evidence shorter than the first of the three rulers indicates that the improvement is "barely worth a mention".}
     \label{fig:ranking}
@@
     \variable{output/results_table.tex}
-    \caption{Model comparison results from the base graph and the extended graph we chose as the ``winning'' model.
+    \caption[x]{Model comparison results from the base graph and the extended graph we chose as the ``winning'' model.
 For both \glspl{dag} we show the log-evidence, computed via thermodynamic integration, the negative one half of the \gls{bic}, as well as the maximum log-likelihood that was encountered during the final \gls{mcmc} sampling round.}
@@
 
-The most likely model parameters for the winning graph, corresponding to the mean of the marginals of the sampled posterior distributions, are tabulated in \cref{table:params}.
+The most likely model parameters for the winning graph, corresponding to the mean of the marginals of the sampled posterior distributions, are tabulated in \autoref{table:params}.
 We have fixed $p_\text{early}=0.3$ for early T-category tumors (i.e., T0, T1, and T2), and $t_\text{max}=10$ time steps.
@@
 %
-which agrees with the second panel from the top in \cref{fig:bg_prevalences}. 
+which agrees with the second panel from the top in \autoref{fig:bg_prevalences}. 
 The large value for the parameter $t_{\text{I} \rightarrow \text{II}}$ reflects the observation that in almost all patients with level I involvement, level II is also involved.
@@
     \variable{output/wg_means.tex}
-    \caption{Mean of sampled parameters in percent.}
+    \caption[x]{Mean of sampled parameters in percent.}
     \label{table:params}
@@
         \includegraphics[width=\linewidth]{figures/comp_IIIandIVandV_prevs.png}
-        \caption{Observed (Beta posteriors as lines) vs.
+        \caption[x]{Observed (Beta posteriors as lines) vs.
 predicted (histograms) prevalences of involvement combinations that include \gls{lnl} V.
@@
 
-{\it Level V:} In \cref{fig:IIIandIVandV_prevs} we compare the base graph's and the winning graph's estimations for prevalences of involvement patterns that include \gls{lnl} V.
+{\it Level V:} In \autoref{fig:IIIandIVandV_prevs} we compare the base graph's and the winning graph's estimations for prevalences of involvement patterns that include \gls{lnl} V.
 The base graph underestimates the probability that level IV and V are simultaneously involved, and overestimates that probability the level V but not IV is involved.
@@
 
-{\it Level I:} In \cref{fig:IandII_prevs}, analogous comparisons are shown for involvement patterns that include \gls{lnl} I.
+{\it Level I:} In \autoref{fig:IandII_prevs}, analogous comparisons are shown for involvement patterns that include \gls{lnl} I.
 The base graph overestimates the probability of level I involvement without simultaneous involvement of level II.
@@
         \includegraphics[width=\linewidth]{figures/comp_IandII_prevs.png}
-        \caption{Comparison of observed and predicted prevalences of \gls{lnl} I and II involvement patterns.
+        \caption[x]{Comparison of observed and predicted prevalences of \gls{lnl} I and II involvement patterns.
 The top and bottom panels show the prevalences for early and late T-category, respectively.
@@
 In this section, the model corresponding to the winning graph is applied to estimating the risk of occult metastases in clinically negative \glspl{lnl}.
-We assume a sensitivity of 0.76 and a specificity of 0.81 for the clinical diagnosis of lymph node metastases, corresponding to CT imaging in \cref{table:sens_spec}.
+We assume a sensitivity of 0.76 and a specificity of 0.81 for the clinical diagnosis of lymph node metastases, corresponding to CT imaging in \autoref{table:sens_spec}.
 
-{\it Level II:} As can be seen in \cref{table:params}, spread from the tumor to \gls{lnl} II to be the most probable transition at any given time step.
+{\it Level II:} As can be seen in \autoref{table:params}, spread from the tumor to \gls{lnl} II to be the most probable transition at any given time step.
 As a consequence, even for an early T-category patient that presents with a clinical N0 neck, our model predicts a \variable{output/wg_II_risks.tex} risk for microscopic metastases in \gls{lnl} II.
 
-{\it Level III:} \Cref{fig:wg_III_risks} compares the risk of occult disease in level III between patients that are clinically N0 (orange) and patients with clinically diagnosed involvement of only level II (red), for early T-category (upper panel) and advanced T-category (bottom panel).
+{\it Level III:} \autoref{fig:wg_III_risks} compares the risk of occult disease in level III between patients that are clinically N0 (orange) and patients with clinically diagnosed involvement of only level II (red), for early T-category (upper panel) and advanced T-category (bottom panel).
 The histograms represent the uncertainty in model's risk prediction arising from the uncertainty in the model parameters and are generated by randomly drawing 1000 samples from the model parameter's joint posterior distribution.
@@
         \includegraphics[width=\linewidth]{figures/wg_III_risks.png}
-        \caption{Histograms over the risk for microscopic involvement in \gls{lnl} III, given that a patient presents as clinically N0 (green), or given that \gls{lnl} II shows clinical involvement (blue).
+        \caption[x]{Histograms over the risk for microscopic involvement in \gls{lnl} III, given that a patient presents as clinically N0 (green), or given that \gls{lnl} II shows clinical involvement (blue).
 The top panel shows these risks for early T-category, the bottom panel for late.}
@@
 
-{\it Level IV:} \Cref{fig:wg_IV_risks} compares the risk of occult disease in level IV for the typical clinical presentations: clinically N0 (green), metastases in level II (blue), and metastases in levels II and III (orange).
+{\it Level IV:} \autoref{fig:wg_IV_risks} compares the risk of occult disease in level IV for the typical clinical presentations: clinically N0 (green), metastases in level II (blue), and metastases in levels II and III (orange).
 The model predicts a low risk of 1-2\% in level IV for patients with clinically healthy level III.
@@
         \includegraphics[width=\linewidth]{figures/wg_IV_risks.png}
-        \caption{Distributions over the risk for microscopic involvement in \gls{lnl} IV as predicted by the winning graph model, given early (top panel) or late T-category (bottom panel), and three different CT-based diagnoses:
+        \caption[x]{Distributions over the risk for microscopic involvement in \gls{lnl} IV as predicted by the winning graph model, given early (top panel) or late T-category (bottom panel), and three different CT-based diagnoses:
         \begin{enumerate*}[label=\arabic*)]
@@
 
-{\it Level V:} \Cref{fig:wg_V_risks} shows the risk of occult disease in level V depending on T-category and the clinical involvement of levels II-IV.
+{\it Level V:} \autoref{fig:wg_V_risks} shows the risk of occult disease in level V depending on T-category and the clinical involvement of levels II-IV.
 For clinically N0 patients, the risk in level V is estimated to be around 1\%.
@@
         \includegraphics[width=\linewidth]{figures/wg_V_risks.png}
-        \caption{The winning graph's predictions over the risk for occult disease in \gls{lnl} V for early (top panel) and late T-category (bottom panel).
+        \caption[x]{The winning graph's predictions over the risk for occult disease in \gls{lnl} V for early (top panel) and late T-category (bottom panel).
 Each of the risks is computed, given one of three diagnostic scenarios that were obtained by a CT scan:
@@
 
-{\it Level I:} \Cref{fig:wg_I_risks} shows the risk of occult disease in level I depending on T-category and the clinical involvement of levels II-IV.
+{\it Level I:} \autoref{fig:wg_I_risks} shows the risk of occult disease in level I depending on T-category and the clinical involvement of levels II-IV.
 For clinically N0 patients, the risk in level I is estimated to be in the order of 1-2\%.
@@
         \includegraphics[width=\linewidth]{figures/wg_I_risks.png}
-        \caption{Distributions over predicted risk for involvement in \gls{lnl} I, given different clinical diagnosis scenarios: For N0 patients (green), patients with macroscopic involvement in \gls{lnl} II (blue), and for the case where the \glspl{lnl} II, III, and IV show involvement.
+        \caption[x]{Distributions over predicted risk for involvement in \gls{lnl} I, given different clinical diagnosis scenarios: For N0 patients (green), patients with macroscopic involvement in \gls{lnl} II (blue), and for the case where the \glspl{lnl} II, III, and IV show involvement.
 The top panel shows these risks for early T-category and the bottom row for late T-category.}
@@
         \includegraphics[width=\linewidth]{figures/wg_any_risks.png}
-        \caption{Shown are the histograms over the predicted risk for involvement in any of the \glspl{lnl} I, IV, V, or VII.
+        \caption[x]{Shown are the histograms over the predicted risk for involvement in any of the \glspl{lnl} I, IV, V, or VII.
 The risk is plotted given a clinical N0 diagnosis (green), macroscopically detected metastases in \gls{lnl} II (blue), and lastly given visible involvement in both \gls{lnl} II and III (orange).
@@
 In the context of radiotherapy, this corresponds to excluding \glspl{lnl} from the \gls{ctv-n}, which are irradiated according to the current guidelines.
-The list below should be seen as a summary of \cref{subsec:results:risk_prediction} and the limitations discussed in \cref{subsec:disc:limitation} should be taken into account in its interpretation.
+The list below should be seen as a summary of \autoref{subsec:results:risk_prediction} and the limitations discussed in \autoref{subsec:disc:limitation} should be taken into account in its interpretation.
 Assuming that one accepts approximately a 5\% risk of occult metastases per \gls{lnl}, the statistical model presented in this paper would suggest to:
@@
 
-{\it T-category dependence of level I involvement:} As shown in \cref{fig:bg_prevalences}, the model describes the involvement of \glspl{lnl} II, III, and IV depending on T-stage very well despite having only a single parameter related to T-stage.
-\Cref{fig:IandII_prevs} shows that the model does not perfectly describe the T-stage dependence of level I.
+{\it T-category dependence of level I involvement:} As shown in \autoref{fig:bg_prevalences}, the model describes the involvement of \glspl{lnl} II, III, and IV depending on T-stage very well despite having only a single parameter related to T-stage.
+\autoref{fig:IandII_prevs} shows that the model does not perfectly describe the T-stage dependence of level I.
 It adjusts the parameters such that level I involvement is correctly described for the set of all patients combined, but it overestimates level I involvement for early T-stage and underestimates it for advanced T-stage.
@@
 Consequently, these values need to be considered with caution.
-\Cref{fig:wg_sens_spec_risks} illustrates for one example how the risk of occult disease depends on sensitivity and specificity.
+\autoref{fig:wg_sens_spec_risks} illustrates for one example how the risk of occult disease depends on sensitivity and specificity.
 Here, we consider the risk in level IV in patients with clinically detected metastases in levels II and III.
@@
         \includegraphics[width=\linewidth]{figures/wg_sens_spec_risks.png}
-        \caption{Dependence of risk of occult disease on sensitivity and specificity.
+        \caption[x]{Dependence of risk of occult disease on sensitivity and specificity.
 The figure shows the risk of involvement in level IV, given the mean of the parameter samples, drawn during sampling, in patients with clinical involvement of levels II and III.}

### 2023-12-14 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     \includegraphics[width=0.6\linewidth]{figures/graph_with_obs.png}
-    \caption[x]{\Gls{dag} representing a possible abstraction of the lymphatic network comprising the tumor (red shaded circle) and \glspl{lnl} II through IV as hidden binary \glspl{rv} (blue outlined circles).
+    \caption{\Gls{dag} representing a possible abstraction of the lymphatic network comprising the tumor (red shaded circle) and \glspl{lnl} II through IV as hidden binary \glspl{rv} (blue outlined circles).
 Attached to each of these is the corresponding observed \gls{rv} (orange shaded squares).
@@
     \includegraphics[width=0.6\linewidth]{figures/extended_graph.png}
-    \caption[x]{Extended \gls{dag} representing different possible spread graphs underlying the \gls{hmm}.
+    \caption{Extended \gls{dag} representing different possible spread graphs underlying the \gls{hmm}.
 As in \autoref{fig:graph_with_obs}, red arcs are parametrized with probabilities of spread from the tumor (red circle) to the \glspl{lnl} (blue circles).
@@
     \includegraphics[width=\linewidth]{figures/thermo_int.png}
-    \caption[x]{The expected log-accuracy under the power posterior plotted against the value of the inverse temperature $\beta$.
+    \caption{The expected log-accuracy under the power posterior plotted against the value of the inverse temperature $\beta$.
 These values were computed during a \gls{ti} run for the loosing graph model (red line), as well as the winning graph model (green line).
@@
     \end{tabular}
-    \caption[x]{Interpretation of Bayes factors and their natural logarithms in terms of their support for or against one of the two compared models as introduced by Harold Jeffreys \cite{jeffreys_theory_1998}.}
+    \caption{Interpretation of Bayes factors and their natural logarithms in terms of their support for or against one of the two compared models as introduced by Harold Jeffreys \cite{jeffreys_theory_1998}.}
     \label{table:bayes_factor}
@@
 An approximation to the evidence that also attempts to balance accuracy and complexity against each other is the heuristic called \gls{bic}.
-The negative one half of the \gls{bic} approximates the $\ln E$ via Lagrange's method \cite{bhat_\text{D}erivation_2010} and yields an easy to compute estimate that may also be used to compare models, as long as its underlying assumptions are valid:
+The negative one half of the \gls{bic} approximates the $\ln E$ via Lagrange's method \cite{bhat_Derivation_2010} and yields an easy to compute estimate that may also be used to compare models, as long as its underlying assumptions are valid:
 %
@@
     \variable{output/data_table.tex}
-    \caption[x]{Prevalence of involvement patterns in the multi-centric dataset.
+    \caption{Prevalence of involvement patterns in the multi-centric dataset.
 An involvement pattern is characterized by the state of the six \glspl{lnl}: A red dot means the \gls{lnl} was reported to be metastatic, a green dot means it was determined to be healthy and a question mark means that the prevalence was marginalized over the state of this \gls{lnl}.}
@@
 When different modalities were conflicting, the conflicts were resolved by inferring the most likely state (healthy or metastatic) for each \gls{lnl} separately.
-To do so, we used literature values for the sensitivity and specificity of the diagnostic modalities \cite{de_bondt_\text{D}etection_2007,kyzas_18f-fluorodeoxyglucose_2008}, which we also tabulated in \autoref{table:sens_spec}.
+To do so, we used literature values for the sensitivity and specificity of the diagnostic modalities \cite{de_bondt_Detection_2007,kyzas_18f-fluorodeoxyglucose_2008}, which we also tabulated in \autoref{table:sens_spec}.
 Practically, this means that, whenever pathology after neck dissection was available, the pathology result was taken as the consensus, overruling any other clinical diagnostic modality.
@@
     \end{tabular}
-    \caption[x]{Literature sensitivity and specificity values that we used to infer the most likely involvement for a patient when multiple diagnostic modalities reported conflicting nodal involvement \cite{de_bondt_\text{D}etection_2007,kyzas_18f-fluorodeoxyglucose_2008}.}
+    \caption{Literature sensitivity and specificity values that we used to infer the most likely involvement for a patient when multiple diagnostic modalities reported conflicting nodal involvement \cite{de_bondt_Detection_2007,kyzas_18f-fluorodeoxyglucose_2008}.}
     \label{table:sens_spec}
@@
         \includegraphics[width=\textwidth]{figures/bg_core_prevs.png}
-        \caption[x]{Prevalence of involvement as predicted by the base graph model for different scenarios involving the most commonly metastatic \glspl{lnl} II, III and IV (shaded histograms).
+        \caption{Prevalence of involvement as predicted by the base graph model for different scenarios involving the most commonly metastatic \glspl{lnl} II, III and IV (shaded histograms).
 The model's predictions are compared to Beta posteriors over the prevalence based on the frequency of the same scenarios and a uniform prior (slid lines).
@@
     \includegraphics[width=\linewidth]{figures/ranking.pdf}
-    \caption[x]{Visual ranking of the investigated graphs w.r.t.
+    \caption{Visual ranking of the investigated graphs w.r.t.
 their model evidence, computed via thermodynamic integration.
@@
     \variable{output/results_table.tex}
-    \caption[x]{Model comparison results from the base graph and the extended graph we chose as the ``winning'' model.
+    \caption{Model comparison results from the base graph and the extended graph we chose as the ``winning'' model.
 For both \glspl{dag} we show the log-evidence, computed via thermodynamic integration, the negative one half of the \gls{bic}, as well as the maximum log-likelihood that was encountered during the final \gls{mcmc} sampling round.}
@@
     \variable{output/wg_means.tex}
-    \caption[x]{Mean of sampled parameters in percent.}
+    \caption{Mean of sampled parameters in percent.}
     \label{table:params}
@@
         \includegraphics[width=\linewidth]{figures/comp_IIIandIVandV_prevs.png}
-        \caption[x]{Observed (Beta posteriors as lines) vs.
+        \caption{Observed (Beta posteriors as lines) vs.
 predicted (histograms) prevalences of involvement combinations that include \gls{lnl} V.
@@
         \includegraphics[width=\linewidth]{figures/comp_IandII_prevs.png}
-        \caption[x]{Comparison of observed and predicted prevalences of \gls{lnl} I and II involvement patterns.
+        \caption{Comparison of observed and predicted prevalences of \gls{lnl} I and II involvement patterns.
 The top and bottom panels show the prevalences for early and late T-category, respectively.
@@
         \includegraphics[width=\linewidth]{figures/wg_III_risks.png}
-        \caption[x]{Histograms over the risk for microscopic involvement in \gls{lnl} III, given that a patient presents as clinically N0 (green), or given that \gls{lnl} II shows clinical involvement (blue).
+        \caption{Histograms over the risk for microscopic involvement in \gls{lnl} III, given that a patient presents as clinically N0 (green), or given that \gls{lnl} II shows clinical involvement (blue).
 The top panel shows these risks for early T-category, the bottom panel for late.}
@@
         \includegraphics[width=\linewidth]{figures/wg_IV_risks.png}
-        \caption[x]{Distributions over the risk for microscopic involvement in \gls{lnl} IV as predicted by the winning graph model, given early (top panel) or late T-category (bottom panel), and three different CT-based diagnoses:
+        \caption{Distributions over the risk for microscopic involvement in \gls{lnl} IV as predicted by the winning graph model, given early (top panel) or late T-category (bottom panel), and three different CT-based diagnoses:
         \begin{enumerate*}[label=\arabic*)]
@@
         \includegraphics[width=\linewidth]{figures/wg_V_risks.png}
-        \caption[x]{The winning graph's predictions over the risk for occult disease in \gls{lnl} V for early (top panel) and late T-category (bottom panel).
+        \caption{The winning graph's predictions over the risk for occult disease in \gls{lnl} V for early (top panel) and late T-category (bottom panel).
 Each of the risks is computed, given one of three diagnostic scenarios that were obtained by a CT scan:
@@
         \includegraphics[width=\linewidth]{figures/wg_I_risks.png}
-        \caption[x]{Distributions over predicted risk for involvement in \gls{lnl} I, given different clinical diagnosis scenarios: For N0 patients (green), patients with macroscopic involvement in \gls{lnl} II (blue), and for the case where the \glspl{lnl} II, III, and IV show involvement.
+        \caption{Distributions over predicted risk for involvement in \gls{lnl} I, given different clinical diagnosis scenarios: For N0 patients (green), patients with macroscopic involvement in \gls{lnl} II (blue), and for the case where the \glspl{lnl} II, III, and IV show involvement.
 The top panel shows these risks for early T-category and the bottom row for late T-category.}
@@
         \includegraphics[width=\linewidth]{figures/wg_any_risks.png}
-        \caption[x]{Shown are the histograms over the predicted risk for involvement in any of the \glspl{lnl} I, IV, V, or VII.
+        \caption{Shown are the histograms over the predicted risk for involvement in any of the \glspl{lnl} I, IV, V, or VII.
 The risk is plotted given a clinical N0 diagnosis (green), macroscopically detected metastases in \gls{lnl} II (blue), and lastly given visible involvement in both \gls{lnl} II and III (orange).
@@
         \includegraphics[width=\linewidth]{figures/wg_sens_spec_risks.png}
-        \caption[x]{Dependence of risk of occult disease on sensitivity and specificity.
+        \caption{Dependence of risk of occult disease on sensitivity and specificity.
 The figure shows the risk of involvement in level IV, given the mean of the parameter samples, drawn during sampling, in patients with clinical involvement of levels II and III.}

### 2023-12-15 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 This work concerns itself with \gls{opscc}, where approximately 70-80\% of patients present with lymph node metastases at the time of diagnosis.
-In clinical practice, \gls{ctv-n} definition in radiotherapy is based on guidelines \cite{gregoire_ct-based_2003,gregoire_delineation_2014,gregoire_delineation_2018,eisbruch_intensity-modulated_2002,biau_selection_2019,chao_determination_2002,vorwerk_guidelines_2011,ferlito_elective_2009} that are mostly derived from the observed prevalence of involvement in an \gls{lnl} for a given tumor location.
+In clinical practice, \gls{ctv-n} definition in radiotherapy is based on guidelines \citep{gregoire_ct-based_2003,gregoire_delineation_2014,gregoire_delineation_2018,eisbruch_intensity-modulated_2002,biau_selection_2019,chao_determination_2002,vorwerk_guidelines_2011,ferlito_elective_2009} that are mostly derived from the observed prevalence of involvement in an \gls{lnl} for a given tumor location.
 These guidelines currently suggest extensive irradiation of both sides of the neck for most patients.
@@
 We previously developed a model of lymphatic metastatic progression for estimating the risk of microscopic disease, given a patient's personal diagnosis.
-The initial model was based on the methodology of \glspl{bn} \cite{pouymayou_bayesian_2019}.
-It was subsequently extended and formulated as an \gls{hmm} to include T-category in an intuitive manner \cite{ludwig_hidden_2021}.
-However, these models were introduced based on only a small dataset of approximately 100 early T-category patients available at that time \cite{sanguineti_defining_2009}.
+The initial model was based on the methodology of \glspl{bn} \citep{pouymayou_bayesian_2019}.
+It was subsequently extended and formulated as an \gls{hmm} to include T-category in an intuitive manner \citep{ludwig_hidden_2021}.
+However, these models were introduced based on only a small dataset of approximately 100 early T-category patients available at that time \citep{sanguineti_defining_2009}.
 The limited data did not allow us to quantify the probability of metastases in the rarely involved \glspl{lnl} I, V, and VII, nor did the data allow us to verify that the \gls{hmm} is adequate to describe the dependence on lymph node involvement on T-category.
-In this paper, we extend the previous work \cite{ludwig_hidden_2021} by making the following contributions:
+In this paper, we extend the previous work \citep{ludwig_hidden_2021} by making the following contributions:
 \begin{enumerate}
@@
 To determine the optimal underlying \gls{dag} we compare different graphs by calculating the model evidence through \gls{ti}.
-    \item We collect a multi-centric dataset consisting of \variable{output/num_patients.txt} patients from three institutions, allowing us to train the model based on a sizable dataset \cite{ludwig_dataset_2022,ludwig_multi-centric_2023}.
+    \item We collect a multi-centric dataset consisting of \variable{output/num_patients.txt} patients from three institutions, allowing us to train the model based on a sizable dataset \citep{ludwig_dataset_2022,ludwig_multi-centric_2023}.
     \item We use the trained model to provide personalized risk estimations for occult metastases for typical clinical states of tumor progression at diagnosis, illustrating its potential for guiding volume-deescalated treatment strategies in the future.
@@
 
-We have introduced a probabilistic model for lymph node involvement based on \glspl{bn} in \cite{pouymayou_bayesian_2019}.
-The model was extended using \glspl{hmm} in \cite{ludwig_hidden_2021}.
+We have introduced a probabilistic model for lymph node involvement based on \glspl{bn} in \citep{pouymayou_bayesian_2019}.
+The model was extended using \glspl{hmm} in \citep{ludwig_hidden_2021}.
 We will briefly recap the \acrlong{hmm} to introduce the notation used throughout the work.
@@
 
-\noindent
-\begin{center}
-    \begin{tabular}{|cc|cc|}
-        \hline
-        & & \multicolumn{2}{c|}{$X_2$} \\
-        & & 0 & 1 \\
-        \hline
-        \multirow{2}{*}{$X_3$} & 0 & $1 - b_3$ & $(1 - b_3)(1 - t_{23})$ \\
-        & 1 & $b_3$ & $1 - b_3 - t_{23} + b_3 t_{23}$ \\
-        \hline
-    \end{tabular}
-\end{center}
+\begin{deluxetable}{cc|cc}
+\tablecaption{$P\left( X_3 \mid X_2 \right)$ for all possible combinations of $X_3$ and $X_2$.}
+\tablehead{& & \colhead{$X_2=0$} & \colhead{$X_2=1$}}
+\startdata
+\begin{tabular}{@{}c@{}}
+\multirow{2}{*}{$X_3$}
+\end{tabular} & $=0$ & $1-b_3$ & $(1-b_3)(1-t_{23})$ \\
+& $=1$ & $b_3$ & $1-b_3-t_{23}+b_3 t_{23}$ \\
+\enddata
+\end{deluxetable}
 
@@
 Lymphatic flow is depicted in the form of parameterized arrows (red and blue) that represent the probability of spread along the respective arc per time-step.
-Sensitivity and specificity (orange arrows) connect the hidden \glspl{rv} to the diagnosis.}
-    \label{fig:graph_with_obs}
+Sensitivity and specificity (orange arrows) connect the hidden \glspl{rv} to the diagnosis.
+\label{fig:graph_with_obs}}
 \end{figure}
@@
 
-\noindent
-\begin{center}
-    \begin{tabular}{|cc|cc|}
-        \hline
-        & & \multicolumn{2}{c|}{$X$} \\
-        & & 0 & 1 \\
-        \hline
-        \multirow{2}{*}{$Z$} & 0 & $s_p$ & $1 - s_N$ \\
-        & 1 & $1 - s_P$ & $s_N$ \\
-        \hline
-    \end{tabular}
-\end{center}
+\begin{deluxetable}{cc|cc}
+\tablecaption{$P\left( Z \mid X \right)$ for all possible values of $X$ and $Z$.}
+\tablehead{& & \colhead{$X=0$} & \colhead{$X=1$}}
+\startdata
+\begin{tabular}{@{}c@{}}
+\multirow{2}{*}{$Z$}
+\end{tabular} & $=0$ & $s_P$ & $1-s_N$ \\
+& $=1$ & $1-s_P$ & $s_N$ \\
+\enddata
+\end{deluxetable}
 
@@
 In this paper, for some patients involvement information of level VII was missing and hence marginalized over.
-A detailed explanation of this formalism can be found in \cite{zora231470}, section 6.2.7.
+A detailed explanation of this formalism can be found in \citep{zora231470}, section 6.2.7.
 
@@
 
-More specifically, we use the Python implementation \texttt{emcee} \cite{foreman-mackey_emcee_2013} and two sample proposal mechanisms based on differential evolution moves \cite{ter_braak_differential_2008,nelson_run_2013} for sampling.
+More specifically, we use the Python implementation \texttt{emcee} \citep{foreman-mackey_emcee_2013} and two sample proposal mechanisms based on differential evolution moves \citep{ter_braak_differential_2008,nelson_run_2013} for sampling.
 Instead of proposing and then accepting or rejecting individual parameter samples one after the other (as in the classical Metropolis-Hastings algorithm), the \texttt{emcee} implementation makes use of an ensemble of $W$ so-called ``walkers''.
@@
 One after the other, each of the green arcs was added to the base graph.
-Subsequently, the performance of the resulting models in terms of its \gls{bic} was compared to the base graph to assess whether the additional edge should be kept in the \emph{winning graph} or not.}
-    \label{fig:extended_graph}
+Subsequently, the performance of the resulting models in terms of its \gls{bic} was compared to the base graph to assess whether the additional edge should be kept in the \emph{winning graph} or not.
+\label{fig:extended_graph}}
 \end{figure}
@@
 It includes the arcs from II to III and from III to IV, representing the main direction of lymphatic drainage, which is well motivated anatomically and by the data on lymph node involvement.
-Previous publications  \cite{pouymayou_bayesian_2019,ludwig_hidden_2021} focused on these levels because they relied on a limited reconstructed dataset of \gls{opscc} patients \cite{sanguineti_defining_2009}.
+Previous publications  \citep{pouymayou_bayesian_2019,ludwig_hidden_2021} focused on these levels because they relied on a limited reconstructed dataset of \gls{opscc} patients \citep{sanguineti_defining_2009}.
 Now, with the datasets available for this work, we can extend the graph to include random variables for all \glspl{lnl} that are relevant for \gls{opscc}: I, II, III, IV, V, and VII.
@@
 To investigate which connections to add we start by establishing a baseline from a model using a minimal \emph{base graph}.
-It contains only the connections from \gls{lnl} II to III and an arc from \gls{lnl} III to IV, as motivated by the \emph{main lymphatic pathway} \cite{lengele_anatomical_2007}.
+It contains only the connections from \gls{lnl} II to III and an arc from \gls{lnl} III to IV, as motivated by the \emph{main lymphatic pathway} \citep{lengele_anatomical_2007}.
 The base graph is illustrated in \autoref{fig:extended_graph} via the red and blue arcs.
@@
 To connect level I, we investigate two candidate arcs: from I to II and from II to I.
-An arc from I to II was used in \cite{pouymayou_bayesian_2019,ludwig_hidden_2021} and is anatomically motivated.
+An arc from I to II was used in \citep{pouymayou_bayesian_2019,ludwig_hidden_2021} and is anatomically motivated.
 However, since \gls{lnl} I is rarely involved compared to level II, the associated parameter $t_{12}$ is mostly undetermined.
@@
 
-Anatomically, the \emph{posterior accessory pathway} that drains \gls{lnl} II through \gls{lnl} V motivates investigating and arc from $X_2$ into $X_5$ \cite{lengele_anatomical_2007}.
+Anatomically, the \emph{posterior accessory pathway} that drains \gls{lnl} II through \gls{lnl} V motivates investigating and arc from $X_2$ into $X_5$ \citep{lengele_anatomical_2007}.
 And, although no lymphatic pathway is described that directly drains \gls{lnl} III or IV into level V, due to their proximity to each other, we will also investigate additional arcs from \gls{lnl} III and IV into level V.
@@
 However, already before $\beta=0.031$ the model based on the winning graph outperforms the loosing graph model.
-Ultimately, the winning graph's log-evidence is better than the loosing graph model's by a value of around \variable{output/evidence_diff.txt} (see \autoref{table:evidence}).}
-    \label{fig:thermo_int}
+Ultimately, the winning graph's log-evidence is better than the loosing graph model's by a value of around \variable{output/evidence_diff.txt} (see \autoref{table:evidence}).
+\label{fig:thermo_int}}
 \end{figure}
 
-\begin{table}
-    \centering
-    \begin{tabular}{ | c | c | c | }
-        \hline
-        $K_\text{1v2}$ & $\ln{K_\text{1v2}}$ & support for $\mathcal{M}_1$ \\
-        \hline
+\begin{deluxetable}{c|c|c}
+\tablecaption{
+    Interpretation of Bayes factors and their natural logarithms in terms of their support for or against one of the two compared models as introduced by \citet{jeffreys_theory_1998}.%
+    \label{table:bayes_factor}%
+}
+\tablehead{
+    \colhead{$K_\text{1v2}$} & \colhead{$\ln{K_\text{1v2}}$} & \colhead{support for $\mathcal{M}_1$}
+}
+    \startdata
     $< 10^0$ & $< 0$ & negative (supports $\mathcal{M}_2$) \\
@@
     $> 10^2$ & $> 4.6$ & decisive \\
-        \hline
-    \end{tabular}
-    \caption{Interpretation of Bayes factors and their natural logarithms in terms of their support for or against one of the two compared models as introduced by Harold Jeffreys \cite{jeffreys_theory_1998}.}
-    \label{table:bayes_factor}
-\end{table}
+    \enddata
+\end{deluxetable}
 
@@
 
-Below, we will briefly outline the main concept behind this algorithm. An intuitive and extensive derivation of \gls{ti} is given by \cite{aponte_introduction_2022}.
+Below, we will briefly outline the main concept behind this algorithm. An intuitive and extensive derivation of \gls{ti} is given by \citep{aponte_introduction_2022}.
 
@@
 An approximation to the evidence that also attempts to balance accuracy and complexity against each other is the heuristic called \gls{bic}.
-The negative one half of the \gls{bic} approximates the $\ln E$ via Lagrange's method \cite{bhat_Derivation_2010} and yields an easy to compute estimate that may also be used to compare models, as long as its underlying assumptions are valid:
+The negative one half of the \gls{bic} approximates the $\ln E$ via Lagrange's method \citep{bhat_Derivation_2010} and yields an easy to compute estimate that may also be used to compare models, as long as its underlying assumptions are valid:
 %
@@
 
-\begin{table}
-    \centering
+% \begin{table}
+%     \centering
+%     \variable{output/data_table.tex}
+%     \caption{Prevalence of involvement patterns in the multi-centric dataset.
+An involvement pattern is characterized by the state of the six \glspl{lnl}: A red dot means the \gls{lnl} was reported to be metastatic, a green dot means it was determined to be healthy and a question mark means that the prevalence was marginalized over the state of this \gls{lnl}.}
+%     \label{table:data_prevalence}
+% \end{table}
+
+\begin{deluxetable}{|cccccc|rr|rr|}
+    \tablecaption{Prevalence of involvement patterns in the multi-centric dataset.
+An involvement pattern is characterized by the state of the six \glspl{lnl}: A red dot means the \gls{lnl} was reported to be metastatic, a green dot means it was determined to be healthy and a question mark means that the prevalence was marginalized over the state of this \gls{lnl}.
+\label{table:data_prevalence}}
+    \tablehead{
+        \multicolumn{6}{|c|}{\textbf{LNL involvement} } & \multicolumn{4}{c|}{\textbf{T-category} } \\
+        I & II & III & IV & V & VII & \multicolumn{2}{c}{\textbf{early} } & \multicolumn{2}{c|}{\textbf{late} } \\
+    }
+    \startdata
     \variable{output/data_table.tex}
-    \caption{Prevalence of involvement patterns in the multi-centric dataset.
-An involvement pattern is characterized by the state of the six \glspl{lnl}: A red dot means the \gls{lnl} was reported to be metastatic, a green dot means it was determined to be healthy and a question mark means that the prevalence was marginalized over the state of this \gls{lnl}.}
-    \label{table:data_prevalence}
-\end{table}
+    \enddata
+\end{deluxetable}
 
 The dataset $\boldsymbol{\mathcal{D}}$ that we used for inference is comprised of the detailed reports on lymph node involvement patterns in \gls{opscc} patients treated at three different institutions in France and Switzerland: The \gls{clb} in Lyon (France), the \gls{isb} in Bern (Switzerland), and the \gls{usz} in Zürich (Switzerland).
-We have previously published the patterns of nodal involvement for the \gls{usz} cohort (287 patients) \cite{ludwig_dataset_2022,ludwig_lydata_2023} and described its characteristics in detail \cite{ludwig_detailed_2022}.
-The first \gls{clb} dataset (263 patients) \cite{ludwig_lydata_2022} underlies a publication on \gls{hpv} status in \gls{opscc} \cite{bauwens_prevalence_2021} and is made available in a separate ``Data in Brief'' article alongside the second dataset from France \cite{ludwig_lydata_2023-2} and the lymphatic progression patterns from the \gls{isb} \cite{ludwig_lydata_2023-1,roman_ludwig_multi-centric_nodate}.
+We have previously published the patterns of nodal involvement for the \gls{usz} cohort (287 patients) \citep{ludwig_dataset_2022,ludwig_lydata_2023} and described its characteristics in detail \citep{ludwig_detailed_2022}.
+The first \gls{clb} dataset (263 patients) \citep{ludwig_lydata_2022} underlies a publication on \gls{hpv} status in \gls{opscc} \citep{bauwens_prevalence_2021} and is made available in a separate ``Data in Brief'' article alongside the second dataset from France \citep{ludwig_lydata_2023-2} and the lymphatic progression patterns from the \gls{isb} \citep{ludwig_lydata_2023-1,roman_ludwig_multi-centric_nodate}.
 
@@
 When different modalities were conflicting, the conflicts were resolved by inferring the most likely state (healthy or metastatic) for each \gls{lnl} separately.
-To do so, we used literature values for the sensitivity and specificity of the diagnostic modalities \cite{de_bondt_Detection_2007,kyzas_18f-fluorodeoxyglucose_2008}, which we also tabulated in \autoref{table:sens_spec}.
+To do so, we used literature values for the sensitivity and specificity of the diagnostic modalities \citep{de_bondt_Detection_2007,kyzas_18f-fluorodeoxyglucose_2008}, which we also tabulated in \autoref{table:sens_spec}.
 Practically, this means that, whenever pathology after neck dissection was available, the pathology result was taken as the consensus, overruling any other clinical diagnostic modality.
@@
 
-\begin{table}
-\centering
-    \begin{tabular}{|l|rr|}
-        \hline
-        \textbf{Modality} & \textbf{Specificity} & \textbf{Sensitivity} \\
-        \hline
+\begin{deluxetable}{l|rr}
+    \tablecaption{
+        Literature sensitivity and specificity values that we used to infer the most likely involvement for a patient when multiple diagnostic modalities reported conflicting nodal involvement \citep{de_bondt_Detection_2007,kyzas_18f-fluorodeoxyglucose_2008}.
+        \label{table:sens_spec}
+    }
+    \tablehead{
+        \textbf{Modality} & \textbf{Specificity} & \textbf{Sensitivity}
+    }
+    \startdata
         \acrshort{ct} & 76\% & 81\% \\
@@
         Pathology & $\approx$ 100\% & $\approx$ 100\% \\
-        \hline
-    \end{tabular}
-    \caption{Literature sensitivity and specificity values that we used to infer the most likely involvement for a patient when multiple diagnostic modalities reported conflicting nodal involvement \cite{de_bondt_Detection_2007,kyzas_18f-fluorodeoxyglucose_2008}.}
-    \label{table:sens_spec}
-\end{table}
+    \enddata
+\end{deluxetable}
 
@@
 The model's predictions are compared to Beta posteriors over the prevalence based on the frequency of the same scenarios and a uniform prior (slid lines).
-The top panel shows some selected scenarios with early T-category tumors and the bottom panel with late T-category.}
-        \label{fig:bg_prevalences}
+The top panel shows some selected scenarios with early T-category tumors and the bottom panel with late T-category.
+\label{fig:bg_prevalences}}
     \end{centering}
@@
 Their respective model evidence fell below $-1528$ and the two graphs would appear far left in the figure.
-In the bottom left corner, we provide a visual reference in analogy to \autoref{table:bayes_factor}: E.g., any difference in the model evidence shorter than the first of the three rulers indicates that the improvement is "barely worth a mention".}
-    \label{fig:ranking}
+In the bottom left corner, we provide a visual reference in analogy to \autoref{table:bayes_factor}: E.g., any difference in the model evidence shorter than the first of the three rulers indicates that the improvement is "barely worth a mention".
+\label{fig:ranking}}
 \end{figure*}
@@
 The left two panels consider combinations of \gls{lnl} III and V involvement, while the right two panels consider combinations of \gls{lnl} IV and V.
-The colored lines show the Beta posterior over the prevalence of the respective involvement pattern, given the data}
-        \label{fig:IIIandIVandV_prevs}
+The colored lines show the Beta posterior over the prevalence of the respective involvement pattern, given the data.
+\label{fig:IIIandIVandV_prevs}}
     \end{centering}
@@
 Green plots indicate \gls{lnl} I involvement without level II, while orange plots indicate the opposite (\gls{lnl} II without level I).
-The winning graph has an added edge from \gls{lnl} I to II, which improves the prediction of the rare green pattern.}
-        \label{fig:IandII_prevs}
+The winning graph has an added edge from \gls{lnl} I to II, which improves the prediction of the rare green pattern.
+\label{fig:IandII_prevs}}
     \end{centering}
@@
         \caption{Histograms over the risk for microscopic involvement in \gls{lnl} III, given that a patient presents as clinically N0 (green), or given that \gls{lnl} II shows clinical involvement (blue).
-The top panel shows these risks for early T-category, the bottom panel for late.}
-        \label{fig:wg_III_risks}
+The top panel shows these risks for early T-category, the bottom panel for late.
+\label{fig:wg_III_risks}}
     \end{centering}
@@
         \end{enumerate*}
-        }
-        \label{fig:wg_IV_risks}
+        \label{fig:wg_IV_risks}}
     \end{centering}
@@
             \item observed involvement in the levels II, III, and IV (red).
-        \end{enumerate*}}
-        \label{fig:wg_V_risks}
+        \end{enumerate*}
+        \label{fig:wg_V_risks}}
     \end{centering}
@@
         \caption{Distributions over predicted risk for involvement in \gls{lnl} I, given different clinical diagnosis scenarios: For N0 patients (green), patients with macroscopic involvement in \gls{lnl} II (blue), and for the case where the \glspl{lnl} II, III, and IV show involvement.
-The top panel shows these risks for early T-category and the bottom row for late T-category.}
-        \label{fig:wg_I_risks}
+The top panel shows these risks for early T-category and the bottom row for late T-category.
+\label{fig:wg_I_risks}}
     \end{centering}
@@
 The risk is plotted given a clinical N0 diagnosis (green), macroscopically detected metastases in \gls{lnl} II (blue), and lastly given visible involvement in both \gls{lnl} II and III (orange).
-The top row shows these risks for early T-category diagnoses and the bottom row for late T-category.}
-        \label{fig:wg_any_risks}
+The top row shows these risks for early T-category diagnoses and the bottom row for late T-category.
+\label{fig:wg_any_risks}}
     \end{centering}
@@
 In this publication we present a statistical model of ipsilateral lymph node involvement in oropharyngeal SCC patients.
-Although the basic \gls{hmm} of lymphatic progression has been conceptually introduced in a previous work \cite{ludwig_hidden_2021}, this is the first publication that evaluates the model based on a large multi-institutional dataset containing \variable{output/num_patients.txt} patients.
+Although the basic \gls{hmm} of lymphatic progression has been conceptually introduced in a previous work \citep{ludwig_hidden_2021}, this is the first publication that evaluates the model based on a large multi-institutional dataset containing \variable{output/num_patients.txt} patients.
 It is demonstrated for the first time that the model can accurately describe the patterns of lymph node involvement observed in the data, including the correlations between levels and its dependence on T-stage.
@@
         \caption{Dependence of risk of occult disease on sensitivity and specificity.
-The figure shows the risk of involvement in level IV, given the mean of the parameter samples, drawn during sampling, in patients with clinical involvement of levels II and III.}
-        \label{fig:wg_sens_spec_risks}
+The figure shows the risk of involvement in level IV, given the mean of the parameter samples, drawn during sampling, in patients with clinical involvement of levels II and III.
+\label{fig:wg_sens_spec_risks}}
     \end{centering}
@@
 
-This work was supported by the Swiss Cancer Research Foundation under grant \href{https://www.krebsforschung.ch/unterstuetzen-sie-uns/stiftungen/-dl-/fileadmin/downloads/unterstuetzen-sie-uns/projekte-der-stiftung-krebsforschung-schweiz-2023.pdf}{KFS-5645-08-2022} and by the University Zürich under the Clinical Research Priority Program \href{https://www.crpp-ai-oncology.uzh.ch/en/Projects/Project-5.html}{Artificial Intelligence in Oncological Imaging}.
+This work was supported by the Swiss Cancer Research Foundation under grant \href{https://www.krebsforschung.ch/unterstuetzen-sie-uns/stiftungen/-dl-/fileadmin/downloads/unterstuetzen-sie-uns/projekte-der-stiftung-krebsforschung-schweiz-2023.pdf}{KFS 5645-08-2022} and by the University Zürich under the Clinical Research Priority Program \href{https://www.crpp-ai-oncology.uzh.ch/en/Projects/Project-5.html}{Artificial Intelligence in Oncological Imaging}.
 

### 2023-12-15 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\begin{deluxetable}{|cccccc|rr|rr|}
+\begin{deluxetable}{ccccccrrrr}
     \tablecaption{Prevalence of involvement patterns in the multi-centric dataset.
@@
     \tablehead{
-        \multicolumn{6}{|c|}{\textbf{LNL involvement} } & \multicolumn{4}{c|}{\textbf{T-category} } \\
-        I & II & III & IV & V & VII & \multicolumn{2}{c}{\textbf{early} } & \multicolumn{2}{c|}{\textbf{late} } \\
+        \multicolumn{6}{c}{\textbf{LNL involvement} } & \multicolumn{4}{c}{\textbf{T-category} } \\
+        I & II & III & IV & V & VII & \multicolumn{2}{c}{\textbf{early} } & \multicolumn{2}{c}{\textbf{advanced} }
     }

### 2023-12-15 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\begin{table}
-    \centering
+% \begin{table}
+%     \centering
+%     \variable{output/results_table.tex}
+%     \caption{Model comparison results from the base graph and the extended graph we chose as the ``winning'' model.
+For both \glspl{dag} we show the log-evidence, computed via thermodynamic integration, the negative one half of the \gls{bic}, as well as the maximum log-likelihood that was encountered during the final \gls{mcmc} sampling round.}
+%     \label{table:evidence}
+% \end{table}
+
+\begin{deluxetable}{lr}
+    \tablecaption{Model comparison results from the base graph and the extended graph we chose as the ``winning'' model.
+For both \glspl{dag} we show the log-evidence, computed via thermodynamic integration, the negative one half of the \gls{bic}, as well as the maximum log-likelihood that was encountered during the final \gls{mcmc} sampling round.
+\label{table:evidence}}
+    \tablehead{
+        \textbf{graph}  & \textbf{log-evidence}
+    }
+    \startdata
     \variable{output/results_table.tex}
-    \caption{Model comparison results from the base graph and the extended graph we chose as the ``winning'' model.
-For both \glspl{dag} we show the log-evidence, computed via thermodynamic integration, the negative one half of the \gls{bic}, as well as the maximum log-likelihood that was encountered during the final \gls{mcmc} sampling round.}
-    \label{table:evidence}
-\end{table}
+    \enddata
+\end{deluxetable}
 

### 2023-12-15 src/scripts/compile_metrics_table.py
update metrics table to work with deluxetable

@@ -2,6 +2,7 @@
 Populate a Jinja2 LaTeX table template with the metrics from the experiment runs.
 """
 import json
+from jinja2 import Environment, FileSystemLoader
 
 import pandas as pd
 import matplotlib.pyplot as plt
@@ -14,27 +15,28 @@ import paths
 REVISIONS = {
     "base-graph-v2": "base graph",
     "add-I-to-II": "add I → II",
-    "add-I-to-II-and-III-to-V": "add I → II & III → V",
-    "add-I-to-II-and-III-to-V-and-IV-to-V": "add I → II & III → V & IV → V",
-    "add-I-to-II-and-IV-to-V": "add I → II & IV → V",
+    "add-I-to-II-and-III-to-V": "add I → II \& III → V",
+    "add-I-to-II-and-III-to-V-and-IV-to-V": "add I → II \& III → V \& IV → V",
+    "add-I-to-II-and-IV-to-V": "add I → II \& IV → V",
     "add-II-to-I": "add II → I",
     "add-II-to-V": "add II → V",
     "add-II-to-VII": "add II → VII",
     "add-III-to-V": "add III → V",
     "add-IV-to-V": "add IV → V",
     "del-II-to-III": "remove II → III",
-    "loose-graph-v1": "remove II → III & III → IV",
+    "loose-graph-v1": "remove II → III \& III → IV",
     "del-III-to-IV": "remove III → IV",
 }
 
+TEMPLATE_NAME = "results_table.temp"
 TABLE_OUTPUT = paths.output / "results_table.tex"
 VISUAL_OUTPUT = paths.figures / "visual_ranking_guide.svg"
 
 
 if __name__ == "__main__":
-    evidences = []
+    results = []
 
-    for revision in REVISIONS.keys():
+    for revision, graph in REVISIONS.items():
         dvc_filesystem = DVCFileSystem(
             url="https://github.com/rmnldwg/lynference",
             rev=revision,
@@ -43,29 +45,30 @@ if __name__ == "__main__":
         with dvc_filesystem.open("metrics.json") as f:
             metrics = json.load(f)
 
-        evidences.append(metrics["evidence"])
+        results.append({
+            "graph": graph,
+            "evidence": metrics["evidence"],
+        })
 
-    results = pd.DataFrame.from_dict({
-        "graph": REVISIONS.values(),
-        "evidence": evidences,
-    }).sort_values(by="evidence", ascending=False)
[…]

### 2023-12-15 src/scripts/compile_means_table.py
update means table to work with deluxetable

@@ -1,13 +1,14 @@
 """
 Write all mean parameter samples to a markdown table.
 """
-import pandas as pd
+from jinja2 import Environment, FileSystemLoader
 from emcee.backends import HDFBackend
 
 import paths
 
 
 INPUT = paths.data / "wg_samples.hdf5"
+TEMPLATE_NAME = "means_table.temp"
 OUTPUT = paths.output
 NAMES = [
     r"b_\text{I}",
@@ -31,12 +32,16 @@ if __name__ == "__main__":
     means = samples.mean(axis=0)
     stddevs = samples.std(axis=0)
 
-    means_dict = {
-        f"${name}$": [f"{100 * mean:.2f} \%", f"$\pm$ {100 * std:.2f} \%"]
-        for name, mean, std in zip(NAMES, means, stddevs)
-    }
-    df = pd.DataFrame.from_dict(means_dict, orient="index", columns=["mean", "std. dev."])
-    latex_table = df.to_latex(column_format="lrr")
+    lines = []
+    for name, mean, std in zip(NAMES, means, stddevs):
+        lines.append({
+            "name": name,
+            "mean": 100 * mean,
+            "std": 100 * std,
+        })
 
-    with open(OUTPUT / "wg_means.tex", "w") as f:
-        f.write(latex_table)
+    environment = Environment(loader=FileSystemLoader(paths.scripts))
+    template = environment.get_template(TEMPLATE_NAME)
+
+    with open(OUTPUT / "means_table.tex", "w") as f:
+        f.write(template.render(lines=lines))

### 2023-12-15 src/scripts/compile_metrics_table.py

@@ -28,8 +28,8 @@ REVISIONS = {
     "del-III-to-IV": "remove III → IV",
 }
 
-TEMPLATE_NAME = "results_table.temp"
-TABLE_OUTPUT = paths.output / "results_table.tex"
+TEMPLATE_NAME = "metrics_table.temp"
+TABLE_OUTPUT = paths.output / "metrics_table.tex"
 VISUAL_OUTPUT = paths.figures / "visual_ranking_guide.svg"
 
 

### 2023-12-15 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \newacronym{isb}{ISB}{Inselspital Bern}
-
-\newcommand{\ju}[1]{{\textcolor{red}{#1}}}
-\newcommand{\rl}[1]{{\textcolor{blue}{#1}}}
-
-% Cross-referencing
-% \usepackage{cleveref}
 
@@
     \startdata
-    \variable{output/results_table.tex}
+    \variable{output/metrics_table.tex}
     \enddata
@@
 
-\begin{table}[]
-    \centering
-    \variable{output/wg_means.tex}
-    \caption{Mean of sampled parameters in percent.}
-    \label{table:params}
-\end{table}
+% \begin{table}[]
+%     \centering
+%     \variable{output/means_table.tex}
+%     \caption{Mean of sampled parameters in percent.}
+%     \label{table:params}
+% \end{table}
+
+\begin{deluxetable}{lrr}
+    \tablecaption{Mean and standard deviation of parameters sampled for the winning graph in percent. \label{table:params}}
+    \tablehead{
+        \textbf{parameter} & \textbf{mean} & \textbf{std. dev.}
+    }
+    \startdata
+    \variable{output/means_table.tex}
+    \enddata
+\end{deluxetable}
 
@@
 For clinically N0 patients, the risk in level I is estimated to be in the order of 1-2\%.
-Extensive nodal involvement of levels II-IV increases the risk in level I to just below 4\% for advanced T-stage tumors.
-It is pointed out that the winning graph does not contain arcs from levels III or IV to level I (and anatomically we do not assume that there is lymphatic drainage from levels III or IV to level I).
-Thus, the increased risk in level I is related to the time evolution: Involvement of levels III and IV corresponds to a more advances state of disease that is likely diagnosed at a later time step, such that the tumor had more time to spread to level I.
+Extensive nodal involvement of levels II-IV increases the risk in level I to just below 4\% for advanced T-category tumors.
+It is pointed out that the winning graph does not contain arcs from levels III or IV to \gls{lnl} I (and anatomically we do not assume that there is lymphatic drainage from levels III or IV to level I).
+Thus, the increased risk in level I is related to the time evolution: Getting diagnosed at a later time during the disease's evolution probably correlates with more advanced nodal metastasis.
+And involvement in the levels III and IV corresponds to a more advances state of disease that is likely diagnosed at a later time step, such that the tumor also had more time to spread to level I.
 The correlation between the clinical involvement pattern, the likely time of diagnosis in the tumor's time frame based on it, and from that the risk of involvement is another benefit of the formulation of the model as a \gls{hmm}.
@@
 \label{sec:discussion}
+
 \subsection{Summary}
+
 In this publication we present a statistical model of ipsilateral lymph node involvement in oropharyngeal SCC patients.

### 2023-12-18 src/scripts/compile_metrics_table.py
sort results in metrics table

@@ -50,11 +50,13 @@ if __name__ == "__main__":
             "evidence": metrics["evidence"],
         })
 
+    sorted_results = sorted(results, key=lambda x: x["evidence"], reverse=True)
+
     environment = Environment(loader=FileSystemLoader(paths.scripts))
     template = environment.get_template(TEMPLATE_NAME)
 
     with open(TABLE_OUTPUT, "w") as f:
-        f.write(template.render(results=results))
+        f.write(template.render(results=sorted_results))
 
     # fig, ax = plt.subplots(figsize=(10,1))
     # ax.plot(

### 2023-12-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 The separation between involvement prevalences of early and late T-category tumors is also reproduced well by the model.
-This is remarkable because the model introduces only a single parameter to describe the differences between early and late T-category for all involvement patterns.
+This is remarkable because the model introduces only a single parameter to describe the differences between early and advanced T-category for all involvement patterns.
 This shows that expecting later diagnosis times, on average, for patients with advanced T-category tumors can explain more severe lymphatic involvement.
@@
 Based on these results, we consider three additional candidates for the optimal graph that combine the added arc from \gls{lnl} I to II with the arc(s) $\text{III} \rightarrow \text{V}$ and/or $\text{IV} \rightarrow \text{V}$.
-The model evidence for these three graphs is reported in \autoref{table:evidence}.
+The model evidence for these three graphs is also reported in \autoref{table:evidence}.
 The best performing graph with decisive evidence over the base graph turned out to be the one which combines the arcs from \gls{lnl} I to II and IV to V.
@@
 
-\autoref{table:evidence} additionally shows the evidence of graphs in which the arc from level II to III and from level III to IV is removed.
+\autoref{table:evidence} additionally shows the evidence of graphs in which the arc from level II to III or from level III to IV is removed.
 The low model evidence for these graphs confirms the importance of these connections and is consistent with the anatomical motivation.

### 2023-12-18 src/scripts/wg_I_risks.py
improve risk plot xlims

@@ -52,12 +52,12 @@ if __name__ == "__main__":
             ),
         ]
 
-    draw(axes[0], contents=panels["early"], xlims=(0., 7.), hist_kwargs={"nbins": 40})
+    draw(axes[0], contents=panels["early"], xlims=(0., 5.), hist_kwargs={"nbins": 40})
     axes[0].set_ylabel("early T-cat.", fontweight="bold")
     axes[0].set_yticks([])
     axes[0].legend()
 
-    draw(axes[1], contents=panels["late"], xlims=(0., 7.), hist_kwargs={"nbins": 40})
+    draw(axes[1], contents=panels["late"], xlims=(0., 5.), hist_kwargs={"nbins": 40})
     axes[1].set_ylabel("late T-cat.", fontweight="bold")
     axes[1].set_yticks([])
     axes[1].set_xlabel("risk [%]")

### 2023-12-18 src/scripts/wg_any_risks.py

@@ -52,12 +52,12 @@ if __name__ == "__main__":
             ),
         ]
 
-    draw(axes[0], contents=panels["early"], xlims=(0., 16.), hist_kwargs={"nbins": 40})
+    draw(axes[0], contents=panels["early"], xlims=(2., 16.), hist_kwargs={"nbins": 40})
     axes[0].set_ylabel("early T-cat.", fontweight="bold")
     axes[0].set_yticks([])
     axes[0].legend()
 
-    draw(axes[1], contents=panels["late"], xlims=(0., 16.), hist_kwargs={"nbins": 40})
+    draw(axes[1], contents=panels["late"], xlims=(2., 16.), hist_kwargs={"nbins": 40})
     axes[1].set_ylabel("late T-cat.", fontweight="bold")
     axes[1].set_yticks([])
     axes[1].set_xlabel("risk [%]")

### 2023-12-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 However, using this definition, we cannot know how many time-steps have passed until $t_\text{D}$, when the patient was diagnosed with cancer.
-We can only make the assumption that a patients with an earlier T-category tumor was \emph{probably} diagnosed after fewer time-steps than a patient with a late T-category tumor.
+We can only make the assumption that a patients with an earlier T-category tumor was \emph{probably} diagnosed after fewer time-steps than a patient with a advanced T-category tumor.
 
 We can use this assumption by marginalizing over the diagnose times $t_\text{D}$ of patients in different T-categories using different prior distributions over the diagnose time.
-E.g., $P \left( t=t_\text{D} \mid \text{early} \right)$ for early T-category patients (T1 \& T2) and $P \left( t=t_\text{D} \mid \text{late} \right)$ for late T-category patients (T3 \& T4).
+E.g., $P \left( t=t_\text{D} \mid \text{early} \right)$ for early T-category patients (T1 \& T2) and $P \left( t=t_\text{D} \mid \text{late} \right)$ for advanced T-category patients (T3 \& T4).
 Throughout this work we will use binomial distributions for these probability mass functions, mainly because they have a plausible shape for this purpose and only a single parameter.
@@
 The model's predictions are compared to Beta posteriors over the prevalence based on the frequency of the same scenarios and a uniform prior (slid lines).
-The top panel shows some selected scenarios with early T-category tumors and the bottom panel with late T-category.
+The top panel shows some selected scenarios with early T-category tumors and the bottom panel with advanced T-category.
 \label{fig:bg_prevalences}}
@@
 Notably, the model is not only accurate in its predictions, it also correctly estimates the variance stemming from the limited amount of data.
-The separation between involvement prevalences of early and late T-category tumors is also reproduced well by the model.
+The separation between involvement prevalences of early and advanced T-category tumors is also reproduced well by the model.
 This is remarkable because the model introduces only a single parameter to describe the differences between early and advanced T-category for all involvement patterns.
@@
 Their respective model evidence fell below $-1528$ and the two graphs would appear far left in the figure.
-In the bottom left corner, we provide a visual reference in analogy to \autoref{table:bayes_factor}: E.g., any difference in the model evidence shorter than the first of the three rulers indicates that the improvement is "barely worth a mention".
+In the bottom left corner, we provide a visual reference in analogy to \autoref{table:bayes_factor}: E.g., any difference in the model evidence shorter than the first of the three rulers indicates that the improvement is ``barely worth a mention''.
 \label{fig:ranking}}
@@
 The probability for no involvement of level II when the patient is diagnosed after $t$ time steps is $\left(1-b_\text{II}\right)^t$.
-The prevalence of level II involvement for late T-category patients is thus
+The prevalence of level II involvement for advanced T-category patients is thus
 %
@@
 {\it Level V:} In \autoref{fig:IIIandIVandV_prevs} we compare the base graph's and the winning graph's estimations for prevalences of involvement patterns that include \gls{lnl} V.
-The base graph underestimates the probability that level IV and V are simultaneously involved, and overestimates that probability the level V but not IV is involved.
+The base graph underestimates the probability that level IV and V are simultaneously involved, and overestimates the probability that level V but not IV is involved.
 By introducing the arc from level IV to V, the winning graph can describe the observation that level V involvement is typically associated with severe involvement of level II-IV.
-In the dataset, 15 patients out of 68 patients with level IV involvement have metastases in level V.
-Instead only 40 patients out of 618 patients without level IV involvement have metastases in level V.
+In the dataset, 14 patients out of 62 patients with level IV involvement have metastases in level V (22 \%).
+Instead, only 40 patients out of 624 patients without level IV involvement have metastases in level V (6 \%).
 
@@
 By introducing the arc from level I to II, the winning graph can capture the correlations between levels I and II.
-It can also be noted that both models overestimate level I involvement for early T-category patients and underestimate level I involvement for late T-category patients.
+It can also be noted that both models overestimate level I involvement for early T-category patients and underestimate level I involvement for advanced T-category patients.
 This is further described in the discussion section below.
@@
         \caption{Comparison of observed and predicted prevalences of \gls{lnl} I and II involvement patterns.
-The top and bottom panels show the prevalences for early and late T-category, respectively.
+The top and bottom panels show the prevalences for early and advanced T-category, respectively.
 The solid lines are Beta posteriors from the data, while the histograms are predicted prevalences (colored: winning graph, gray-hatched: base graph).
@@
 {\it Level III:} \autoref{fig:wg_III_risks} compares the risk of occult disease in level III between patients that are clinically N0 (orange) and patients with clinically diagnosed involvement of only level II (red), for early T-category (upper panel) and advanced T-category (bottom panel).
-The histograms represent the uncertainty in model's risk prediction arising from the uncertainty in the model parameters and are generated by randomly drawing 1000 samples from the model parameter's joint posterior distribution.
-The model predicts a risk of 5\% for early T-stage tumors and 8\% for advanced T-stage.
+The histograms represent the uncertainty in the model's risk prediction arising from the uncertainty in the model parameters and are generated by randomly drawing a tenth of the samples from the model parameter's joint posterior distribution.
+This amounts to $S = 20 \cdot W$ samples, as described in \autoref{subsec:complete_model:comparison}.
+The model predicts a risk of just below 6\% for early T-category tumors and 8\% for T3 or T4 ones.
 For patients with involvement of level II, the risk in level III increases to approximately 9\% and 12\%, respectively.
@@
 The model predicts a low risk of 1-2\% in level IV for patients with clinically healthy level III.
-For patients with clinical involvement of level III, the risk of occult disease in level IV increases to approximately 3\% for early T-stage and 5\% for advanced T-stage tumors.
+For patients with clinical involvement of level III, the risk of occult disease in level IV increases to approximately 3\% for early T-category and 5\% for advanced T-category tumors.
 
@@
         \includegraphics[width=\linewidth]{figures/wg_IV_risks.png}
-        \caption{Distributions over the risk for microscopic involvement in \gls{lnl} IV as predicted by the winning graph model, given early (top panel) or late T-category (bottom panel), and three different CT-based diagnoses:
-        \begin{enumerate*}[label=\arabic*)]
-            \item A clinical N0 patient (gray histogram),
-            \item visible metastases in \gls{lnl} II, but otherwise healthy-looking lymph nodes (blue histogram), and
-            \item macroscopic metastases in the \glspl{lnl} II \& III, with the rest of the neck still being clinically node negative (orange histogram).
-        \end{enumerate*}
-        \label{fig:wg_IV_risks}}
+        \caption{Distributions over the risk for microscopic involvement in \gls{lnl} IV as predicted by the winning graph model, given early (top panel) or advanced T-category (bottom panel), and three different CT-based diagnoses: 1) A clinical N0 patient (gray histogram), 2) visible metastases in \gls{lnl} II, but otherwise healthy-looking lymph nodes (blue histogram), and 3) macroscopic metastases in the \glspl{lnl} II \& III, with the rest of the neck still being clinically node negative (orange histogram).\label{fig:wg_IV_risks}}
     \end{centering}
@@
 {\it Level V:} \autoref{fig:wg_V_risks} shows the risk of occult disease in level V depending on T-category and the clinical involvement of levels II-IV.
-For clinically N0 patients, the risk in level V is estimated to be around 1\%.
-Extensive nodal involvement of levels II-IV increases the risk in level V to more than 4\% for advanced T-stage tumors.
+For clinically N0 patients, the risk in level V is estimated to be just above 1\%.
+Extensive nodal involvement of levels II-IV increases the risk in level V to more than 4\% for advanced T-category tumors.
 
@@
         \includegraphics[width=\linewidth]{figures/wg_V_risks.png}
-        \caption{The winning graph's predictions over the risk for occult disease in \gls{lnl} V for early (top panel) and late T-category (bottom panel).
-Each of the risks is computed, given one of three diagnostic scenarios that were obtained by a CT scan:
-        \begin{enumerate*}[label=\arabic*)]
-            \item A clinically healthy neck (green),
-            \item on CT imaging visible involvement in the \glspl{lnl} II \& III (orange) with the remaining levels still clinically negative, and
-            \item observed involvement in the levels II, III, and IV (red).
-        \end{enumerate*}
-        \label{fig:wg_V_risks}}
+        \caption{The winning graph's predictions over the risk for occult disease in \gls{lnl} V for early (top panel) and advanced T-category (bottom panel).
+Each of the risks is computed, given one of three diagnostic scenarios that were obtained by a CT scan: 1) A clinically healthy neck (green), 2) on CT imaging visible involvement in the \glspl{lnl} II \& III (orange) with the remaining levels still clinically negative, and 3) observed involvement in the levels II, III, and IV (red).\label{fig:wg_V_risks}}
     \end{centering}
@@
         \caption{Distributions over predicted risk for involvement in \gls{lnl} I, given different clinical diagnosis scenarios: For N0 patients (green), patients with macroscopic involvement in \gls{lnl} II (blue), and for the case where the \glspl{lnl} II, III, and IV show involvement.
-The top panel shows these risks for early T-category and the bottom row for late T-category.
+The top panel shows these risks for early T-category and the bottom row for advanced T-category.
 \label{fig:wg_I_risks}}
@@
 The risk is plotted given a clinical N0 diagnosis (green), macroscopically detected metastases in \gls{lnl} II (blue), and lastly given visible involvement in both \gls{lnl} II and III (orange).
-The top row shows these risks for early T-category diagnoses and the bottom row for late T-category.
+The top row shows these risks for early T-category diagnoses and the bottom row for advanced T-category.
 \label{fig:wg_any_risks}}
@@
 Although the basic \gls{hmm} of lymphatic progression has been conceptually introduced in a previous work \citep{ludwig_hidden_2021}, this is the first publication that evaluates the model based on a large multi-institutional dataset containing \variable{output/num_patients.txt} patients.
-It is demonstrated for the first time that the model can accurately describe the patterns of lymph node involvement observed in the data, including the correlations between levels and its dependence on T-stage.
+It is demonstrated for the first time that the model can accurately describe the patterns of lymph node involvement observed in the data, including the correlations between levels and its dependence on T-category.
 Furthermore, techniques from statistical physics are applied to calculate the model evidence for Bayesian model comparison.
@@
 
-{\it T-category dependence of level I involvement:} As shown in \autoref{fig:bg_prevalences}, the model describes the involvement of \glspl{lnl} II, III, and IV depending on T-stage very well despite having only a single parameter related to T-stage.
-\autoref{fig:IandII_prevs} shows that the model does not perfectly describe the T-stage dependence of level I.
-It adjusts the parameters such that level I involvement is correctly described for the set of all patients combined, but it overestimates level I involvement for early T-stage and underestimates it for advanced T-stage.
-A possible explanation is that advanced T-stage tumors are more likely to have grown into regions with direct lymph drainage to level I.
-The more severe involvement in levels II, III, IV for advanced T-stage can be explained by tumors having more time to spread while keeping the spread probability rates $b_2$, $b_3$, $b_4$ constant.
-Regarding level I, early versus advanced T-stage tumors may have different spread probability rates $b_1$.
+{\it T-category dependence of level I involvement:} As shown in \autoref{fig:bg_prevalences}, the model describes the involvement of \glspl{lnl} II, III, and IV depending on T-category very well despite having only a single parameter related to T-category.
+\autoref{fig:IandII_prevs} shows that the model does not perfectly describe the T-category dependence of level I.
+It adjusts the parameters such that level I involvement is correctly described for the set of all patients combined, but it overestimates level I involvement for early T-category and underestimates it for advanced T-category.
+A possible explanation is that advanced T-category tumors are more likely to have grown into regions with direct lymph drainage to level I.
+The more severe involvement in levels II, III, IV for advanced T-category can be explained by tumors having more time to spread while keeping the spread probability rates $b_2$, $b_3$, $b_4$ constant.
+Regarding level I, early versus advanced T-category tumors may have different spread probability rates $b_1$.
 

### 2023-12-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Anatomically, the \emph{posterior accessory pathway} that drains \gls{lnl} II through \gls{lnl} V motivates investigating and arc from $X_2$ into $X_5$ \citep{lengele_anatomical_2007}.
-And, although no lymphatic pathway is described that directly drains \gls{lnl} III or IV into level V, due to their proximity to each other, we will also investigate additional arcs from \gls{lnl} III and IV into level V.
-Finally, an arc from \gls{lnl} II to \gls{lnl} VII is considered.
+And, although no lymphatic pathway is described that directly drains \gls{lnl} III or IV into level V, due to their proximity to each other, we will also investigate additional edges from the levels III and IV into \gls{lnl} V.
+Finally, we look at adding an arc from \gls{lnl} II to \gls{lnl} VII also due to their anatomical proximity.
 
 To determine the optimal graph, we first consider six models, each with one of the six green candidate arcs of \autoref{fig:extended_graph} added to the base graph.
-Every one of these models was evaluated by computing its evidence via thermodynamic integration as described below in \autoref{subsec:complete_model:comparison}.
-%These evidences are shown, among others, in \autoref{table:evidence}. 
+Every one of these models was evaluated by computing an approximation to its evidence via thermodynamic integration as described below in \autoref{subsec:complete_model:comparison}.
 Subsequently, graphs combining multiple arcs that individually improve the model evidence are considered.
-Thereby, the ``winning graph'' is determined, which yields the highest value of the model evidence.
-%which is in great detail compared to the base graph in the results (\autoref{sec:results}). All considered graphs' performances are visualized in \autoref{fig:ranking}.
+Thereby, the ``winning graph'' is determined, which yields the highest (i.e., the least negative) value of the logarithm of the model evidence.\looseness=-1
 
@@
 
-All figures in this work, like the risk predictions and prevalences we show histograms of, can be reproduced following the instructions of the GitHub repository underlying this publication \href{\GitHubURL}{\texttt{rmnldwg/graph-extension}}.
+All figures in this work, like the risk predictions and prevalences we show histograms of, can be reproduced following the instructions of the GitHub repository underlying this publication \href{https://github.com/rmnldwg/graph-extension-paper}{\texttt{rmnldwg/graph-extension-paper}}.
 It is based off of the \showyourwork project.
@@
 \label{sec:data}
-
-% \begin{table}
-%     \centering
-%     \variable{output/data_table.tex}
-%     \caption{Prevalence of involvement patterns in the multi-centric dataset.
-An involvement pattern is characterized by the state of the six \glspl{lnl}: A red dot means the \gls{lnl} was reported to be metastatic, a green dot means it was determined to be healthy and a question mark means that the prevalence was marginalized over the state of this \gls{lnl}.}
-%     \label{table:data_prevalence}
-% \end{table}
 
@@
 We have previously published the patterns of nodal involvement for the \gls{usz} cohort (287 patients) \citep{ludwig_dataset_2022,ludwig_lydata_2023} and described its characteristics in detail \citep{ludwig_detailed_2022}.
-The first \gls{clb} dataset (263 patients) \citep{ludwig_lydata_2022} underlies a publication on \gls{hpv} status in \gls{opscc} \citep{bauwens_prevalence_2021} and is made available in a separate ``Data in Brief'' article alongside the second dataset from France \citep{ludwig_lydata_2023-2} and the lymphatic progression patterns from the \gls{isb} \citep{ludwig_lydata_2023-1,roman_ludwig_multi-centric_nodate}.
+The first \gls{clb} dataset (263 patients) \citep{ludwig_lydata_2022} underlies a publication on \gls{hpv} status in \gls{opscc} \citep{bauwens_prevalence_2021} and is made available in a separate ``Data in Brief'' article alongside the second dataset from France \citep{ludwig_lydata_2023-2} and the lymphatic progression patterns from the \gls{isb} \citep{ludwig_lydata_2023-1,ludwig_multi-centric_2023}.
+All datasets may be explored online in our web-based interface \href{https://lyprox.org}{LyProX}.
 
@@
 
-% \rl{\textbf{NOTE:} This subsection is still unfinished an will be completed when the respective computations are ready. Qualitatively, we don't expect any major changes though.}
-
 The model evidences of all candidate graphs are reported in \autoref{table:evidence}.
@@
 The connection from level III to IV is crucial for describing the observation that metastases in level IV are extremely rare without simultaneous involvement of the upstream level III.
+\looseness=-1
 
@@
 \label{subsec:results:optimal_graph}
-
-% \begin{table}
-%     \centering
-%     \variable{output/results_table.tex}
-%     \caption{Model comparison results from the base graph and the extended graph we chose as the ``winning'' model.
-For both \glspl{dag} we show the log-evidence, computed via thermodynamic integration, the negative one half of the \gls{bic}, as well as the maximum log-likelihood that was encountered during the final \gls{mcmc} sampling round.}
-%     \label{table:evidence}
-% \end{table}
 
@@
 The large uncertainty in $t_{\text{I} \rightarrow \text{II}}$ is related to the fact that level I involvement is rare compared to level II.
-
-% \begin{table}[]
-%     \centering
-%     \variable{output/means_table.tex}
-%     \caption{Mean of sampled parameters in percent.}
-%     \label{table:params}
-% \end{table}
 
@@
 
+To illustrate the flexibility of the model in predicting various risks, we have plotted the risk for occult disease in \textit{any} of the \glspl{lnl} I, IV, and/or V, given different clinical diagnoses in \autoref{fig:wg_any_risks}.
+Similar to this, we may compute the risk for an arbitrary combination of involved levels, given a similarly arbitrary clinical diagnosis.
+For the base graph (\href{https://lyprox.org/riskpredictor/9}{\texttt{base-graph-v2}}) and the winning graph (\href{https://lyprox.org/riskpredictor/8}{\texttt{win-graph-v3}}), one may also interactively explore these risks in our web-based interface \href{https://lyprox.org/riskpredictor/list}{LyProX}, similar to how it is possible to explore the underlying data in an interactive way.
+
 \begin{figure}
@@
     \item[$\bullet$] Exclude level I from the \gls{ctv-n} for early T-category patients with limited metastatic disease.
-For advanced T-category patients with extensive nodal involvement of levels II, III, and IV, level I should be irradiated (see also the limitations discussed below).
+For advanced T-category patients with extensive nodal involvement of levels II, III, and IV, level I should be irradiated (see also the limitations discussed in \autoref{subsec:disc:limitation}).
     \item[$\bullet$] Exclude level VII in all patients, unless it is clinically involved.
@@
 The more severe involvement in levels II, III, IV for advanced T-category can be explained by tumors having more time to spread while keeping the spread probability rates $b_2$, $b_3$, $b_4$ constant.
-Regarding level I, early versus advanced T-category tumors may have different spread probability rates $b_1$.
+Regarding level I, early versus advanced T-category tumors the model may need different spread probability rates $b_1$ to describe their involvement correctly.
 
@@
 For our default parameters of 81\% specificity and 76 \% sensitivity, the risk is 5\%, but it increases to around 8\% for a a sensitivity of 66\%.
+
+Also, as described in \autoref{sec:data}, we assumed the consensus of the data to represent the true state of nodal involvement.
+This was not strictly necessary: Instead of computing a consensus beforehand and providing that with sensitivity and specificity of 1 to the model, as if it were the ground truth, we could have provided multiple diagnostic modalities per patient to the model directly.
+In fact, for patients with a pathology report available, this would even yield the same results.
+But we also decided to consider the consensus as an observation of the true hidden state for patients without pathologically assessed involvement.
+We did this because the literature values for sensitivity and specificity of around 80\% do not plausibly match the observation that around 78\% of patients in the \gls{usz} cohort showed clinical \gls{lnl} II involvement.
+The most likely true prevalence of involvement in \gls{lnl} II would need to be close to 100\%.
+
+Discussing the possible origins for this discrepancy is beyond the scope of this work.
+Assuming the consensus to represent the true hidden state of a patient nonetheless allowed us to investigate if the model can describe plausible patterns of nodal involvement well.
 Future work may aim at developing new methods to model the difference between pathological and clinical lymph node involvement based on surgically treated patients in whom both is reported.

### 2023-12-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     Methods: We extend a previously published \gls{hmm}, which models the \glspl{lnl} (I, II, III, IV, V, and VII) as hidden binary \glspl{rv}.
-Each represents a patient's true state of lymphatic involvement corresponding to pathological examinations in that level.
+Each represents a patient's true state of lymphatic involvement.
 Clinical involvement at diagnosis represents the observed binary \glspl{rv} linked to the true state via sensitivity and specificity.
@@
 
-In total, the dataset contains \variable{output/num_patients.txt}patients with newly diagnosed \gls{opscc}.
+In total, the dataset contains \variable{output/num_patients.txt} patients with newly diagnosed \gls{opscc}.
 It includes patients treated with definitive (chemo)radiotherapy, adjuvant (chemo)radiotherapy following neck dissection, or neck dissection alone.

### 2023-12-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \begin{abstract}
-    Objective: The \gls{ctv-n} in \gls{opscc} is currently based mostly on the prevalence of lymph node metastases in different \glspl{lnl} for a given primary tumor location.
-We present a probabilistic model for ipsilateral lymphatic spread that can quantify the risk for microscopic nodal involvement based on an individual patient's T-category and clinical involvement of \glspl{lnl} at diagnosis.
+    The \gls{ctv-n} in \gls{opscc} is currently based mostly on the prevalence of lymph node metastases in different \glspl{lnl} for a given primary tumor location.
+We present a probabilistic model for ipsilateral lymphatic spread that can quantify the microscopic nodal involvement risk based on an individual patient's T-category and clinical involvement of \glspl{lnl} at diagnosis.
 
-    Methods: We extend a previously published \gls{hmm}, which models the \glspl{lnl} (I, II, III, IV, V, and VII) as hidden binary \glspl{rv}.
+    We extend a previously published \gls{hmm}, which models the \glspl{lnl} (I, II, III, IV, V, and VII) as hidden binary \glspl{rv}.
 Each represents a patient's true state of lymphatic involvement.
@@
 The primary tumor and the hidden \glspl{rv} are connected in a graph.
-Each edge represents the conditional probability of metastatic spread per abstract time-step, given involvement at the edge's starting node.
+Each edge represents the conditional probability of metastatic spread per abstract time-step, given disease at the edge's starting node.
 To learn these probabilities, we draw \acrlong{mcmc} samples from the likelihood of a dataset (\variable{output/num_patients.txt} \gls{opscc} patients) from three institutions.
@@
 
-    Results: The graph maximizing the model evidence connects the tumor to each \gls{lnl} and the \glspl{lnl} I through V in order.
+    The graph maximizing the model evidence connects the tumor to each \gls{lnl} and the \glspl{lnl} I through V in order.
 It predicts the risk of occult disease in level IV is below 5\% if level III is clinically negative, and that the risk of occult disease in level V is below 5\% except for advanced T-category (T3 and T4) patients with clinical involvement of levels II, III, and IV.
 
-    Conclusion: The provided statistical model of nodal involvement in \gls{opscc} patients trained on multi-institutional data may guide the design of clinical trials on volume-deescalated treatment of \gls{opscc} and contribute to further personalized guidelines on elective nodal treatment.
+    The provided statistical model of nodal involvement in \gls{opscc} patients trained on multi-institutional data may guide the design of clinical trials on volume-deescalated treatment of \gls{opscc} and contribute to more personal guidelines on elective nodal treatment.
 \end{abstract}

### 2023-12-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Title
-\title{Modelling the Lymphatic Metastatic Progression Pathways of OPSCC from Multi-Institutional Datasets}
+\title{Modelling the Lymphatic Metastatic Progression Pathways of OPSCC \\ from Multi-Institutional Datasets}
+
+% Affiliations
+\def \USZ {University Hospital Zurich, Rämistrasse 100, 8091 Zurich, Switzerland}
+\def \UZH {University of Zurich, Rämistrasse 71, 8006 Zurich, Switzerland}
+\def \ISB {Inselspital Bern, Freiburgstrasse 18, 3010 Bern, Switzerland}
+\def \CLB {Centre L\'eon B\'erard, 28 Rue Laennec, 69008 Lyon, France}
+\def \KS  {Klinikum Stuttgart, Kriegsbergstraße 60, 70174 Stuttgart, Germany}
 
 % Author list
-\author{Roman Ludwig}
+\author[0000-0001-9434-328X]{Roman Ludwig}
+\affiliation{\USZ}
+\affiliation{\UZH}
+
 \author{Adrian Schubert}
+\affiliation{\ISB}
+
 \author{Dorothea Barbatei}
+\affiliation{\CLB}
+
 \author{Lauence Bauwens}
-\author{Jean-Marc Hoffmann}
+\affiliation{\CLB}
+
+\author[0000-0003-1193-0863]{Jean-Marc Hoffmann}
+\affiliation{\USZ}
+
 \author{Sandrine Werlen}
-\author{Olgun Elicin}
-\author{Matthias Dettmer}
-\author{Philippe Zrounba}
-\author{Bertrand Pouymayou}
-\author{Panagiotis Balermpas}
+\affiliation{\ISB}
+
+\author[0000-0002-6996-0646]{Olgun Elicin}
+\affiliation{\ISB}
+
+\author[0000-0003-0948-1392]{Matthias Dettmer}
+\affiliation{\KS}
+
+\author[0000-0001-6223-6113]{Philippe Zrounba}
+\affiliation{\CLB}
+
+\author[0000-0002-8060-7827]{Bertrand Pouymayou}
+\affiliation{\USZ}
+
+\author[0000-0001-5261-6446]{Panagiotis Balermpas}
+\affiliation{\USZ}
+
 \author{Vincent Grégoire}
-\author{Roland Giger}
-\author{Jan Unkelbach}
+\affiliation{\CLB}
+
+\author[0000-0002-5574-3210]{Roland Giger}
+\affiliation{\ISB}
+
+\author[0000-0002-4275-990X]{Jan Unkelbach}
+\affiliation{\USZ}
+\affiliation{\UZH}
+
+\correspondingauthor{Roman Ludwig}
+\email{roman.ludwig@usz.ch}
 

### 2023-12-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-{\it Level IV:} \autoref{fig:wg_IV_risks} compares the risk of occult disease in level IV for the typical clinical presentations: clinically N0 (green), metastases in level II (blue), and metastases in levels II and III (orange).
+{\it Level IV:} \autoref{fig:wg_IVandV_risks} (left panels) compares the risk of occult disease in level IV for the typical clinical presentations: clinically N0 (green), metastases in level II (blue), and metastases in levels II and III (orange).
 The model predicts a low risk of 1-2\% in level IV for patients with clinically healthy level III.
@@
 
-\begin{figure}
-    \script{wg_IV_risks.py}
+\begin{figure*}
+    \script{wg_IVandV_risks.py}
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/wg_IV_risks.png}
-        \caption{Distributions over the risk for microscopic involvement in \gls{lnl} IV as predicted by the winning graph model, given early (top panel) or advanced T-category (bottom panel), and three different CT-based diagnoses: 1) A clinical N0 patient (gray histogram), 2) visible metastases in \gls{lnl} II, but otherwise healthy-looking lymph nodes (blue histogram), and 3) macroscopic metastases in the \glspl{lnl} II \& III, with the rest of the neck still being clinically node negative (orange histogram).\label{fig:wg_IV_risks}}
+        \includegraphics{figures/wg_IVandV_risks.png}
+        \caption{Distributions over the risk for microscopic involvement in \gls{lnl} IV (left panels) and in \gls{lnl} V (right panels) as predicted by the winning graph model, given early (top panels) or advanced T-category (bottom panels), and different CT-based diagnoses: 1) A clinical N0 patient (green histograms), 2) visible metastases in \gls{lnl} II, but otherwise healthy-looking lymph nodes (blue histograms), 3) macroscopic metastases in the \glspl{lnl} II \& III, with the rest of the neck still being clinically node negative (orange histograms), and finally 4) extensive clinical involvement in the levels II, III, and IV (red histograms).\label{fig:wg_IVandV_risks}}
     \end{centering}
-\end{figure}
+\end{figure*}
 
-{\it Level V:} \autoref{fig:wg_V_risks} shows the risk of occult disease in level V depending on T-category and the clinical involvement of levels II-IV.
+{\it Level V:} The right panels in \autoref{fig:wg_IVandV_risks} shows the risk of occult disease in level V depending on T-category and the clinical involvement of levels II-IV.
 For clinically N0 patients, the risk in level V is estimated to be just above 1\%.
 Extensive nodal involvement of levels II-IV increases the risk in level V to more than 4\% for advanced T-category tumors.
-
-\begin{figure}
-    \script{wg_V_risks.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/wg_V_risks.png}
-        \caption{The winning graph's predictions over the risk for occult disease in \gls{lnl} V for early (top panel) and advanced T-category (bottom panel).
-Each of the risks is computed, given one of three diagnostic scenarios that were obtained by a CT scan: 1) A clinically healthy neck (green), 2) on CT imaging visible involvement in the \glspl{lnl} II \& III (orange) with the remaining levels still clinically negative, and 3) observed involvement in the levels II, III, and IV (red).\label{fig:wg_V_risks}}
-    \end{centering}
-\end{figure}
 

### 2023-12-19 src/scripts/wg_IVandV_risks.py
fuse two risks & don't thin

@@ -1,0 +1,104 @@
+"""
+Plot histograms of predicted risks, given different diagnoses.
+"""
+from pathlib import Path
+
+import matplotlib.pyplot as plt
+from lyscripts.plot.utils import Histogram, draw
+from lyscripts.plot.utils import COLORS as USZ
+from tueplots import figsizes, fontsizes
+
+import paths
+
+
+INPUT = paths.data / "wg_risks.hdf5"
+OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
+NROWS, NCOLS = 2, 2
+
+
+if __name__ == "__main__":
+    plt.rcParams.update(figsizes.icml2022_full(
+        nrows=NROWS,
+        ncols=NCOLS,
+        height_to_width_ratio=0.45,
+    ))
+    plt.rcParams.update(fontsizes.icml2022())
+    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
+
+    left_panels = {
+        "early": [],
+        "late": [],
+    }
+
+    for stage in ["early", "late"]:
+        left_panels[stage] = [
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"{stage}/IV/N0",
+                label="given N0",
+                color=USZ["green"],
+            ),
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"{stage}/IV/II",
+                label="given LNL II involved",
+                color=USZ["blue"],
+            ),
+            Histogram.from_hdf5(
+                filename=INPUT,
+                dataname=f"{stage}/IV/IIandIII",
+                label="given LNL II & III involved",
+                color=USZ["orange"],
+            ),
+        ]
+
+    axes[0,0].set_title("risk for LNL IV involvement", fontweight="bold")
+
+    draw(axes[0,0], contents=left_panels["early"], xlims=(0., 7.), hist_kwargs={"nbins": 40})
+    axes[0,0].set_ylabel("early T-cat.", fontweight="bold")
+    axes[0,0].set_yticks([])
+    axes[0,0].legend()
+
+    draw(axes[1,0], contents=left_panels["late"], xlims=(0., 7.), hist_kwargs={"nbins": 40})
+    axes[1,0].set_ylabel("late T-cat.", fontweight="bold")
+    axes[1,0].set_yticks([])
+    axes[1,0].set_xlabel("risk [%]")
+
+    right_panels = {
+        "early": [],
+        "late": [],
+    }
+
[…]

### 2023-12-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-{\it Level II:} As can be seen in \autoref{table:params}, spread from the tumor to \gls{lnl} II to be the most probable transition at any given time step.
-As a consequence, even for an early T-category patient that presents with a clinical N0 neck, our model predicts a \variable{output/wg_II_risks.tex} risk for microscopic metastases in \gls{lnl} II.
-
-{\it Level III:} \autoref{fig:wg_III_risks} compares the risk of occult disease in level III between patients that are clinically N0 (orange) and patients with clinically diagnosed involvement of only level II (red), for early T-category (upper panel) and advanced T-category (bottom panel).
-The histograms represent the uncertainty in the model's risk prediction arising from the uncertainty in the model parameters and are generated by randomly drawing a tenth of the samples from the model parameter's joint posterior distribution.
-This amounts to $S = 20 \cdot W$ samples, as described in \autoref{subsec:complete_model:comparison}.
-The model predicts a risk of just below 6\% for early T-category tumors and 8\% for T3 or T4 ones.
-For patients with involvement of level II, the risk in level III increases to approximately 9\% and 12\%, respectively.
-
 \begin{figure}
@@
 
+{\it Level II:} As can be seen in \autoref{table:params}, spread from the tumor to \gls{lnl} II to be the most probable transition at any given time step.
+As a consequence, even for an early T-category patient that presents with a clinical N0 neck, our model predicts a \variable{output/wg_II_risks.tex} risk for microscopic metastases in \gls{lnl} II.
+
+{\it Level III:} \autoref{fig:wg_III_risks} compares the risk of occult disease in level III between patients that are clinically N0 (orange) and patients with clinically diagnosed involvement of only level II (red), for early T-category (upper panel) and advanced T-category (bottom panel).
+The histograms represent the uncertainty in the model's risk prediction arising from the uncertainty in the model parameters and are generated by randomly drawing a tenth of the samples from the model parameter's joint posterior distribution.
+This amounts to $S = 20 \cdot W$ samples, as described in \autoref{subsec:complete_model:comparison}.
+The model predicts a risk of just below 6\% for early T-category tumors and 8\% for T3 or T4 ones.
+For patients with involvement of level II, the risk in level III increases to approximately 9\% and 12\%, respectively.
+
+\begin{figure*}
+    \script{wg_IVandV_risks.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/wg_IVandV_risks.png}
+        \caption{Distributions over the risk for microscopic involvement in \gls{lnl} IV (left panels) and in \gls{lnl} V (right panels) as predicted by the winning graph model, given early (top panels) or advanced T-category (bottom panels), and different CT-based diagnoses: 1) A clinical N0 patient (green histograms), 2) visible metastases in \gls{lnl} II, but otherwise healthy-looking lymph nodes (blue histograms), 3) macroscopic metastases in the \glspl{lnl} II \& III, with the rest of the neck still being clinically node negative (orange histograms), and finally 4) extensive clinical involvement in the levels II, III, and IV (red histograms).\label{fig:wg_IVandV_risks}}
+    \end{centering}
+\end{figure*}
+
 {\it Level IV:} \autoref{fig:wg_IVandV_risks} (left panels) compares the risk of occult disease in level IV for the typical clinical presentations: clinically N0 (green), metastases in level II (blue), and metastases in levels II and III (orange).
@@
 
-\begin{figure*}
-    \script{wg_IVandV_risks.py}
-    \begin{centering}
-        \includegraphics{figures/wg_IVandV_risks.png}
-        \caption{Distributions over the risk for microscopic involvement in \gls{lnl} IV (left panels) and in \gls{lnl} V (right panels) as predicted by the winning graph model, given early (top panels) or advanced T-category (bottom panels), and different CT-based diagnoses: 1) A clinical N0 patient (green histograms), 2) visible metastases in \gls{lnl} II, but otherwise healthy-looking lymph nodes (blue histograms), 3) macroscopic metastases in the \glspl{lnl} II \& III, with the rest of the neck still being clinically node negative (orange histograms), and finally 4) extensive clinical involvement in the levels II, III, and IV (red histograms).\label{fig:wg_IVandV_risks}}
-    \end{centering}
-\end{figure*}
-
-{\it Level V:} The right panels in \autoref{fig:wg_IVandV_risks} shows the risk of occult disease in level V depending on T-category and the clinical involvement of levels II-IV.
+{\it Level V:} The right panels in \autoref{fig:wg_IVandV_risks} show the risk of occult disease in level V depending on T-category and the clinical involvement of levels II-IV.
 For clinically N0 patients, the risk in level V is estimated to be just above 1\%.

### 2023-12-19 src/scripts/wg_IVandV_risks.py
remove titles again

@@ -35,24 +35,22 @@ if __name__ == "__main__":
             Histogram.from_hdf5(
                 filename=INPUT,
                 dataname=f"{stage}/IV/N0",
-                label="given N0",
+                label="LNL IV, given N0",
                 color=USZ["green"],
             ),
             Histogram.from_hdf5(
                 filename=INPUT,
                 dataname=f"{stage}/IV/II",
-                label="given LNL II involved",
+                label="LNL IV, given LNL II involved",
                 color=USZ["blue"],
             ),
             Histogram.from_hdf5(
                 filename=INPUT,
                 dataname=f"{stage}/IV/IIandIII",
-                label="given LNL II & III involved",
+                label="LNL IV, given LNL II & III involved",
                 color=USZ["orange"],
             ),
         ]
-
-    axes[0,0].set_title("risk for LNL IV involvement", fontweight="bold")
 
     draw(axes[0,0], contents=left_panels["early"], xlims=(0., 7.), hist_kwargs={"nbins": 40})
     axes[0,0].set_ylabel("early T-cat.", fontweight="bold")
@@ -74,24 +72,22 @@ if __name__ == "__main__":
             Histogram.from_hdf5(
                 filename=INPUT,
                 dataname=f"{stage}/V/N0",
-                label="given N0",
+                label="LNL V, given N0",
                 color=USZ["green"],
             ),
             Histogram.from_hdf5(
                 filename=INPUT,
                 dataname=f"{stage}/V/IIandIII",
-                label="given LNL II & III involved",
+                label="LNL V, given LNL II & III involved",
                 color=USZ["orange"],
             ),
             Histogram.from_hdf5(
                 filename=INPUT,
                 dataname=f"{stage}/V/IIandIIIandIV",
-                label="given LNL II, III, IV involved",
+                label="LNL V, given LNL II, III, IV involved",
                 color=USZ["red"],
             ),
         ]
-
-    axes[0,1].set_title("risk for LNL V involvement", fontweight="bold")
 
     draw(axes[0,1], contents=right_panels["early"], xlims=(0., 6.), hist_kwargs={"nbins": 40})
     axes[0,1].set_yticks([])

### 2023-12-19 src/scripts/wg_IV_risks.py
remove unused risk scripts

@@ -1,65 +1,0 @@
-"""
-Plot histograms of predicted risks, given different diagnoses.
-"""
-from pathlib import Path
-
-import matplotlib.pyplot as plt
-from lyscripts.plot.utils import Histogram, draw
-from lyscripts.plot.utils import COLORS as USZ
-from tueplots import figsizes, fontsizes
-
-import paths
-
-
-INPUT = paths.data / "wg_risks.hdf5"
-OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
-NROWS, NCOLS = 2, 1
-
-
-if __name__ == "__main__":
-    plt.rcParams.update(figsizes.icml2022_half(
-        nrows=NROWS,
-        ncols=NCOLS,
-        height_to_width_ratio=0.45,
-    ))
-    plt.rcParams.update(fontsizes.icml2022())
-    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
-
-    panels = {
-        "early": [],
-        "late": [],
-    }
-
-    for stage in ["early", "late"]:
-        panels[stage] = [
-            Histogram.from_hdf5(
-                filename=INPUT,
-                dataname=f"{stage}/IV/N0",
-                label="LNL IV, given N0",
-                color=USZ["green"],
-            ),
-            Histogram.from_hdf5(
-                filename=INPUT,
-                dataname=f"{stage}/IV/II",
-                label="LNL IV, given LNL II involved",
-                color=USZ["blue"],
-            ),
-            Histogram.from_hdf5(
-                filename=INPUT,
-                dataname=f"{stage}/IV/IIandIII",
-                label="LNL IV, given LNL II & III involved",
-                color=USZ["orange"],
-            ),
-        ]
-
-    draw(axes[0], contents=panels["early"], xlims=(0., 7.), hist_kwargs={"nbins": 40})
-    axes[0].set_ylabel("early T-cat.", fontweight="bold")
-    axes[0].set_yticks([])
-    axes[0].legend()
-
-    draw(axes[1], contents=panels["late"], xlims=(0., 7.), hist_kwargs={"nbins": 40})
-    axes[1].set_ylabel("late T-cat.", fontweight="bold")
-    axes[1].set_yticks([])
-    axes[1].set_xlabel("risk [%]")
-
-    plt.savefig(OUTPUT, dpi=300)

### 2023-12-19 src/scripts/wg_V_risks.py

@@ -1,65 +1,0 @@
-"""
-Plot histograms of predicted risks, given different diagnoses.
-"""
-from pathlib import Path
-
-import matplotlib.pyplot as plt
-from lyscripts.plot.utils import Histogram, draw
-from lyscripts.plot.utils import COLORS as USZ
-from tueplots import figsizes, fontsizes
-
-import paths
-
-
-INPUT = paths.data / "wg_risks.hdf5"
-OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
-NROWS, NCOLS = 2, 1
-
-
-if __name__ == "__main__":
-    plt.rcParams.update(figsizes.icml2022_half(
-        nrows=NROWS,
-        ncols=NCOLS,
-        height_to_width_ratio=0.45,
-    ))
-    plt.rcParams.update(fontsizes.icml2022())
-    fig, axes = plt.subplots(nrows=NROWS, ncols=NCOLS, sharex="col", sharey="row")
-
-    panels = {
-        "early": [],
-        "late": [],
-    }
-
-    for stage in ["early", "late"]:
-        panels[stage] = [
-            Histogram.from_hdf5(
-                filename=INPUT,
-                dataname=f"{stage}/V/N0",
-                label="LNL V, given N0",
-                color=USZ["green"],
-            ),
-            Histogram.from_hdf5(
-                filename=INPUT,
-                dataname=f"{stage}/V/IIandIII",
-                label="LNL V, given LNL II & III involved",
-                color=USZ["orange"],
-            ),
-            Histogram.from_hdf5(
-                filename=INPUT,
-                dataname=f"{stage}/V/IIandIIIandIV",
-                label="LNL V, given LNL II, III, IV involved",
-                color=USZ["red"],
-            ),
-        ]
-
-    draw(axes[0], contents=panels["early"], xlims=(0., 6.), hist_kwargs={"nbins": 40})
-    axes[0].set_ylabel("early T-cat.", fontweight="bold")
-    axes[0].set_yticks([])
-    axes[0].legend()
-
-    draw(axes[1], contents=panels["late"], xlims=(0., 6.), hist_kwargs={"nbins": 40})
-    axes[1].set_ylabel("late T-cat.", fontweight="bold")
-    axes[1].set_yticks([])
-    axes[1].set_xlabel("risk [%]")
-
-    plt.savefig(OUTPUT, dpi=300)

### 2023-12-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Each edge represents the conditional probability of metastatic spread per abstract time-step, given disease at the edge's starting node.
-To learn these probabilities, we draw \acrlong{mcmc} samples from the likelihood of a dataset (\variable{output/num_patients.txt} \gls{opscc} patients) from three institutions.
+To learn these probabilities, we draw \acrlong{mcmc} samples from the likelihood of a dataset (\variable{output/num_patients.txt}\gls{opscc} patients) from three institutions.
 We compute the model evidence using \acrlong{ti} for different graphs to determine which describes the data best.
@@
 To determine the optimal underlying \gls{dag} we compare different graphs by calculating the model evidence through \gls{ti}.
-    \item We collect a multi-centric dataset consisting of \variable{output/num_patients.txt} patients from three institutions, allowing us to train the model based on a sizable dataset \citep{ludwig_dataset_2022,ludwig_multi-centric_2023}.
+    \item We collect a multi-centric dataset consisting of \variable{output/num_patients.txt}patients from three institutions, allowing us to train the model based on a sizable dataset \citep{ludwig_dataset_2022,ludwig_multi-centric_2023}.
     \item We use the trained model to provide personalized risk estimations for occult metastases for typical clinical states of tumor progression at diagnosis, illustrating its potential for guiding volume-deescalated treatment strategies in the future.
@@
 However, already before $\beta=0.031$ the model based on the winning graph outperforms the loosing graph model.
-Ultimately, the winning graph's log-evidence is better than the loosing graph model's by a value of around \variable{output/evidence_diff.txt} (see \autoref{table:evidence}).
+Ultimately, the winning graph's log-evidence is better than the loosing graph model's by a value of around \variable{output/evidence_diff.txt}(see \autoref{table:evidence}).
 \label{fig:thermo_int}}
@@
 All figures in this work, like the risk predictions and prevalences we show histograms of, can be reproduced following the instructions of the GitHub repository underlying this publication \href{https://github.com/rmnldwg/graph-extension-paper}{\texttt{rmnldwg/graph-extension-paper}}.
-It is based off of the \showyourwork project.
+It is based off of the \href{https://show-your.work}{showyourwork} project that aims to make scientific papers easily and fully reproducible.
 
@@
 
-In total, the dataset contains \variable{output/num_patients.txt} patients with newly diagnosed \gls{opscc}.
+In total, the dataset contains \variable{output/num_patients.txt}patients with newly diagnosed \gls{opscc}.
 It includes patients treated with definitive (chemo)radiotherapy, adjuvant (chemo)radiotherapy following neck dissection, or neck dissection alone.
@@
 
-The dataset containing the consensus decision for the involvement of each level in every patient was then used for model parameter learning, assuming that it represents an observation of the true hidden state $\boldsymbol{\xi}_k$.
+The dataset containing the consensus decision for the involvement of each level in every patient was then used for model parameter learning.
+We assumed that it represents an observation of the true hidden state $\boldsymbol{\xi}_k$, which corresponds to setting the sensitivity and specificity both to one.
 The frequencies of some of the most important combinations of involved lymph node levels are listed in \autoref{table:data_prevalence}.
@@
 {\it Level II:} As can be seen in \autoref{table:params}, spread from the tumor to \gls{lnl} II to be the most probable transition at any given time step.
-As a consequence, even for an early T-category patient that presents with a clinical N0 neck, our model predicts a \variable{output/wg_II_risks.tex} risk for microscopic metastases in \gls{lnl} II.
+As a consequence, even for an early T-category patient that presents with a clinical N0 neck, our model predicts a \variable{output/wg_II_risks.tex}risk for microscopic metastases in \gls{lnl} II.
 
@@
 For patients with involvement of level II, the risk in level III increases to approximately 9\% and 12\%, respectively.
-
-\begin{figure*}
-    \script{wg_IVandV_risks.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/wg_IVandV_risks.png}
-        \caption{Distributions over the risk for microscopic involvement in \gls{lnl} IV (left panels) and in \gls{lnl} V (right panels) as predicted by the winning graph model, given early (top panels) or advanced T-category (bottom panels), and different CT-based diagnoses: 1) A clinical N0 patient (green histograms), 2) visible metastases in \gls{lnl} II, but otherwise healthy-looking lymph nodes (blue histograms), 3) macroscopic metastases in the \glspl{lnl} II \& III, with the rest of the neck still being clinically node negative (orange histograms), and finally 4) extensive clinical involvement in the levels II, III, and IV (red histograms).\label{fig:wg_IVandV_risks}}
-    \end{centering}
-\end{figure*}
 
@@
 
+\begin{figure*}
+    \script{wg_IVandV_risks.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/wg_IVandV_risks.png}
+        \caption{Distributions over the risk for microscopic involvement in \gls{lnl} IV (left panels) and in \gls{lnl} V (right panels) as predicted by the winning graph model, given early (top panels) or advanced T-category (bottom panels), and different CT-based diagnoses: 1) A clinical N0 patient (green histograms), 2) visible metastases in \gls{lnl} II, but otherwise healthy-looking lymph nodes (blue histograms), 3) macroscopic metastases in the \glspl{lnl} II \& III, with the rest of the neck still being clinically node negative (orange histograms), and finally 4) extensive clinical involvement in the levels II, III, and IV (red histograms).\label{fig:wg_IVandV_risks}}
+    \end{centering}
+\end{figure*}
+
+To illustrate the flexibility of the model in predicting various risks, we have plotted the risk for occult disease in \textit{any} of the \glspl{lnl} I, IV, and/or V, given different clinical diagnoses in \autoref{fig:wg_any_risks}.
+Similar to this, we may compute the risk for an arbitrary combination of involved levels, given a similarly arbitrary clinical diagnosis.
+For the base graph (\href{https://lyprox.org/riskpredictor/9}{\texttt{base-graph-v2}}) and the winning graph (\href{https://lyprox.org/riskpredictor/8}{\texttt{win-graph-v3}}), one may also interactively explore these risks in our web-based interface \href{https://lyprox.org/riskpredictor/list}{LyProX}, similar to how it is possible to explore the underlying data in an interactive way.
+
 \begin{figure}
@@
 
-To illustrate the flexibility of the model in predicting various risks, we have plotted the risk for occult disease in \textit{any} of the \glspl{lnl} I, IV, and/or V, given different clinical diagnoses in \autoref{fig:wg_any_risks}.
-Similar to this, we may compute the risk for an arbitrary combination of involved levels, given a similarly arbitrary clinical diagnosis.
-For the base graph (\href{https://lyprox.org/riskpredictor/9}{\texttt{base-graph-v2}}) and the winning graph (\href{https://lyprox.org/riskpredictor/8}{\texttt{win-graph-v3}}), one may also interactively explore these risks in our web-based interface \href{https://lyprox.org/riskpredictor/list}{LyProX}, similar to how it is possible to explore the underlying data in an interactive way.
-
-\begin{figure}
-    \script{wg_any_risks.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/wg_any_risks.png}
-        \caption{Shown are the histograms over the predicted risk for involvement in any of the \glspl{lnl} I, IV, V, or VII.
-The risk is plotted given a clinical N0 diagnosis (green), macroscopically detected metastases in \gls{lnl} II (blue), and lastly given visible involvement in both \gls{lnl} II and III (orange).
-The top row shows these risks for early T-category diagnoses and the bottom row for advanced T-category.
-\label{fig:wg_any_risks}}
-    \end{centering}
-\end{figure}
-
 \section{Discussion}
@@
 In this publication we present a statistical model of ipsilateral lymph node involvement in oropharyngeal SCC patients.
-Although the basic \gls{hmm} of lymphatic progression has been conceptually introduced in a previous work \citep{ludwig_hidden_2021}, this is the first publication that evaluates the model based on a large multi-institutional dataset containing \variable{output/num_patients.txt} patients.
+Although the basic \gls{hmm} of lymphatic progression has been conceptually introduced in a previous work \citep{ludwig_hidden_2021}, this is the first publication that evaluates the model based on a large multi-institutional dataset containing \variable{output/num_patients.txt}patients.
 It is demonstrated for the first time that the model can accurately describe the patterns of lymph node involvement observed in the data, including the correlations between levels and its dependence on T-category.
@@
 \end{itemize}
+
+\begin{figure}
+    \script{wg_any_risks.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/wg_any_risks.png}
+        \caption{Shown are the histograms over the predicted risk for involvement in any of the \glspl{lnl} I, IV, V, or VII.
+The risk is plotted given a clinical N0 diagnosis (green), macroscopically detected metastases in \gls{lnl} II (blue), and lastly given visible involvement in both \gls{lnl} II and III (orange).
+The top row shows these risks for early T-category diagnoses and the bottom row for advanced T-category.
+\label{fig:wg_any_risks}}
+    \end{centering}
+\end{figure}
 

### 2023-12-20 src/scripts/bg_core_prevs.py
fix ordering of legends

@@ -36,11 +36,6 @@ def draw_early_late_panels(panels, lnl, xlims, ylims=None, idx=0):
 
     draw(early_ax, contents=panels[lnl]["early"], xlims=xlims, hist_kwargs={"nbins": 80})
     draw(late_ax, contents=panels[lnl]["late"], xlims=xlims, hist_kwargs={"nbins": 80})
-    num = len(panels[lnl]["early"]) // 2
-
-    early_handles, early_labels = early_ax.get_legend_handles_labels()
-    sorted_early_handles = [early_handles[i * num + j] for j in range(num) for i in [1,0]]
-    sorted_early_labels = [early_labels[i * num + j] for j in range(num) for i in [1,0]]
 
     if ylims:
         early_ax.set_ylim(ylims)
@@ -48,18 +43,15 @@ def draw_early_late_panels(panels, lnl, xlims, ylims=None, idx=0):
     early_ax.get_xaxis().set_visible(False)
     early_ax.set_yticks([])
     early_ax.set_ylabel("early T-cat.", fontweight="bold")
-    early_ax.legend(sorted_early_handles, sorted_early_labels, labelspacing=0.3)
+    early_ax.legend(labelspacing=0.3)
 
     if ylims:
         late_ax.set_ylim(ylims)
 
-    late_handles, late_labels = late_ax.get_legend_handles_labels()
-    post_late_handles = late_handles[:num]
-    post_late_labels = late_labels[:num]
     late_ax.set_yticks([])
     late_ax.set_xlabel("prevalence [%]")
     late_ax.set_ylabel("late T-cat.", fontweight="bold")
-    late_ax.legend(post_late_handles, post_late_labels, labelspacing=0.3)
+    late_ax.legend(labelspacing=0.3)
 
 
 if __name__ == "__main__":

### 2023-12-20 src/scripts/comp_IandII_prevs.py

@@ -80,23 +80,20 @@ if __name__ == "__main__":
         ]
 
     draw(axes[0,0], contents=left_panels["early"], xlims=(0., 16.), hist_kwargs={"nbins": 40})
-    h, l = axes[0,0].get_legend_handles_labels()
-    handles = [*h[2:4], h[0], *h[4:6], h[1]]
-    labels = [*l[2:4], l[0], *l[4:6], l[1]]
     axes[0,0].set_ylim(bottom=0., top=1.5)
     axes[0,0].set_ylabel("early T-cat.", fontweight="bold")
     axes[0,0].set_yticks([])
-    axes[0,0].legend(handles=handles, labels=labels)
+    axes[0,0].legend()
 
     draw(axes[1,0], contents=left_panels["late"], xlims=(0., 16.), hist_kwargs={"nbins": 40})
     h, l = axes[1,0].get_legend_handles_labels()
-    handles = h[:2]
-    labels = l[:2]
+    h = h[2:6:3]
+    l = l[2:6:3]
     axes[1,0].set_ylim(bottom=0., top=1.2)
     axes[1,0].set_ylabel("late T-cat.", fontweight="bold")
     axes[1,0].set_yticks([])
     axes[1,0].set_xlabel("prevalence [%]")
-    axes[1,0].legend(handles=handles, labels=labels)
+    axes[1,0].legend(handles=h, labels=l)
 
 
     right_panels = {
@@ -150,18 +147,15 @@ if __name__ == "__main__":
         ]
 
     draw(axes[0,1], contents=right_panels["early"], xlims=(50., 90.), hist_kwargs={"nbins": 40})
-    h, l = axes[0,1].get_legend_handles_labels()
-    handles = [*h[2:4], h[0], *h[4:6], h[1]]
-    labels = [*l[2:4], l[0], *l[4:6], l[1]]
     axes[0,1].set_yticks([])
-    axes[0,1].legend(handles=handles, labels=labels)
+    axes[0,1].legend()
 
     draw(axes[1,1], contents=right_panels["late"], xlims=(50., 90.), hist_kwargs={"nbins": 40})
     h, l = axes[1,1].get_legend_handles_labels()
-    handles = h[:2]
-    labels = l[:2]
+    h = h[2:6:3]
+    l = l[2:6:3]
     axes[1,1].set_yticks([])
     axes[1,1].set_xlabel("prevalence [%]")
[…]

### 2023-12-20 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.6\linewidth]{figures/graph_with_obs.png}
+    \caption{\Gls{dag} representing a possible abstraction of the lymphatic network comprising the tumor (red shaded circle) and \glspl{lnl} II through IV as hidden binary \glspl{rv} (blue outlined circles).
+Attached to each of these is the corresponding observed \gls{rv} (orange shaded squares).
+Lymphatic flow is depicted in the form of parameterized arrows (red and blue) that represent the probability of spread along the respective arc per time-step.
+Sensitivity and specificity (orange arrows) connect the hidden \glspl{rv} to the diagnosis.
+\label{fig:graph_with_obs}}
+\end{figure}
+
 \subsection{State of the Hidden Markov Model}
@@
 The variable $b_3$ denotes the probability of lymphatic spread from the tumor to \gls{lnl} III during one time-step, and $t_{23}$ is the probability of spread from an involved level II further down the lymphatic chain into \gls{lnl} III.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.6\linewidth]{figures/graph_with_obs.png}
-    \caption{\Gls{dag} representing a possible abstraction of the lymphatic network comprising the tumor (red shaded circle) and \glspl{lnl} II through IV as hidden binary \glspl{rv} (blue outlined circles).
-Attached to each of these is the corresponding observed \gls{rv} (orange shaded squares).
-Lymphatic flow is depicted in the form of parameterized arrows (red and blue) that represent the probability of spread along the respective arc per time-step.
-Sensitivity and specificity (orange arrows) connect the hidden \glspl{rv} to the diagnosis.
-\label{fig:graph_with_obs}}
-\end{figure}
 
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.6\linewidth]{figures/extended_graph.png}
+    \caption{Extended \gls{dag} representing different possible spread graphs underlying the \gls{hmm}.
+As in \autoref{fig:graph_with_obs}, red arcs are parametrized with probabilities of spread from the tumor (red circle) to the \glspl{lnl} (blue circles).
+These red arcs, together with the blue arcs fron \gls{lnl} to \gls{lnl}, make up the \emph{base graph}.
+One after the other, each of the green arcs was added to the base graph.
+Subsequently, the performance of the resulting models in terms of its \gls{bic} was compared to the base graph to assess whether the additional edge should be kept in the \emph{winning graph} or not.
+\label{fig:extended_graph}}
+\end{figure}
+
 We use \gls{mcmc} sampling to draw parameter samples $\boldsymbol{\hat{\theta}}_i$ for $i \leq S$ from the likelihood described in \autoref{eq:log_likelihood} (i.e.
@@
 \begin{figure}
+    \script{thermo_int.py}
     \centering
-    \includegraphics[width=0.6\linewidth]{figures/extended_graph.png}
-    \caption{Extended \gls{dag} representing different possible spread graphs underlying the \gls{hmm}.
-As in \autoref{fig:graph_with_obs}, red arcs are parametrized with probabilities of spread from the tumor (red circle) to the \glspl{lnl} (blue circles).
-These red arcs, together with the blue arcs fron \gls{lnl} to \gls{lnl}, make up the \emph{base graph}.
-One after the other, each of the green arcs was added to the base graph.
-Subsequently, the performance of the resulting models in terms of its \gls{bic} was compared to the base graph to assess whether the additional edge should be kept in the \emph{winning graph} or not.
-\label{fig:extended_graph}}
+    \includegraphics[width=\linewidth]{figures/thermo_int.png}
+    \caption{The expected log-accuracy under the power posterior plotted against the value of the inverse temperature $\beta$.
+These values were computed during a \gls{ti} run for the loosing graph model (red line), as well as the winning graph model (green line).
+Note that the scaling of the x-axis is chosen such that the 64 points on the temperature ladder appear evenly spaced.
+This is to stress how the log-accuracy develops in the range from $\beta=0$ to around $\beta=0.1$.
+The area under these two curves yields an approximation of the log-evidence of the respective model.
+It is barely visible that due to the simplicity of the loosing graph model its accuracy increases earlier and faster that the winning graph model's.
+However, already before $\beta=0.031$ the model based on the winning graph outperforms the loosing graph model.
+Ultimately, the winning graph's log-evidence is better than the loosing graph model's by a value of around \variable{output/evidence_diff.txt}(see \autoref{table:evidence}).
+\label{fig:thermo_int}}
 \end{figure}
@@
 \label{subsec:complete_model:comparison}
-
-\begin{figure}
-    \script{thermo_int.py}
-    \centering
-    \includegraphics[width=\linewidth]{figures/thermo_int.png}
-    \caption{The expected log-accuracy under the power posterior plotted against the value of the inverse temperature $\beta$.
-These values were computed during a \gls{ti} run for the loosing graph model (red line), as well as the winning graph model (green line).
-Note that the scaling of the x-axis is chosen such that the 64 points on the temperature ladder appear evenly spaced.
-This is to stress how the log-accuracy develops in the range from $\beta=0$ to around $\beta=0.1$.
-The area under these two curves yields an approximation of the log-evidence of the respective model.
-This plot one indicates that due to the simplicity of the loosing graph model its accuracy increases earlier and faster that the winning graph model's.
-However, already before $\beta=0.031$ the model based on the winning graph outperforms the loosing graph model.
-Ultimately, the winning graph's log-evidence is better than the loosing graph model's by a value of around \variable{output/evidence_diff.txt}(see \autoref{table:evidence}).
-\label{fig:thermo_int}}
-\end{figure}
 
@@
 
+\begin{deluxetable}{l|rr}
+    \tablecaption{
+        Literature sensitivity and specificity values that we used to infer the most likely involvement for a patient when multiple diagnostic modalities reported conflicting nodal involvement \citep{de_bondt_Detection_2007,kyzas_18f-fluorodeoxyglucose_2008}.
+        \label{table:sens_spec}
+    }
+    \tablehead{
+        \textbf{Modality} & \textbf{Specificity} & \textbf{Sensitivity}
+    }
+    \startdata
+        \acrshort{ct} & 76\% & 81\% \\
+        \acrshort{pet} & 86\% & 79\% \\
+        \gls{mri} & 63\% & 81\% \\
+        \acrshort{fna} & 98\% & 80\% \\
+        Pathology & $\approx$ 100\% & $\approx$ 100\% \\
+    \enddata
+\end{deluxetable}
+
 The derivatives of the log-evidences $\ln E_\beta$ are essentially expectation values of the data log-likelihood under the power posteriors of the corresponding value for $\beta$.
@@
 For the \glspl{ti} that were performed in this work we used such a fifth order power rule with 64 steps, meaning that $R=63$.
+
+\begin{deluxetable}{ccccccrrrr}
+    \tablecaption{Prevalence of involvement patterns in the multi-centric dataset.
+An involvement pattern is characterized by the state of the six \glspl{lnl}: A red dot means the \gls{lnl} was reported to be metastatic, a green dot means it was determined to be healthy and a question mark means that the prevalence was marginalized over the state of this \gls{lnl}.
+\label{table:data_prevalence}}
+    \tablehead{
+        \multicolumn{6}{c}{\textbf{LNL involvement} } & \multicolumn{4}{c}{\textbf{T-category} } \\
+        I & II & III & IV & V & VII & \multicolumn{2}{c}{\textbf{early} } & \multicolumn{2}{c}{\textbf{advanced} }
+    }
+    \startdata
+    \variable{output/data_table.tex}%
+    \enddata
+\end{deluxetable}
+
+\begin{figure*}[h]
+    \script{bg_core_prevs.py}
+    \begin{centering}
+        \includegraphics[width=\textwidth]{figures/bg_core_prevs.png}
+        \caption{Prevalence of involvement as predicted by the base graph model for different scenarios involving the most commonly metastatic \glspl{lnl} II, III and IV (shaded histograms).
+The model's predictions are compared to Beta posteriors over the prevalence based on the frequency of the same scenarios and a uniform prior (slid lines).
+The top panel shows some selected scenarios with early T-category tumors and the bottom panel with advanced T-category.
+\label{fig:bg_prevalences}}
+    \end{centering}
+\end{figure*}
 
@@
 
-\begin{deluxetable}{ccccccrrrr}
-    \tablecaption{Prevalence of involvement patterns in the multi-centric dataset.
-An involvement pattern is characterized by the state of the six \glspl{lnl}: A red dot means the \gls{lnl} was reported to be metastatic, a green dot means it was determined to be healthy and a question mark means that the prevalence was marginalized over the state of this \gls{lnl}.
-\label{table:data_prevalence}}
-    \tablehead{
-        \multicolumn{6}{c}{\textbf{LNL involvement} } & \multicolumn{4}{c}{\textbf{T-category} } \\
-        I & II & III & IV & V & VII & \multicolumn{2}{c}{\textbf{early} } & \multicolumn{2}{c}{\textbf{advanced} }
-    }
-    \startdata
-    \variable{output/data_table.tex}
-    \enddata
-\end{deluxetable}
-
 The dataset $\boldsymbol{\mathcal{D}}$ that we used for inference is comprised of the detailed reports on lymph node involvement patterns in \gls{opscc} patients treated at three different institutions in France and Switzerland: The \gls{clb} in Lyon (France), the \gls{isb} in Bern (Switzerland), and the \gls{usz} in Zürich (Switzerland).
@@
 The dataset containing the consensus decision for the involvement of each level in every patient was then used for model parameter learning.
-We assumed that it represents an observation of the true hidden state $\boldsymbol{\xi}_k$, which corresponds to setting the sensitivity and specificity both to one.
+We assumed that it represents an observation of the true hidden state $\boldsymbol{\xi}_k$.
 The frequencies of some of the most important combinations of involved lymph node levels are listed in \autoref{table:data_prevalence}.
+\clearpage
 
-\begin{deluxetable}{l|rr}
-    \tablecaption{
-        Literature sensitivity and specificity values that we used to infer the most likely involvement for a patient when multiple diagnostic modalities reported conflicting nodal involvement \citep{de_bondt_Detection_2007,kyzas_18f-fluorodeoxyglucose_2008}.
-        \label{table:sens_spec}
-    }
-    \tablehead{
-        \textbf{Modality} & \textbf{Specificity} & \textbf{Sensitivity}
-    }
-    \startdata
-        \acrshort{ct} & 76\% & 81\% \\
-        \acrshort{pet} & 86\% & 79\% \\
-        \gls{mri} & 63\% & 81\% \\
-        \acrshort{fna} & 98\% & 80\% \\
-        Pathology & $\approx$ 100\% & $\approx$ 100\% \\
-    \enddata
-\end{deluxetable}
+\begin{figure*}[!ht]
+    \centering
+    \includegraphics[width=\linewidth]{figures/ranking.pdf}
+    \caption{Visual ranking of the investigated graphs w.r.t.
+their model evidence, computed via thermodynamic integration.
+Not shown are the two graphs where the arc from \gls{lnl} III to IV was removed.
+Their respective model evidence fell below $-1528$ and the two graphs would appear far left in the figure.
+In the bottom left corner, we provide a visual reference in analogy to \autoref{table:bayes_factor}: E.g., any difference in the model evidence shorter than the first of the three rulers indicates that the improvement is ``barely worth a mention''.
+\label{fig:ranking}}
+\end{figure*}
 
-\section{Results} \label{sec:results}
-
-\begin{figure*}
-    \script{bg_core_prevs.py}
-    \begin{centering}
-        \includegraphics[width=\textwidth]{figures/bg_core_prevs.png}
-        \caption{Prevalence of involvement as predicted by the base graph model for different scenarios involving the most commonly metastatic \glspl{lnl} II, III and IV (shaded histograms).
-The model's predictions are compared to Beta posteriors over the prevalence based on the frequency of the same scenarios and a uniform prior (slid lines).
-The top panel shows some selected scenarios with early T-category tumors and the bottom panel with advanced T-category.
-\label{fig:bg_prevalences}}
-    \end{centering}
-\end{figure*}
+\section{Results}
+\label{sec:results}
 
@@
 
+\begin{deluxetable}{lr}[t]
+    \tablecaption{Model comparison results from the base graph and the extended graph we chose as the ``winning'' model.
+For both \glspl{dag} we show the log-evidence, computed via thermodynamic integration, the negative one half of the \gls{bic}, as well as the maximum log-likelihood that was encountered during the final \gls{mcmc} sampling round.
+\label{table:evidence}}
+    \tablehead{
+        \textbf{graph}  & \textbf{log-evidence}
+    }
+    \startdata
+    \variable{output/metrics_table.tex}%
+    \enddata
+\end{deluxetable}
+
 The model evidences of all candidate graphs are reported in \autoref{table:evidence}.
@@
 This indicates that the two additional parameters are largely independent of each other and manage to describe different aspects of the data.
-
-\autoref{table:evidence} additionally shows the evidence of graphs in which the arc from level II to III or from level III to IV is removed.
-The low model evidence for these graphs confirms the importance of these connections and is consistent with the anatomical motivation.
-The connection from level III to IV is crucial for describing the observation that metastases in level IV are extremely rare without simultaneous involvement of the upstream level III.
-\looseness=-1
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{figures/ranking.pdf}
-    \caption{Visual ranking of the investigated graphs w.r.t.
-their model evidence, computed via thermodynamic integration.
-Not shown are the two graphs where the arc from \gls{lnl} III to IV was removed.
-Their respective model evidence fell below $-1528$ and the two graphs would appear far left in the figure.
-In the bottom left corner, we provide a visual reference in analogy to \autoref{table:bayes_factor}: E.g., any difference in the model evidence shorter than the first of the three rulers indicates that the improvement is ``barely worth a mention''.
-\label{fig:ranking}}
-\end{figure*}
-
-\subsection{The winning graph}
-\label{subsec:results:optimal_graph}
-
-\begin{deluxetable}{lr}
-    \tablecaption{Model comparison results from the base graph and the extended graph we chose as the ``winning'' model.
-For both \glspl{dag} we show the log-evidence, computed via thermodynamic integration, the negative one half of the \gls{bic}, as well as the maximum log-likelihood that was encountered during the final \gls{mcmc} sampling round.
-\label{table:evidence}}
-    \tablehead{
-        \textbf{graph}  & \textbf{log-evidence}
-    }
-    \startdata
-    \variable{output/metrics_table.tex}
-    \enddata
-\end{deluxetable}
-
-The most likely model parameters for the winning graph, corresponding to the mean of the marginals of the sampled posterior distributions, are tabulated in \autoref{table:params}.
-We have fixed $p_\text{early}=0.3$ for early T-category tumors (i.e., T0, T1, and T2), and $t_\text{max}=10$ time steps.
-The result that $t_{\text{II} \rightarrow \text{III}}$ and $t_{\text{III} \rightarrow \text{IV}}$ are relatively large compared to $b_\text{III}$ and $b_\text{IV}$ reflects the observation that skip metastases in levels III and IV without involvement of the upstream level are rare.
-Since level II's parent node (level I) is rarely involved, $b_\text{II}$ can approximately be related to the prevalence of involvement in level II.
-The probability for no involvement of level II when the patient is diagnosed after $t$ time steps is $\left(1-b_\text{II}\right)^t$.
-The prevalence of level II involvement for advanced T-category patients is thus
-%
-\begin{equation}
-    \begin{aligned}
-        \text{prev}_\text{late}^\text{II} &= 1-\sum_{t=0}^{10} \left(1-b_\text{II}\right)^t \cdot p_\text{late}^t \, (1-p_\text{late})^{(10-t)}\binom{10}{t} \\
-        &\approx 79 \%
-    \end{aligned}
-\end{equation}
-%
-which agrees with the second panel from the top in \autoref{fig:bg_prevalences}. 
-The large value for the parameter $t_{\text{I} \rightarrow \text{II}}$ reflects the observation that in almost all patients with level I involvement, level II is also involved.
-The large uncertainty in $t_{\text{I} \rightarrow \text{II}}$ is related to the fact that level I involvement is rare compared to level II.
-
-\begin{deluxetable}{lrr}
-    \tablecaption{Mean and standard deviation of parameters sampled for the winning graph in percent. \label{table:params}}
-    \tablehead{
-        \textbf{parameter} & \textbf{mean} & \textbf{std. dev.}
-    }
-    \startdata
-    \variable{output/means_table.tex}
-    \enddata
-\end{deluxetable}
-
-\subsection{Involvement of levels I and V}
-
-We can observe that the winning graph describes the involvement of levels II, III, and IV equally well as the base graph, a result that is expected and not further shown.
-We thus focus on the improvements w.r.t.
-involvement patterns that include the \glspl{lnl} I and V, that more rarely harbor metastases.
 
@@
 
-{\it Level V:} In \autoref{fig:IIIandIVandV_prevs} we compare the base graph's and the winning graph's estimations for prevalences of involvement patterns that include \gls{lnl} V.
-The base graph underestimates the probability that level IV and V are simultaneously involved, and overestimates the probability that level V but not IV is involved.
-By introducing the arc from level IV to V, the winning graph can describe the observation that level V involvement is typically associated with severe involvement of level II-IV.
-In the dataset, 14 patients out of 62 patients with level IV involvement have metastases in level V (22 \%).
-Instead, only 40 patients out of 624 patients without level IV involvement have metastases in level V (6 \%).
+\autoref{table:evidence} additionally shows the evidence of graphs in which the arc from level II to III or from level III to IV is removed.
+The low model evidence for these graphs confirms the importance of these connections and is consistent with the anatomical motivation.
+The connection from level III to IV is crucial for describing the observation that metastases in level IV are extremely rare without simultaneous involvement of the upstream level III.
+\looseness=-1
 
-{\it Level I:} In \autoref{fig:IandII_prevs}, analogous comparisons are shown for involvement patterns that include \gls{lnl} I.
-The base graph overestimates the probability of level I involvement without simultaneous involvement of level II.
-By introducing the arc from level I to II, the winning graph can capture the correlations between levels I and II.
-It can also be noted that both models overestimate level I involvement for early T-category patients and underestimate level I involvement for advanced T-category patients.
-This is further described in the discussion section below.
+\begin{deluxetable}{lrr}[t]
+    \tablecaption{Mean and standard deviation of parameters sampled for the winning graph in percent. \label{table:params}}
+    \tablehead{
+        \textbf{parameter} & \textbf{mean} & \textbf{std. dev.}
+    }
+    \startdata
+    \variable{output/means_table.tex}%
+    \enddata
+\end{deluxetable}
+
+\subsection{The winning graph}
+\label{subsec:results:optimal_graph}
+
+The most likely model parameters for the winning graph, corresponding to the mean of the marginals of the sampled posterior distributions, are tabulated in \autoref{table:params}.
+We have fixed $p_\text{early}=0.3$ for early T-category tumors (i.e., T0, T1, and T2), and $t_\text{max}=10$ time steps.
+The result that $t_{\text{II} \rightarrow \text{III}}$ and $t_{\text{III} \rightarrow \text{IV}}$ are relatively large compared to $b_\text{III}$ and $b_\text{IV}$ reflects the observation that skip metastases in levels III and IV without involvement of the upstream level are rare.
+Since level II's parent node (level I) is rarely involved, $b_\text{II}$ can approximately be related to the prevalence of involvement in level II.
+The probability for no involvement of level II when the patient is diagnosed after $t$ time steps is $\left(1-b_\text{II}\right)^t$.
+The prevalence of level II involvement for advanced T-category patients is thus
+
+\begin{equation}
+    \begin{aligned}
+        \text{prev}_\text{late}^\text{II} &= 1-\sum_{t=0}^{10} \left(1-b_\text{II}\right)^t \cdot p_\text{late}^t \, (1-p_\text{late})^{(10-t)}\binom{10}{t} \\
+        &\approx 79 \%
+    \end{aligned}
+\end{equation}
+%
+which agrees with the second panel from the top in \autoref{fig:bg_prevalences}. 
+The large value for the parameter $t_{\text{I} \rightarrow \text{II}}$ reflects the observation that in almost all patients with level I involvement, level II is also involved.
+The large uncertainty in $t_{\text{I} \rightarrow \text{II}}$ is related to the fact that level I involvement is rare compared to level II.
+
+\subsection{Involvement of levels I and V}
+
+We can observe that the winning graph describes the involvement of levels II, III, and IV equally well as the base graph, a result that is expected and not further shown.
+We thus focus on the improvements w.r.t.
+involvement patterns that include the \glspl{lnl} I and V, that more rarely harbor metastases.
 
@@
 
+{\it Level V:} In \autoref{fig:IIIandIVandV_prevs} we compare the base graph's and the winning graph's estimations for prevalences of involvement patterns that include \gls{lnl} V.
+The base graph underestimates the probability that level IV and V are simultaneously involved, and overestimates the probability that level V but not IV is involved.
+By introducing the arc from level IV to V, the winning graph can describe the observation that level V involvement is typically associated with severe involvement of level II-IV.
+In the dataset, 14 patients out of 62 patients with level IV involvement have metastases in level V (22 \%).
+Instead, only 40 patients out of 624 patients without level IV involvement have metastases in level V (6 \%).
+
+{\it Level I:} In \autoref{fig:IandII_prevs}, analogous comparisons are shown for involvement patterns that include \gls{lnl} I.
+The base graph overestimates the probability of level I involvement without simultaneous involvement of level II.
+By introducing the arc from level I to II, the winning graph can capture the correlations between levels I and II.
+It can also be noted that both models overestimate level I involvement for early T-category patients and underestimate level I involvement for advanced T-category patients.
+This is further described in the discussion section below.
+
 \subsection{Risk Prediction for Occult Disease}
@@
 We assume a sensitivity of 0.76 and a specificity of 0.81 for the clinical diagnosis of lymph node metastases, corresponding to CT imaging in \autoref{table:sens_spec}.
+
+{\it Level II:} As can be seen in \autoref{table:params}, spread from the tumor to \gls{lnl} II to be the most probable transition at any given time step.
+As a consequence, even for an early T-category patient that presents with a clinical N0 neck, our model predicts a \variable{output/wg_II_risks.tex}risk for microscopic metastases in \gls{lnl} II.
 
@@
 
-{\it Level II:} As can be seen in \autoref{table:params}, spread from the tumor to \gls{lnl} II to be the most probable transition at any given time step.
-As a consequence, even for an early T-category patient that presents with a clinical N0 neck, our model predicts a \variable{output/wg_II_risks.tex}risk for microscopic metastases in \gls{lnl} II.
-
 {\it Level III:} \autoref{fig:wg_III_risks} compares the risk of occult disease in level III between patients that are clinically N0 (orange) and patients with clinically diagnosed involvement of only level II (red), for early T-category (upper panel) and advanced T-category (bottom panel).
@@
 For patients with involvement of level II, the risk in level III increases to approximately 9\% and 12\%, respectively.
+
+\begin{figure*}
+    \script{wg_IVandV_risks.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/wg_IVandV_risks.png}
+        \caption{Distributions over the risk for microscopic involvement in \gls{lnl} IV (left panels) and in \gls{lnl} V (right panels) as predicted by the winning graph model, given early (top panels) or advanced T-category (bottom panels), and different CT-based diagnoses: 1) A clinical N0 patient (green histograms), 2) visible metastases in \gls{lnl} II, but otherwise healthy-looking lymph nodes (blue histograms), 3) macroscopic metastases in the \glspl{lnl} II \& III, with the rest of the neck still being clinically node negative (orange histograms), and finally 4) extensive clinical involvement in the levels II, III, and IV (red histograms).\label{fig:wg_IVandV_risks}}
+    \end{centering}
+\end{figure*}
 
@@
 The correlation between the clinical involvement pattern, the likely time of diagnosis in the tumor's time frame based on it, and from that the risk of involvement is another benefit of the formulation of the model as a \gls{hmm}.
-
-\begin{figure*}
-    \script{wg_IVandV_risks.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/wg_IVandV_risks.png}
-        \caption{Distributions over the risk for microscopic involvement in \gls{lnl} IV (left panels) and in \gls{lnl} V (right panels) as predicted by the winning graph model, given early (top panels) or advanced T-category (bottom panels), and different CT-based diagnoses: 1) A clinical N0 patient (green histograms), 2) visible metastases in \gls{lnl} II, but otherwise healthy-looking lymph nodes (blue histograms), 3) macroscopic metastases in the \glspl{lnl} II \& III, with the rest of the neck still being clinically node negative (orange histograms), and finally 4) extensive clinical involvement in the levels II, III, and IV (red histograms).\label{fig:wg_IVandV_risks}}
-    \end{centering}
-\end{figure*}
 
@@
 
-\begin{figure}
+\begin{figure}[b]
     \script{wg_any_risks.py}
@@
 
+\begin{figure}[t]
+    \script{wg_sens_spec_risks.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/wg_sens_spec_risks.png}
+        \caption{Dependence of risk of occult disease on sensitivity and specificity.
+The figure shows the risk of involvement in level IV, given the mean of the parameter samples, drawn during sampling, in patients with clinical involvement of levels II and III.
+\label{fig:wg_sens_spec_risks}}
+    \end{centering}
+\end{figure}
+
 Also, as described in \autoref{sec:data}, we assumed the consensus of the data to represent the true state of nodal involvement.
@@
 
-\begin{figure}
-    \script{wg_sens_spec_risks.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/wg_sens_spec_risks.png}
-        \caption{Dependence of risk of occult disease on sensitivity and specificity.
-The figure shows the risk of involvement in level IV, given the mean of the parameter samples, drawn during sampling, in patients with clinical involvement of levels II and III.
-\label{fig:wg_sens_spec_risks}}
-    \end{centering}
-\end{figure}
-
 \section{Acknowledgments}

### 2023-12-21 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Affiliations
-\def \USZ {University Hospital Zurich, Rämistrasse 100, 8091 Zurich, Switzerland}
-\def \UZH {University of Zurich, Rämistrasse 71, 8006 Zurich, Switzerland}
-\def \ISB {Inselspital Bern, Freiburgstrasse 18, 3010 Bern, Switzerland}
-\def \CLB {Centre L\'eon B\'erard, 28 Rue Laennec, 69008 Lyon, France}
-\def \KS  {Klinikum Stuttgart, Kriegsbergstraße 60, 70174 Stuttgart, Germany}
+\def \USZ {Dep. of Radiation Oncology, University Hospital Zurich, Rämistrasse 100, 8091 Zurich (CH)}
+\def \UZH {Dep. of Physics, University of Zurich,  Rämistrasse 71, 8006 Zurich (CH)}
+
+\def \ISBrao {Dep. of Radiation Oncology, Bern University Hospital, University of Bern, Freiburgstrasse 18, 3010, Bern (CH)}
+\def \ISBsurgery {Dep. of ENT, Head \& Neck Surgery, Inselspital, Bern University Hospital, University of Bern, Freiburgstrasse 18, 3010 Bern (CH)}
+\def \ISBcancer {Head and Neck Anticancer Center, Bern University Hospital, University of Bern, Freiburgstrasse 18, 3010, Bern (CH)}
+\def \ISBpath {Institute of Tissue Medicine and Pathology, Bern University Hospital, University of Bern, Murtenstrasse 31, 3008 Bern (CH)}
+
+\def \RHNe {Dep. of ENT, Head \& Neck Surgery, R\'eseau Hospitalier Neuch\^atelois (RHNe),  Neuch\^atel (CH)}
+
+\def \CLBrao {Dep. of Radiation Oncology, Centre L\'eon B\'erard, 28 Rue Laennec, 69008 Lyon (FR)}
+\def \CLBsurgery {Dep. of Head and Neck surgery, Centre L\'eon B\'erard, 28 Rue Laennec, 69008 Lyon (FR)}
+
+\def \KS {Institute of Pathology, Klinikum Stuttgart, Kriegsbergstr. 60c, 70174 Stuttgart (GER)}
 
@@
 \author{Adrian Schubert}
-\affiliation{\ISB}
+\affiliation{\RHNe}
+\affiliation{\ISBsurgery}
+\affiliation{\ISBcancer}
 
 \author{Dorothea Barbatei}
-\affiliation{\CLB}
+\affiliation{\CLBrao}
 
 \author{Lauence Bauwens}
-\affiliation{\CLB}
+\affiliation{\CLBrao}
 
@@
 \author{Sandrine Werlen}
-\affiliation{\ISB}
+\affiliation{\ISBsurgery}
+\affiliation{\ISBcancer}
 
 \author[0000-0002-6996-0646]{Olgun Elicin}
-\affiliation{\ISB}
+\affiliation{\ISBrao}
 
@@
 \affiliation{\KS}
+\affiliation{\ISBpath}
 
 \author[0000-0001-6223-6113]{Philippe Zrounba}
-\affiliation{\CLB}
+\affiliation{\CLBsurgery}
 
@@
 \author{Vincent Grégoire}
-\affiliation{\CLB}
+\affiliation{\CLBrao}
 
 \author[0000-0002-5574-3210]{Roland Giger}
-\affiliation{\ISB}
+\affiliation{\ISBsurgery}
+\affiliation{\ISBcancer}
 
@@
 These $S$ parameter estimates are then used to compute expectation values of estimates that depend on the parameters $\boldsymbol{\theta}$ through an integral over the parameter space $\boldsymbol{\Theta}$:
-%
+
 \begin{equation}
@@
 
+\begin{figure}[t]
+    \script{thermo_int.py}
+    \centering
+    \includegraphics[width=\linewidth]{figures/thermo_int.png}
+    \caption{The expected log-accuracy under the power posterior plotted against the value of the inverse temperature $\beta$.
+These values were computed during a \gls{ti} run for the loosing graph model (red line), as well as the winning graph model (green line).
+Note that the scaling of the x-axis is chosen such that the 64 points on the temperature ladder appear evenly spaced.
+This is to stress how the log-accuracy develops in the range from $\beta=0$ to around $\beta=0.1$.
+The area under these two curves yields an approximation of the log-evidence of the respective model.
+It is barely visible that due to the simplicity of the loosing graph model its accuracy increases earlier and faster that the winning graph model's.
+However, already before $\beta=0.031$ the model based on the winning graph outperforms the loosing graph model.
+Ultimately, the winning graph's log-evidence is better than the loosing graph model's by a value of around \variable{output/evidence_diff.txt}(see \autoref{table:evidence}).
+\label{fig:thermo_int}}
+\end{figure}
+
 \subsection{Risk estimation}
@@
 Using Bayes' theorem, we get
-%
+
 \begin{equation}
@@
 \label{subsec:complete_model:graphs}
-
-\begin{figure}
-    \script{thermo_int.py}
-    \centering
-    \includegraphics[width=\linewidth]{figures/thermo_int.png}
-    \caption{The expected log-accuracy under the power posterior plotted against the value of the inverse temperature $\beta$.
-These values were computed during a \gls{ti} run for the loosing graph model (red line), as well as the winning graph model (green line).
-Note that the scaling of the x-axis is chosen such that the 64 points on the temperature ladder appear evenly spaced.
-This is to stress how the log-accuracy develops in the range from $\beta=0$ to around $\beta=0.1$.
-The area under these two curves yields an approximation of the log-evidence of the respective model.
-It is barely visible that due to the simplicity of the loosing graph model its accuracy increases earlier and faster that the winning graph model's.
-However, already before $\beta=0.031$ the model based on the winning graph outperforms the loosing graph model.
-Ultimately, the winning graph's log-evidence is better than the loosing graph model's by a value of around \variable{output/evidence_diff.txt}(see \autoref{table:evidence}).
-\label{fig:thermo_int}}
-\end{figure}
 
@@
 Subsequently, we can write the logarithm of the evidence as an integral over a derivative:
-%
+
 \begin{equation} \label{eq:ti}
@@
 The frequencies of some of the most important combinations of involved lymph node levels are listed in \autoref{table:data_prevalence}.
-\clearpage
 
@@
 
-\begin{deluxetable}{lrr}[t]
-    \tablecaption{Mean and standard deviation of parameters sampled for the winning graph in percent. \label{table:params}}
-    \tablehead{
-        \textbf{parameter} & \textbf{mean} & \textbf{std. dev.}
-    }
-    \startdata
-    \variable{output/means_table.tex}%
-    \enddata
-\end{deluxetable}
-
 \subsection{The winning graph}
@@
 The prevalence of level II involvement for advanced T-category patients is thus
-
+%
 \begin{equation}
@@
 \end{equation}
-%
+
 which agrees with the second panel from the top in \autoref{fig:bg_prevalences}. 
@@
 
+\begin{deluxetable}{lrr}[b]
+    \tablecaption{Mean and standard deviation of parameters sampled for the winning graph in percent. \label{table:params}}
+    \tablehead{
+        \textbf{parameter} & \textbf{mean} & \textbf{std. dev.}
+    }
+    \startdata
+    \variable{output/means_table.tex}%
+    \enddata
+\end{deluxetable}
+
 \subsection{Risk Prediction for Occult Disease}
@@
 
-\begin{figure}
+\begin{figure}[b]
     \script{wg_III_risks.py}
@@
 
-\section{Acknowledgments}
+\begin{acknowledgments}
 
 This work was supported by the Swiss Cancer Research Foundation under grant \href{https://www.krebsforschung.ch/unterstuetzen-sie-uns/stiftungen/-dl-/fileadmin/downloads/unterstuetzen-sie-uns/projekte-der-stiftung-krebsforschung-schweiz-2023.pdf}{KFS 5645-08-2022} and by the University Zürich under the Clinical Research Priority Program \href{https://www.crpp-ai-oncology.uzh.ch/en/Projects/Project-5.html}{Artificial Intelligence in Oncological Imaging}.
+
+\end{acknowledgments}
+
+\clearpage
 

### 2024-02-05 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
+\vspace{5mm}
+
 \subsection{Reproducibility}
@@
 
-{\it Level III:} \autoref{fig:wg_III_risks} compares the risk of occult disease in level III between patients that are clinically N0 (orange) and patients with clinically diagnosed involvement of only level II (red), for early T-category (upper panel) and advanced T-category (bottom panel).
-The histograms represent the uncertainty in the model's risk prediction arising from the uncertainty in the model parameters and are generated by randomly drawing a tenth of the samples from the model parameter's joint posterior distribution.
-This amounts to $S = 20 \cdot W$ samples, as described in \autoref{subsec:complete_model:comparison}.
-The model predicts a risk of just below 6\% for early T-category tumors and 8\% for T3 or T4 ones.
-For patients with involvement of level II, the risk in level III increases to approximately 9\% and 12\%, respectively.
-
 \begin{figure*}
@@
 \end{figure*}
+
+{\it Level III:} \autoref{fig:wg_III_risks} compares the risk of occult disease in level III between patients that are clinically N0 (orange) and patients with clinically diagnosed involvement of only level II (red), for early T-category (upper panel) and advanced T-category (bottom panel).
+The histograms represent the uncertainty in the model's risk prediction arising from the uncertainty in the model parameters and are generated by randomly drawing a tenth of the samples from the model parameter's joint posterior distribution.
+This amounts to $S = 20 \cdot W$ samples, as described in \autoref{subsec:complete_model:comparison}.
+The model predicts a risk of just below 6\% for early T-category tumors and 8\% for T3 or T4 ones.
+For patients with involvement of level II, the risk in level III increases to approximately 9\% and 12\%, respectively.
 
@@
 For the base graph (\href{https://lyprox.org/riskpredictor/9}{\texttt{base-graph-v2}}) and the winning graph (\href{https://lyprox.org/riskpredictor/8}{\texttt{win-graph-v3}}), one may also interactively explore these risks in our web-based interface \href{https://lyprox.org/riskpredictor/list}{LyProX}, similar to how it is possible to explore the underlying data in an interactive way.
+
+\clearpage
 
@@
 
+\section*{Author Contributions}
+
+RL and JU wrote the main manuscript.
+RL created all figures and tables.
+AS, SW, OE, MD, and RG collected the \gls{isb} patient data.
+DB, LB, PZ, and VG collected the \gls{clb} datasets.
+JMH, BP, PB, RL, and JU collected the table of \gls{usz} patients.
+All authors reviewed the manuscript.
+
+\section*{Data Availability}
+
+The patient data detailing lymphatic involvement is publicly available in the form of CSV tables in the GitHub repository \href{https://github.com/rmnldwg/lydata}{\texttt{rmnldwg/lydata}}.
+
+In another GitHub repository, \href{https://github.com/rmnldwg/lynference}{\texttt{rmnldwg/lynference}}, we define the experiments based on different graph structures.
+These experiments are reproducible via the tool \href{https://dvc.org}{DVC} and their locally computed results are uploaded to an Azure blob storage container.
+
+Lastly, the \LaTeX\ source code, all Python scripts to generate the figures and tables in this work, and the pipeline definition to build the document (using \showyourwork) are made public in the GitHub repository \href{https://github.com/rmnldwg/graph-extension-paper}{\texttt{rmnldwg/graph-extension-paper}}.
+To rebuild it, the \showyourwork tool pulls the patient data from the \href{https://github.com/rmnldwg/lynference}{\texttt{rmnldwg/lynference}} repository and the trained models in the form of \gls{mcmc} samples from the Azure blob storage container, then runs the Python scripts and compiles the \LaTeX\ article.
+
+Detailed instructions on how to reproduce the entire pipeline, including the individual experiments, are gladly given upon request.
+They may also be found in the respective repository's \texttt{README.md} file.
+
+\section*{Competing Interests}
+
+The authors declare no competing interests.
+
 \begin{acknowledgments}

### 2024-02-05 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-Lastly, the \LaTeX\ source code, all Python scripts to generate the figures and tables in this work, and the pipeline definition to build the document (using \showyourwork) are made public in the GitHub repository \href{https://github.com/rmnldwg/graph-extension-paper}{\texttt{rmnldwg/graph-extension-paper}}.
+Lastly, the \LaTeX\ source code, all Python scripts to generate the figures and tables in this work, and the pipeline definition to build the document (using \showyourwork \cite{luger_mapping_2021}) are made public in the GitHub repository \href{https://github.com/rmnldwg/graph-extension-paper}{\texttt{rmnldwg/graph-extension-paper}}.
 To rebuild it, the \showyourwork tool pulls the patient data from the \href{https://github.com/rmnldwg/lynference}{\texttt{rmnldwg/lynference}} repository and the trained models in the form of \gls{mcmc} samples from the Azure blob storage container, then runs the Python scripts and compiles the \LaTeX\ article.

### 2024-02-05 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\vspace{5mm}
-
-\subsection{Reproducibility}
-\label{subsec:complete_model:reproducibility}
-
-The entire methodology used in this work is publicly available in the GitHub repository \href{https://github.com/rmnldwg/lynference}{\texttt{rmnldwg/lynference}}.
-Tagged references to specific versions of an inference pipeline allow reproducing the inferred parameters of the models described here.
-Every parameter necessary to reproduce such a pipeline is specified there in designated configuration files.
-It also defines the sequences of computations that constitute the pipeline that are mostly calls to commands of the \href{https://pypi.org/project/lyscripts/}{\texttt{lyscripts}} we published.
-
-All figures in this work, like the risk predictions and prevalences we show histograms of, can be reproduced following the instructions of the GitHub repository underlying this publication \href{https://github.com/rmnldwg/graph-extension-paper}{\texttt{rmnldwg/graph-extension-paper}}.
-It is based off of the \href{https://show-your.work}{showyourwork} project that aims to make scientific papers easily and fully reproducible.
-
 \section{Multicentric Dataset}
@@
 
-\clearpage
-
 \begin{figure}

### 2024-02-05 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Define document class
-\documentclass[twocolumn]{aastex631}
+\documentclass[twocolumn]{article}
+\usepackage[a4paper, total={190mm, 270mm}]{geometry}
 \usepackage{showyourwork}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{wasysym}
+\usepackage{xcolor}
+\usepackage{nicefrac}
 \usepackage{multirow}
-\usepackage{nicefrac}
-\usepackage{wasysym}
-\usepackage{amssymb}
-\usepackage[inline]{enumitem}
-\usepackage{booktabs}
+\usepackage[affil-it]{authblk}
+\usepackage{graphicx}
+\usepackage{biblatex}
+\addbibresource{bib.bib}
 
@@
 
-% Begin!
-\begin{document}
+\usepackage{hyperref}
+\usepackage{cleveref}
 
@@
 
+% Author list
+\author[1,2,*]{Roman Ludwig}
+\author[4,5,7]{Adrian Schubert}
+\author[8]{Dorothea Barbatei}
+\author[8]{Lauence Bauwens}
+\author[1]{Jean-Marc Hoffmann}
+\author[4,5]{Sandrine Werlen}
+\author[3]{Olgun Elicin}
+
+\author[6,10]{Matthias Dettmer}
+\author[9]{Philippe Zrounba}
+\author[1]{Bertrand Pouymayou}
+\author[1]{Panagiotis Balermpas}
+\author[8]{Vincent Grégoire}
+\author[4,5]{Roland Giger}
+\author[1,2]{Jan Unkelbach}
+
 % Affiliations
-\def \USZ {Dep. of Radiation Oncology, University Hospital Zurich, Rämistrasse 100, 8091 Zurich (CH)}
-\def \UZH {Dep. of Physics, University of Zurich,  Rämistrasse 71, 8006 Zurich (CH)}
+\affil[1]{\footnotesize Dep. of Radiation Oncology, University Hospital Zurich, Rämistrasse 100, 8091 Zurich (CH)}
+\affil[2]{\footnotesize Dep. of Physics, University of Zurich,  Rämistrasse 71, 8006 Zurich (CH)}
+\affil[3]{\footnotesize Dep. of Radiation Oncology, Bern University Hospital, University of Bern, Freiburgstrasse 18, 3010, Bern (CH)}
+\affil[4]{\footnotesize Dep. of ENT, Head \& Neck Surgery, Inselspital, Bern University Hospital, University of Bern, Freiburgstrasse 18, 3010 Bern (CH)}
+\affil[5]{\footnotesize Head and Neck Anticancer Center, Bern University Hospital, University of Bern, Freiburgstrasse 18, 3010, Bern (CH)}
+\affil[6]{\footnotesize Institute of Tissue Medicine and Pathology, Bern University Hospital, University of Bern, Murtenstrasse 31, 3008 Bern (CH)}
+\affil[7]{\footnotesize Dep. of ENT, Head \& Neck Surgery, R\'eseau Hospitalier Neuch\^atelois (RHNe),  Neuch\^atel (CH)}
+\affil[8]{\footnotesize Dep. of Radiation Oncology, Centre L\'eon B\'erard, 28 Rue Laennec, 69008 Lyon (FR)}
+\affil[9]{\footnotesize Dep. of Head and Neck surgery, Centre L\'eon B\'erard, 28 Rue Laennec, 69008 Lyon (FR)}
+\affil[10]{\footnotesize Institute of Pathology, Klinikum Stuttgart, Kriegsbergstr. 60c, 70174 Stuttgart (GER)}
+\affil[*]{\footnotesize roman.ludwig@usz.ch}
 
-\def \ISBrao {Dep. of Radiation Oncology, Bern University Hospital, University of Bern, Freiburgstrasse 18, 3010, Bern (CH)}
-\def \ISBsurgery {Dep. of ENT, Head \& Neck Surgery, Inselspital, Bern University Hospital, University of Bern, Freiburgstrasse 18, 3010 Bern (CH)}
-\def \ISBcancer {Head and Neck Anticancer Center, Bern University Hospital, University of Bern, Freiburgstrasse 18, 3010, Bern (CH)}
-\def \ISBpath {Institute of Tissue Medicine and Pathology, Bern University Hospital, University of Bern, Murtenstrasse 31, 3008 Bern (CH)}
-
-\def \RHNe {Dep. of ENT, Head \& Neck Surgery, R\'eseau Hospitalier Neuch\^atelois (RHNe),  Neuch\^atel (CH)}
-
-\def \CLBrao {Dep. of Radiation Oncology, Centre L\'eon B\'erard, 28 Rue Laennec, 69008 Lyon (FR)}
-\def \CLBsurgery {Dep. of Head and Neck surgery, Centre L\'eon B\'erard, 28 Rue Laennec, 69008 Lyon (FR)}
-
-\def \KS {Institute of Pathology, Klinikum Stuttgart, Kriegsbergstr. 60c, 70174 Stuttgart (GER)}
-
-% Author list
-\author[0000-0001-9434-328X]{Roman Ludwig}
-\affiliation{\USZ}
-\affiliation{\UZH}
-
-\author{Adrian Schubert}
-\affiliation{\RHNe}
-\affiliation{\ISBsurgery}
-\affiliation{\ISBcancer}
-
-\author{Dorothea Barbatei}
-\affiliation{\CLBrao}
-
-\author{Lauence Bauwens}
-\affiliation{\CLBrao}
-
-\author[0000-0003-1193-0863]{Jean-Marc Hoffmann}
-\affiliation{\USZ}
-
-\author{Sandrine Werlen}
-\affiliation{\ISBsurgery}
-\affiliation{\ISBcancer}
-
-\author[0000-0002-6996-0646]{Olgun Elicin}
-\affiliation{\ISBrao}
-
-\author[0000-0003-0948-1392]{Matthias Dettmer}
-\affiliation{\KS}
-\affiliation{\ISBpath}
-
-\author[0000-0001-6223-6113]{Philippe Zrounba}
-\affiliation{\CLBsurgery}
-
-\author[0000-0002-8060-7827]{Bertrand Pouymayou}
-\affiliation{\USZ}
-
-\author[0000-0001-5261-6446]{Panagiotis Balermpas}
-\affiliation{\USZ}
-
-\author{Vincent Grégoire}
-\affiliation{\CLBrao}
-
-\author[0000-0002-5574-3210]{Roland Giger}
-\affiliation{\ISBsurgery}
-\affiliation{\ISBcancer}
-
-\author[0000-0002-4275-990X]{Jan Unkelbach}
-\affiliation{\USZ}
-\affiliation{\UZH}
-
-\correspondingauthor{Roman Ludwig}
-\email{roman.ludwig@usz.ch}
+% Begin!
+\begin{document}
 
 % Abstract
+\twocolumn[
+  \begin{@twocolumnfalse}
+    \maketitle
         \begin{abstract}
@@
         \end{abstract}
+        \vspace{1cm}
+  \end{@twocolumnfalse}
+]
 
@@
 This work concerns itself with \gls{opscc}, where approximately 70-80\% of patients present with lymph node metastases at the time of diagnosis.
-In clinical practice, \gls{ctv-n} definition in radiotherapy is based on guidelines \citep{gregoire_ct-based_2003,gregoire_delineation_2014,gregoire_delineation_2018,eisbruch_intensity-modulated_2002,biau_selection_2019,chao_determination_2002,vorwerk_guidelines_2011,ferlito_elective_2009} that are mostly derived from the observed prevalence of involvement in an \gls{lnl} for a given tumor location.
+In clinical practice, \gls{ctv-n} definition in radiotherapy is based on guidelines \cite{gregoire_ct-based_2003,gregoire_delineation_2014,gregoire_delineation_2018,eisbruch_intensity-modulated_2002,biau_selection_2019,chao_determination_2002,vorwerk_guidelines_2011,ferlito_elective_2009} that are mostly derived from the observed prevalence of involvement in an \gls{lnl} for a given tumor location.
 These guidelines currently suggest extensive irradiation of both sides of the neck for most patients.
@@
 We previously developed a model of lymphatic metastatic progression for estimating the risk of microscopic disease, given a patient's personal diagnosis.
-The initial model was based on the methodology of \glspl{bn} \citep{pouymayou_bayesian_2019}.
-It was subsequently extended and formulated as an \gls{hmm} to include T-category in an intuitive manner \citep{ludwig_hidden_2021}.
-However, these models were introduced based on only a small dataset of approximately 100 early T-category patients available at that time \citep{sanguineti_defining_2009}.
+The initial model was based on the methodology of \glspl{bn} \cite{pouymayou_bayesian_2019}.
+It was subsequently extended and formulated as an \gls{hmm} to include T-category in an intuitive manner \cite{ludwig_hidden_2021}.
+However, these models were introduced based on only a small dataset of approximately 100 early T-category patients available at that time \cite{sanguineti_defining_2009}.
 The limited data did not allow us to quantify the probability of metastases in the rarely involved \glspl{lnl} I, V, and VII, nor did the data allow us to verify that the \gls{hmm} is adequate to describe the dependence on lymph node involvement on T-category.
-In this paper, we extend the previous work \citep{ludwig_hidden_2021} by making the following contributions:
+In this paper, we extend the previous work \cite{ludwig_hidden_2021} by making the following contributions:
 \begin{enumerate}
@@
 To determine the optimal underlying \gls{dag} we compare different graphs by calculating the model evidence through \gls{ti}.
-    \item We collect a multi-centric dataset consisting of \variable{output/num_patients.txt}patients from three institutions, allowing us to train the model based on a sizable dataset \citep{ludwig_dataset_2022,ludwig_multi-centric_2023}.
+    \item We collect a multi-centric dataset consisting of \variable{output/num_patients.txt}patients from three institutions, allowing us to train the model based on a sizable dataset \cite{ludwig_dataset_2022,ludwig_multi-centric_2023}.
     \item We use the trained model to provide personalized risk estimations for occult metastases for typical clinical states of tumor progression at diagnosis, illustrating its potential for guiding volume-deescalated treatment strategies in the future.
@@
 
-We have introduced a probabilistic model for lymph node involvement based on \glspl{bn} in \citep{pouymayou_bayesian_2019}.
-The model was extended using \glspl{hmm} in \citep{ludwig_hidden_2021}.
+We have introduced a probabilistic model for lymph node involvement based on \glspl{bn} in \cite{pouymayou_bayesian_2019}.
+The model was extended using \glspl{hmm} in \cite{ludwig_hidden_2021}.
 We will briefly recap the \acrlong{hmm} to introduce the notation used throughout the work.
@@
 The term $P\left( \boldsymbol{\xi}_i \mid \boldsymbol{\xi}_j \right)$ describes the probability to transition from the hidden state of lymphatic involvement $\boldsymbol{\xi}_j$ to the state $\boldsymbol{\xi}_i$ between the time $t$ and $t+1$.
-Using a \gls{dag} as depicted in \autoref{fig:graph_with_obs}, we can formulate this transition probability in the following way:
+Using a \gls{dag} as depicted in \cref{fig:graph_with_obs}, we can formulate this transition probability in the following way:
 %
@@
 
-The terms of the form $P \left( \xi_{iv} \mid \left\{ \xi_{jr} \right\}_{r \in \operatorname{pa}(v)} \right)$ implicitly depend on how we parameterize the arcs of \autoref{fig:graph_with_obs}.
+The terms of the form $P \left( \xi_{iv} \mid \left\{ \xi_{jr} \right\}_{r \in \operatorname{pa}(v)} \right)$ implicitly depend on how we parameterize the arcs of \cref{fig:graph_with_obs}.
 For example, if we look at the probability of spread to \gls{lnl} III ($X_3$) depending on the state of that level's parent -- which, in this case, is $\operatorname{pa}(3) = 2$ -- we can write the different combinations into a conditional probability table as below.
 
-\begin{deluxetable}{cc|cc}
-\tablecaption{$P\left( X_3 \mid X_2 \right)$ for all possible combinations of $X_3$ and $X_2$.}
-\tablehead{& & \colhead{$X_2=0$} & \colhead{$X_2=1$}}
-\startdata
-\begin{tabular}{@{}c@{}}
-\multirow{2}{*}{$X_3$}
-\end{tabular} & $=0$ & $1-b_3$ & $(1-b_3)(1-t_{23})$ \\
+\begin{table}
+\centering
+\begin{tabular}{|cc|cc|}
+    \hline
+    & & $X_2=0$ & $X_2=1$ \\
+    \hline
+    \multirow{2}{*}{$X_3$} & $=0$ & $1-b_3$ & $(1-b_3)(1-t_{23})$ \\
     & $=1$ & $b_3$ & $1-b_3-t_{23}+b_3 t_{23}$ \\
-\enddata
-\end{deluxetable}
+    \hline
+\end{tabular}
+\caption{$P\left( X_3 \mid X_2 \right)$ for all four possible combinations of the states of $X_3$ and $X_2$.}
+\end{table}
 
@@
 
-\begin{deluxetable}{cc|cc}
-\tablecaption{$P\left( Z \mid X \right)$ for all possible values of $X$ and $Z$.}
-\tablehead{& & \colhead{$X=0$} & \colhead{$X=1$}}
-\startdata
-\begin{tabular}{@{}c@{}}
-\multirow{2}{*}{$Z$}
-\end{tabular} & $=0$ & $s_P$ & $1-s_N$ \\
+\begin{table}
+\centering
+\begin{tabular}{|cc|cc|}
+    \hline
+    & & $X=0$ & $X=1$ \\
+    \hline
+    \multirow{2}{*}{$Z$} & $=0$ & $s_P$ & $1-s_N$ \\
     & $=1$ & $1-s_P$ & $s_N$ \\
-\enddata
-\end{deluxetable}
+    \hline
+\end{tabular}
+\caption{$P\left( Z \mid X \right)$ for all possible values of $X$ and $Z$.}
+\end{table}
 
@@
 %
-This likelihood implicitly depends on how we parameterize the arcs of the \gls{dag} underlying the model -- see \autoref{eq:transition_prob} -- and the parameterization of the distribution over diagnosis times -- e.g., as in \autoref{eq:time_dist}.
+This likelihood implicitly depends on how we parameterize the arcs of the \gls{dag} underlying the model -- see \cref{eq:transition_prob} -- and the parameterization of the distribution over diagnosis times -- e.g., as in \cref{eq:time_dist}.
 
@@
 %
-Which effectively amounts to computing the element-wise logarithm of the likelihood vector $\boldsymbol{\ell}$ from \autoref{eq:likelihood_vec} and summing up the entries that correspond to each of the patients $d_i$ for $i\leq N$.
+Which effectively amounts to computing the element-wise logarithm of the likelihood vector $\boldsymbol{\ell}$ from \cref{eq:likelihood_vec} and summing up the entries that correspond to each of the patients $d_i$ for $i\leq N$.
 
@@
 In this paper, for some patients involvement information of level VII was missing and hence marginalized over.
-A detailed explanation of this formalism can be found in \citep{zora231470}, section 6.2.7.
+A detailed explanation of this formalism can be found in \cite{zora231470}, section 6.2.7.
 
@@
 
-\begin{figure}
+\begin{figure}[h]
     \centering
@@
     \caption{Extended \gls{dag} representing different possible spread graphs underlying the \gls{hmm}.
-As in \autoref{fig:graph_with_obs}, red arcs are parametrized with probabilities of spread from the tumor (red circle) to the \glspl{lnl} (blue circles).
+As in \cref{fig:graph_with_obs}, red arcs are parametrized with probabilities of spread from the tumor (red circle) to the \glspl{lnl} (blue circles).
 These red arcs, together with the blue arcs fron \gls{lnl} to \gls{lnl}, make up the \emph{base graph}.
@@
 
-We use \gls{mcmc} sampling to draw parameter samples $\boldsymbol{\hat{\theta}}_i$ for $i \leq S$ from the likelihood described in \autoref{eq:log_likelihood} (i.e.
+We use \gls{mcmc} sampling to draw parameter samples $\boldsymbol{\hat{\theta}}_i$ for $i \leq S$ from the likelihood described in \cref{eq:log_likelihood} (i.e.
 the unnormalized posterior distribution over the parameters $\boldsymbol{\theta}$, since we used a uniform prior in this work).
 
-More specifically, we use the Python implementation \texttt{emcee} \citep{foreman-mackey_emcee_2013} and two sample proposal mechanisms based on differential evolution moves \citep{ter_braak_differential_2008,nelson_run_2013} for sampling.
+More specifically, we use the Python implementation \texttt{emcee} \cite{foreman-mackey_emcee_2013} and two sample proposal mechanisms based on differential evolution moves \cite{ter_braak_differential_2008,nelson_run_2013} for sampling.
 Instead of proposing and then accepting or rejecting individual parameter samples one after the other (as in the classical Metropolis-Hastings algorithm), the \texttt{emcee} implementation makes use of an ensemble of $W$ so-called ``walkers''.
@@
 Alternatively, the individual $\hat{f}_i = f\big( \boldsymbol{\hat{\theta}}_i \big)$ can be used to plot histograms over the distribution of $f$.
-We will do so in \autoref{sec:results} to show distributions over prevalence predictions and risk computations.
+We will do so in \cref{sec:results} to show distributions over prevalence predictions and risk computations.
 
@@
 
-\begin{figure}[t]
+\begin{figure}[ht]
     \script{thermo_int.py}
@@
 However, already before $\beta=0.031$ the model based on the winning graph outperforms the loosing graph model.
-Ultimately, the winning graph's log-evidence is better than the loosing graph model's by a value of around \variable{output/evidence_diff.txt}(see \autoref{table:evidence}).
-\label{fig:thermo_int}}
+Ultimately, the winning graph's log-evidence is better than the loosing graph model's by a value of around \variable{output/evidence_diff.txt}(see \cref{table:evidence}).}
+    \label{fig:thermo_int}
 \end{figure}
 
-\subsection{Risk estimation}
+\subsection{Risk Estimation}
 \label{subsec:formalism:risk}
@@
 
-\section{Complete Model of Ipsilateral Spread in OPSCC}
+\section{Complete Model\\ of Ipsilateral Spread in OPSCC}
 \label{sec:complete_model}
@@
 
-The \gls{dag} shown in \autoref{fig:graph_with_obs} includes the \glspl{lnl} II, III, and IV, which represent the most relevant lymph node levels for \gls{opscc}.
+The \gls{dag} shown in \cref{fig:graph_with_obs} includes the \glspl{lnl} II, III, and IV, which represent the most relevant lymph node levels for \gls{opscc}.
 It includes the arcs from II to III and from III to IV, representing the main direction of lymphatic drainage, which is well motivated anatomically and by the data on lymph node involvement.
-Previous publications  \citep{pouymayou_bayesian_2019,ludwig_hidden_2021} focused on these levels because they relied on a limited reconstructed dataset of \gls{opscc} patients \citep{sanguineti_defining_2009}.
+Previous publications  \cite{pouymayou_bayesian_2019,ludwig_hidden_2021} focused on these levels because they relied on a limited reconstructed dataset of \gls{opscc} patients \cite{sanguineti_defining_2009}.
 Now, with the datasets available for this work, we can extend the graph to include random variables for all \glspl{lnl} that are relevant for \gls{opscc}: I, II, III, IV, V, and VII.
@@
 To investigate which connections to add we start by establishing a baseline from a model using a minimal \emph{base graph}.
-It contains only the connections from \gls{lnl} II to III and an arc from \gls{lnl} III to IV, as motivated by the \emph{main lymphatic pathway} \citep{lengele_anatomical_2007}.
-The base graph is illustrated in \autoref{fig:extended_graph} via the red and blue arcs.
+It contains only the connections from \gls{lnl} II to III and an arc from \gls{lnl} III to IV, as motivated by the \emph{main lymphatic pathway} \cite{lengele_anatomical_2007}.
+The base graph is illustrated in \cref{fig:extended_graph} via the red and blue arcs.
 The lymphatic drainage to or from levels I, V, and VII is not as clearly defined.
-Therefore, we define a set of candidate arcs (green arcs in \autoref{fig:extended_graph}) and use the model comparison methodology described in \autoref{subsec:complete_model:comparison} to determine which graph is most supported our data.
+Therefore, we define a set of candidate arcs (green arcs in \cref{fig:extended_graph}) and use the model comparison methodology described in \cref{subsec:complete_model:comparison} to determine which graph is most supported our data.
 
 To connect level I, we investigate two candidate arcs: from I to II and from II to I.
-An arc from I to II was used in \citep{pouymayou_bayesian_2019,ludwig_hidden_2021} and is anatomically motivated.
+An arc from I to II was used in \cite{pouymayou_bayesian_2019,ludwig_hidden_2021} and is anatomically motivated.
 However, since \gls{lnl} I is rarely involved compared to level II, the associated parameter $t_{12}$ is mostly undetermined.
@@
 
-Anatomically, the \emph{posterior accessory pathway} that drains \gls{lnl} II through \gls{lnl} V motivates investigating and arc from $X_2$ into $X_5$ \citep{lengele_anatomical_2007}.
+Anatomically, the \emph{posterior accessory pathway} that drains \gls{lnl} II through \gls{lnl} V motivates investigating and arc from $X_2$ into $X_5$ \cite{lengele_anatomical_2007}.
 And, although no lymphatic pathway is described that directly drains \gls{lnl} III or IV into level V, due to their proximity to each other, we will also investigate additional edges from the levels III and IV into \gls{lnl} V.
@@
 
-To determine the optimal graph, we first consider six models, each with one of the six green candidate arcs of \autoref{fig:extended_graph} added to the base graph.
-Every one of these models was evaluated by computing an approximation to its evidence via thermodynamic integration as described below in \autoref{subsec:complete_model:comparison}.
+To determine the optimal graph, we first consider six models, each with one of the six green candidate arcs of \cref{fig:extended_graph} added to the base graph.
+Every one of these models was evaluated by computing an approximation to its evidence via thermodynamic integration as described below in \cref{subsec:complete_model:comparison}.
 Subsequently, graphs combining multiple arcs that individually improve the model evidence are considered.
@@
 
-\begin{deluxetable}{c|c|c}
-\tablecaption{
-    Interpretation of Bayes factors and their natural logarithms in terms of their support for or against one of the two compared models as introduced by \citet{jeffreys_theory_1998}.%
-    \label{table:bayes_factor}%
-}
-\tablehead{
-    \colhead{$K_\text{1v2}$} & \colhead{$\ln{K_\text{1v2}}$} & \colhead{support for $\mathcal{M}_1$}
-}
-    \startdata
+\begin{table}[h]
+\centering
+\begin{tabular}{|c|c|c|}
+    \hline
+    $K_\text{1v2}$ & $\ln{K_\text{1v2}}$ & support for $\mathcal{M}_1$ \\
+    \hline
     $< 10^0$ & $< 0$ & negative (supports $\mathcal{M}_2$) \\
@@
     $> 10^2$ & $> 4.6$ & decisive \\
-    \enddata
-\end{deluxetable}
+    \hline
+\end{tabular}
+\caption{Interpretation of Bayes factors and their natural logarithms in terms of their support for or against one of the two compared models as introduced by \cite{jeffreys_theory_1998}.}
+\label{table:bayes_factor}
+\end{table}
 
@@
 If we assume all models $\mathcal{M}_i$ for $i \in \{ 1,2 \}$ to have the same \emph{a priori} probability -- meaning in this case $P (\mathcal{M}_1) = P (\mathcal{M}_2)$ -- then we can compute the so-called \emph{Bayes factor} of the two models as the ratio of their likelihoods.
-The interpretation of the values for different Bayes factors is given in \autoref{table:bayes_factor}.
+The interpretation of the values for different Bayes factors is given in \cref{table:bayes_factor}.
 It is defined as follows:
@@
 
-Below, we will briefly outline the main concept behind this algorithm. An intuitive and extensive derivation of \gls{ti} is given by \citep{aponte_introduction_2022}.
+Below, we will briefly outline the main concept behind this algorithm. An intuitive and extensive derivation of \gls{ti} is given by \cite{aponte_introduction_2022}.
 
@@
 
-\begin{deluxetable}{l|rr}
-    \tablecaption{
-        Literature sensitivity and specificity values that we used to infer the most likely involvement for a patient when multiple diagnostic modalities reported conflicting nodal involvement \citep{de_bondt_Detection_2007,kyzas_18f-fluorodeoxyglucose_2008}.
-        \label{table:sens_spec}
-    }
-    \tablehead{
-        \textbf{Modality} & \textbf{Specificity} & \textbf{Sensitivity}
-    }
-    \startdata
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|rr|}
+    \hline
+    \textbf{Modality} & \textbf{Specificity} & \textbf{Sensitivity} \\
+    \hline
     \acrshort{ct} & 76\% & 81\% \\
@@
     Pathology & $\approx$ 100\% & $\approx$ 100\% \\
-    \enddata
-\end{deluxetable}
+    \hline
+\end{tabular}
+\caption{Literature sensitivity and specificity values that we used to infer the most likely involvement for a patient when multiple diagnostic modalities reported conflicting nodal involvement \cite{de_bondt_detection_2007,kyzas_18f-fluorodeoxyglucose_2008}.}
+\label{table:sens_spec}
+\end{table}
 
@@
 %
-The integral in \autoref{eq:ti} can then be computed via a trapezoidal rule using the $\mathcal{A}_\text{MC}$ to yield a  numerical approximation of the model evidence:
+The integral in \cref{eq:ti} can then be computed via a trapezoidal rule using the $\mathcal{A}_\text{MC}$ to yield a  numerical approximation of the model evidence:
 %
@@
 
-\begin{deluxetable}{ccccccrrrr}
-    \tablecaption{Prevalence of involvement patterns in the multi-centric dataset.
-An involvement pattern is characterized by the state of the six \glspl{lnl}: A red dot means the \gls{lnl} was reported to be metastatic, a green dot means it was determined to be healthy and a question mark means that the prevalence was marginalized over the state of this \gls{lnl}.
-\label{table:data_prevalence}}
-    \tablehead{
-        \multicolumn{6}{c}{\textbf{LNL involvement} } & \multicolumn{4}{c}{\textbf{T-category} } \\
-        I & II & III & IV & V & VII & \multicolumn{2}{c}{\textbf{early} } & \multicolumn{2}{c}{\textbf{advanced} }
-    }
-    \startdata
-    \variable{output/data_table.tex}%
-    \enddata
-\end{deluxetable}
+\begin{table}[h]
+\centering
+\begin{tabular}{|ccccccrrrr|}
+    \hline
+    \multicolumn{6}{|c}{\textbf{LNL involvement} } & \multicolumn{4}{c|}{\textbf{T-category} } \\
+    I & II & III & IV & V & VII & \multicolumn{2}{c}{\textbf{early} } & \multicolumn{2}{c|}{\textbf{advanced} } \\
+    \hline
+    \variable{output/data_table.tex} \\
+    \hline
+\end{tabular}
+\caption{Prevalence of involvement patterns in the multi-centric dataset.
+An involvement pattern is characterized by the state of the six \glspl{lnl}: A red dot means the \gls{lnl} was reported to be metastatic, a green dot means it was determined to be healthy and a question mark means that the prevalence was marginalized over the state of this \gls{lnl}.}
+\label{table:data_prevalence}
+\end{table}
 
-\begin{figure*}[h]
+\begin{figure*}[!h]
     \script{bg_core_prevs.py}
@@
 The process of computing the log-evidence using \gls{ti} was as follows: We randomly initialized the starting positions of the $W$ samplers in the ensemble within the $k$ dimensional unit cube $\boldsymbol{\Theta}$.
-Subsequently, for each $j \in \{ 0, 1, 2, \ldots, R=63 \}$ we drew samples from the corresponding power posterior with the value of $\beta_j$ set according to the power rule in \autoref{eq:power_rule}.
+Subsequently, for each $j \in \{ 0, 1, 2, \ldots, R=63 \}$ we drew samples from the corresponding power posterior with the value of $\beta_j$ set according to the power rule in \cref{eq:power_rule}.
 This sampling at point $j$ consisted of 1000 burn-in steps, followed by 200 steps, of which only every tenth was kept.
@@
 An approximation to the evidence that also attempts to balance accuracy and complexity against each other is the heuristic called \gls{bic}.
-The negative one half of the \gls{bic} approximates the $\ln E$ via Lagrange's method \citep{bhat_Derivation_2010} and yields an easy to compute estimate that may also be used to compare models, as long as its underlying assumptions are valid:
+The negative one half of the \gls{bic} approximates the $\ln E$ via Lagrange's method \cite{bhat_Derivation_2010} and yields an easy to compute estimate that may also be used to compare models, as long as its underlying assumptions are valid:
 %
@@
 The dataset $\boldsymbol{\mathcal{D}}$ that we used for inference is comprised of the detailed reports on lymph node involvement patterns in \gls{opscc} patients treated at three different institutions in France and Switzerland: The \gls{clb} in Lyon (France), the \gls{isb} in Bern (Switzerland), and the \gls{usz} in Zürich (Switzerland).
-We have previously published the patterns of nodal involvement for the \gls{usz} cohort (287 patients) \citep{ludwig_dataset_2022,ludwig_lydata_2023} and described its characteristics in detail \citep{ludwig_detailed_2022}.
-The first \gls{clb} dataset (263 patients) \citep{ludwig_lydata_2022} underlies a publication on \gls{hpv} status in \gls{opscc} \citep{bauwens_prevalence_2021} and is made available in a separate ``Data in Brief'' article alongside the second dataset from France \citep{ludwig_lydata_2023-2} and the lymphatic progression patterns from the \gls{isb} \citep{ludwig_lydata_2023-1,ludwig_multi-centric_2023}.
+We have previously published the patterns of nodal involvement for the \gls{usz} cohort (287 patients) \cite{ludwig_dataset_2022,ludwig_lydata_2023} and described its characteristics in detail \cite{ludwig_detailed_2022}.
+The first \gls{clb} dataset (263 patients) \cite{ludwig_lydata_2022} underlies a publication on \gls{hpv} status in \gls{opscc} \cite{bauwens_prevalence_2021} and is made available in a separate ``Data in Brief'' article alongside the second dataset from France \cite{ludwig_lydata_2023-2} and the lymphatic progression patterns from the \gls{isb} \cite{ludwig_lydata_2023-1,ludwig_multi-centric_2023}.
 All datasets may be explored online in our web-based interface \href{https://lyprox.org}{LyProX}.
@@
 When different modalities were conflicting, the conflicts were resolved by inferring the most likely state (healthy or metastatic) for each \gls{lnl} separately.
-To do so, we used literature values for the sensitivity and specificity of the diagnostic modalities \citep{de_bondt_Detection_2007,kyzas_18f-fluorodeoxyglucose_2008}, which we also tabulated in \autoref{table:sens_spec}.
+To do so, we used literature values for the sensitivity and specificity of the diagnostic modalities \cite{de_bondt_Detection_2007,kyzas_18f-fluorodeoxyglucose_2008}, which we also tabulated in \cref{table:sens_spec}.
 Practically, this means that, whenever pathology after neck dissection was available, the pathology result was taken as the consensus, overruling any other clinical diagnostic modality.
@@
 We assumed that it represents an observation of the true hidden state $\boldsymbol{\xi}_k$.
-The frequencies of some of the most important combinations of involved lymph node levels are listed in \autoref{table:data_prevalence}.
+The frequencies of some of the most important combinations of involved lymph node levels are listed in \cref{table:data_prevalence}.
 
@@
 Their respective model evidence fell below $-1528$ and the two graphs would appear far left in the figure.
-In the bottom left corner, we provide a visual reference in analogy to \autoref{table:bayes_factor}: E.g., any difference in the model evidence shorter than the first of the three rulers indicates that the improvement is ``barely worth a mention''.
+In the bottom left corner, we provide a visual reference in analogy to \cref{table:bayes_factor}: E.g., any difference in the model evidence shorter than the first of the three rulers indicates that the improvement is ``barely worth a mention''.
 \label{fig:ranking}}
@@
 
-For the base graph, we have plotted the predicted prevalence of involvement patterns in the investigated patient cohort for scenarios involving the most commonly metastatic \glspl{lnl} II, III and IV in \autoref{fig:bg_prevalences}.
+For the base graph, we have plotted the predicted prevalence of involvement patterns in the investigated patient cohort for scenarios involving the most commonly metastatic \glspl{lnl} II, III and IV in \cref{fig:bg_prevalences}.
 It shows -- for each pattern of lymphatic involvement -- two plots overlaid:
@@
 
-\autoref{fig:bg_prevalences} shows that this minimal graph is already capable of describing the most important parts of the observed data very well.
+\cref{fig:bg_prevalences} shows that this minimal graph is already capable of describing the most important parts of the observed data very well.
 Notably, the model is not only accurate in its predictions, it also correctly estimates the variance stemming from the limited amount of data.
@@
 
-\subsection{Comparison of candidates graphs}
+\subsection{Comparison of Candidate Graphs}
 \label{subsec:results:graph_candidates}
 
-\begin{deluxetable}{lr}[t]
-    \tablecaption{Model comparison results from the base graph and the extended graph we chose as the ``winning'' model.
-For both \glspl{dag} we show the log-evidence, computed via thermodynamic integration, the negative one half of the \gls{bic}, as well as the maximum log-likelihood that was encountered during the final \gls{mcmc} sampling round.
-\label{table:evidence}}
-    \tablehead{
-        \textbf{graph}  & \textbf{log-evidence}
-    }
-    \startdata
-    \variable{output/metrics_table.tex}%
-    \enddata
-\end{deluxetable}
+\begin{table}
+\centering
+\begin{tabular}{|lr|}
+    \hline
+    \textbf{graph}  & \textbf{log-evidence} \\
+    \hline
+    \variable{output/metrics_table.tex}
+    \hline
+\end{tabular}
+\caption{Model comparison results from the base graph and the extended graph we chose as the ``winning'' model.
+For both \glspl{dag} we show the log-evidence, computed via thermodynamic integration, the negative one half of the \gls{bic}, as well as the maximum log-likelihood that was encountered during the final \gls{mcmc} sampling round.}
+\label{table:evidence}
+\end{table}
 
-The model evidences of all candidate graphs are reported in \autoref{table:evidence}.
-A visual ranking is provided in \autoref{fig:ranking}.
+The model evidences of all candidate graphs are reported in \cref{table:evidence}.
+A visual ranking is provided in \cref{fig:ranking}.
 Let us first consider the six models in which one of the candidate arcs is added to the base graph.
@@
 Based on these results, we consider three additional candidates for the optimal graph that combine the added arc from \gls{lnl} I to II with the arc(s) $\text{III} \rightarrow \text{V}$ and/or $\text{IV} \rightarrow \text{V}$.
-The model evidence for these three graphs is also reported in \autoref{table:evidence}.
+The model evidence for these three graphs is also reported in \cref{table:evidence}.
 The best performing graph with decisive evidence over the base graph turned out to be the one which combines the arcs from \gls{lnl} I to II and IV to V.
@@
 
-\autoref{table:evidence} additionally shows the evidence of graphs in which the arc from level II to III or from level III to IV is removed.
+\cref{table:evidence} additionally shows the evidence of graphs in which the arc from level II to III or from level III to IV is removed.
 The low model evidence for these graphs confirms the importance of these connections and is consistent with the anatomical motivation.
@@
 
-\subsection{The winning graph}
+\subsection{The Winning Graph}
 \label{subsec:results:optimal_graph}
 
-The most likely model parameters for the winning graph, corresponding to the mean of the marginals of the sampled posterior distributions, are tabulated in \autoref{table:params}.
+The most likely model parameters for the winning graph, corresponding to the mean of the marginals of the sampled posterior distributions, are tabulated in \cref{table:params}.
 We have fixed $p_\text{early}=0.3$ for early T-category tumors (i.e., T0, T1, and T2), and $t_\text{max}=10$ time steps.
@@
 
-which agrees with the second panel from the top in \autoref{fig:bg_prevalences}. 
+which agrees with the second panel from the top in \cref{fig:bg_prevalences}. 
 The large value for the parameter $t_{\text{I} \rightarrow \text{II}}$ reflects the observation that in almost all patients with level I involvement, level II is also involved.
@@
 
-\subsection{Involvement of levels I and V}
+\subsection{Involvement of Levels I and V}
 
@@
 
-{\it Level V:} In \autoref{fig:IIIandIVandV_prevs} we compare the base graph's and the winning graph's estimations for prevalences of involvement patterns that include \gls{lnl} V.
+{\it Level V:} In \cref{fig:IIIandIVandV_prevs} we compare the base graph's and the winning graph's estimations for prevalences of involvement patterns that include \gls{lnl} V.
 The base graph underestimates the probability that level IV and V are simultaneously involved, and overestimates the probability that level V but not IV is involved.
@@
 
-{\it Level I:} In \autoref{fig:IandII_prevs}, analogous comparisons are shown for involvement patterns that include \gls{lnl} I.
+{\it Level I:} In \cref{fig:IandII_prevs}, analogous comparisons are shown for involvement patterns that include \gls{lnl} I.
 The base graph overestimates the probability of level I involvement without simultaneous involvement of level II.
@@
 
-\begin{deluxetable}{lrr}[b]
-    \tablecaption{Mean and standard deviation of parameters sampled for the winning graph in percent. \label{table:params}}
-    \tablehead{
-        \textbf{parameter} & \textbf{mean} & \textbf{std. dev.}
-    }
-    \startdata
-    \variable{output/means_table.tex}%
-    \enddata
-\end{deluxetable}
+\begin{table}
+\centering
+\begin{tabular}{|lrr|}
+    \hline
+    \textbf{parameter} & \textbf{mean} & \textbf{std. dev.} \\
+    \hline
+    \variable{output/means_table.tex} \\
+    \hline
+\end{tabular}
+\caption{Mean and standard deviation of parameters sampled for the winning graph in percent.}
+\label{table:params}
+\end{table}
 
@@
 In this section, the model corresponding to the winning graph is applied to estimating the risk of occult metastases in clinically negative \glspl{lnl}.
-We assume a sensitivity of 0.76 and a specificity of 0.81 for the clinical diagnosis of lymph node metastases, corresponding to CT imaging in \autoref{table:sens_spec}.
+We assume a sensitivity of 0.76 and a specificity of 0.81 for the clinical diagnosis of lymph node metastases, corresponding to CT imaging in \cref{table:sens_spec}.
 
-{\it Level II:} As can be seen in \autoref{table:params}, spread from the tumor to \gls{lnl} II to be the most probable transition at any given time step.
+{\it Level II:} As can be seen in \cref{table:params}, spread from the tumor to \gls{lnl} II to be the most probable transition at any given time step.
 As a consequence, even for an early T-category patient that presents with a clinical N0 neck, our model predicts a \variable{output/wg_II_risks.tex}risk for microscopic metastases in \gls{lnl} II.
 
-\begin{figure}[b]
+\begin{figure}
     \script{wg_III_risks.py}
@@
 
-{\it Level III:} \autoref{fig:wg_III_risks} compares the risk of occult disease in level III between patients that are clinically N0 (orange) and patients with clinically diagnosed involvement of only level II (red), for early T-category (upper panel) and advanced T-category (bottom panel).
+{\it Level III:} \cref{fig:wg_III_risks} compares the risk of occult disease in level III between patients that are clinically N0 (orange) and patients with clinically diagnosed involvement of only level II (red), for early T-category (upper panel) and advanced T-category (bottom panel).
 The histograms represent the uncertainty in the model's risk prediction arising from the uncertainty in the model parameters and are generated by randomly drawing a tenth of the samples from the model parameter's joint posterior distribution.
-This amounts to $S = 20 \cdot W$ samples, as described in \autoref{subsec:complete_model:comparison}.
+This amounts to $S = 20 \cdot W$ samples, as described in \cref{subsec:complete_model:comparison}.
 The model predicts a risk of just below 6\% for early T-category tumors and 8\% for T3 or T4 ones.
@@
 
-{\it Level IV:} \autoref{fig:wg_IVandV_risks} (left panels) compares the risk of occult disease in level IV for the typical clinical presentations: clinically N0 (green), metastases in level II (blue), and metastases in levels II and III (orange).
+{\it Level IV:} \cref{fig:wg_IVandV_risks} (left panels) compares the risk of occult disease in level IV for the typical clinical presentations: clinically N0 (green), metastases in level II (blue), and metastases in levels II and III (orange).
 The model predicts a low risk of 1-2\% in level IV for patients with clinically healthy level III.
@@
 
-{\it Level V:} The right panels in \autoref{fig:wg_IVandV_risks} show the risk of occult disease in level V depending on T-category and the clinical involvement of levels II-IV.
+{\it Level V:} The right panels in \cref{fig:wg_IVandV_risks} show the risk of occult disease in level V depending on T-category and the clinical involvement of levels II-IV.
 For clinically N0 patients, the risk in level V is estimated to be just above 1\%.
@@
 
-{\it Level I:} \autoref{fig:wg_I_risks} shows the risk of occult disease in level I depending on T-category and the clinical involvement of levels II-IV.
+{\it Level I:} \cref{fig:wg_I_risks} shows the risk of occult disease in level I depending on T-category and the clinical involvement of levels II-IV.
 For clinically N0 patients, the risk in level I is estimated to be in the order of 1-2\%.
@@
 
-To illustrate the flexibility of the model in predicting various risks, we have plotted the risk for occult disease in \textit{any} of the \glspl{lnl} I, IV, and/or V, given different clinical diagnoses in \autoref{fig:wg_any_risks}.
+To illustrate the flexibility of the model in predicting various risks, we have plotted the risk for occult disease in \textit{any} of the \glspl{lnl} I, IV, and/or V, given different clinical diagnoses in \cref{fig:wg_any_risks}.
 Similar to this, we may compute the risk for an arbitrary combination of involved levels, given a similarly arbitrary clinical diagnosis.
@@
 In this publication we present a statistical model of ipsilateral lymph node involvement in oropharyngeal SCC patients.
-Although the basic \gls{hmm} of lymphatic progression has been conceptually introduced in a previous work \citep{ludwig_hidden_2021}, this is the first publication that evaluates the model based on a large multi-institutional dataset containing \variable{output/num_patients.txt}patients.
+Although the basic \gls{hmm} of lymphatic progression has been conceptually introduced in a previous work \cite{ludwig_hidden_2021}, this is the first publication that evaluates the model based on a large multi-institutional dataset containing \variable{output/num_patients.txt}patients.
 It is demonstrated for the first time that the model can accurately describe the patterns of lymph node involvement observed in the data, including the correlations between levels and its dependence on T-category.
@@
 
-\subsection{Implications for elective nodal treatment} 
+\subsection{Implications for Elective\\Nodal Treatment} 
 Risk predictions obtained by the model may be used to design clinical trials on volume-deescalated treatment of \gls{opscc}.
 In the context of radiotherapy, this corresponds to excluding \glspl{lnl} from the \gls{ctv-n}, which are irradiated according to the current guidelines.
-The list below should be seen as a summary of \autoref{subsec:results:risk_prediction} and the limitations discussed in \autoref{subsec:disc:limitation} should be taken into account in its interpretation.
+The list below should be seen as a summary of \cref{subsec:results:risk_prediction} and the limitations discussed in \cref{subsec:disc:limitation} should be taken into account in its interpretation.
 Assuming that one accepts approximately a 5\% risk of occult metastases per \gls{lnl}, the statistical model presented in this paper would suggest to:
@@
     \item[$\bullet$] Exclude level I from the \gls{ctv-n} for early T-category patients with limited metastatic disease.
-For advanced T-category patients with extensive nodal involvement of levels II, III, and IV, level I should be irradiated (see also the limitations discussed in \autoref{subsec:disc:limitation}).
+For advanced T-category patients with extensive nodal involvement of levels II, III, and IV, level I should be irradiated (see also the limitations discussed in \cref{subsec:disc:limitation}).
     \item[$\bullet$] Exclude level VII in all patients, unless it is clinically involved.
@@
 
-\begin{figure}[b]
+\begin{figure}
     \script{wg_any_risks.py}
@@
 
-\subsection{Limitations and future work}
+\subsection{Limitations and Future Work}
 \label{subsec:disc:limitation}
 
-{\it T-category dependence of level I involvement:} As shown in \autoref{fig:bg_prevalences}, the model describes the involvement of \glspl{lnl} II, III, and IV depending on T-category very well despite having only a single parameter related to T-category.
-\autoref{fig:IandII_prevs} shows that the model does not perfectly describe the T-category dependence of level I.
+{\it T-category dependence of level I involvement:} As shown in \cref{fig:bg_prevalences}, the model describes the involvement of \glspl{lnl} II, III, and IV depending on T-category very well despite having only a single parameter related to T-category.
+\cref{fig:IandII_prevs} shows that the model does not perfectly describe the T-category dependence of level I.
 It adjusts the parameters such that level I involvement is correctly described for the set of all patients combined, but it overestimates level I involvement for early T-category and underestimates it for advanced T-category.
@@
 Consequently, these values need to be considered with caution.
-\autoref{fig:wg_sens_spec_risks} illustrates for one example how the risk of occult disease depends on sensitivity and specificity.
+\cref{fig:wg_sens_spec_risks} illustrates for one example how the risk of occult disease depends on sensitivity and specificity.
 Here, we consider the risk in level IV in patients with clinically detected metastases in levels II and III.
@@
 
-\begin{figure}[t]
+\begin{figure}
     \script{wg_sens_spec_risks.py}
@@
 
-Also, as described in \autoref{sec:data}, we assumed the consensus of the data to represent the true state of nodal involvement.
+Also, as described in \cref{sec:data}, we assumed the consensus of the data to represent the true state of nodal involvement.
 This was not strictly necessary: Instead of computing a consensus beforehand and providing that with sensitivity and specificity of 1 to the model, as if it were the ground truth, we could have provided multiple diagnostic modalities per patient to the model directly.
@@
 
-\begin{acknowledgments}
+\section*{Acknowledgments}
 
@@
 
-\end{acknowledgments}
-
-\clearpage
-
-\bibliography{bib}
+\printbibliography
 

### 2024-02-05 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{graphicx}
+\graphicspath{ {./figures/} }
 \usepackage{biblatex}
@@
 
-\section{Complete Model\\ of Ipsilateral Spread in OPSCC}
+\section{Complete Model\\of Ipsilateral Spread in OPSCC}
 \label{sec:complete_model}
@@
 
-\begin{figure*}[!h]
+\begin{figure*}
     \script{bg_core_prevs.py}
@@
 An approximation to the evidence that also attempts to balance accuracy and complexity against each other is the heuristic called \gls{bic}.
-The negative one half of the \gls{bic} approximates the $\ln E$ via Lagrange's method \cite{bhat_Derivation_2010} and yields an easy to compute estimate that may also be used to compare models, as long as its underlying assumptions are valid:
+The negative one half of the \gls{bic} approximates the $\ln E$ via Lagrange's method \cite{bhat_derivation_2010} and yields an easy to compute estimate that may also be used to compare models, as long as its underlying assumptions are valid:
 %
@@
 When different modalities were conflicting, the conflicts were resolved by inferring the most likely state (healthy or metastatic) for each \gls{lnl} separately.
-To do so, we used literature values for the sensitivity and specificity of the diagnostic modalities \cite{de_bondt_Detection_2007,kyzas_18f-fluorodeoxyglucose_2008}, which we also tabulated in \cref{table:sens_spec}.
+To do so, we used literature values for the sensitivity and specificity of the diagnostic modalities \cite{de_bondt_detection_2007,kyzas_18f-fluorodeoxyglucose_2008}, which we also tabulated in \cref{table:sens_spec}.
 Practically, this means that, whenever pathology after neck dissection was available, the pathology result was taken as the consensus, overruling any other clinical diagnostic modality.
@@
     \hline
-    \variable{output/metrics_table.tex}
+    \variable{output/metrics_table.tex} \\
     \hline

### 2024-02-05 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{graphicx}
-\graphicspath{ {./figures/} }
 \usepackage{biblatex}
@@
     \caption{The expected log-accuracy under the power posterior plotted against the value of the inverse temperature $\beta$.
-These values were computed during a \gls{ti} run for the loosing graph model (red line), as well as the winning graph model (green line).
+These values were computed during a TI run for the loosing graph model (red line), as well as the winning graph model (green line).
 Note that the scaling of the x-axis is chosen such that the 64 points on the temperature ladder appear evenly spaced.
@@
 However, already before $\beta=0.031$ the model based on the winning graph outperforms the loosing graph model.
-Ultimately, the winning graph's log-evidence is better than the loosing graph model's by a value of around \variable{output/evidence_diff.txt}(see \cref{table:evidence}).}
+Ultimately, the winning graph's log-evidence is better than the loosing graph model's.}
     \label{fig:thermo_int}

### 2024-02-05 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage[a4paper, total={190mm, 270mm}]{geometry}
+\usepackage{caption}
+\DeclareCaptionFormat{custom}{{\small\textbf{#1#2}\textit{#3}}}
+\captionsetup{format=custom}
+
 \usepackage{showyourwork}
@@
 Each edge represents the conditional probability of metastatic spread per abstract time-step, given disease at the edge's starting node.
-To learn these probabilities, we draw \acrlong{mcmc} samples from the likelihood of a dataset (\variable{output/num_patients.txt}\gls{opscc} patients) from three institutions.
+To learn these probabilities, we draw \acrlong{mcmc} samples from the likelihood of a dataset (\variable{output/num_patients.txt} \gls{opscc} patients) from three institutions.
 We compute the model evidence using \acrlong{ti} for different graphs to determine which describes the data best.
@@
 
-\begin{table}
+\begin{table}[!h]
 \centering
@@
 \end{tabular}
-\caption{$P\left( X_3 \mid X_2 \right)$ for all four possible combinations of the states of $X_3$ and $X_2$.}
+% \caption{$P\left( X_3 \mid X_2 \right)$ for all four possible combinations of the states of $X_3$ and $X_2$.}
 \end{table}
@@
 \end{tabular}
-\caption{$P\left( Z \mid X \right)$ for all possible values of $X$ and $Z$.}
+% \caption{$P\left( Z \mid X \right)$ for all possible values of $X$ and $Z$.}
 \end{table}
@@
 
-\begin{figure}[h]
+\begin{figure}[!b]
     \centering
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
     \script{thermo_int.py}
@@
 
-\begin{table}[h]
+\begin{table}[b]
 \centering
@@
 
-\begin{table}[h]
+\begin{table}[t]
 \centering
@@
 
-\begin{figure*}[!ht]
+\begin{figure*}[t]
     \centering
@@
 
+\subsection{Risk Prediction for Occult Disease}
+\label{subsec:results:risk_prediction}
+
+In this section, the model corresponding to the winning graph is applied to estimating the risk of occult metastases in clinically negative \glspl{lnl}.
+We assume a sensitivity of 0.76 and a specificity of 0.81 for the clinical diagnosis of lymph node metastases, corresponding to CT imaging in \cref{table:sens_spec}.
+
+{\it Level II:} As can be seen in \cref{table:params}, spread from the tumor to \gls{lnl} II to be the most probable transition at any given time step.
+As a consequence, even for an early T-category patient that presents with a clinical N0 neck, our model predicts a \variable{output/wg_II_risks.tex}risk for microscopic metastases in \gls{lnl} II.
+
 \begin{table}
@@
 
-\subsection{Risk Prediction for Occult Disease}
-\label{subsec:results:risk_prediction}
-
-In this section, the model corresponding to the winning graph is applied to estimating the risk of occult metastases in clinically negative \glspl{lnl}.
-We assume a sensitivity of 0.76 and a specificity of 0.81 for the clinical diagnosis of lymph node metastases, corresponding to CT imaging in \cref{table:sens_spec}.
-
-{\it Level II:} As can be seen in \cref{table:params}, spread from the tumor to \gls{lnl} II to be the most probable transition at any given time step.
-As a consequence, even for an early T-category patient that presents with a clinical N0 neck, our model predicts a \variable{output/wg_II_risks.tex}risk for microscopic metastases in \gls{lnl} II.
-
-\begin{figure}
-    \script{wg_III_risks.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/wg_III_risks.png}
-        \caption{Histograms over the risk for microscopic involvement in \gls{lnl} III, given that a patient presents as clinically N0 (green), or given that \gls{lnl} II shows clinical involvement (blue).
-The top panel shows these risks for early T-category, the bottom panel for late.
-\label{fig:wg_III_risks}}
-    \end{centering}
-\end{figure}
-
-\begin{figure*}
-    \script{wg_IVandV_risks.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/wg_IVandV_risks.png}
-        \caption{Distributions over the risk for microscopic involvement in \gls{lnl} IV (left panels) and in \gls{lnl} V (right panels) as predicted by the winning graph model, given early (top panels) or advanced T-category (bottom panels), and different CT-based diagnoses: 1) A clinical N0 patient (green histograms), 2) visible metastases in \gls{lnl} II, but otherwise healthy-looking lymph nodes (blue histograms), 3) macroscopic metastases in the \glspl{lnl} II \& III, with the rest of the neck still being clinically node negative (orange histograms), and finally 4) extensive clinical involvement in the levels II, III, and IV (red histograms).\label{fig:wg_IVandV_risks}}
-    \end{centering}
-\end{figure*}
-
 {\it Level III:} \cref{fig:wg_III_risks} compares the risk of occult disease in level III between patients that are clinically N0 (orange) and patients with clinically diagnosed involvement of only level II (red), for early T-category (upper panel) and advanced T-category (bottom panel).
@@
 For patients with involvement of level II, the risk in level III increases to approximately 9\% and 12\%, respectively.
+
+\begin{figure}[!b]
+    \script{wg_III_risks.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/wg_III_risks.png}
+        \caption{Histograms over the risk for microscopic involvement in \gls{lnl} III, given that a patient presents as clinically N0 (green), or given that the patient's \gls{lnl} II shows clinical involvement (blue).
+The top panel displays these risks for early T-category, the bottom panel for advanced T-category.}
+        \label{fig:wg_III_risks}
+    \end{centering}
+\end{figure}
 
@@
 
-\begin{figure}
-    \script{wg_I_risks.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/wg_I_risks.png}
-        \caption{Distributions over predicted risk for involvement in \gls{lnl} I, given different clinical diagnosis scenarios: For N0 patients (green), patients with macroscopic involvement in \gls{lnl} II (blue), and for the case where the \glspl{lnl} II, III, and IV show involvement.
-The top panel shows these risks for early T-category and the bottom row for advanced T-category.
-\label{fig:wg_I_risks}}
-    \end{centering}
-\end{figure}
-
 \section{Discussion}
@@
 
+\begin{figure*}
+    \script{wg_IVandV_risks.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/wg_IVandV_risks.png}
+        \caption{Distributions over the risk for microscopic involvement in \gls{lnl} IV (left panels) and in \gls{lnl} V (right panels) as predicted by the winning graph model, given early (top panels) or advanced T-category (bottom panels), and different CT-based diagnoses: 1) A clinical N0 patient (green histograms), 2) visible metastases in \gls{lnl} II, but otherwise healthy-looking lymph nodes (blue histograms), 3) macroscopic metastases in the \glspl{lnl} II \& III, with the rest of the neck still being clinically node negative (orange histograms), and finally 4) extensive clinical involvement in the levels II, III, and IV (red histograms).\label{fig:wg_IVandV_risks}}
+    \end{centering}
+\end{figure*}
+
 \subsection{Implications for Elective\\Nodal Treatment}
+
 Risk predictions obtained by the model may be used to design clinical trials on volume-deescalated treatment of \gls{opscc}.
@@
 
-\begin{figure}
-    \script{wg_any_risks.py}
+\subsection{Limitations and Future Work}
+\label{subsec:disc:limitation}
+
+\begin{figure}[!hb]
+    \script{wg_I_risks.py}
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/wg_any_risks.png}
-        \caption{Shown are the histograms over the predicted risk for involvement in any of the \glspl{lnl} I, IV, V, or VII.
-The risk is plotted given a clinical N0 diagnosis (green), macroscopically detected metastases in \gls{lnl} II (blue), and lastly given visible involvement in both \gls{lnl} II and III (orange).
-The top row shows these risks for early T-category diagnoses and the bottom row for advanced T-category.
-\label{fig:wg_any_risks}}
+        \includegraphics[width=\linewidth]{figures/wg_I_risks.png}
+        \caption{Distributions over predicted risk for involvement in \gls{lnl} I, given different clinical diagnosis scenarios: For N0 patients (green), patients with macroscopic involvement in \gls{lnl} II (blue), and for the case where the \glspl{lnl} II, III, and IV show involvement.
+The top panel shows these risks for early T-category and the bottom row for advanced T-category.
+\label{fig:wg_I_risks}}
     \end{centering}
 \end{figure}
-
-\subsection{Limitations and Future Work}
-\label{subsec:disc:limitation}
 
@@
 \begin{figure}
-    \script{wg_sens_spec_risks.py}
+    \script{wg_any_risks.py}
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/wg_sens_spec_risks.png}
-        \caption{Dependence of risk of occult disease on sensitivity and specificity.
-The figure shows the risk of involvement in level IV, given the mean of the parameter samples, drawn during sampling, in patients with clinical involvement of levels II and III.
-\label{fig:wg_sens_spec_risks}}
+        \includegraphics[width=\linewidth]{figures/wg_any_risks.png}
+        \caption{Shown are the histograms over the predicted risk for involvement in any of the \glspl{lnl} I, IV, V, or VII.
+The risk is plotted given a clinical N0 diagnosis (green), macroscopically detected metastases in \gls{lnl} II (blue), and lastly given visible involvement in both \gls{lnl} II and III (orange).
+The top row shows these risks for early T-category diagnoses and the bottom row for advanced T-category.
+\label{fig:wg_any_risks}}
     \end{centering}
@@
 
+\begin{figure}[!b]
+    \script{wg_sens_spec_risks.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/wg_sens_spec_risks.png}
+        \caption{Dependence of risk of occult disease on sensitivity and specificity.
+The figure shows the risk of involvement in level IV, given the mean of the parameter samples, drawn during sampling, in patients with clinical involvement of levels II and III.
+\label{fig:wg_sens_spec_risks}}
+    \end{centering}
+\end{figure}
+
 The patient data detailing lymphatic involvement is publicly available in the form of CSV tables in the GitHub repository \href{https://github.com/rmnldwg/lydata}{\texttt{rmnldwg/lydata}}.
-
 In another GitHub repository, \href{https://github.com/rmnldwg/lynference}{\texttt{rmnldwg/lynference}}, we define the experiments based on different graph structures.
 These experiments are reproducible via the tool \href{https://dvc.org}{DVC} and their locally computed results are uploaded to an Azure blob storage container.
-
 Lastly, the \LaTeX\ source code, all Python scripts to generate the figures and tables in this work, and the pipeline definition to build the document (using \showyourwork \cite{luger_mapping_2021}) are made public in the GitHub repository \href{https://github.com/rmnldwg/graph-extension-paper}{\texttt{rmnldwg/graph-extension-paper}}.
 To rebuild it, the \showyourwork tool pulls the patient data from the \href{https://github.com/rmnldwg/lynference}{\texttt{rmnldwg/lynference}} repository and the trained models in the form of \gls{mcmc} samples from the Azure blob storage container, then runs the Python scripts and compiles the \LaTeX\ article.
-
 Detailed instructions on how to reproduce the entire pipeline, including the individual experiments, are gladly given upon request.

### 2024-02-07 src/tex/ms.tex
[showyourwork] overleaf sync

@@
             The graph maximizing the model evidence connects the tumor to each \gls{lnl} and the \glspl{lnl} I through V in order.
-It predicts the risk of occult disease in level IV is below 5\% if level III is clinically negative, and that the risk of occult disease in level V is below 5\% except for advanced T-category (T3 and T4) patients with clinical involvement of levels II, III, and IV.
+It predicts the risk of occult disease in level IV is below 5\% if level III is clinically negative, and that the risk of occult disease in level V is below 5\% except for advanced T-category (T3 and T4) patients with clinical involvement of levels II, III, and IV.\looseness=-1
 
@@
 The dataset $\boldsymbol{\mathcal{D}}$ that we used for inference is comprised of the detailed reports on lymph node involvement patterns in \gls{opscc} patients treated at three different institutions in France and Switzerland: The \gls{clb} in Lyon (France), the \gls{isb} in Bern (Switzerland), and the \gls{usz} in Zürich (Switzerland).
-We have previously published the patterns of nodal involvement for the \gls{usz} cohort (287 patients) \cite{ludwig_dataset_2022,ludwig_lydata_2023} and described its characteristics in detail \cite{ludwig_detailed_2022}.
-The first \gls{clb} dataset (263 patients) \cite{ludwig_lydata_2022} underlies a publication on \gls{hpv} status in \gls{opscc} \cite{bauwens_prevalence_2021} and is made available in a separate ``Data in Brief'' article alongside the second dataset from France \cite{ludwig_lydata_2023-2} and the lymphatic progression patterns from the \gls{isb} \cite{ludwig_lydata_2023-1,ludwig_multi-centric_2023}.
+We have previously published the patterns of nodal involvement for the \gls{usz} cohort (287 patients) \cite{ludwig_dataset_2022} and described its characteristics in detail \cite{ludwig_detailed_2022}.
+The first \gls{clb} dataset (263 patients) underlies a publication on \gls{hpv} status in \gls{opscc} \cite{bauwens_prevalence_2021} and is made available in a separate ``Data in Brief'' article alongside the second dataset from France and the lymphatic progression patterns from the \gls{isb} \cite{ludwig_multi-centric_2023}.
 All datasets may be explored online in our web-based interface \href{https://lyprox.org}{LyProX}.
@@
 RL and JU wrote the main manuscript.
-RL created all figures and tables.
+RL created all results, figures, and tables.
 AS, SW, OE, MD, and RG collected the \gls{isb} patient data.
@@
 Lastly, the \LaTeX\ source code, all Python scripts to generate the figures and tables in this work, and the pipeline definition to build the document (using \showyourwork \cite{luger_mapping_2021}) are made public in the GitHub repository \href{https://github.com/rmnldwg/graph-extension-paper}{\texttt{rmnldwg/graph-extension-paper}}.
-To rebuild it, the \showyourwork tool pulls the patient data from the \href{https://github.com/rmnldwg/lynference}{\texttt{rmnldwg/lynference}} repository and the trained models in the form of \gls{mcmc} samples from the Azure blob storage container, then runs the Python scripts and compiles the \LaTeX\ article.
-Detailed instructions on how to reproduce the entire pipeline, including the individual experiments, are gladly given upon request.
-They may also be found in the respective repository's \texttt{README.md} file.
+To rebuild it, the \showyourwork tool first pulls the patient data from the \href{https://github.com/rmnldwg/lydata}{\texttt{rmnldwg/lydata}} repository and the trained models in the form of \gls{mcmc} samples from the Azure blob storage container, then runs the Python scripts, and finally compiles the \LaTeX\ article.
+Detailed instructions on how to reproduce the entire pipeline, including the individual experiments, may be found in the respective repository's \texttt{README.md} file.
+We also gladly provide support for any effort to repr
 

### 2024-02-20 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{graphicx}
-\usepackage{biblatex}
+\usepackage[
+    maxalphanames=6,
+    style=nature,
+    backend=biber
+]{biblatex}
 \addbibresource{bib.bib}
@@
 A patient's state of (hidden) lymphatic involvement at time $t$ is described as a collection of binary \glspl{rv}, one for each of the $V$ \glspl{lnl}:
-%
+
 \begin{equation}
@@
 The model evidences of all candidate graphs are reported in \cref{table:evidence}.
+For the graph with the highest and lowest model evidence, the expected log-accuracy is plotted against the inverse temperature in \cref{fig:thermo_int}.
 A visual ranking is provided in \cref{fig:ranking}.
@@
 Detailed instructions on how to reproduce the entire pipeline, including the individual experiments, may be found in the respective repository's \texttt{README.md} file.
-We also gladly provide support for any effort to repr
+We also gladly provide support for any effort to reproduce our results.
 

### 2024-05-15 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-We define the time $t=0$ to be the moment just before a patient's tumor formed, and hence $X_v[t=0]=0 \,\,\, \forall v$.
+We define the time $t=0$ to be the moment a patient's tumor formed, and hence $X_v[t=0]=0 \,\,\, \forall v$.
 However, using this definition, we cannot know how many time-steps have passed until $t_\text{D}$, when the patient was diagnosed with cancer.
@@
 E.g., $P \left( t=t_\text{D} \mid \text{early} \right)$ for early T-category patients (T1 \& T2) and $P \left( t=t_\text{D} \mid \text{late} \right)$ for advanced T-category patients (T3 \& T4).
-Throughout this work we will use binomial distributions for these probability mass functions, mainly because they have a plausible shape for this purpose and only a single parameter.
+Throughout this work we will use binomial distributions for these probability mass functions.
+They depend only on a single parameter, which when multiplied with the number of considered time-steps, conveniently also represents the distribution's mean.
 %
@@
     \includegraphics[width=\linewidth]{figures/thermo_int.png}
-    \caption{The expected log-accuracy under the power posterior plotted against the value of the inverse temperature $\beta$.
-These values were computed during a TI run for the loosing graph model (red line), as well as the winning graph model (green line).
-Note that the scaling of the x-axis is chosen such that the 64 points on the temperature ladder appear evenly spaced.
+    \caption{The expected log-accuracy under the power posterior plotted against the value of the inverse temperature $\beta$ for the winning graph, computed during thermodynamic integration (top panel).
+The are under this curve yields the winning graph's model evidence.
+Also shown is the difference of the winning graph's accuracy over the base graph's (bottom panel).
+The scaling of the x-axis is chosen such that the 64 points on the temperature ladder appear evenly spaced.
 This is to stress how the log-accuracy develops in the range from $\beta=0$ to around $\beta=0.1$.
-The area under these two curves yields an approximation of the log-evidence of the respective model.
-It is barely visible that due to the simplicity of the loosing graph model its accuracy increases earlier and faster that the winning graph model's.
-However, already before $\beta=0.031$ the model based on the winning graph outperforms the loosing graph model.
-Ultimately, the winning graph's log-evidence is better than the loosing graph model's.}
-    \label{fig:thermo_int}
+Note how the winning graph's advantage begins to show already shortly after $\beta=0.031$.
+    \label{fig:thermo_int}}
 \end{figure}
@@
 
-The main task for personalizing the \gls{ctv-n} definition is to predict the probability of the hidden possible states $\boldsymbol{\xi}_k$ given the diagnosis $d^\star=\boldsymbol{\zeta}_\ell$ of a new patient.
+The main task for personalizing the \gls{ctv-n} definition is to predict the probability of the hidden possible states $\boldsymbol{\xi}_k$ given the diagnosis $d^\star=\boldsymbol{\zeta}_\ell$ of a new patient at the time of diagnosis.
 Using Bayes' theorem, we get
@@
         \includegraphics[width=\textwidth]{figures/bg_core_prevs.png}
-        \caption{Prevalence of involvement as predicted by the base graph model for different scenarios involving the most commonly metastatic \glspl{lnl} II, III and IV (shaded histograms).
-The model's predictions are compared to Beta posteriors over the prevalence based on the frequency of the same scenarios and a uniform prior (slid lines).
-The top panel shows some selected scenarios with early T-category tumors and the bottom panel with advanced T-category.
+        \caption{Prevalence of involvement as predicted by the base graph model for different scenarios involving the most commonly metastatic \glspl{lnl} II, III, and IV (shaded histograms).
+The model's predictions are compared to Beta posteriors over the prevalence based on the frequency of the same scenarios in the data and a uniform prior (solid lines).
+The top panels of each of the three subfigures shows some selected scenarios with early T-category tumors and the bottom panel the same scenarios for advanced T-category.
+These figures show that for the most important \glspl{lnl} II, III, and IV the base graph model already fits the data well.
 \label{fig:bg_prevalences}}
@@
 predicted (histograms) prevalences of involvement combinations that include \gls{lnl} V.
-We have plotted the predictions from the winning graph (colored histograms) and those of the base graph (gray, hatched histograms).
+We have plotted the predictions from the winning graph (colored histograms) and those of the base graph (black, hatched histograms).
 The top two panels show scenarios for early T-category patients, the bottom two panels for advanced T-category.
@@
 The colored lines show the Beta posterior over the prevalence of the respective involvement pattern, given the data.
+Especially the right two panels indicate the winning graph model's better fit to the data over the base graph's model.
 \label{fig:IIIandIVandV_prevs}}
@@
 The top and bottom panels show the prevalences for early and advanced T-category, respectively.
-The solid lines are Beta posteriors from the data, while the histograms are predicted prevalences (colored: winning graph, gray-hatched: base graph).
+The solid lines are Beta posteriors from the data, while the histograms are predicted prevalences (colored: winning graph, black-hatched: base graph).
 Blue and red plots indicate overall \gls{lnl} I and II involvement, respectively.
@@
 The winning graph has an added edge from \gls{lnl} I to II, which improves the prediction of the rare green pattern.
+Otherwise, the winnings graph does not meaningfully improve the model's fit to the data.
 \label{fig:IandII_prevs}}
@@
 
+{\it Contralateral Spread:} So far, our model concerns itself with the ipsilateral lymphatic spread.
+However, our data suggests that at least contralateral \gls{lnl} II is frequently involved in \gls{opscc} patients.
+It is also of interest clinically to be able to decide when the contralateral side may be spared entirely, as this could meaningfully reduce side effects such as xerostomia and dysphagia [citation needed].
+Therefore extending the model to include some or all contralateral \glspl{lnl} would represent an important addition on its way to a clinical validation and ultimately implementation.
+
 {\it Sensitivity and Specificity:} Estimating the risk of occult metastases depends on the assumed parameter values for sensitivity and specificity of clinical detection of lymph node metastases.
@@
 The figure shows the risk of involvement in level IV, given the mean of the parameter samples, drawn during sampling, in patients with clinical involvement of levels II and III.
+This illustrates the dependence of the risk prediction on the provided sensitivity and specificity of the diagnostic modality.
 \label{fig:wg_sens_spec_risks}}

### 2024-05-16 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Not shown are the two graphs where the arc from \gls{lnl} III to IV was removed.
-Their respective model evidence fell below $-1528$ and the two graphs would appear far left in the figure.
-In the bottom left corner, we provide a visual reference in analogy to \cref{table:bayes_factor}: E.g., any difference in the model evidence shorter than the first of the three rulers indicates that the improvement is ``barely worth a mention''.
+Their respective log model evidence differed from the base graph by more than $-19$ and the two graphs would appear far left in the figure.
+In the bottom left corner, we provide a visual reference in analogy to \cref{table:bayes_factor}: E.g., any difference in the model evidence shorter than the first of the three rulers indicates that the improvement is ``barely worth a mention'' Anything longer than it, but shorter than the second ruler from the top indicates a ``substantial'' improvement.
 \label{fig:ranking}}

### 2024-05-22 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Throughout this work we will use binomial distributions for these probability mass functions.
-They depend only on a single parameter, which when multiplied with the number of considered time-steps, conveniently also represents the distribution's mean.
+They depend only on a single parameter, which when multiplied with the number of considered time-steps, conveniently also represents the distribution's mean.\footnote{In our previous publication on the lymphatic progression model, we present some experiments showing that the shape of the time-prior and number of time-steps have no impact as long as only one T-category is considered \cite{ludwig_hidden_2021}.}
 %
@@
 However, our data suggests that at least contralateral \gls{lnl} II is frequently involved in \gls{opscc} patients.
-It is also of interest clinically to be able to decide when the contralateral side may be spared entirely, as this could meaningfully reduce side effects such as xerostomia and dysphagia [citation needed].
-Therefore extending the model to include some or all contralateral \glspl{lnl} would represent an important addition on its way to a clinical validation and ultimately implementation.
+It is also of interest clinically as it presents the opportunity for further reducing the \gls{ctv-n} when the predicted risk is low enough.
+Therefore extending the model to include the contralateral side could be an important next step on its way to a clinical validation and ultimately implementation.
 

### 2024-05-23 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2024-05-16 src/tex/ms.tex
@@
 In clinical practice, these modalities are \gls{ct}, \gls{mri}, or \gls{pet} scan, but it may also include information from biopsies after a \gls{fna} or other techniques to detect lymphatic metastases.
-For each \gls{lnl} $v$ the conditional probability table of $P\left( Z_v \mid X_v \right)$ looks like this:
+For each \gls{lnl} $v$ the conditional probability table of $P\left( Z_v \mid X_v \right)$ is given by:
 
@@
 However, using this definition, we cannot know how many time-steps have passed until $t_\text{D}$, when the patient was diagnosed with cancer.
-We can only make the assumption that a patients with an earlier T-category tumor was \emph{probably} diagnosed after fewer time-steps than a patient with a advanced T-category tumor.
+We can only make the assumption that a patient with an earlier T-category tumor was \emph{probably} diagnosed after fewer time-steps than a patient with a advanced T-category tumor.
 
@@
 Throughout this work we will use binomial distributions for these probability mass functions.
-They depend only on a single parameter, which when multiplied with the number of considered time-steps, conveniently also represents the distribution's mean.
 %
@@
 Here, the parameter $p_{\text{T}x}$ can be interpreted as the probability that the patient with a tumor of T-category $x$ will be diagnosed at time-step $t+1$ given they are in time-step $t$.
-We will use as the latest time-step $t_\text{max} = 10$, which will therefore give us a distribution over the diagnosis time that has its mean at $\mathbb{E}[t_\text{D}] = 10 \cdot p_{\text{T}x}$.
+We will use as the latest time-step $t_\text{max} = 10$.
+Binomial distributions depend only on a single parameter, which when multiplied with the number of considered time-steps, conveniently also represents the distribution's mean.\footnote{In our previous publication on the lymphatic progression model, we have shown that the shape of the time-prior and number of time-steps have no impact as long as only one T-category is considered \cite{ludwig_hidden_2021}.} $t_\text{max}=10$ will therefore give us a distribution over the diagnosis time that has its mean at $\mathbb{E}[t_\text{D}] = 10 \cdot p_{\text{T}x}$.
 
@@
         \caption{Prevalence of involvement as predicted by the base graph model for different scenarios involving the most commonly metastatic \glspl{lnl} II, III, and IV (shaded histograms).
-The model's predictions are compared to Beta posteriors over the prevalence based on the frequency of the same scenarios in the data and a uniform prior (solid lines).
+The model's predictions are compared to Beta posteriors over the prevalence based on the frequency of the same scenarios in the data assuming a uniform prior (solid lines).
+E.g., out 423 early T-category patients, 100 (23.6\%) presented with involvement of the \glspl{lnl} II and III, as shown by the blue line in the uppermost panel.
 The top panels of each of the three subfigures shows some selected scenarios with early T-category tumors and the bottom panel the same scenarios for advanced T-category.
@@
 
-{\it Contralateral Spread:} So far, our model concerns itself with the ipsilateral lymphatic spread.
-However, our data suggests that at least contralateral \gls{lnl} II is frequently involved in \gls{opscc} patients.
-It is also of interest clinically to be able to decide when the contralateral side may be spared entirely, as this could meaningfully reduce side effects such as xerostomia and dysphagia [citation needed].
-Therefore extending the model to include some or all contralateral \glspl{lnl} would represent an important addition on its way to a clinical validation and ultimately implementation.
+{\it Contralateral Spread:} The work presented in this paper considers only ipsilateral lymphatic spread.
+To guide the elective CTV-N definition for the contralateral neck, the model must be extended to the contralateral side.
+Contralateral lymph node involvement is strongly dependent on whether the primary tumor extends over the midsaggital plane, but also on T-category and the extent of ipsilateral involvement \cite{ludwig_detailed_2022}.
+A comprehensive model accounting for these risk factors on contralateral spread is the subject of a follow-up publication.
 

### 2024-05-23 src/scripts/bg_core_prevs.py
consistently color core prevalences

@@ -20,7 +20,7 @@ OUTPUT = (paths.figures / Path(__file__).name).with_suffix(".png")
 NROWS, NCOLS = 6, 1
 LNLS_COMBOS = list(product(["II", "III", "IV"], ["and", "not"], ["II", "III", "IV"]))
 LNLS_COMBOS = [*LNLS_COMBOS, "II", "III", "IV"]
-COMBOS_TO_SKIP = ["IVandII", "IIInotIV"]
+COMBOS_TO_SKIP = ["IVandII"]
 
 
 def get_lnl_label(match):
@@ -29,6 +29,26 @@ def get_lnl_label(match):
         return f"LNL {match[1]} {filler} {match[3]}"
 
     return f"LNL {match[1]} overall"
+
+
+def get_color(match) -> str:
+    if match[0] == "IVnotII":
+        return "black"
+
+    if match[2] is None:
+        return COLORS["red"]
+
+    if match[2] == "and":
+        return COLORS["blue"]
+
+    if match[2] == "not":
+        if (
+            (match[1] == "II" and match[3] == "III")
+            or (match[1] == "III" and match[3] == "IV")
+        ):
+            return COLORS["orange"]
+
+    return COLORS["green"]
 
 
 def draw_early_late_panels(panels, lnl, xlims, ylims=None, idx=0):
@@ -84,7 +104,7 @@ def create_legends(ax, positions):
         ax.add_artist(legend)
 
 
-if __name__ == "__main__":
+def main():
     plt.rcParams.update(figsizes.icml2022_full(
         nrows=NROWS,
         ncols=NCOLS,
@@ -94,12 +114,6 @@ if __name__ == "__main__":
     fig = plt.figure()
 
     panels = {lnl: {"early": [], "late": []} for lnl in ["II", "III", "IV"]}
-    colors = {
-        lnl: {
-            "early": cycle(COLORS.values()),
-            "late": cycle(COLORS.values()),
-        } for lnl in ["II", "III", "IV"]
-    }
     for lnl_combo in LNLS_COMBOS:
         lnl_combo = "".join(lnl_combo)
 
@@ -119,29 +133,47 @@ if __name__ == "__main__":
                     filename=INPUT,
                     dataname=f"{lnl_combo}/{stage}",
                 )
-                color = next(colors[lnl_block][stage])
+                color = get_color(match)
                 _histo.kwargs["color"] = color
                 _post.kwargs["color"] = color
[…]

### 2024-05-23 src/scripts/thermo_int.py
change red/green to red/blue in thermo int

@@ -55,7 +55,7 @@ if __name__ == "__main__":
         x="β",
         y=["lose graph", "win graph"],
         ax=axes,
-        color=[USZ["red"], USZ["green"]],
+        color=[USZ["red"], USZ["blue"]],
         label=["remove II → III & III → IV", "add I → II & IV → V"],
     )
 

### 2024-05-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     \includegraphics[width=\linewidth]{figures/thermo_int.png}
-    \caption{The expected log-accuracy under the power posterior plotted against the value of the inverse temperature $\beta$ for the winning graph, computed during thermodynamic integration (top panel).
-The are under this curve yields the winning graph's model evidence.
-Also shown is the difference of the winning graph's accuracy over the base graph's (bottom panel).
+    \caption{The expected log-accuracy under the power posterior plotted against the value of the inverse temperature $\beta$ for the graph with the best (blue) and worst (red) log-evidence.
+The are under this curve yields the respective graph's model evidence.
 The scaling of the x-axis is chosen such that the 64 points on the temperature ladder appear evenly spaced.
@@
 E.g., out 423 early T-category patients, 100 (23.6\%) presented with involvement of the \glspl{lnl} II and III, as shown by the blue line in the uppermost panel.
-The top panels of each of the three subfigures shows some selected scenarios with early T-category tumors and the bottom panel the same scenarios for advanced T-category.
-These figures show that for the most important \glspl{lnl} II, III, and IV the base graph model already fits the data well.
+The top panels of each of the three subfigures show some selected scenarios with early T-category tumors and the bottom panel the same scenarios for advanced T-category.
+Similar scenarios are color-coded: Blue for joint involvement with downstream \gls{lnl}, orange for a level's involvement without the downstream \gls{lnl}, red for an \gls{lnl}'s overall involvement prevalence, and green for involvement without direct upstream metastases.
+The black scenario in the bottom subfigures only appears once and did not fit into these categories.
+This figure shows that for the most important \glspl{lnl} II, III, and IV the base graph model already fits the data well.
 \label{fig:bg_prevalences}}
@@
     \hline
-    \textbf{graph}  & \textbf{log-evidence} \\
+    \textbf{graph}  & $\Delta$ \textbf{log-evidence} \\
     \hline
@@
 \end{tabular}
-\caption{Model comparison results from the base graph and the extended graph we chose as the ``winning'' model.
-For both \glspl{dag} we show the log-evidence, computed via thermodynamic integration, the negative one half of the \gls{bic}, as well as the maximum log-likelihood that was encountered during the final \gls{mcmc} sampling round.}
+\caption{Model comparison results for all compared graph structures.
+For all \glspl{dag} we show the difference of the log-evidence to the base graph, as in \cref{fig:ranking}, computed via thermodynamic integration.}
 \label{table:evidence}
